### 2023-08-08 pool_model.tex
create draft of the paper repository

\documentclass[10pt,A4paper]{article}
\usepackage[a4paper,
            inner=2.5cm,
            outer=3cm
            ]{geometry}
\input{additional_tex/packages}

\title{Pool Model Paper}
\author{Polina Gaindrik}

\begin{document}
\maketitle

\tableofcontents

\newpage

\section{Introduction}

...

- See Papers.

- Different processes in bacterial communities. 
Differents states, lag-phase, growth, stationary state. What processes behind.
Activation/inhibition/waste production, resource competition, back-flow to lag-phase.

- About known models for Predictive Microbiology Modelling, e.g., Baranyi/Roberts

\section{Motivating Example}

...

\cmtb{Maybe Example from Seminar by Jens? 2 species,  1 resource, 1 waste production.}

\section{General Mathemetical Formulation}
Consider the system conssisting of $N$ species in $M$ different states.
In the general representation the system can be described with the Ordinary Differential Equations (ODEs) using a non-linear function $\mbF$:
\begin{equation}
   \dot{\mbx} = \mbF(\mbx, \mbm, \mbp, \mbu),
   \label{eq:model_ODE}
\end{equation}
where  $\mbx \in \mathbb{R}^{N}  \otimes \mathbb{R}^{M}$ is a vector descibing bacterial count.
In a matrix form it can be written as
\begin{equation}
    \mbx = \begin{pmatrix}
        x^1_1  & \dots & x^1_M  \\
        \vdots &       & \vdots \\
        x^N_1  & \dots & x^N_M  \\
            \end{pmatrix}
    \label{eq:model_bact}
\end{equation}
with $x_{ij}$ describing the bacterial counts for species $i$ in state $j$.
$\mbu$ is the vector of different inputs of the model, e.g., temperature, pressure etc.
The vector $\mbm \in \mathbb{R}^{K}$ includes all variables corresponding to the microenvironment of the system which can be concentartion of the inhibitors or activators:
\begin{equation}
    \dot{\mbm} = \mbH (\mbx, \mbm, \mbp, \mbu, ...).
    \label{eq:model_microenv}
\end{equation}
Here function $\mbH$ is in general case also non-linear.

The advantage of the following formulation is its generality. 
Depending on the knowledge about system as well as the needed level of accuracy one can define arbitrary number of states $M$ and control the complexity of teh model.
Moreover, opposite to the already known Predictive Microbiology like Baranyi and Roberts one can easily change.

To calculate the bacterial count for one species $i$, we simply need to sum the tensor/vector $\mbx$ over all possible states:
\begin{equation}
    N_n = \sum_m \mbx_i^n.
\end{equation}
While the total bacterial count of all species of our system is
\begin{equation}
    N_t = \sum_j \sum_m \mbx_i^j.
\end{equation}
In the following section we demonstate the application of this approach for different toy models describing the processes within and between bacteria communities.

\newpage

\section{One Species Modelling}
\subsection{Three-Pool Model}

Consider for simplicity the system where only one bacteria species is present $N=1$.
According to the growth curves (\cmt{some references on bacteria growth curves??}) quite often at the beginning no increase in bacterial count is present and so-called lag-phase is observed.
This phenomenon appears due to adaptation of the microbial community, for example, to the new environmental conditions.
After this some bacteria enter the exponential growth phase, while some of it may not be able to adjust and experience the cell death.
Some bacteria even may not be able to adjust and die other will enter the exponential growth phase after a lag phase. 
After exponential growth the system reaches the saturation where the balance between the number of dying bacteria is in balance with the growing one leads to the stationary state.
In the described simplified version of the real-life system three possible bacteria states are mentioned corresponding to which process is happeng: adaplation, growth or death.
Motivated by this we divide a bacterial population into three pools (states) (\textbf{Figure \ref{SchematicRep}}): 
\begin{itemize}
\item $L(t)$: the fraction of the population in the lag phase
\item $G(t)$: the fraction of the population growing and dividing
\item $D(t)$: the fraction of the population undergoing cell death
\end{itemize}
Here the concentrations of the pools are components of our variable vector $\mbx (t) =\{L(t), G(t), D(t)\}$.

\begin{figure}[t]
    \begin{center}
    \includegraphics[width=0.7\textwidth]{Figures/TPM_fig.pdf}
    \caption{Schematic representation of exchange between three states or pools.}
    \label{SchematicRep}
    \end{center}
    \end{figure}

To detremine the function $\mbF(\mbx, \mbm, \mbp, \mbu)$ consider the flows between different states.
The pools $G$ and $D$ are unidirectionally connected with the pool $L$. 
For now we assume for simplicity that there is no back flow between growth state $G$ and lag-state $L$. 
The suggested model can be cast in the following reactions:
\begin{eqnarray}
L &\stackrel{\lambda}{\longrightarrow} & G\\
L &\stackrel{\mu}{\longrightarrow} & D\\
G &\stackrel{\alpha}{\longrightarrow} & 2G\\
G &\stackrel{\mu'}{\longrightarrow} & D\\
D &\stackrel{\beta}{\longrightarrow} & \emptyset
\end{eqnarray}.

Then the ODEs describing such a model can be written as 
\begin{equation}
\begin{cases}
    \dot{L} = -(\mu + \lambda) L\\
    \dot{G} = \lambda L + \alpha G - \mu' G\\
    \dot{D} = \mu  L + \mu' G- \beta D  
\end{cases}.
\end{equation}

\cmt{Add plots ?? Or already after resource is added?}

\subsubsection{Resource Dependence}

Based on experimental observations, it was shown that the bacterial growth is strongly influenced by the limiting resource. \cmt{ref review}
In the system where all needed nutritients are in excess except for one the total growth is linear to the conentration of this liminting nutritient. \cmt{ref Monod}
\cmt{Ref about Quorum sensing. Signalling mechanism that controll gene expression (bacteria growth within species) in response to environment. }

According to this, we aim to include a resource pool which is used by the growing bacteria using the second order kinetic reaction:
\begin{equation}
    G + R_0  \stackrel{\alpha_0}{\longrightarrow} 2G.
\end{equation}

Note that the pools $L$, $G$ and $D$ represent number of bacteria (or number of bacteria per unit area/volume) while $R_0$ represents an abstract resource pool. 
Let say $G=N_B/V$ and $[G]=1/V$, where $N_B$ is number of bacteria and $V$ is the volume. 
Then $R_0=N_R/V$ and $[R_0]=1/V$, where $N_R$ is the number of resource molecules (or any appropriate unit, e.g., mol). 
Then $[\alpha_0]=V/s$, $[N_t]=1/V$, and $[\alpha]=1/s$. 
For homogeneous densities it is equivalent to consider the total amount, i.e., multiplying $L$ and $G$ by the volume. 
In this case $[L]=[G]=[N_t]=1$.
For simplicity we rather make a transition to the dimensionless constant $\alpha:=\alpha_0 N_t$ and diminsioneless resource pool $R(t) = R_0 / N_t$ ($0 \leqslant R \leqslant 1$).
Here the full capacity of the resource pool $N_t$ was used as a scaling constant.
Then the growth term can be simply written as $\alpha G R$, and the microenvironment of the system introduced by \ref{eq:model_microenv} is defined by ODE:
\begin{equation}
\dot{R} = -\frac{\alpha}{N_t} R G
\end{equation}

Explointing that limitation of the resource pool also limits the total becterial count of the system  we find:
\begin{equation}
    R = 1 - \frac{L+G+D}{N_t}
\end{equation}
And, hence, the full system can be written as
\begin{equation}
    \begin{cases}
        \dot{L} = -(\mu + \lambda) L\\
        \dot{G} = \lambda L + \alpha G\left(1-\frac{L+G+D}{N_t}\right)-\mu' G\\
        \dot{D} = \mu  L + \mu' G- \beta D  
    \end{cases}.
\end{equation}

\cmt{Plots / results}

\subsubsection{Lag-time Estimation (??)}

Note that one could further simplify this model to a two pool or state model by omitting the pool $D$ and removing cells with rate $\beta$ directly from pool $L$. 
The relevant question here is whether cells in the pool $D$ still contribute to the limitation of the resources by, e.g., occupying space or consuming important nutrients. 
If yes, the three pool model would be advantageous, if not the two pool model is sufficient with the advantages of having one parameter less.\\

The dynamics of the three pool model can be captures by the coupled system of ordinary differential equations (the two pool model can be achieved by setting $D\equiv 0$ and replacing $\mu$ with $\beta$):
to be solved with the initial conditions $L(0)=n_0$, $G(0)=0$, and $D(0)=0$, where $n_0$ is the initial bacterial population. 
$N_t$ denotes the maximal size of the total bacterial population due to environmental limitations (\textbf{Figure \ref{GrowthDip}}). 
For $\mu'=0$ the equations for $L$ and $D$ can be readily integrated resulting in:
\begin{eqnarray}
L(t) &=& n_0 e^{-(\mu+\lambda)t}\\
D(t) &=& n_0 \mu e^{-\beta t}\frac{1-e^{-(\mu+\lambda-\beta)t}}{\mu+\lambda-\beta}.
\end{eqnarray}
For constant parameters $\mu$, $\lambda$, and $\beta$ the integrals for $L$ and $D$ can be put into the differential equation for $G$. 
However, for time dependent parameters, e.g. for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$. 
It is interesting to see what type of equations result for the total bacterial population $n=L+G+D$:
\begin{eqnarray}
\dot{G} &=& \lambda L + \alpha G\left(1-\frac{n}{N_t}\right)\\
\dot{n} &=& - \beta D + \alpha G\left(1-\frac{n}{N_t}\right). 
\end{eqnarray}
with the initial conditions $G(0)=0$ and $n(0)=n_0$. 
These equations have a clear interpretation, are derived from simple principles, and are different from the popular Barayni-Roberts and modified-Gompertz models. 
A notable difference is that the growth rate $\alpha$ is time independent. 
For the functions $L$ and $D$ it holds: $\lim_{t\to\infty} L(t) = \lim_{t\to\infty} D(t) = 0$. 
It follows immediately that $\lim_{t\to\infty} G(t) = N_t$. 
The equation exhibit a lag phase and an initial drop due to cell death. 
To find a substantial initial drop one needs $\beta \mu \gg \alpha \lambda$. 
The lag time $t_L$ can be approximated by the time at which half of the initial population is in the growth phase, i.e., $n_0=2G(t_L)$. 
Assuming $L\ll N_t$ we can ignore the non-linear term in the equation for $G$, which gives rise to:
\begin{eqnarray}
1&=& 2\int_0^{t_L} \lambda L(t)e^{\alpha(t_L-t)}dt\\
1  &=& 2\lambda e^{\alpha t_L}\frac{1-e^{-(\mu+\lambda+\alpha)t_L}}{\mu+\lambda+\alpha}.
\end{eqnarray}
For $\alpha > \mu+\lambda$ $t_L$ can be approximated by:
\begin{eqnarray}
t_L &\approx& -\frac{1}{\alpha}\ln\left(\frac{2\lambda}{\mu+\lambda+\alpha}\right).
\end{eqnarray}
The lag time $t_L$ diverges very slowly, logarithmically with $\lambda\to 0$.

\cmt{Add/make plots to the model 3 pool/ 2 pool.}

\cmt{Add limitation due to resource}

\subsection{Resource-dependent lag-phase}

Another way to encorporate the resource dependence is to assume that the bacteria only leave the lag-phase if there are nutrients, i.e., resources. 
Otherwise they stay in the lag-phase.
Then using the model consisted of two bacteria states $L$ and $G$ we can define the system as
\begin{eqnarray}
\dot{L} &=& - \lambda R L\\
\dot{G} &=& \lambda R L +\alpha R G\\
\dot{R} &=&-\frac{\alpha}{N_t} R G.
\end{eqnarray}

Turning back to the equations above we can state: $L+G+N_tR=N_t$. For $t=0$ we have $n_0+N_tR_0=N_t$ (given $L_0=n_0$ and $G_0=0$) and from this: $N_tR_0=N_t-n_0$. Using this we eliminate $R$ and arrive at: 
\begin{eqnarray}
\dot{L} &=& - \lambda \left(1-\frac{L+G}{N_t}\right)L\\
\dot{G} &=& \lambda \left(1-\frac{L+G}{N_t}\right)L +\alpha \left(1-\frac{L+G}{N_t}\right) G
\end{eqnarray}
The steady state is $L_s+G_s=N_t$. However, it depends on the initial conditions and the parameters what $L_s$ and $G_s$ are. There is no way to determine $L_s$ and $G_s$ besides solving the dynamical equations, i.e., the steady state cannot be determined directly by solving a set of algebraic equations. To see this we determine the nullclines: the first nullcline is $G=N_t-L$ and the second is $G=N_t-L$, i.e., the nullclines coincides. Every point on the line $G=N_t-L$ is a fixed point. 
This can be seen in Fig. \ref{Phaseplot}. 
For $\lambda\ll \alpha$ the lag-pool is not emptied before all resources are consumed. As a consequence the lag-pool is non-zero in steady-state. 
In contrast, for $\lambda\gg \alpha$ the lag-pool is depleted before the resources are consumed, i.e., the lag-pool is zero in steady state.

The fraction $G$ of the exponentially growing cells will go to steady state due to limitations in resources. 
$\lambda$ is the rate by which cells go from the lag phase to the growth phase; $\mu$ is the rate by which cells go from the lag phase to the state in which they die; $\alpha$ is the growth rate of the healthy population; $\beta$ is the rate by which the cells are removed from the $D$ pool.

\cmt{Plots / results}

\subsection{Back flow from G to L}

We explore now the idea that under certain conditions, as sudden change in the environment, part of the population $G$ will enter a lag-phase, i.e., there is a back-flow from $G$ to $L$. 
\begin{eqnarray}
L &\stackrel{\lambda}{\longrightarrow} & G\\
L &\stackrel{\mu}{\longrightarrow} & D\\
G &\stackrel{\alpha}{\longrightarrow} & 2G\\
G &\stackrel{\gamma(t)}{\longrightarrow} & L\\
D &\stackrel{\beta}{\longrightarrow} & \emptyset
\end{eqnarray}
The rate $\gamma(t)$ depends on the rate of change of the environment. Stronger and quicker changes will lead to a higher rate $\gamma$. 
Let the change in the environment be in temperature $T$ (\textbf{Figure \ref{default}}).

\subsubsection{Maxwell type of stress - strain relation}
We propose a kind of visco-elastic Maxwell type of stress-strain relation and write down the ordinary differential equation for $\gamma$ to be:
\begin{eqnarray}
\dot{\gamma} &=& \Gamma \left |\dot{T}\right |-\delta \gamma.
\end{eqnarray}
We assumed that the direction of the temperature change does not matter, i.e., a change from $T=2$ to $T=12$ has the same effect as a change from $T=14$ to $T=4$. 
$\Gamma$ is a scaling factor and $\delta$ is the relaxation time for the temperature disturbance. 
Integration of this equation yields (with $\gamma(0)=0$):
\begin{eqnarray}
\gamma(t) &=& \Gamma \int_0^t \left |\dot{T}(t^{\prime})\right |e^{-\delta (t-t^{\prime})}dt^{\prime}.
\end{eqnarray}
After some manipulation of this equation we find:
\begin{eqnarray}
\gamma(t) &=& \Gamma\delta\left |\left[T[t]-T[0]e^{-\delta t}-\delta \int_0^t T(t^{\prime})e^{-\delta (t-t^{\prime})}dt^{\prime}\right]\right |.
\end{eqnarray}
This representation of the rate $\gamma(t)$ has the advantage that it only involves $T(t)$ and does not require the calculation of the derivative of $T$. 
If $T$ exhibits $n$ temperature jumps, one finds:
\begin{eqnarray}
\gamma(t) &=& \Gamma\sum_{i=1}^n \left |\Delta T_i \right |\theta(t-t_i)e^{-\delta(t-t_i)}.
\end{eqnarray}

The ordinary differential equations for the pool model with environmental-shock based back-flow read:
\begin{eqnarray}
\dot{L} &=&\gamma(t) G - (\mu + \lambda) L\\
\dot{G} &=& \lambda L -\gamma(t) G + \alpha G\left(1-\frac{L+G+D}{N_t}\right)\\
\dot{D} &=& \mu  L - \beta D
\end{eqnarray}

-  \cmt{Ref to temperature jumps effect. Plots / results.}

\newpage

\section{Interaction between species}

As it was mentioned above, bacteria usually coexist in large multispecies interacting communities whose growth is vastly determined by the developed survival mechanisms.
These survival mechanisms mostly aim to secure as much available nutrition which lead to competitive strategies or, if it is advantageous, cooperative ones.
Further we show that the TPM can be readily extended to the systems with more than one bacteria an take describe more complex relationships between species.

\subsection{Resource Competition}

One on the ways to descibe competition between bacteria populations is as a function of limiting resources ratio. \cmt{Ref Tilman 1977, Ref Smith 2002}
To descibe such a situation we consider for simplicity only one common limiting resource pool. 
The simplest pool model for bacterial species $A$ and $B$ encompassing lag-phases, growth-phases and a common resource pool is given by (no back-flow to the lag-pool):

\begin{eqnarray}
\label{Model_Comp1}
\dot{L}_A &=& - \lambda_A R L_A\\
\dot{L}_B &=& - \lambda_B R L_B \\
\dot{G}_A &=& \lambda_A R L_A +\alpha_A R G_A\\
\label{Model_Comp4}
\dot{G}_B &=& \lambda_B R L_B +\alpha_B R G_B\\
\dot{R} &=&-\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B
\end{eqnarray}
We again rescaled the resource pool with $N_t$ and find
\begin{eqnarray}
\label{Resource}
R &=&1-\frac{L_A+G_A+L_B+G_B}{N_t}\\
R_0&=&1-\frac{n_A+n_B}{N_t}. 
\end{eqnarray}
Inserting Eq. \ref{Resource} into Eqs. \ref{Model_Comp1} - \ref{Model_Comp4} yields:
\begin{eqnarray*}
\dot{L}_A &=& - \lambda_A  L_A + \frac{\lambda_A}{N_t}\left(L_A+G_A+L_B+G_B\right )L_A\\
\dot{L}_B &=& - \lambda_B L_B + \frac{\lambda_B}{N_t}\left(L_A+G_A+L_B+G_B\right )L_B \\
\dot{G}_A &=&  \lambda_A  L_A - \frac{\lambda_A}{N_t}\left(L_A+G_A+L_B+G_B\right )L_A +\alpha_A G_A - \frac{\alpha_A}{N_t}\left(L_A+G_A+L_B+G_B\right )G_A\\
\dot{G}_B &=& \lambda_B L_B - \frac{\lambda_B}{N_t}\left(L_A+G_A+L_B+G_B\right )L_B  +\alpha_B G_B -\frac{\alpha_B}{N_t}\left(L_A+G_A+L_B+G_B\right )G_B
\end{eqnarray*}

which has the structure of a generalised Lotka-Volterra system. In order to better see this, we set $L_A=L_B=0$: 
\begin{eqnarray}
\label{Comp_Model}
\dot{G}_A &=& \alpha_A G_A\left(1 - \frac{G_A}{N_t}\right ) - \frac{\alpha_A}{N_t}G_AG_B\\
\dot{G}_B &=& \alpha_B G_B\left(1-\frac{G_B}{N_t}\right ) -\frac{\alpha_B}{N_t}G_AG_B. 
\end{eqnarray}

For $N_t\gg\alpha_{A/B}$ the equations decouple. 
The null-clines of the system given by Eqs. \ref{Model_Comp1} - \ref{Model_Comp4} coincide and are given by $N_t=L_A+G_A+L_B+G_B$. 
If $\lambda_{A/B}>\alpha_{A/B}$ the lag-pools are completely depleted before the depletion of the resource pool stops the dynamics. 
To simplify the investigation of the systems dynamics we therefore ignore the $L_A$ and $L_B$ dimension and only consider the dynamics on the centre-manifold given by the $G_A\times G_B$ plane. 
The corresponding phase-plot is shown in Fig. \ref{Competition}. 
The advantage in using the pool model in contrast to using Lotka-Volterra directly is that all parameters have a clear meaning and consistent parameter relations arrive naturally. 
And a species can have more than one internal state. 
However, it is instructive to rewrite the system by setting $x_1\equiv G_A$, $x_2\equiv G_B$, and $\mathbf{x}=(x_1,x_2)^T$:

\begin{eqnarray}
\dot{\mathbf{x}}&=&A\mathbf{x}+\frac{1}{2}\mathbf{x}B\mathbf{x}\\
A&=&
\begin{bmatrix}
\alpha_A&0\\
0 &\alpha_B
\end{bmatrix}\\
B&=&
\frac{-2}{N_t}\begin{bmatrix}
\alpha_A&\alpha_A\\
\alpha_B &\alpha_B
\end{bmatrix}
\end{eqnarray}

The pool model, in contrast to the classical Lotka-Volterra approach, allows for far more flexibility. 
Non-linear functional responses can be introduced without any problem (see how section \ref{Gen_Lot_Vol} for an extended discussion).

\cmt{Plots / results}

\subsection{Interspecies Competition}

However, except the inhibition due to resource competition, bacteria developed differnt others active competition mechanisms.
One of the examples of such mechanisms is a production of the antimicrobial compounds, which can be also called antibiotics or inhibitors.

\cmt{Read how this antibiotics affect the system. Which eq form to use?}

Now extend our model where bacteria not only consume nutritions from resource pool $R$ but one of the species also actively produce the inhibiting compounds $I$. 
It is the accumulation of the bacteria waste which is responsible for the spoilage of food products. 
In order to take this into account we add an equation for the waste production:
\begin{eqnarray}
\dot{L} &=& - \lambda_f R L + \mathcal{B}(R)G\\
\dot{G} &=& \lambda_f R L +\alpha R G-\mathcal{B}(R)G\\
\dot{R} &=&-\frac{\alpha}{N_t} R G\\
\dot{I} &=& \mathcal{F}(R,G).
\end{eqnarray}
The function $\mathcal{F}$ denotes the model for the antibiotic production rate. 
We shall discuss two possible models:
\begin{eqnarray}
\mathcal{F}_1(R,G)&=&\kappa\alpha R G=-\kappa N_t\dot{R}\\
\mathcal{F}_2(R,G)&=&\kappa G
\end{eqnarray}
Model $\mathcal{F}_1$ results in waste accumulation in steady state $I_1^s=\lim_{t\to\infty}I_1(t)=\kappa(N_t-n_0)$. 
To see this we note $\dot{I}_1=\kappa(\dot{L}+\dot{G})$, which yields: $I_1(t)=\kappa(L(t)+G(t)-n_0)$ where we used $I_1(0)=G(0)=0$, $L(0)=n_0$. 
This model assumes that inhibitor is only produced when the bacteria grow and divide. \\

Model $\mathcal{F}_2$ results in $I_2(t)=\kappa\int_{t_0}^tG(t')dt'$. 
As long as $G>0$ inhibitor is produced, no matter whether the amount of bacteria is still growing. 
The amount of inhibitor depends therefore on the back-flow model employed, which can be seen in Fig. \ref{Waste}.

\cmt{Look is this inhibition the same as waste production. Or change it.}

The simplest way to introduce inhibition of species A by species B is via a growth rate of species A depending on the abundance of species B. 
For sake of simplicity we ignore the lag pool and write for a mutual inhibition:
\begin{eqnarray}
\dot{G}_A &=& \mathcal{B}_A(G_B)G_A\left(1 - \frac{G_A+G_B}{N_t}\right )\\
\dot{G}_B &=& \mathcal{B}_B(G_A) G_B\left(1-\frac{G_A+G_B}{N_t}\right ). 
\end{eqnarray}
The function $\mathcal{B}_{A/B}$ captures the inhibiting effects in an effective way. It is straight forward to make the mechanism of inhibition more explicit by, e.g., introducing a waste pool. One possible choice for $\mathcal{B}$ is: 
\begin{eqnarray}
\mathcal{B}(G)&=&\frac{\alpha}{1+KG}. 
\end{eqnarray}
This corresponds to the case $W\sim G$, where $W$ is the waste (or whatever one chooses to call it) of species A affecting species B. If the waste accumulates one needs to replace $G$ by $\int G dt$. Keeping it simple yields:
\begin{eqnarray}
\dot{G}_A &=& \frac{\alpha_A G_A}{1+K_BG_B}\left(1 - \frac{G_A+G_B}{N_t}\right )\\
\dot{G}_B &=& \frac{\alpha_B G_B}{1+K_AG_A}\left(1-\frac{G_A+G_B}{N_t}\right ). 
\end{eqnarray}
One can distinguish two limiting cases: {\it i}) $K_{A/B}\ll N_t$ and {\it ii}) $K_{A/B}\gg N_t$. In the first case the inhibition is of minor importance and could be ignored. In the latter case one can simplify the equations, given $n_A$ and $n_B$ are non-zero:
\begin{eqnarray}
\label{Mut_Inhib}
\dot{G}_A &=&\tilde{\alpha}_A\frac{G_A}{G_B}\left(1 - G_A-G_B\right )\\
\dot{G}_B &=& \tilde{\alpha}_B\frac{G_B}{G_A}\left(1-G_A-G_B\right ). 
\end{eqnarray}
We introduced the rescaled growth rates $\tilde{\alpha}_A=\alpha_A/(N_tK_B)$ and $\tilde{\alpha}_B=\alpha_B/(N_tK_A)$. Note that $[\tilde{\alpha}]=\mathrm{s}^{-1}$. Rescaling time with $\tilde{\alpha}_A$ and defining the dimensionless parameter 
\begin{eqnarray}
\psi&=&\tilde{\alpha}_B/\tilde{\alpha}_A\\
\psi&=&\frac{K_B\alpha_B}{K_A\alpha_A}
\end{eqnarray}
results in: 
\begin{eqnarray}
\dot{G}_A &=&\frac{G_A}{G_B}\left(1 - G_A-G_B\right )\\
\dot{G}_B &=&\psi\frac{G_B}{G_A}\left(1-G_A-G_B\right ). 
\end{eqnarray}
We can separate three cases: 
\begin{eqnarray}
\psi&>&1:\quad\text{Species B growth faster and/or inhibits A stronger}\\
\psi&=&1:\quad\text{A and B are equivalent}\\
\psi&<&1:\quad\text{Species A growth faster and/or inhibits B stronger}
\end{eqnarray}

\subsection{Interspecies Cooperation}

On the other hand, one can also observe some cooperative behaviours between species in case when it facilitates the competition with others.
Quite often it happens between the species that are close relatives or clonal populations and such process is called kin selection. \cmt{Ref 25 from review}
Sometimes neighbours, so-called 'social cheaters', can benefit from a certain behaviour without expressing it by itself.  \cmt{Ref 29 from review}

The benefit, or mutual activation, of the two species from the presence of the other can be simply introduced by mathematical functions. 
For sake of clarity we omit again the lag pool and write:
\begin{eqnarray}
\dot{G}_A &=& \mathcal{B}_A(W_B,R)G_A\\
\dot{G}_B &=& \mathcal{B}_B(W_A,R)G_B\\
\dot{R} &=&-\mathcal{B}_A(W_B,R)G_A-\mathcal{B}_B(W_A,R)G_B\\
\dot{W}_A &=& \mathcal{F}_A(R,G_A)\\
\dot{W}_B &=& \mathcal{F}_B(R,G_B).
\end{eqnarray}
$W_A$ and $W_B$ denote the substance produced by species A and B, respectively. The functions $\mathcal{B}_A(W_B,R)$ and $\mathcal{B}_B(W_A,R)$ capture the resource dependent growth. Setting $\mathcal{B}_A(W_B,R)=\alpha_A W_BR$, $\mathcal{B}_B(W_A,R)=\alpha_B W_AR$, $\mathcal{F}_A(R,G_A)=\kappa_A RG_A$, $\mathcal{F}_B(R,G_B)=\kappa_B RG_B$, rescaling $G_A$, $G_B$, and $W_B$ with $N_t$, $W_A$ with $\kappa_A/\alpha_A$, time with $\alpha_AN_t^2$, defining $\phi=\kappa_A\alpha_B/(\alpha_A^2N_t)$, and $\psi=\kappa_B/(\alpha_A N_t)$ results in: 
\begin{eqnarray}
\dot{G}_A &=& W_BG_A\left(1 - G_A-G_B\right )\\
\dot{G}_B &=& \phi W_AG_B\left(1 - G_A-G_B\right )\\
\dot{W}_A &=& G_A\left(1 - G_A-G_B\right )\\
\dot{W}_B &=&\psi G_B\left(1 - G_A-G_B\right ).
\end{eqnarray}

\section{Conclusion}

\end{document}

### 2023-08-08 pool_model.tex
small changes
add citations

@@
 Consider for simplicity the system where only one bacteria species is present $N=1$.
-According to the growth curves (\cmt{some references on bacteria growth curves??}) quite often at the beginning no increase in bacterial count is present and so-called lag-phase is observed.
+According to the growth curves (\cmt{some references on bacteria growth curves??}) \cite{buchananLifePhasesBacterial1918, rolfeLagPhaseDistinct2012} quite often at the beginning no increase in bacterial count is present and so-called lag-phase is observed.
 This phenomenon appears due to adaptation of the microbial community, for example, to the new environmental conditions.
+\cmt{\cite{rolfeLagPhaseDistinct2012}: About Lag-phase machanisms.}
 After this some bacteria enter the exponential growth phase, while some of it may not be able to adjust and experience the cell death.
@@
 In the system where all needed nutritients are in excess except for one the total growth is linear to the conentration of this liminting nutritient. \cmt{ref Monod}
-\cmt{Ref about Quorum sensing. Signalling mechanism that controll gene expression (bacteria growth within species) in response to environment. }
+\cite{monodGrowthBacterialCultures1949}
+\cmt{Ref about Quorum sensing. Signalling mechanism that controll gene expression (bacteria growth within species) in response to environment. \cite{ngBacterialQuorumsensingNetwork2009}}
 
@@
 These survival mechanisms mostly aim to secure as much available nutrition which lead to competitive strategies or, if it is advantageous, cooperative ones.
+\cite{hibbingBacterialCompetitionSurviving2010, stubbendieckBacterialCommunitiesInteractions2016} \cmt{Both good reviews on bacterial interaction}
 Further we show that the TPM can be readily extended to the systems with more than one bacteria an take describe more complex relationships between species.
+
+\cmt{\cite{h.tusonBacteriaSurfaceInteractions2013}: Attachment of bacteria to surfaces. (One of the mechanisms to stick to the areas with nutritients)}
 
@@
 
-One on the ways to descibe competition between bacteria populations is as a function of limiting resources ratio. \cmt{Ref Tilman 1977, Ref Smith 2002}
+One on the ways to descibe competition between bacteria populations is as a function of limiting resources ratio. 
+\cmt{Ref Tilman 1977 \cite{tilmanResourceCompetitionPlankton1977}, Ref Smith 2002 \cite{smithEffectsResourceSupplies2002}}
 To descibe such a situation we consider for simplicity only one common limiting resource pool. 
@@
 
-which has the structure of a generalised Lotka-Volterra system. In order to better see this, we set $L_A=L_B=0$: 
+which has the structure of a generalised Lotka-Volterra system. \cmt{Add reference about Lotka-Volterra. Compare with different states LV.}
+In order to better see this, we set $L_A=L_B=0$: 
 \begin{eqnarray}
@@
 \cmt{Read how this antibiotics affect the system. Which eq form to use?}
+
+\cmt{\cite{lenskiCoexistenceTwoCompetitors1986}: Modelling of the 2 sp. system with 1resource and 1 inhibitor.}
 
@@
 
+\cmt{\cite{rochaBaseCompositionBias2002}: Symbiosis host/bacteria. bacterial genomes suffering reductive evolution due to this. Maybe it's connected with 'social cheaters'?}
+
 The benefit, or mutual activation, of the two species from the presence of the other can be simply introduced by mathematical functions. 
@@
 
+\bibliography{references}
+
 \end{document}

### 2023-08-10 pool_model.tex
add text about cell death

--- 2023-08-08 pool_model.tex
@@
 ...
+
+\cmtb{Maybe add here motivations about bacteria states: So toy system where we can eliminate states (sum and get F(n1, n2)).
+And describe it initially without states/1state. Get the difference.}
 
@@
 Consider for simplicity the system where only one bacteria species is present $N=1$.
-According to the growth curves (\cmt{some references on bacteria growth curves??}) quite often at the beginning no increase in bacterial count is present and so-called lag-phase is observed.
+The bacterial growth curves usually consists of different consecutive phases~\cite{buchanan_when_1997}.
+At the beginning no increase in bacterial count is present and so-called lag-phase is observed.
 This phenomenon appears due to adaptation of the microbial community, for example, to the new environmental conditions.
-After this some bacteria enter the exponential growth phase, while some of it may not be able to adjust and experience the cell death.
-Some bacteria even may not be able to adjust and die other will enter the exponential growth phase after a lag phase. 
-After exponential growth the system reaches the saturation where the balance between the number of dying bacteria is in balance with the growing one leads to the stationary state.
+This so-called preparation to the exponential growth includes synthesys of the needed compounds, initiation of the transciption~\cite{rolfe_lag_2012} and poorly studied physiological and regulatory mechanisms~\cite{monod_growth_1949,}
+\cmt{For example, in Baranyi and Roberts modelling of lag-phase ???(maybe add).}
+After this, the succeeded in adaptation bacteria enter the exponential growth phase, where the process of cell division is directly happening \cmt{Add Ref}.
+When the nutrition is exhausted by the growing bacteria, the stationary state of the growth curve occurres.
+In this state the cultures accumulate waste products of metabolitic activity that leads to cell death of the bacteria.
+At the same time dying cells release the nutritions for survivors to continue growing.
+As a result the balance between the growing and dyuing processes leads to the constant with time bacterial count~\cite{navarro_llorens_stationary_2010}.
 In the described simplified version of the real-life system three possible bacteria states are mentioned corresponding to which process is happeng: adaplation, growth or death.
-Motivated by this we divide a bacterial population into three pools (states) (\textbf{Figure \ref{SchematicRep}}): 
+Motivated by this we divide a bacterial population into three pools (states) (\textbf{Figure \ref{fig:SchematicRep}}): 
 \begin{itemize}
@@
     \caption{Schematic representation of exchange between three states or pools.}
-    \label{SchematicRep}
+    \label{fig:SchematicRep}
     \end{center}
@@
 
-Based on experimental observations, it was shown that the bacterial growth is strongly influenced by the limiting resource. \cmt{ref review}
-In the system where all needed nutritients are in excess except for one the total growth is linear to the conentration of this liminting nutritient. \cmt{ref Monod}
-\cmt{Ref about Quorum sensing. Signalling mechanism that controll gene expression (bacteria growth within species) in response to environment. }
+Based on experimental observations, it was shown that the bacterial growth is strongly influenced by the limiting resource.
+In the system where all needed nutritients are in excess except for one the total growth is linear to the conentration of this liminting nutritient~\cite{hibbing_bacterial_2010, monod_growth_1949}.
+The controll of the population density with respect to the available nutritions is implemented using cell communication process called quorum sensing. 
+Using release of the extracellular signalling molecules (autoinducers) it is possible to synchronize the gene expression of the whole group~\cite{ng_bacterial_2009}.
 
+We would not dive into the processes of quorum sensing and try to keep it as simple as possible.
 According to this, we aim to include a resource pool which is used by the growing bacteria using the second order kinetic reaction:
@@
 
+\cmt{Look into this!}
+
 We explore now the idea that under certain conditions, as sudden change in the environment, part of the population $G$ will enter a lag-phase, i.e., there is a back-flow from $G$ to $L$. 
@@
 As it was mentioned above, bacteria usually coexist in large multispecies interacting communities whose growth is vastly determined by the developed survival mechanisms.
-These survival mechanisms mostly aim to secure as much available nutrition which lead to competitive strategies or, if it is advantageous, cooperative ones.
+These survival mechanisms mostly aim to secure as much available nutrition which lead to competitive strategies or, if it is advantageous, cooperative ones~\cite{hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
 Further we show that the TPM can be readily extended to the systems with more than one bacteria an take describe more complex relationships between species.
@@
 
-One on the ways to descibe competition between bacteria populations is as a function of limiting resources ratio. \cmt{Ref Tilman 1977, Ref Smith 2002}
+One on the ways to descibe competition between bacteria populations is as a function of limiting resources ratio~\cite{tilman_resource_1977, smith_effects_2002}.
 To descibe such a situation we consider for simplicity only one common limiting resource pool. 
@@
 \dot{G}_B &=& \alpha_B G_B\left(1-\frac{G_B}{N_t}\right ) -\frac{\alpha_B}{N_t}G_AG_B. 
+\label{eq:LV_simple}
 \end{eqnarray}
+
+Equations (\ref{eq:LV_simple}) are basically competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competive excursion.
+That principle claims that two species in competition for the same resource cannot coexist~\cite{wangersky_lotka-volterra_1978}.
 
@@
 
+\cmtb{Look papers on modelling of the LV in bacterial communities. Did anyone do the same as we??}
+
 \subsection{Interspecies Competition}
@@
 \cmt{Read how this antibiotics affect the system. Which eq form to use?}
+
+\cmt{\cite{lenski_coexistence_1986}: Modelling of the 2 sp. system with 1resource and 1 inhibitor.}
 
@@
 On the other hand, one can also observe some cooperative behaviours between species in case when it facilitates the competition with others.
-Quite often it happens between the species that are close relatives or clonal populations and such process is called kin selection. \cmt{Ref 25 from review}
-Sometimes neighbours, so-called 'social cheaters', can benefit from a certain behaviour without expressing it by itself.  \cmt{Ref 29 from review}
+Quite often it happens between the species that are close relatives or clonal populations and such process is called kin selection \cite{west_social_2007}.
+Sometimes neighbours, so-called 'social cheaters', can benefit from a certain behaviour without expressing it by itself \cite{rainey_evolution_2003}.
+
+\cmt{\cite{rocha_base_2002}: Symbiosis host/bacteria. bacterial genomes suffering reductive evolution due to this. 
+Maybe it's connected with 'social cheaters'?}
 
@@
 
+\section{Spatial Limitations}
+
+\cmtb{Maybe as motivation can be spatial competitive mechanisms as adhesion to survaces and motility. 
+So how much including these effects will affect system.}
+
+\cmt{\cite{htuson_bacteriasurface_2013}: Attachment of bacteria to surfaces. (One of the mechanisms to stick to the areas with nutritients)}
+
+\cmtb{We can take some toy model (e.g the one from group seminar) and do agent-based simulation. (add spatial component)
+Then from this simulation we sum bacteria and get growth curves.
+And using our model equations we get also growth curves.
+And find prob diffusion coef, maybe rates where model works fine.
+For sim maybe we will need to name specific bacteria species. (maybe nicer)}
+
 \section{Conclusion}
 
+\newpage
+\printbibliography
+
 \end{document}

### 2023-08-11 pool_model.tex
add LV examples/references

@@
 
-which has the structure of a generalised Lotka-Volterra system. \cmt{Add reference about Lotka-Volterra. Compare with different states LV.}
+which has the structure of a generalised Lotka-Volterra system (gLV). \cmt{Add reference about Lotka-Volterra. Compare with different states LV.}
 In order to better see this, we set $L_A=L_B=0$: 
@@
 That principle claims that two species in competition for the same resource cannot coexist~\cite{wangersky_lotka-volterra_1978}. 
+In previous studies it was shown that gLV equations can be sucessfully applied to predict the microbial community dynamics, determine different types of the pairwise interspecies interactions~\cite{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
+In food studies the gLV was applied by Mounier \etal~\cite{mounier_microbial_2008} to study interactions of 6 bacteria species  with yeast omissin in cheese.
 
@@
 This corresponds to the case $W\sim G$, where $W$ is the waste (or whatever one chooses to call it) of species A affecting species B. 
-If the waste accumulates one needs to replace $G$ by $\int G dt$.
-Keeping it simple yields:
+If the waste accumulates one needs to replace $G$ by $\int G dt$. Keeping it simple yields:
 \begin{eqnarray}

### 2023-08-16 pool_model.tex
Add ref for dormancy

@@
 In the system where all needed nutritients are in excess except for one the total growth is linear to the conentration of this liminting nutritient~\cite{hibbing_bacterial_2010, monod_growth_1949}.
-The controll of the population density with respect to the available nutritions is implemented using cell communication process called quorum sensing. 
+The control of the population density with respect to the available nutritions is implemented using cell communication process called quorum sensing. 
 Using release of the extracellular signalling molecules (autoinducers) it is possible to synchronize the gene expression of the whole group~\cite{ng_bacterial_2009}.
@@
 
-\cmt{Look into this!}
+Due to the life in rather unfavourable conditions bacteria developed different surviving mechanisms, one of which is adapting their metabolism and transitioning into dormant state.
+By defeinition dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\cite{kaprelyants_dormancy_1993}. 
+This survival strategy can appear in response to different external stresses, e.g., nutritient deficiency, temperature stress or extreme fluctuations of environment parameters.
+And after "difficult times" bacteria, also called viable but non-culturable(VBNC), can continue cell division~\cite{kell_viability_1998}.
+The models including transition of bacteria to the dormant state were proposed and studied by Bär \etal~\cite{bar_modelling_2002} and Jones and Lennon~\cite{jones_dormancy_2010}.
+
+\cmt{So now is it better to use flow to dormancy instead?}
 
@@
 \subsubsection{Maxwell type of stress - strain relation}
+
+One of the reasons of the dormant state appearence can be also the temperature shift~\cite{oliver_viable_1995}.
+
 We propose a kind of visco-elastic Maxwell type of stress-strain relation and write down the ordinary differential equation for $\gamma$ to be:

### 2023-08-23 pool_model.tex
some mistakes correction

@@
 Consider the system conssisting of $N$ species in $M$ different states.
-In the general representation the system can be described with the Ordinary Differential Equations (ODEs) using a non-linear function $\mbF$:
+In the general representation, the system can be described with the Ordinary Differential Equations (ODEs) using a non-linear function $\mbF$:
 \begin{equation}
@@
 where  $\mbx \in \mathbb{R}^{N}  \otimes \mathbb{R}^{M}$ is a vector descibing bacterial count.
-In a matrix form it can be written as
+In matrix form it can be written as
 \begin{equation}
@@
 The advantage of the following formulation is its generality. 
-Depending on the knowledge about system as well as the needed level of accuracy one can define arbitrary number of states $M$ and control the complexity of teh model.
+Depending on the knowledge about the system as well as the needed level of accuracy one can define an arbitrary number of states $M$ and control the complexity of the model.
 Moreover, opposite to the already known Predictive Microbiology like Baranyi and Roberts one can easily change.
@@
 
-Due to the life in rather unfavourable conditions bacteria developed different surviving mechanisms, one of which is adapting their metabolism and transitioning into dormant state.
+Due to its life in rather unfavourable conditions bacteria developed different surviving mechanisms, one of which is adapting their metabolism and transitioning into dormant state.
 By defeinition dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\cite{kaprelyants_dormancy_1993}. 
 This survival strategy can appear in response to different external stresses, e.g., nutritient deficiency, temperature stress or extreme fluctuations of environment parameters.
-And after "difficult times" bacteria, also called viable but non-culturable(VBNC), can continue cell division~\cite{kell_viability_1998}.
+And after "difficult times" bacteria, also called viable but non-culturable (VBNC), can continue cell division~\cite{kell_viability_1998}.
 The models including transition of bacteria to the dormant state were proposed and studied by Bär \etal~\cite{bar_modelling_2002} and Jones and Lennon~\cite{jones_dormancy_2010}.
@@
 One of the reasons of the dormant state appearence can be also the temperature shift~\cite{oliver_viable_1995}.
-
-We propose a kind of visco-elastic Maxwell type of stress-strain relation and write down the ordinary differential equation for $\gamma$ to be:
+Moreover, the rapid shift in temperature is associated with the appearence of an additional lag phase~\cite{zwietering_modeling_1994}.
+To model it we propose a kind of visco-elastic Maxwell type of stress-strain relation and write down the ordinary differential equation for $\gamma$ to be:
 \begin{eqnarray}

### 2023-08-24 pool_model.tex
Add System/ODEs/Parameters  for the ABM.

@@
 
-\cmtb{We can take some toy model (e.g the one from group seminar) and do agent-based simulation. (add spatial component)
-Then from this simulation we sum bacteria and get growth curves.
-And using our model equations we get also growth curves.
-And find prob diffusion coef, maybe rates where model works fine.
-For sim maybe we will need to name specific bacteria species. (maybe nicer)}
+To test the spatial limitations of the described above approach consider the system consisting from two bacterial species thet can be described bt two internal states: lag and growth states. 
+Both of these species grow using limiting resource pool $R$ with different growth constants $\alpha$ and different lengths of lag phase determined by parameter $\lambda$.
+Moreover, the species 1 is able to release a certain antimicrobial component $I$ that shows inhibitory effect on species 2.
+The ODEs the bacterial population can be written as: 
+\begin{equation}
+    \begin{cases}
+        \dot{L_1} = -\lambda_1 R L_1\\
+        \dot{G_1} = \lambda_1 R L_1 + \alpha_1 R G_1\\
+        \dot{L_2} = -\lambda_2 R L_2\\
+        \dot{G_2} = \lambda_2 R L_2 + \frac{\alpha_2}{1 + \mu_I I} R G_2\\
+    \end{cases}.
+    \label{eq:spatial_limit_F}
+\end{equation}
+
+The microenvironment of our system contains the resource and the inhibitor molecules:
+\begin{equation}
+    \begin{cases}
+        \dot{R} = -\frac{\alpha_1}{N_t} R G_1-\frac{\alpha_2}{N_t} R G_2 \\
+        \dot{I} = \kappa G_1
+    \end{cases}.
+    \label{eq:spatial_limit_H}
+\end{equation}
+
+\begin{table}[H]
+    \centering
+    \begin{tabular}{ccc}
+    \specialrule{.1em}{.01em}{.05em} 
+    \textbf{Parameter} \hspace{3mm} & \textbf{Value} \hspace{3mm} & \textbf{Units}\\
+    \toprule
+    $\lambda_1$ & 0.008  & [time$^{-1}$]                 \\
+    $\lambda_2$ & 0.005  & [time$^{-1}$]                 \\
+    $\alpha_1$  & 4.0    & [time$^{-1}$]                 \\
+    $\alpha_2$  & 5.0    & [time$^{-1}$]                 \\
+    $\kappa$    & 0.1    & [(cell $\cdot$ time)$^{-1}$]  \\
+    $\mu_I$     & 0.1    & [cell$^{-1}$]                 \\
+    $N_t$       & $10^4$ & [cell]                        \\
+    \bottomrule
+    \end{tabular}
+    \caption{{\footnotesize The list of the parameters of the system.}}
+    \label{tab:spatial_limit_param}
+\end{table}
+
+\begin{table}[H]
+    \centering
+    \begin{tabular}{ccc}
+    \specialrule{.1em}{.01em}{.05em} 
+    \textbf{Variable} \hspace{3mm} & \textbf{Initial Value} \hspace{3mm} & \textbf{Units}\\
+    \toprule
+    $L_1$ & 5.0 & [cell]        \\
+    $G_1$ & 0.0 & [cell]        \\
+    $L_2$ & 5.0 & [cell]        \\
+    $G_2$ & 0.0 & [cell]        \\
+    $R$   & 1.0 & [cell/cell]   \\
+    $I$   & 0.0 & [cell]        \\
+    \bottomrule
+    \end{tabular}
+    \caption{{\footnotesize The list of the initial values of the system.}}
+    \label{tab:spatial_limit_init}
+\end{table}
+
+\begin{figure}[H]
+    \begin{center}
+    \includegraphics[width=0.67\textwidth]{Figures/pool_model_spatial.pdf}
+    \caption{{\footnotesize The output of the model (\ref{eq:spatial_limit_F}, \ref{eq:spatial_limit_H}) for given parameters and initial valued presented in Tables \ref{tab:spatial_limit_param}, \ref{tab:spatial_limit_init}, respectively.
+    The observables of the system are the bacterial count of each species $L+G$ (blue and orange lines) and the total bacterial count of the system (green).}}
+    \label{fig:spatial_limit_plot1}
+    \end{center}
+\end{figure}
+
+\cmtb{The ABM additional parameters $D_R$, $D_I$, $q$ (At which level of $R_i$ cell division happens) and more parameters describing bacterial division.}
 

### 2023-08-31 pool_model.tex
spatial limitation: correct mistake in ODEs
3 pool model, 1 species corrections

@@
     \mbx = \begin{pmatrix}
-        x^1_1  & \dots & x^1_M  \\
+        x^1_1  & \dots & x^1_N  \\
         \vdots &       & \vdots \\
-        x^N_1  & \dots & x^N_M  \\
+        x^M_1  & \dots & x^M_N  \\
             \end{pmatrix}
@@
 \end{equation}
-with $x_{ij}$ describing the bacterial counts for species $i$ in state $j$.
+with $x_{j}^{i}$ describing the bacterial counts for species $i$ in state $j$.
 $\mbu$ is the vector of different inputs of the model, e.g., temperature, pressure etc.
@@
 \begin{equation}
-    \dot{\mbm} = \mbH (\mbx, \mbm, \mbp, \mbu, ...).
+    \dot{\mbm} = \mbH (\mbx, \mbm, \mbp, \mbu).
     \label{eq:model_microenv}
@@
 \begin{equation}
-    N_n = \sum_m \mbx_i^n.
+    N_i = \sum_k \mbx_i^k.
 \end{equation}
@@
 \begin{equation}
-    N_t = \sum_j \sum_m \mbx_i^j.
+    N = \sum_i \sum_k \mbx_i^k.
 \end{equation}
@@
 \begin{itemize}
-\item $L(t)$: the fraction of the population in the lag phase
-\item $G(t)$: the fraction of the population growing and dividing
-\item $D(t)$: the fraction of the population undergoing cell death
+\item $x^1_1 (t) = L(t)$: the fraction of the population in the lag phase
+\item $x^2_1 (t) = G(t)$: the fraction of the population growing and dividing
+\item $x^3_1 (t) = D(t)$: the fraction of the population undergoing cell death
 \end{itemize}
-Here the concentrations of the pools are components of our variable vector $\mbx (t) =\{L(t), G(t), D(t)\}$.
+Here the concentrations of the pools are components of our variable vector $\mbx (t) = (L(t), G(t), D(t))^T$.
 
@@
     \includegraphics[width=0.7\textwidth]{Figures/TPM_fig.pdf}
-    \caption{Schematic representation of exchange between three states or pools.}
+    \caption{{\footnotesize Schematic representation of exchange between three states or pools.}}
     \label{fig:SchematicRep}
@@
 D &\stackrel{\beta}{\longrightarrow} & \emptyset
-\end{eqnarray}.
+\end{eqnarray}
 
@@
     \dot{D} = \mu  L + \mu' G- \beta D  
-\end{cases}.
+\end{cases}
+\end{equation}
+or in matrix form it shows the linear dependency
+\begin{equation}
+    \dot \mbx  = \begin{pmatrix}
+                    -(\mu + \lambda) & 0             & 0      \\
+                    \lambda          & \alpha - \mu' & 0      \\
+                    \mu              & \mu'          & -\beta 
+                 \end{pmatrix} \mbx .
 \end{equation}
 
-\cmt{Add plots ?? Or already after resource is added?}
-
-\subsubsection{Resource Dependence}
+\subsection{Resource Dependence}
 
 Based on experimental observations, it was shown that the bacterial growth is strongly influenced by the limiting resource.
-In the system where all needed nutritients are in excess except for one the total growth is linear to the conentration of this liminting nutritient~\cite{hibbing_bacterial_2010, monod_growth_1949}.
+In the system where all needed nutritients are in excess except for one the total growth is linear to the concentration of this liminting nutritient~\cite{hibbing_bacterial_2010, monod_growth_1949}.
 The control of the population density with respect to the available nutritions is implemented using cell communication process called quorum sensing. 
@@
 Here the full capacity of the resource pool $N_t$ was used as a scaling constant.
-Then the growth term can be simply written as $\alpha G R$, and the microenvironment of the system introduced by \ref{eq:model_microenv} is defined by ODE:
-\begin{equation}
-\dot{R} = -\frac{\alpha}{N_t} R G
-\end{equation}
+In this case the growth term can be simply written as $\alpha G R$ and the model is fully formulated in the following way: 
+\begin{eqnarray}
+    \dot \mbx  &=& \mbF(\mbx, \mbm, \mbp, \mbu) = \begin{pmatrix}
+                                                    -(\mu + \lambda) & 0               & 0      \\
+                                                    \lambda          & \alpha R - \mu' & 0      \\
+                                                    \mu              & \mu'            & -\beta 
+                                                \end{pmatrix} \mbx  \\
+    \dot \mbm &=& \mbH (\mbx, \mbm, \mbp, \mbu) = -\frac{\alpha}{N_t} R G.
+\end{eqnarray}
 
-Explointing that limitation of the resource pool also limits the total becterial count of the system  we find:
+Such model can be simplified by explointing that limitation of the resource pool also limits the total bacterial count of the system:
 \begin{equation}
@@
 \end{equation}
-And, hence, the full system can be written as
+and, hence, the full system can be written as
 \begin{equation}
@@
         \dot{D} = \mu  L + \mu' G- \beta D 
+        \label{eq:3pool_resource} 
     \end{cases}.
@@
 
-\cmt{Plots / results}
+\begin{figure}[t]
+    \begin{center}
+    \includegraphics[width=0.67\textwidth]{Figures/pool_model_3pools_resource.pdf}
+    \caption{{\footnotesize The time dependence of the total bacterial count $N(t) = \sum_i \sum_k \mbx_i^k = L(t) + G(t) + D(t)$ is obtained as a sum of bacteria in all states.
+    All curves were calculated using the model (\ref{eq:3pool_resource}) with parameters $\alpha=2.5$, $\mu'=0.5$ and $\beta=5$.
+    The purple curve shows significant dip in behaviour of the system where the death of the bacteria is present $\mu = 0.5$.
+    Other curves show the case of no bacteria dying $\mu=0$ and represent different length of lag phases:
+    short $\lambda=10^{-2}$ (red), intermediate $\lambda=10^{-3}$ (blue) and long one $\lambda=10^{-4}$ (green).
+}}
+    \label{fig:3pool_resource_plots}
+    \end{center}
+\end{figure}
 
-\subsubsection{Lag-time Estimation (??)}
+\subsection{Lag-time Estimation}
 
@@
 The relevant question here is whether cells in the pool $D$ still contribute to the limitation of the resources by, e.g., occupying space or consuming important nutrients. 
-If yes, the three pool model would be advantageous, if not the two pool model is sufficient with the advantages of having one parameter less.\\
+If yes, the three pool model would be advantageous, if not the two pool model is sufficient with the advantages of having one parameter less.
 
@@
 to be solved with the initial conditions $L(0)=n_0$, $G(0)=0$, and $D(0)=0$, where $n_0$ is the initial bacterial population. 
-$N_t$ denotes the maximal size of the total bacterial population due to environmental limitations (\textbf{Figure \ref{GrowthDip}}). 
+$N_t$ denotes the maximal size of the total bacterial population due to environmental limitations (\textbf{Figure \ref{fig:3pool_resource_plots}}). 
 For $\mu'=0$ the equations for $L$ and $D$ can be readily integrated resulting in:
-\begin{eqnarray}
-L(t) &=& n_0 e^{-(\mu+\lambda)t}\\
-D(t) &=& n_0 \mu e^{-\beta t}\frac{1-e^{-(\mu+\lambda-\beta)t}}{\mu+\lambda-\beta}.
-\end{eqnarray}
+\begin{equation}
+    \begin{cases}
+        L(t) = n_0 e^{-(\mu+\lambda)t}\\
+        D(t) = n_0 \mu e^{-\beta t}\frac{1-e^{-(\mu+\lambda-\beta)t}}{\mu+\lambda-\beta}
+    \end{cases}.
+\end{equation}
 For constant parameters $\mu$, $\lambda$, and $\beta$ the integrals for $L$ and $D$ can be put into the differential equation for $G$. 
-However, for time dependent parameters, e.g.
-for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$.
+However, for time dependent parameters, e.g., for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$.
 It is interesting to see what type of equations result for the total bacterial population $n=L+G+D$:
-\begin{eqnarray}
-\dot{G} &=& \lambda L + \alpha G\left(1-\frac{n}{N_t}\right)\\
-\dot{n} &=& - \beta D + \alpha G\left(1-\frac{n}{N_t}\right). 
-\end{eqnarray}
+\begin{equation}
+    \begin{cases}
+        \dot{G} = \lambda L + \alpha G\left(1-\frac{n}{N_t}\right)\\
+        \dot{n} = - \beta D + \alpha G\left(1-\frac{n}{N_t}\right)
+    \end{cases}
+\end{equation}
 with the initial conditions $G(0)=0$ and $n(0)=n_0$. 
@@
 
-\cmt{Add/make plots to the model 3 pool/ 2 pool.}
-
-\cmt{Add limitation due to resource}
-
 \subsection{Resource-dependent lag-phase}
@@
 Then using the model consisted of two bacteria states $L$ and $G$ we can define the system as
-\begin{eqnarray}
-\dot{L} &=& - \lambda R L\\
-\dot{G} &=& \lambda R L +\alpha R G\\
-\dot{R} &=&-\frac{\alpha}{N_t} R G.
-\end{eqnarray}
+\begin{equation}
+    \begin{cases}
+        \dot{L} = - \lambda R L\\
+        \dot{G} = \lambda R L +\alpha R G\\
+        \dot{R} =-\frac{\alpha}{N_t} R G.
+    \end{cases}
+\end{equation}
 
@@
 Using this we eliminate $R$ and arrive at:
-\begin{eqnarray}
-\dot{L} &=& - \lambda \left(1-\frac{L+G}{N_t}\right)L\\
-\dot{G} &=& \lambda \left(1-\frac{L+G}{N_t}\right)L +\alpha \left(1-\frac{L+G}{N_t}\right) G
-\end{eqnarray}
+\begin{equation}
+    \begin{cases}
+        \dot{L} = - \lambda \left(1-\frac{L+G}{N_t}\right)L\\
+        \dot{G} = \lambda \left(1-\frac{L+G}{N_t}\right)L +\alpha \left(1-\frac{L+G}{N_t}\right) G
+\end{cases}
+\end{equation}
 The steady state is $L_s+G_s=N_t$.
@@
 
-\subsection{Back flow from G to L}
+\subsection{Back flow from Growth to Lag pool}
 
@@
 The ordinary differential equations for the pool model with environmental-shock based back-flow read:
-\begin{eqnarray}
-\dot{L} &=&\gamma(t) G - (\mu + \lambda) L\\
-\dot{G} &=& \lambda L -\gamma(t) G + \alpha G\left(1-\frac{L+G+D}{N_t}\right)\\
-\dot{D} &=& \mu  L - \beta D
-\end{eqnarray}
+\begin{equation}
+    \begin{cases}
+    \dot{L} =\gamma(t) G - (\mu + \lambda) L\\
+    \dot{G} = \lambda L -\gamma(t) G + \alpha G\left(1-\frac{L+G+D}{N_t}\right)\\
+    \dot{D} = \mu  L - \beta D
+    \end{cases}
+\end{equation}
 
@@
 However, except the inhibition due to resource competition, bacteria developed differnt others active competition mechanisms.
-One of the examples of such mechanisms is a production of the antimicrobial compounds, which can be also called antibiotics or inhibitors.
+One of the examples of such mechanisms is a production of the antimicrobial compounds, which can be also called antibiotics or inhibitors. \cmt{toxins}
 
@@
 
-\cmt{\cite{lenski_coexistence_1986}: Modelling of the 2 sp. system with 1resource and 1 inhibitor.}
-
-Now extend our model where bacteria not only consume nutritions from resource pool $R$ but one of the species also actively produce the inhibiting compounds $I$. 
+\subsubsection{Toxins Production}
+Now extend our model where bacteria not only consume nutritions from resource pool $R$ but one of the species also actively produce the inhibiting compounds $T$. 
 It is the accumulation of the bacteria waste which is responsible for the spoilage of food products. 
@@
 \dot{R} &=&-\frac{\alpha}{N_t} R G\\
-\dot{I} &=& \mathcal{F}(R,G).
+\dot{T} &=& \mathcal{F}(R,G).
 \end{eqnarray}
@@
 Model $\mathcal{F}_1$ results in waste accumulation in steady state $I_1^s=\lim_{t\to\infty}I_1(t)=\kappa(N_t-n_0)$. 
-To see this we note $\dot{I}_1=\kappa(\dot{L}+\dot{G})$, which yields: $I_1(t)=\kappa(L(t)+G(t)-n_0)$ where we used $I_1(0)=G(0)=0$, $L(0)=n_0$. 
+To see this we note $\dot{T}_1=\kappa(\dot{L}+\dot{G})$, which yields: $I_1(t)=\kappa(L(t)+G(t)-n_0)$ where we used $T_1(0)=G(0)=0$, $L(0)=n_0$. 
 This model assumes that inhibitor is only produced when the bacteria grow and divide. \\
@@
 Model $\mathcal{F}_2$ results in $I_2(t)=\kappa\int_{t_0}^tG(t')dt'$. 
-As long as $G>0$ inhibitor is produced, no matter whether the amount of bacteria is still growing. 
-The amount of inhibitor depends therefore on the back-flow model employed, which can be seen in Fig. \ref{Waste}.
+As long as $G>0$ toxin is produced, no matter whether the amount of bacteria is still growing. 
+The amount of toxin depends therefore on the back-flow model employed, which can be seen in Fig. \ref{Waste}.
 
-\cmt{Look is this inhibition the same as waste production. Or change it.}
-
+\subsubsection{Species Inhibition}
+\cmt{\cite{lenski_coexistence_1986}: Modelling of the 2 sp. system with 1resource and 1 inhibitor.}
 The simplest way to introduce inhibition of species A by species B is via a growth rate of species A depending on the abundance of species B. 
@@
 
+\cmt{Production of public goods!}
+
 The benefit, or mutual activation, of the two species from the presence of the other can be simply introduced by mathematical functions. 
@@
 Both of these species grow using limiting resource pool $R$ with different growth constants $\alpha$ and different lengths of lag phase determined by parameter $\lambda$.
-Moreover, the species 1 is able to release a certain antimicrobial component $I$ that shows inhibitory effect on species 2.
+Moreover, the species 1 is able to release a certain antimicrobial component $I$ that shows inhibitory effect on species 2 (amensalism).
 The ODEs the bacterial population can be written as: 
@@
     \begin{cases}
-        \dot{R} = -\frac{\alpha_1}{N_t} R G_1-\frac{\alpha_2}{N_t} R G_2 \\
+        \dot{R} = -\frac{\alpha_1}{N_t} R G_1-\frac{\alpha_2}{(1 + \mu_I I) N_t} R G_2 \\
         \dot{I} = \kappa G_1

### 2023-09-01 pool_model.tex
add Temprture stress Figures

@@
 \gamma(t) &=& \Gamma\sum_{i=1}^n \left |\Delta T_i \right |\theta(t-t_i)e^{-\delta(t-t_i)}.
+\label{eq:gamma_tempshift}
 \end{eqnarray}
+
+\begin{figure}[t]
+    \begin{center}
+    \includegraphics[width=0.67\textwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}
+    \caption{{\footnotesize (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stres (\ref{eq:ODE_tempshift}).
+    (b) The value of the corresponding back-flow rate $\gamma$ calculated according to (\ref{eq:gamma_tempshift}).
+    The curves represent no temperature shift (blue), one temperature jump $\Delta T = 10$ at time $t=4$ (red) 
+    and the series of different temperature shifts of value 10, 5 and 15 at the tme points 1, 3 asnd 7, respectively (green).
+    The paremeters of the back-flow rate are $\Gamma=1$ and $\delta=5$.
+    The other model parameters are $\alpha=2.5$, $\mu'=0.5$, $\beta=5$, $\lambda=10^{-2}$, $\mu = 0$ fe each curve.
+    }}
+    \label{fig:TempJump}
+    \end{center}
+\end{figure}
 
@@
     \dot{D} = \mu  L - \beta D
+    \label{eq:ODE_tempshift}
     \end{cases}
 \end{equation}
-
--  \cmt{Ref to temperature jumps effect. Plots / results.}
 

### 2023-10-19 pool_model.tex
resource comperition and toxin production, insert results

@@
 \section{General Mathemetical Formulation}
-Consider the system conssisting of $N$ species in $M$ different states.
+Consider the system consisting of $N$ species in $M$ different states.
 In the general representation, the system can be described with the Ordinary Differential Equations (ODEs) using a non-linear function $\mbF$:
@@
 
-\subsection{Resource-dependent lag-phase}
-
-Another way to encorporate the resource dependence is to assume that the bacteria only leave the lag-phase if there are nutrients, i.e., resources. 
-Otherwise they stay in the lag-phase.
-Then using the model consisted of two bacteria states $L$ and $G$ we can define the system as
-\begin{equation}
-    \begin{cases}
-        \dot{L} = - \lambda R L\\
-        \dot{G} = \lambda R L +\alpha R G\\
-        \dot{R} =-\frac{\alpha}{N_t} R G.
-    \end{cases}
-\end{equation}
-
-Turning back to the equations above we can state: $L+G+N_tR=N_t$.
-For $t=0$ we have $n_0+N_tR_0=N_t$ (given $L_0=n_0$ and $G_0=0$) and from this: $N_tR_0=N_t-n_0$.
-Using this we eliminate $R$ and arrive at:
-\begin{equation}
-    \begin{cases}
-        \dot{L} = - \lambda \left(1-\frac{L+G}{N_t}\right)L\\
-        \dot{G} = \lambda \left(1-\frac{L+G}{N_t}\right)L +\alpha \left(1-\frac{L+G}{N_t}\right) G
-\end{cases}
-\end{equation}
-The steady state is $L_s+G_s=N_t$.
-However, it depends on the initial conditions and the parameters what $L_s$ and $G_s$ are.
-There is no way to determine $L_s$ and $G_s$ besides solving the dynamical equations, i.e., the steady state cannot be determined directly by solving a set of algebraic equations.
-To see this we determine the nullclines: the first nullcline is $G=N_t-L$ and the second is $G=N_t-L$, i.e., the nullclines coincides.
-Every point on the line $G=N_t-L$ is a fixed point.
-This can be seen in Fig. \ref{Phaseplot}. 
-For $\lambda\ll \alpha$ the lag-pool is not emptied before all resources are consumed. As a consequence the lag-pool is non-zero in steady-state. 
-In contrast, for $\lambda\gg \alpha$ the lag-pool is depleted before the resources are consumed, i.e., the lag-pool is zero in steady state.
-
-The fraction $G$ of the exponentially growing cells will go to steady state due to limitations in resources. 
-$\lambda$ is the rate by which cells go from the lag phase to the growth phase; $\mu$ is the rate by which cells go from the lag phase to the state in which they die; $\alpha$ is the growth rate of the healthy population; $\beta$ is the rate by which the cells are removed from the $D$ pool.
-
-\cmt{Plots / results}
-
 \subsection{Back flow from Growth to Lag pool}
@@
 \end{eqnarray}
-The rate $\gamma(t)$ depends on the rate of change of the environment. Stronger and quicker changes will lead to a higher rate $\gamma$. 
-Let the change in the environment be in temperature $T$ (\textbf{Figure \ref{default}}).
-
+The rate $\gamma(t)$ depends on the rate of change of the environment.
+Stronger and quicker changes will lead to a higher rate $\gamma$. 
 \subsubsection{Maxwell type of stress - strain relation}
@@
 
-\begin{eqnarray}
-\label{Model_Comp1}
-\dot{L}_A &=& - \lambda_A R L_A\\
-\dot{L}_B &=& - \lambda_B R L_B \\
-\dot{G}_A &=& \lambda_A R L_A +\alpha_A R G_A\\
-\label{Model_Comp4}
-\dot{G}_B &=& \lambda_B R L_B +\alpha_B R G_B\\
-\dot{R} &=&-\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B
-\end{eqnarray}
+\begin{equation}
+    \begin{cases}
+        \dot{L}_A = - \lambda_A R L_A\\
+        \dot{L}_B = - \lambda_B R L_B \\
+        \dot{G}_A = \lambda_A R L_A +\alpha_A R G_A\\
+        \dot{G}_B = \lambda_B R L_B +\alpha_B R G_B\\
+        \dot{R} =-\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B
+    \end{cases}
+    \label{eq:model_2sp_resource_comp}
+\end{equation}
+
 We again rescaled the resource pool with $N_t$ and find
@@
 \label{Resource}
-R &=&1-\frac{L_A+G_A+L_B+G_B}{N_t}\\
-R_0&=&1-\frac{n_A+n_B}{N_t}. 
+R &=&1-\frac{L_A+G_A+L_B+G_B}{N_t}
 \end{eqnarray}
-Inserting Eq. \ref{Resource} into Eqs. \ref{Model_Comp1} - \ref{Model_Comp4} yields:
+Inserting Eq. \ref{Resource} into Eqs. \ref{eq:model_2sp_resource_comp} yields:
 \begin{eqnarray*}
@@
 
-For $N_t\gg\alpha_{A/B}$ the equations decouple. 
-The null-clines of the system given by Eqs. \ref{Model_Comp1} - \ref{Model_Comp4} coincide and are given by $N_t=L_A+G_A+L_B+G_B$. 
-If $\lambda_{A/B}>\alpha_{A/B}$ the lag-pools are completely depleted before the depletion of the resource pool stops the dynamics. 
-To simplify the investigation of the systems dynamics we therefore ignore the $L_A$ and $L_B$ dimension and only consider the dynamics on the centre-manifold given by the $G_A\times G_B$ plane. 
-The corresponding phase-plot is shown in Fig. \ref{Competition}. 
-The advantage in using the pool model in contrast to using Lotka-Volterra directly is that all parameters have a clear meaning and consistent parameter relations arrive naturally. 
-And a species can have more than one internal state. 
-However, it is instructive to rewrite the system by setting $x_1\equiv G_A$, $x_2\equiv G_B$, and $\mathbf{x}=(x_1,x_2)^T$:
+The pool model, in contrast to the classical Lotka-Volterra approach, allows for far more flexibility.
 
-\begin{eqnarray}
-\dot{\mathbf{x}}&=&A\mathbf{x}+\frac{1}{2}\mathbf{x}B\mathbf{x}\\
-A&=&
-\begin{bmatrix}
-\alpha_A&0\\
-0 &\alpha_B
-\end{bmatrix}\\
-B&=&
-\frac{-2}{N_t}\begin{bmatrix}
-\alpha_A&\alpha_A\\
-\alpha_B &\alpha_B
-\end{bmatrix}
-\end{eqnarray}
-
-The pool model, in contrast to the classical Lotka-Volterra approach, allows for far more flexibility. 
-Non-linear functional responses can be introduced without any problem (see how section \ref{Gen_Lot_Vol} for an extended discussion).
-
-\cmt{Plots / results}
+\begin{figure}[H]
+    \begin{center}
+    \includegraphics[width=0.67\textwidth]{Figures/pool_model_2pools_resource_competition.pdf}
+    \caption{{\footnotesize The competition of the two bacterial species due to the common limiting resource calculated using Eq. (\ref{eq:model_2sp_resource_comp}).
+    The red solid curve describes the growth of the bacteria $A$ $N_A = \sum_{k} x_A^k = L_A+G_A$ with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
+    The blue solid curve describes the growth of the bacteria $B$ $N_B = \sum_{k} x_B^k = L_B+G_B$ with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
+    The purple dashed line shows the total bacterial count of the system  $N(t)=\sum_i\sum_k\mbx_i^k=N_A+N_B$.
+    The gray solid line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.}}
+    \label{fig:2pool_resource_2sp}
+    \end{center}
+\end{figure}
+As one can see in Figure~\ref{fig:2pool_resource_2sp} even without direct interaction between species the bacterial growth was inhibited due to the limitation of the resource.
+Here species $A$ was able to supress species $B$ due to the higher values of the growth constant $\alpha$..
 
@@
 \begin{eqnarray}
-\mathcal{F}_1(R,G)&=&\kappa\alpha R G=-\kappa N_t\dot{R}\\
-\mathcal{F}_2(R,G)&=&\kappa G
+\mathcal{F}_1(R,G)&=&\kappa G\\
+\mathcal{F}_2(R,G)&=&\kappa\alpha R G=-\kappa N_t\dot{R}
 \end{eqnarray}
-Model $\mathcal{F}_1$ results in waste accumulation in steady state $I_1^s=\lim_{t\to\infty}I_1(t)=\kappa(N_t-n_0)$. 
-To see this we note $\dot{T}_1=\kappa(\dot{L}+\dot{G})$, which yields: $I_1(t)=\kappa(L(t)+G(t)-n_0)$ where we used $T_1(0)=G(0)=0$, $L(0)=n_0$. 
+We would like to test the following model:
+\begin{equation}
+    \begin{cases}
+        \dot{L}_A = - \lambda_A R L_A\\
+        \dot{L}_B = - \lambda_B R L_B \\
+        \dot{G}_A = \lambda_A R L_A + \alpha_A R G_A - \frac{\mu}{\kappa N_t} T G_A\\
+        \dot{G}_B = \lambda_B R L_B + \alpha_B R G_B\\
+        \dot{R} = -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B\\
+        \dot{T} = \mathcal{F}_{1,2} (R, G_B) \\
+    \end{cases}
+    \label{eq:model_2sp_toxin}
+\end{equation}
+
+As it can be seen in Figure~\ref{fig:2pool_2sp_toxin}, model $\mathcal{F}_1$ results in $I_1(t)=\kappa\int_{t_0}^tG(t')dt'$. 
+As long as $G>0$ toxin is produced, no matter whether the amount of bacteria is still growing. 
+Model $\mathcal{F}_2$ results in waste accumulation in steady state $I_2^s=\lim_{t\to\infty}I_1(t)=\kappa(N_t-n_0)$. 
+To see this we note $\dot{T}_2=\kappa(\dot{L}+\dot{G})$, which yields: $I_2(t)=\kappa(L(t)+G(t)-n_0)$ where we used $T_2(0)=G(0)=0$, $L(0)=n_0$. 
 This model assumes that inhibitor is only produced when the bacteria grow and divide. \\
 
-Model $\mathcal{F}_2$ results in $I_2(t)=\kappa\int_{t_0}^tG(t')dt'$. 
-As long as $G>0$ toxin is produced, no matter whether the amount of bacteria is still growing. 
-The amount of toxin depends therefore on the back-flow model employed, which can be seen in Fig. \ref{Waste}.
+\begin{figure}[H]
+    \begin{center}
+    \includegraphics[width=1.\textwidth]{Figures/pool_model_2pools_toxin.pdf}
+    \caption{{\footnotesize The two species system competing for the common nutritient $R$ with production of the toxin $T$ (brown solid line) calculated using Eq. (\ref{eq:model_2sp_toxin}).
+    The growing bacteria $B$ produces toxin $T$ (a) constantly $\dot{T} = \mathcal{F}_1 = \kappa G_B$ and (b) only during growth process $\dot{T} = \mathcal{F}_2 = \kappa \alpha_B R G_B$ with 
+    production rate $\kappa=0.2$.
+    The toxin $T$ causes the cell death of bacteria $A$ with rate constant $\mu = 0.2$.
+    The red solid curve describes the growth of the bacteria $A$ $N_A = \sum_{k} x_A^k = L_A+G_A$ with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
+    The blue solid curve describes the growth of the bacteria $B$ $N_B = \sum_{k} x_B^k = L_B+G_B$ with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
+    The gray solid line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.}}
+    \label{fig:2pool_2sp_toxin}
+    \end{center}
+\end{figure}
 

### 2023-10-26 pool_model.tex
species inhibition: add plots

@@
 \begin{eqnarray}
-\dot{G}_A &=&\frac{G_A}{G_B}\left(1 - G_A-G_B\right )\\
+\dot{G}_A &=&\frac{G_A}{G_B}\left(1 - G_A-G_B\right )
+\label{eq:2sp_1pool_inhibit_unitless1}\\
 \dot{G}_B &=&\psi\frac{G_B}{G_A}\left(1-G_A-G_B\right ). 
+\label{eq:2sp_1pool_inhibit_unitless2}
 \end{eqnarray}
@@
 \end{eqnarray}
+
+\begin{figure}[H]
+    \begin{center}
+    \includegraphics[width=0.5\textwidth]{Figures/pool_model_1pool_inhib.pdf}
+    \caption{{\footnotesize The two species system where species inhibit each other defined with equations~(\ref{eq:2sp_1pool_inhibit_unitless1}-\ref{eq:2sp_1pool_inhibit_unitless2}).
+    Here the red lines describes the evolution of the species $A$ and blue lines correspond to species $B$.
+    The relation between inhibition strength of spesies $B$ and $A$ are defined by controll parameter value (a) $\psi$ = 0.5, (b) $\psi$ = 1 or (c) $\psi$ = 2.}}
+    \label{fig:1pool_2sp_inhibit}
+    \end{center}
+\end{figure}
 

### 2023-10-27 pool_model.tex
add matrix forms to equations

@@
 \end{equation}
+Rewriting the equations above, we get the following non-linear system:
+\begin{equation}
+    \dot \mbx = \begin{pmatrix}
+        -(\mu + \lambda) & 0       & 0 \\
+         \lambda         & \alpha  & 0 \\
+         \mu & + \mu'& - \beta 
+    \end{pmatrix} 
+    \mbx + \frac{\alpha}{N_t} x^2 \sum_i^3 x^i \begin{pmatrix} 0 \\ 1 \\ 0  \end{pmatrix}.
+\end{equation}
+Here the first term is the linear part and the second term define the quadratic behaviour of the system.
 
@@
 \end{equation}
-
+This system of equations defines the microenvironment $\mbm$ and bacteria state matrix $\mbx$ as
+\begin{equation}
+    \mbm = R, \hspace{1cm}
+    \mbx = \begin{pmatrix}
+        L_A & L_B \\
+        G_A & G_B 
+    \end{pmatrix}.
+\end{equation}
 We again rescaled the resource pool with $N_t$ and find

### 2023-12-04 pool_model.tex
Add reference for Lotka-Volterra

@@
 This phenomenon appears due to adaptation of the microbial community, for example, to the new environmental conditions.
-This so-called preparation to the exponential growth includes synthesys of the needed compounds, initiation of the transciption~\cite{rolfe_lag_2012} and poorly studied physiological and regulatory mechanisms~\cite{monod_growth_1949,}
-\cmt{For example, in Baranyi and Roberts modelling of lag-phase ???(maybe add).}
+This so-called preparation to the exponential growth includes synthesys of the needed compounds, initiation of the transciption~\cite{rolfe_lag_2012} and poorly studied physiological and regulatory mechanisms~\cite{monod_growth_1949}
 After this, the succeeded in adaptation bacteria enter the exponential growth phase, where the process of cell division is directly happening \cmt{Add Ref}.
@@
 At the same time dying cells release the nutritions for survivors to continue growing.
-As a result the balance between the growing and dyuing processes leads to the constant with time bacterial count~\cite{navarro_llorens_stationary_2010}.
+As a result the balance between the growing and dying processes leads to the constant with time bacterial count~\cite{navarro_llorens_stationary_2010}.
 In the described simplified version of the real-life system three possible bacteria states are mentioned corresponding to which process is happeng: adaplation, growth or death.
@@
 For now we assume for simplicity that there is no back flow between growth state $G$ and lag-state $L$. 
-The suggested model can be cast in the following reactions:
+The suggested model can be cast in the following reactions: \cmt{Use notations x ??}
 \begin{eqnarray}
@@
     \begin{center}
-    \includegraphics[width=0.67\textwidth]{Figures/pool_model_3pools_resource.pdf}
+    \includegraphics[width=0.6\textwidth]{Figures/pool_model_3pools_resource.pdf}
     \caption{{\footnotesize The time dependence of the total bacterial count $N(t) = \sum_i \sum_k \mbx_i^k = L(t) + G(t) + D(t)$ is obtained as a sum of bacteria in all states.
@@
     \begin{center}
-    \includegraphics[width=0.67\textwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}
+    \includegraphics[width=0.6\textwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}
     \caption{{\footnotesize (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stres (\ref{eq:ODE_tempshift}).
@@
 
-which has the structure of a generalised Lotka-Volterra system (gLV). \cmt{Add reference about Lotka-Volterra. Compare with different states LV.}
+which has the structure of a generalised Lotka-Volterra system (gLV).
 In order to better see this, we set $L_A=L_B=0$: 
 \begin{eqnarray}
-\label{Comp_Model}
+\label{eq:LV_simple1}
 \dot{G}_A &=& \alpha_A G_A\left(1 - \frac{G_A}{N_t}\right ) - \frac{\alpha_A}{N_t}G_AG_B\\
 \dot{G}_B &=& \alpha_B G_B\left(1-\frac{G_B}{N_t}\right ) -\frac{\alpha_B}{N_t}G_AG_B. 
-\label{eq:LV_simple}
+\label{eq:LV_simple2}
 \end{eqnarray}
 
-Equations (\ref{eq:LV_simple}) are basically competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competive excursion.
+Equations (\ref{eq:LV_simple1}-\ref{eq:LV_simple2}) are basically competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competive excursion.
 That principle claims that two species in competition for the same resource cannot coexist~\cite{wangersky_lotka-volterra_1978}. 
-In previous studies it was shown that gLV equations can be sucessfully applied to predict the microbial community dynamics, determine different types of the pairwise interspecies interactions~\cite{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
-In food studies the gLV was applied by Mounier \etal~\cite{mounier_microbial_2008} to study interactions of 6 bacteria species  with yeast omissin in cheese.
+In previous studies it was shown that gLV equations with one bacterial state can be sucessfully applied to predict the microbial community dynamics, determine different types of the pairwise interspecies interactions~\cite{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
+In food studies the gLV was applied by Mounier \etal~\cite{mounier_microbial_2008} to study interactions of six bacteria species  with yeast omissin in cheese.
+Cauchie \etal~\cite{cauchie_modeling_2020} applied Lotka-Volterra model to study the three-species system of the bacteria majorly present in minced pork meat.
 
@@
 As one can see in Figure~\ref{fig:2pool_resource_2sp} even without direct interaction between species the bacterial growth was inhibited due to the limitation of the resource.
-Here species $A$ was able to supress species $B$ due to the higher values of the growth constant $\alpha$..
+Here species $A$ was able to supress species $B$ due to the higher values of the growth constant $\alpha$.
 
-\cmtb{Look papers on modelling of the LV in bacterial communities. Did anyone do the same as we??}
+\cmt{But LV for several bacteria states in literature? Didn't find!}
 
@@
 
-However, except the inhibition due to resource competition, bacteria developed differnt others active competition mechanisms.
-One of the examples of such mechanisms is a production of the antimicrobial compounds, which can be also called antibiotics or inhibitors. \cmt{toxins}
-
-\cmt{Read how this antibiotics affect the system. Which eq form to use?}
+However, except the inhibition due to resource competition, bacteria developed different others active competition mechanisms.
+One of the examples of such mechanisms is a production of the antimicrobial compounds, inhibitors or toxins~\ref{wloch-salamon_effect_2008, chao_structured_1981}.
 

### 2023-12-04 pool_model.tex
Add references for spatial bacterial modelling

@@
 
+- \ref{gorochowski_bsim_2012, li_nufeb_2019}: examples of other toolboxes for bacterial communities modelling.\\
+
+In general, bacterial communities in 2D can experience collective motion called swarming~\ref{wu_collective_2015}.
+Some works modelled 2D motile bacteria:\\
+- \ref{decoene_microscopic_2011}: 2D chemostatic self-propelled bacterial motion of the flagellated bactreia (e.g. E. coli, B. subtilis).
+Chemostaxis here was driven by the oxygen concentrations.
+
+- \ref{rosser_modelling_2014}: modelling of the bacterial reorintation, rotation.
+
+- \ref{sokolov_physical_2012, li_amplified_2008}: modelling of the bacteria using hydrodynamic interactions and Brownian motion.
+Fluctuations in velocity and rotations.
+(\ref{li_amplified_2008}: Brownian motion gives right motion observed using microscopy).
+
+- \ref{farrell_mechanically_2013}: 2D modelling using Newtonian dynamics and only mechanical interactions + nutritient diffusion.
+Diff limits where diffusion dominant and limits growth, and where it's irrelevant.
+Mechanical pressure (due to bacteria-bacteria interactions) affects growth, can limit it.
+
+- \ref{blanchard_bacterial_2015}: Modelled bacterial social interactions (toxins/public goods release) and its effect on the spatial structure.
+
 \cmtb{Maybe as motivation can be spatial competitive mechanisms as adhesion to survaces and motility. 

### 2023-12-05 pool_model.tex
Spatial limitation ref

--- 2023-12-04 pool_model.tex
@@
 
-...
-
-- See Papers.
-
-- Different processes in bacterial communities. 
-Differents states, lag-phase, growth, stationary state. What processes behind.
-Activation/inhibition/waste production, resource competition, back-flow to lag-phase.
-
-- About known models for Predictive Microbiology Modelling, e.g., Baranyi/Roberts
+Comprehensive reviews about competition/cooperation bacterial mechanisms: \cite{ghoul_ecology_2016, stubbendieck_bacterial_2016, hibbing_bacterial_2010, west_social_2007}
 
@@
 Consider the system consisting of $N$ species in $M$ different states.
-In the general representation, the system can be described with the Ordinary Differential Equations (ODEs) using a non-linear function $\mbF$:
+In the general representation, the system can be described with the Ordinary Differential Equations (\acp{ode}) using a non-linear function $\mbF$:
 \begin{equation}
@@
 
-Then the ODEs describing such a model can be written as 
+Then the \acp{ode} describing such a model can be written as 
 \begin{equation}
@@
 However, except the inhibition due to resource competition, bacteria developed different others active competition mechanisms.
-One of the examples of such mechanisms is a production of the antimicrobial compounds, inhibitors or toxins~\ref{wloch-salamon_effect_2008, chao_structured_1981}.
+One of the examples of such mechanisms is a production of the antimicrobial compounds, inhibitors or toxins~\cite{wloch-salamon_effect_2008, chao_structured_1981}.
 
@@
 
-\cmtb{Maybe as motivation can be spatial competitive mechanisms as adhesion to survaces and motility. 
-So how much including these effects will affect system.}
+As described above, the proposed above model operates with the total number of the bacterial counts for different species and states in the system not accounting for their spatial distributions.
+Assuming that the bacteria is uniformly distributed we can simplify the problem to the system of the \acp{ode}.
+However, in reality spatial effects play significant role for population survival forcing some species develop such competitive mechanism as cell adhesion to surfaces in the areas rich with nutritients~\cite{htuson_bacteriasurface_2013}.
+Another mechanism is the motility of bacteria, e.g., the flagellates bacteria as \textit{E.
+coli} or \textit{B.
+subtilis} can experience the chemostatic motion of driven by oxygen concentrations~\cite{decoene_microscopic_2011}.
+The bacterial social interactions such as releasing the toxins or public goods can also significantly affect the spatial structure of the community~\cite{blanchard_bacterial_2015}.
+Therefore, it is rather important to establish the limitations for the model application.\\
 
-\cmt{\cite{htuson_bacteriasurface_2013}: Attachment of bacteria to surfaces. (One of the mechanisms to stick to the areas with nutritients)}
+- \cite{gorochowski_bsim_2012, li_nufeb_2019}: examples of another toolboxes for bacterial communities modelling (ABM).\\
+More about the modelling of the motile bacteria can be seen in works:~\cite{decoene_microscopic_2011, rosser_modelling_2014, sokolov_physical_2012, li_amplified_2008}.
+
+In general, bacterial communities in 2D can experience collective motion called swarming~\cite{wu_collective_2015}.
+Some works modelled 2D motile bacteria:
+\begin{itemize}
+    \item \cite{decoene_microscopic_2011}: 2D chemostatic self-propelled bacterial motion of the flagellated bactreia (e.g. E. coli, B. subtilis).
+    Chemostaxis here was driven by the oxygen concentrations.
+    \item \cite{rosser_modelling_2014}: modelling of the bacterial reorintation, rotation.
+    \item \cite{sokolov_physical_2012, li_amplified_2008}: modelling of the bacteria using hydrodynamic interactions and Brownian motion.
+    Fluctuations in velocity and rotations.
+    (\cite{li_amplified_2008}: Brownian motion gives right motion observed using microscopy).
+    \item \cite{farrell_mechanically_2013}: 2D modelling using Newtonian dynamics and only mechanical interactions + nutritient diffusion.
+    Branching parameter defines diff limits where diffusion dominant and limits growth, and where it's irrelevant.
+    Mechanical pressure (due to bacteria-bacteria interactions) affects growth, can limit it. (!!!)
+\end{itemize}
 
@@
 Moreover, the species 1 is able to release a certain antimicrobial component $I$ that shows inhibitory effect on species 2 (amensalism).
-The ODEs the bacterial population can be written as: 
+The \acp{ode} the bacterial population can be written as: 
 \begin{equation}

### 2023-12-05 pool_model.tex
formatting of equations and text

@@
 In the following section we demonstate the application of this approach for different toy models describing the processes within and between bacteria communities.
-
 \newpage
-
+%
+%
 \section{One Species Modelling}
 \subsection{Three-Pool Model}
-
 Consider for simplicity the system where only one bacteria species is present $N=1$.
@@
 Here the concentrations of the pools are components of our variable vector $\mbx (t) = (L(t), G(t), D(t))^T$.
-
 \begin{figure}[t]
@@
 \end{figure}
-
 To detremine the function $\mbF(\mbx, \mbm, \mbp, \mbu)$ consider the flows between different states.
@@
 \end{eqnarray}
-
 Then the \acp{ode} describing such a model can be written as 
@@
 \end{equation}
-
+%
+%
 \subsection{Resource Dependence}
-
 Based on experimental observations, it was shown that the bacterial growth is strongly influenced by the limiting resource.
@@
 Using release of the extracellular signalling molecules (autoinducers) it is possible to synchronize the gene expression of the whole group~\cite{ng_bacterial_2009}.
-
 We would not dive into the processes of quorum sensing and try to keep it as simple as possible.
@@
 \end{equation}
-
 Note that the pools $L$, $G$ and $D$ represent number of bacteria (or number of bacteria per unit area/volume) while $R_0$ represents an abstract resource pool. 
@@
         \mu              & \mu'            & -\beta 
-                                                \end{pmatrix} \mbx  \\
+    \end{pmatrix} \mbx\\
     \dot \mbm &=& \mbH (\mbx, \mbm, \mbp, \mbu) = -\frac{\alpha}{N_t} R G.
@@
 Here the first term is the linear part and the second term define the quadratic behaviour of the system.
-
 \begin{figure}[t]
@@
     \includegraphics[width=0.6\textwidth]{Figures/pool_model_3pools_resource.pdf}
-    \caption{{\footnotesize The time dependence of the total bacterial count $N(t) = \sum_i \sum_k \mbx_i^k = L(t) + G(t) + D(t)$ is obtained as a sum of bacteria in all states.
+    \caption{{
+        \footnotesize The time dependence of the total bacterial count $N(t) = \sum_i \sum_k \mbx_i^k = L(t) + G(t) + D(t)$ is obtained as a sum of bacteria in all states.
         All curves were calculated using the model (\ref{eq:3pool_resource}) with parameters $\alpha=2.5$, $\mu'=0.5$ and $\beta=5$.
@@
 \end{figure}
-
+%
+%
 \subsection{Lag-time Estimation}
-
 Note that one could further simplify this model to a two pool or state model by omitting the pool $D$ and removing cells with rate $\beta$ directly from pool $L$. 
@@
 The lag time $t_L$ diverges very slowly, logarithmically with $\lambda\to 0$.
-
+%
+%
 \subsection{Back flow from Growth to Lag pool}
-
 Due to its life in rather unfavourable conditions bacteria developed different surviving mechanisms, one of which is adapting their metabolism and transitioning into dormant state.
@@
 The models including transition of bacteria to the dormant state were proposed and studied by Bär \etal~\cite{bar_modelling_2002} and Jones and Lennon~\cite{jones_dormancy_2010}.
-
+%
 \cmt{So now is it better to use flow to dormancy instead?}
-
+%
 We explore now the idea that under certain conditions, as sudden change in the environment, part of the population $G$ will enter a lag-phase, i.e., there is a back-flow from $G$ to $L$. 
@@
 Stronger and quicker changes will lead to a higher rate $\gamma$. 
+%
+%
 \subsubsection{Maxwell type of stress - strain relation}
-
 One of the reasons of the dormant state appearence can be also the temperature shift~\cite{oliver_viable_1995}.
@@
 \end{figure}
-
 The ordinary differential equations for the pool model with environmental-shock based back-flow read:
@@
 \end{equation}
-
 \newpage
-
+%
+%
 \section{Interaction between species}
-
 As it was mentioned above, bacteria usually coexist in large multispecies interacting communities whose growth is vastly determined by the developed survival mechanisms.
@@
 Further we show that the TPM can be readily extended to the systems with more than one bacteria an take describe more complex relationships between species. 
-
+%
+%
 \subsection{Resource Competition}
-
 One on the ways to descibe competition between bacteria populations is as a function of limiting resources ratio~\cite{tilman_resource_1977, smith_effects_2002}.
@@
 The simplest pool model for bacterial species $A$ and $B$ encompassing lag-phases, growth-phases and a common resource pool is given by (no back-flow to the lag-pool):
-
 \begin{equation}
@@
 \begin{eqnarray*}
-\dot{L}_A &=& - \lambda_A  L_A + \frac{\lambda_A}{N_t}\left(L_A+G_A+L_B+G_B\right )L_A\\
-\dot{L}_B &=& - \lambda_B L_B + \frac{\lambda_B}{N_t}\left(L_A+G_A+L_B+G_B\right )L_B \\
-\dot{G}_A &=&  \lambda_A  L_A - \frac{\lambda_A}{N_t}\left(L_A+G_A+L_B+G_B\right )L_A +\alpha_A G_A - \frac{\alpha_A}{N_t}\left(L_A+G_A+L_B+G_B\right )G_A\\
-\dot{G}_B &=& \lambda_B L_B - \frac{\lambda_B}{N_t}\left(L_A+G_A+L_B+G_B\right )L_B  +\alpha_B G_B -\frac{\alpha_B}{N_t}\left(L_A+G_A+L_B+G_B\right )G_B
+    \dot{L}_A &=& - \lambda_A  L_A + \frac{\lambda_A}{N_t}\left(L_A+G_A+L_B+G_B\right)L_A\\
+    \dot{L}_B &=& - \lambda_B L_B + \frac{\lambda_B}{N_t}\left(L_A+G_A+L_B+G_B\right)L_B \\
+    \dot{G}_A &=&  \lambda_A  L_A - \frac{\lambda_A}{N_t}\left(L_A+G_A+L_B+G_B\right)L_A +\alpha_A G_A - \frac{\alpha_A}{N_t}\left(L_A+G_A+L_B+G_B\right)G_A\\
+    \dot{G}_B &=& \lambda_B L_B - \frac{\lambda_B}{N_t}\left(L_A+G_A+L_B+G_B\right)L_B  +\alpha_B G_B -\frac{\alpha_B}{N_t}\left(L_A+G_A+L_B+G_B\right)G_B
 \end{eqnarray*}
-
 which has the structure of a generalised Lotka-Volterra system (gLV).
@@
 \begin{eqnarray}
-\label{eq:LV_simple1}
-\dot{G}_A &=& \alpha_A G_A\left(1 - \frac{G_A}{N_t}\right ) - \frac{\alpha_A}{N_t}G_AG_B\\
-\dot{G}_B &=& \alpha_B G_B\left(1-\frac{G_B}{N_t}\right ) -\frac{\alpha_B}{N_t}G_AG_B. 
+    \dot{G}_A &=& \alpha_A G_A\left(1 - \frac{G_A}{N_t}\right) - \frac{\alpha_A}{N_t}G_AG_B
+    \label{eq:LV_simple1}\\
+    \dot{G}_B &=& \alpha_B G_B\left(1-\frac{G_B}{N_t}\right) -\frac{\alpha_B}{N_t}G_AG_B. 
     \label{eq:LV_simple2}
@@
 To see this we note $\dot{T}_2=\kappa(\dot{L}+\dot{G})$, which yields: $I_2(t)=\kappa(L(t)+G(t)-n_0)$ where we used $T_2(0)=G(0)=0$, $L(0)=n_0$. 
-This model assumes that inhibitor is only produced when the bacteria grow and divide. \\
+This model assumes that inhibitor is only produced when the bacteria grow and divide.\\
 
@@
 \begin{eqnarray}
-\dot{G}_A &=& \mathcal{B}_A(G_B)G_A\left(1 - \frac{G_A+G_B}{N_t}\right )\\
-\dot{G}_B &=& \mathcal{B}_B(G_A) G_B\left(1-\frac{G_A+G_B}{N_t}\right ). 
+    \dot{G}_A &=& \mathcal{B}_A(G_B)G_A\left(1 - \frac{G_A+G_B}{N_t}\right)\\
+    \dot{G}_B &=& \mathcal{B}_B(G_A) G_B\left(1-\frac{G_A+G_B}{N_t}\right). 
 \end{eqnarray}
@@
 \begin{eqnarray}
-\dot{G}_A &=& \frac{\alpha_A G_A}{1+K_BG_B}\left(1 - \frac{G_A+G_B}{N_t}\right )\\
-\dot{G}_B &=& \frac{\alpha_B G_B}{1+K_AG_A}\left(1-\frac{G_A+G_B}{N_t}\right ). 
+    \dot{G}_A &=& \frac{\alpha_A G_A}{1+K_BG_B}\left(1 - \frac{G_A+G_B}{N_t}\right)\\
+    \dot{G}_B &=& \frac{\alpha_B G_B}{1+K_AG_A}\left(1-\frac{G_A+G_B}{N_t}\right). 
 \end{eqnarray}
@@
 \label{Mut_Inhib}
-\dot{G}_A &=&\tilde{\alpha}_A\frac{G_A}{G_B}\left(1 - G_A-G_B\right )\\
-\dot{G}_B &=& \tilde{\alpha}_B\frac{G_B}{G_A}\left(1-G_A-G_B\right ). 
+    \dot{G}_A &=&\tilde{\alpha}_A\frac{G_A}{G_B}\left(1 - G_A-G_B\right)\\
+    \dot{G}_B &=& \tilde{\alpha}_B\frac{G_B}{G_A}\left(1-G_A-G_B\right). 
 \end{eqnarray}
-We introduced the rescaled growth rates $\tilde{\alpha}_A=\alpha_A/(N_tK_B)$ and $\tilde{\alpha}_B=\alpha_B/(N_tK_A)$.
-Note that $[\tilde{\alpha}]=\mathrm{s}^{-1}$.
+We introduced the rescaled growth rates $\tilde{\alpha}_A=\alpha_A/(N_tK_B)$ and $\tilde{\alpha}_B=\alpha_B/(N_tK_A)$. Note that $[\tilde{\alpha}]=\mathrm{s}^{-1}$.
 Rescaling time with $\tilde{\alpha}_A$ and defining the dimensionless parameter 
@@
 \begin{eqnarray}
-\dot{G}_A &=&\frac{G_A}{G_B}\left(1 - G_A-G_B\right )
+    \dot{G}_A &=&\frac{G_A}{G_B}\left(1 - G_A-G_B\right)
     \label{eq:2sp_1pool_inhibit_unitless1}\\
-\dot{G}_B &=&\psi\frac{G_B}{G_A}\left(1-G_A-G_B\right ). 
+    \dot{G}_B &=&\psi\frac{G_B}{G_A}\left(1-G_A-G_B\right). 
     \label{eq:2sp_1pool_inhibit_unitless2}
@@
 \end{eqnarray}
-
+%
 \begin{figure}[H]
@@
 \end{figure}
-
+%
+%
 \subsection{Interspecies Cooperation}
-
 On the other hand, one can also observe some cooperative behaviours between species in case when it facilitates the competition with others.
@@
 Sometimes neighbours, so-called 'social cheaters', can benefit from a certain behaviour without expressing it by itself \cite{rainey_evolution_2003}.
-
-\cmt{\cite{rocha_base_2002}: Symbiosis host/bacteria. bacterial genomes suffering reductive evolution due to this. 
+%
+\cmt{\cite{rocha_base_2002}: Symbiosis host/bacteria bacterial genomes suffering reductive evolution due to this. 
 Maybe it's connected with 'social cheaters'?}
-
+%
 \cmt{Production of public goods!}
-
+%
 The benefit, or mutual activation, of the two species from the presence of the other can be simply introduced by mathematical functions. 
@@
 \begin{eqnarray}
-\dot{G}_A &=& W_BG_A\left(1 - G_A-G_B\right )\\
-\dot{G}_B &=& \phi W_AG_B\left(1 - G_A-G_B\right )\\
-\dot{W}_A &=& G_A\left(1 - G_A-G_B\right )\\
-\dot{W}_B &=&\psi G_B\left(1 - G_A-G_B\right ).
+    \dot{G}_A &=& W_BG_A\left(1 - G_A-G_B\right)\\
+    \dot{G}_B &=& \phi W_AG_B\left(1 - G_A-G_B\right)\\
+    \dot{W}_A &=& G_A\left(1 - G_A-G_B\right)\\
+    \dot{W}_B &=&\psi G_B\left(1 - G_A-G_B\right).
 \end{eqnarray}
-
+%
+%
 \section{Spatial Limitations}
-
 As described above, the proposed above model operates with the total number of the bacterial counts for different species and states in the system not accounting for their spatial distributions.
@@
     Branching parameter defines diff limits where diffusion dominant and limits growth, and where it's irrelevant.
-    Mechanical pressure (due to bacteria-bacteria interactions) affects growth, can limit it. (!!!)
+    Mechanical pressure (due to bacteria-bacteria interactions) affects growth, can limit it.(!!!)
 \end{itemize}
-
+%
 To test the spatial limitations of the described above approach consider the system consisting from two bacterial species thet can be described bt two internal states: lag and growth states. 
@@
 \end{equation}
-
+%
 The microenvironment of our system contains the resource and the inhibitor molecules:
@@
 \end{equation}
-
+%
 \begin{table}[H]
@@
 \end{table}
-
+%
 \begin{table}[H]
@@
 \end{table}
-
+%
 \begin{figure}[H]
@@
 \end{figure}
-
+%
 \cmtb{The ABM additional parameters $D_R$, $D_I$, $q$ (At which level of $R_i$ cell division happens) and more parameters describing bacterial division.}
-
+%
+%
 \section{Conclusion}
-
 \newpage
 \printbibliography
-
+%
 \end{document}

### 2023-12-05 pool_model.tex
more formatting to visually distinguish between ((sub)sub)section

@@
 \newpage
-
+%
+%
+%
+%
+% ================================================================================================
 \section{Introduction}
@@
 Comprehensive reviews about competition/cooperation bacterial mechanisms: \cite{ghoul_ecology_2016, stubbendieck_bacterial_2016, hibbing_bacterial_2010, west_social_2007}
-
+%
+%
+%
+%
+% ================================================================================================
 \section{Motivating Example}
@@
 \cmtb{Maybe Example from Seminar by Jens? 2 species,  1 resource, 1 waste production.}
-
+%
+%
+%
+%
+% ================================================================================================
 \section{General Mathemetical Formulation}
@@
 %
+%
+%
+% ================================================================================================
 \section{One Species Modelling}
+% ------------------------------------------------------------------------------------------------
 \subsection{Three-Pool Model}
@@
 %
+%
+% ------------------------------------------------------------------------------------------------
 \subsection{Resource Dependence}
@@
 \end{eqnarray}
-
+%
 Such model can be simplified by explointing that limitation of the resource pool also limits the total bacterial count of the system:
@@
 %
+%
+% ------------------------------------------------------------------------------------------------
 \subsection{Lag-time Estimation}
@@
 %
+%
+% ------------------------------------------------------------------------------------------------
 \subsection{Back flow from Growth to Lag pool}
@@
 %
+% ................................................................................................
 \subsubsection{Maxwell type of stress - strain relation}
@@
 %
+%
+% ================================================================================================
 \section{Interaction between species}
@@
 %
+%
+% ------------------------------------------------------------------------------------------------
 \subsection{Resource Competition}
@@
 Here species $A$ was able to supress species $B$ due to the higher values of the growth constant $\alpha$.
-
+%
 \cmt{But LV for several bacteria states in literature? Didn't find!}
-
+%
+%
+%
+% ------------------------------------------------------------------------------------------------
 \subsection{Interspecies Competition}
-
 However, except the inhibition due to resource competition, bacteria developed different others active competition mechanisms.
 One of the examples of such mechanisms is a production of the antimicrobial compounds, inhibitors or toxins~\cite{wloch-salamon_effect_2008, chao_structured_1981}.
-
+%
+%
+% ................................................................................................
 \subsubsection{Toxins Production}
@@
 \end{figure}
-
+%
+%
+% ................................................................................................
 \subsubsection{Species Inhibition}
@@
 %
+%
+% ------------------------------------------------------------------------------------------------
 \subsection{Interspecies Cooperation}
@@
 %
+%
+%
+% ================================================================================================
 \section{Spatial Limitations}
@@
 %
+% ================================================================================================
 \section{Conclusion}

### 2023-12-05 pool_model.tex
more formatting of equations and some imports

@@
 \documentclass[10pt,A4paper]{article}
-\usepackage[a4paper,
+\usepackage[
+    a4paper,
     inner=2.5cm,
@@
     \includegraphics[width=0.6\textwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}
-    \caption{{\footnotesize (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stres (\ref{eq:ODE_tempshift}).
+    \caption{{
+        \footnotesize (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stres (\ref{eq:ODE_tempshift}).
         (b) The value of the corresponding back-flow rate $\gamma$ calculated according to (\ref{eq:gamma_tempshift}).
@@
     \includegraphics[width=0.67\textwidth]{Figures/pool_model_2pools_resource_competition.pdf}
-    \caption{{\footnotesize The competition of the two bacterial species due to the common limiting resource calculated using Eq. (\ref{eq:model_2sp_resource_comp}).
+    \caption{{
+        \footnotesize The competition of the two bacterial species due to the common limiting resource calculated using Eq. (\ref{eq:model_2sp_resource_comp}).
         The red solid curve describes the growth of the bacteria $A$ $N_A = \sum_{k} x_A^k = L_A+G_A$ with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
@@
         The purple dashed line shows the total bacterial count of the system  $N(t)=\sum_i\sum_k\mbx_i^k=N_A+N_B$.
-    The gray solid line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.}}
+        The gray solid line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
+    }}
     \label{fig:2pool_resource_2sp}
@@
     \includegraphics[width=1.\textwidth]{Figures/pool_model_2pools_toxin.pdf}
-    \caption{{\footnotesize The two species system competing for the common nutritient $R$ with production of the toxin $T$ (brown solid line) calculated using Eq. (\ref{eq:model_2sp_toxin}).
+    \caption{{
+        \footnotesize The two species system competing for the common nutritient $R$ with production of the toxin $T$ (brown solid line) calculated using Eq. (\ref{eq:model_2sp_toxin}).
         The growing bacteria $B$ produces toxin $T$ (a) constantly $\dot{T} = \mathcal{F}_1 = \kappa G_B$ and (b) only during growth process $\dot{T} = \mathcal{F}_2 = \kappa \alpha_B R G_B$ with 
@@
         The blue solid curve describes the growth of the bacteria $B$ $N_B = \sum_{k} x_B^k = L_B+G_B$ with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
-    The gray solid line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.}}
+        The gray solid line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
+    }}
     \label{fig:2pool_2sp_toxin}
@@
     \includegraphics[width=0.5\textwidth]{Figures/pool_model_1pool_inhib.pdf}
-    \caption{{\footnotesize The two species system where species inhibit each other defined with equations~(\ref{eq:2sp_1pool_inhibit_unitless1}-\ref{eq:2sp_1pool_inhibit_unitless2}).
+    \caption{{
+        \footnotesize The two species system where species inhibit each other defined with equations~(\ref{eq:2sp_1pool_inhibit_unitless1}-\ref{eq:2sp_1pool_inhibit_unitless2}).
         Here the red lines describes the evolution of the species $A$ and blue lines correspond to species $B$.
-    The relation between inhibition strength of spesies $B$ and $A$ are defined by controll parameter value (a) $\psi$ = 0.5, (b) $\psi$ = 1 or (c) $\psi$ = 2.}}
+        The relation between inhibition strength of spesies $B$ and $A$ are defined by controll parameter value (a) $\psi$ = 0.5, (b) $\psi$ = 1 or (c) $\psi$ = 2.
+    }}
     \label{fig:1pool_2sp_inhibit}
@@
     \includegraphics[width=0.67\textwidth]{Figures/pool_model_spatial.pdf}
-    \caption{{\footnotesize The output of the model (\ref{eq:spatial_limit_F}, \ref{eq:spatial_limit_H}) for given parameters and initial valued presented in Tables \ref{tab:spatial_limit_param}, \ref{tab:spatial_limit_init}, respectively.
-    The observables of the system are the bacterial count of each species $L+G$ (blue and orange lines) and the total bacterial count of the system (green).}}
+    \caption{{
+        \footnotesize The output of the model (\ref{eq:spatial_limit_F}, \ref{eq:spatial_limit_H}) for given parameters and initial valued presented in Tables \ref{tab:spatial_limit_param}, \ref{tab:spatial_limit_init}, respectively.
+        The observables of the system are the bacterial count of each species $L+G$ (blue and orange lines) and the total bacterial count of the system (green).
+    }}
     \label{fig:spatial_limit_plot1}

### 2023-12-05 pool_model.tex
More citations for Bacteria modelling and ABM

@@
 
-- \cite{gorochowski_bsim_2012, li_nufeb_2019}: examples of another toolboxes for bacterial communities modelling (ABM).\\
+- \cite{emonet_agentcell_2005, gorochowski_bsim_2012, li_nufeb_2019, kreft_bacsim_1998}: examples of another toolboxes for bacterial communities modelling (ABM).\\
 More about the modelling of the motile bacteria can be seen in works:~\cite{decoene_microscopic_2011, rosser_modelling_2014, sokolov_physical_2012, li_amplified_2008}.
@@
     Mechanical pressure (due to bacteria-bacteria interactions) affects growth, can limit it.(!!!)
+\end{itemize}
+%
+For simulation results:
+\begin{itemize}
+    \item \cite{farrell_mechanically_2013}: Showed the influence of mechanic interactions between bacteria in 2D system with bacteria consuming the nutritient.
+    The growth of bacterial colonies where nonmotile microorganisms replicate and push each other away as they grow.
+    Found a transition between two different growth regimes, controlled by the balance between growth and uptake of nutrient:\\
+    Exponential growth where diffusion of the nutritient does not play any role. Sublinear growth where growth is limited by nutritient diffusion.
+    (Structurally in 2D it's regimes: smooth propagating of bacterial front vs. branching.)
+    \item \cite{nagarajan_agent-based_2022-1}: review on the ABM for bacterial population.
+    The ABM represent a microbial colony as a group of discrete agents that are governed by a set of rules, where each agent can be an individual cell or a collection of cells.
+    Can use citation when talk about general processes we are accounting as growth, uptake of nutritient, production of inhibitor, e.g.,:
+    Physical interactions between agents are calculated using excluded volume interactions.
+    Chemical interactions as releasing inhibitors are described using concentrations and solved using PDEs.
 \end{itemize}

### 2023-12-05 pool_model.tex
Change color palette for the Figures

@@
 \end{equation}
-Here the first term is the linear part and the second term define the quadratic behaviour of the system.
+Here the first term is the linear part and the second term define the quadratic behavior of the system.
 \begin{figure}[t]
@@
         All curves were calculated using the model (\ref{eq:3pool_resource}) with parameters $\alpha=2.5$, $\mu'=0.5$ and $\beta=5$.
-        The purple curve shows significant dip in behaviour of the system where the death of the bacteria is present $\mu = 0.5$.
+        The blue curve shows significant dip in behavior of the system where the death of the bacteria is present $\mu = 0.5$.
         Other curves show the case of no bacteria dying $\mu=0$ and represent different length of lag phases:
-        short $\lambda=10^{-2}$ (red), intermediate $\lambda=10^{-3}$ (blue) and long one $\lambda=10^{-4}$ (green).
+        short $\lambda=10^{-2}$ (dark green), intermediate $\lambda=10^{-3}$ (light green) and long one $\lambda=10^{-4}$ (yellow).
     }}
@@
         (b) The value of the corresponding back-flow rate $\gamma$ calculated according to (\ref{eq:gamma_tempshift}).
-        The curves represent no temperature shift (blue), one temperature jump $\Delta T = 10$ at time $t=4$ (red) 
-        and the series of different temperature shifts of value 10, 5 and 15 at the tme points 1, 3 asnd 7, respectively (green).
+        The curves represent no temperature shift (green), one temperature jump $\Delta T = 10$ at time $t=4$ (orange) 
+        and the series of different temperature shifts of value 10, 5 and 15 at the time points 1, 3 and 7, respectively (rose).
         The paremeters of the back-flow rate are $\Gamma=1$ and $\delta=5$.
@@
         \footnotesize The competition of the two bacterial species due to the common limiting resource calculated using Eq. (\ref{eq:model_2sp_resource_comp}).
-        The red solid curve describes the growth of the bacteria $A$ $N_A = \sum_{k} x_A^k = L_A+G_A$ with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
-        The blue solid curve describes the growth of the bacteria $B$ $N_B = \sum_{k} x_B^k = L_B+G_B$ with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
-        The purple dashed line shows the total bacterial count of the system  $N(t)=\sum_i\sum_k\mbx_i^k=N_A+N_B$.
-        The gray solid line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
+        The red curve describes the growth of the bacteria $A$ $N_A = \sum_{k} x_A^k = L_A+G_A$ with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
+        The blue curve describes the growth of the bacteria $B$ $N_B = \sum_{k} x_B^k = L_B+G_B$ with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
+        The purple line shows the total bacterial count of the system  $N(t)=\sum_i\sum_k\mbx_i^k=N_A+N_B$.
+        The gray line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
     }}
@@
     \caption{{
-        \footnotesize The two species system competing for the common nutritient $R$ with production of the toxin $T$ (brown solid line) calculated using Eq. (\ref{eq:model_2sp_toxin}).
+        \footnotesize The two species system competing for the common nutritient $R$ with production of the toxin $T$ (brown line) calculated using Eq. (\ref{eq:model_2sp_toxin}).
         The growing bacteria $B$ produces toxin $T$ (a) constantly $\dot{T} = \mathcal{F}_1 = \kappa G_B$ and (b) only during growth process $\dot{T} = \mathcal{F}_2 = \kappa \alpha_B R G_B$ with 
@@
         The toxin $T$ causes the cell death of bacteria $A$ with rate constant $\mu = 0.2$.
-        The red solid curve describes the growth of the bacteria $A$ $N_A = \sum_{k} x_A^k = L_A+G_A$ with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
-        The blue solid curve describes the growth of the bacteria $B$ $N_B = \sum_{k} x_B^k = L_B+G_B$ with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
-        The gray solid line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
+        The red curve describes the growth of the bacteria $A$ $N_A = \sum_{k} x_A^k = L_A+G_A$ with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
+        The blue curve describes the growth of the bacteria $B$ $N_B = \sum_{k} x_B^k = L_B+G_B$ with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
+        The gray line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
     }}
@@
         Here the red lines describes the evolution of the species $A$ and blue lines correspond to species $B$.
-        The relation between inhibition strength of spesies $B$ and $A$ are defined by controll parameter value (a) $\psi$ = 0.5, (b) $\psi$ = 1 or (c) $\psi$ = 2.
+        The relation between inhibition strength of species $B$ and $A$ are defined by control parameter value (a) $\psi$ = 0.5, (b) $\psi$ = 1 or (c) $\psi$ = 2.
     }}

### 2023-12-05 pool_model.tex
caption size formatting

@@
 As a result the balance between the growing and dying processes leads to the constant with time bacterial count~\cite{navarro_llorens_stationary_2010}.
-In the described simplified version of the real-life system three possible bacteria states are mentioned corresponding to which process is happeng: adaplation, growth or death.
+In the described simplified version of the real-life system three possible bacteria states are mentioned corresponding to which process is happening: adaptation, growth or death.
 Motivated by this we divide a bacterial population into three pools (states) (\textbf{Figure \ref{fig:SchematicRep}}): 
@@
     \includegraphics[width=0.7\textwidth]{Figures/TPM_fig.pdf}
-    \caption{{\footnotesize Schematic representation of exchange between three states or pools.}}
+    \caption{Schematic representation of exchange between three states or pools.}
     \label{fig:SchematicRep}
@@
     \includegraphics[width=0.6\textwidth]{Figures/pool_model_3pools_resource.pdf}
-    \caption{{
-        \footnotesize The time dependence of the total bacterial count $N(t) = \sum_i \sum_k \mbx_i^k = L(t) + G(t) + D(t)$ is obtained as a sum of bacteria in all states.
+    \caption{
+        The time dependence of the total bacterial count $N(t) = \sum_i \sum_k \mbx_i^k = L(t) + G(t) + D(t)$ is obtained as a sum of bacteria in all states.
         All curves were calculated using the model (\ref{eq:3pool_resource}) with parameters $\alpha=2.5$, $\mu'=0.5$ and $\beta=5$.
@@
         short $\lambda=10^{-2}$ (dark green), intermediate $\lambda=10^{-3}$ (light green) and long one $\lambda=10^{-4}$ (yellow).
-    }}
+    }
     \label{fig:3pool_resource_plots}
@@
     \includegraphics[width=0.6\textwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}
-    \caption{{
-        \footnotesize (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stres (\ref{eq:ODE_tempshift}).
+    \caption{
+        (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stres (\ref{eq:ODE_tempshift}).
         (b) The value of the corresponding back-flow rate $\gamma$ calculated according to (\ref{eq:gamma_tempshift}).
@@
         The other model parameters are $\alpha=2.5$, $\mu'=0.5$, $\beta=5$, $\lambda=10^{-2}$, $\mu = 0$ fe each curve.
-    }}
+    }
     \label{fig:TempJump}
@@
     \includegraphics[width=0.67\textwidth]{Figures/pool_model_2pools_resource_competition.pdf}
-    \caption{{
-        \footnotesize The competition of the two bacterial species due to the common limiting resource calculated using Eq. (\ref{eq:model_2sp_resource_comp}).
+    \caption{
+        The competition of the two bacterial species due to the common limiting resource calculated using Eq. (\ref{eq:model_2sp_resource_comp}).
         The red curve describes the growth of the bacteria $A$ $N_A = \sum_{k} x_A^k = L_A+G_A$ with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
@@
         The gray line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
-    }}
+    }
     \label{fig:2pool_resource_2sp}
@@
     \includegraphics[width=1.\textwidth]{Figures/pool_model_2pools_toxin.pdf}
-    \caption{{
-        \footnotesize The two species system competing for the common nutritient $R$ with production of the toxin $T$ (brown line) calculated using Eq. (\ref{eq:model_2sp_toxin}).
+    \caption{
+        The two species system competing for the common nutritient $R$ with production of the toxin $T$ (brown line) calculated using Eq. (\ref{eq:model_2sp_toxin}).
         The growing bacteria $B$ produces toxin $T$ (a) constantly $\dot{T} = \mathcal{F}_1 = \kappa G_B$ and (b) only during growth process $\dot{T} = \mathcal{F}_2 = \kappa \alpha_B R G_B$ with 
@@
         The gray line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
-    }}
+    }
     \label{fig:2pool_2sp_toxin}
@@
     \includegraphics[width=0.5\textwidth]{Figures/pool_model_1pool_inhib.pdf}
-    \caption{{
-        \footnotesize The two species system where species inhibit each other defined with equations~(\ref{eq:2sp_1pool_inhibit_unitless1}-\ref{eq:2sp_1pool_inhibit_unitless2}).
+    \caption{
+        The two species system where species inhibit each other defined with equations~(\ref{eq:2sp_1pool_inhibit_unitless1}-\ref{eq:2sp_1pool_inhibit_unitless2}).
         Here the red lines describes the evolution of the species $A$ and blue lines correspond to species $B$.
         The relation between inhibition strength of species $B$ and $A$ are defined by control parameter value (a) $\psi$ = 0.5, (b) $\psi$ = 1 or (c) $\psi$ = 2.
-    }}
+    }
     \label{fig:1pool_2sp_inhibit}
@@
     \end{tabular}
-    \caption{{\footnotesize The list of the parameters of the system.}}
+    \caption{The list of the parameters of the system.}
     \label{tab:spatial_limit_param}
@@
     \end{tabular}
-    \caption{{\footnotesize The list of the initial values of the system.}}
+    \caption{The list of the initial values of the system.}
     \label{tab:spatial_limit_init}
@@
     \includegraphics[width=0.67\textwidth]{Figures/pool_model_spatial.pdf}
-    \caption{{
-        \footnotesize The output of the model (\ref{eq:spatial_limit_F}, \ref{eq:spatial_limit_H}) for given parameters and initial valued presented in Tables \ref{tab:spatial_limit_param}, \ref{tab:spatial_limit_init}, respectively.
+    \caption{
+        The output of the model (\ref{eq:spatial_limit_F}, \ref{eq:spatial_limit_H}) for given parameters and initial valued presented in Tables \ref{tab:spatial_limit_param}, \ref{tab:spatial_limit_init}, respectively.
         The observables of the system are the bacterial count of each species $L+G$ (blue and orange lines) and the total bacterial count of the system (green).
-    }}
+    }
     \label{fig:spatial_limit_plot1}

### 2023-12-06 pool_model.tex
Correct some small typos

@@
 % ================================================================================================
-\section{One Species Modelling}
+\section{One Species Modeling}
 % ------------------------------------------------------------------------------------------------
@@
 After this, the succeeded in adaptation bacteria enter the exponential growth phase, where the process of cell division is directly happening \cmt{Add Ref}.
-When the nutrition is exhausted by the growing bacteria, the stationary state of the growth curve occurres.
+When the nutrition is exhausted by the growing bacteria, the stationary state of the growth curve occurs.
 In this state the cultures accumulate waste products of metabolitic activity that leads to cell death of the bacteria.
@@
 \end{figure}
-To detremine the function $\mbF(\mbx, \mbm, \mbp, \mbu)$ consider the flows between different states.
+To determine the function $\mbF(\mbx, \mbm, \mbp, \mbu)$ consider the flows between different states.
 The pools $G$ and $D$ are unidirectionally connected with the pool $L$. 
-For now we assume for simplicity that there is no back flow between growth state $G$ and lag-state $L$. 
+For now, we assume for simplicity that there is no back flow between growth state $G$ and lag-state $L$. 
 The suggested model can be cast in the following reactions: \cmt{Use notations x ??}
@@
 \end{equation}
-Note that the pools $L$, $G$ and $D$ represent number of bacteria (or number of bacteria per unit area/volume) while $R_0$ represents an abstract resource pool. 
-Let say $G=N_B/V$ and $[G]=1/V$, where $N_B$ is number of bacteria and $V$ is the volume. 
+Note that the pools $L$, $G$ and $D$ represent the number of bacteria (or number of bacteria per unit area/volume) while $R_0$ represents an abstract resource pool. 
+Let say $G=N_B/V$ and $[G]=1/V$, where $N_B$ is the number of bacteria and $V$ is the volume. 
 Then $R_0=N_R/V$ and $[R_0]=1/V$, where $N_R$ is the number of resource molecules (or any appropriate unit, e.g., mol). 
@@
 Due to its life in rather unfavourable conditions bacteria developed different surviving mechanisms, one of which is adapting their metabolism and transitioning into dormant state.
-By defeinition dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\cite{kaprelyants_dormancy_1993}. 
+By definition dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\cite{kaprelyants_dormancy_1993}. 
 This survival strategy can appear in response to different external stresses, e.g., nutritient deficiency, temperature stress or extreme fluctuations of environment parameters.
@@
 \subsubsection{Maxwell type of stress - strain relation}
-One of the reasons of the dormant state appearence can be also the temperature shift~\cite{oliver_viable_1995}.
-Moreover, the rapid shift in temperature is associated with the appearence of an additional lag phase~\cite{zwietering_modeling_1994}.
+One of the reasons of the dormant state appearance can be also the temperature shift~\cite{oliver_viable_1995}.
+Moreover, the rapid shift in temperature is associated with the appearance of an additional lag phase~\cite{zwietering_modeling_1994}.
 To model it we propose a kind of visco-elastic Maxwell type of stress-strain relation and write down the ordinary differential equation for $\gamma$ to be:
@@
     \caption{
-        (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stres (\ref{eq:ODE_tempshift}).
+        (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stress~(\ref{eq:ODE_tempshift}).
         (b) The value of the corresponding back-flow rate $\gamma$ calculated according to (\ref{eq:gamma_tempshift}).
@@
         and the series of different temperature shifts of value 10, 5 and 15 at the time points 1, 3 and 7, respectively (rose).
-        The paremeters of the back-flow rate are $\Gamma=1$ and $\delta=5$.
+        The parameters of the back-flow rate are $\Gamma=1$ and $\delta=5$.
         The other model parameters are $\alpha=2.5$, $\mu'=0.5$, $\beta=5$, $\lambda=10^{-2}$, $\mu = 0$ fe each curve.
@@
 These survival mechanisms mostly aim to secure as much available nutrition which lead to competitive strategies or, if it is advantageous, cooperative ones~\cite{hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
-Further we show that the TPM can be readily extended to the systems with more than one bacteria an take describe more complex relationships between species. 
+Further we show that the TPM can be readily extended to the systems with more than one bacteria and describe more complex relationships between species. 
 %
@@
 \subsection{Resource Competition}
-One on the ways to descibe competition between bacteria populations is as a function of limiting resources ratio~\cite{tilman_resource_1977, smith_effects_2002}.
-To descibe such a situation we consider for simplicity only one common limiting resource pool. 
+One on the ways to describe competition between bacteria populations is as a function of limiting resources ratio~\cite{tilman_resource_1977, smith_effects_2002}.
+To describe such a situation we consider for simplicity only one common limiting resource pool. 
 The simplest pool model for bacterial species $A$ and $B$ encompassing lag-phases, growth-phases and a common resource pool is given by (no back-flow to the lag-pool):
@@
 \end{eqnarray*}
-which has the structure of a generalised Lotka-Volterra system (gLV).
+which has the structure of a \ac{glv}.
 In order to better see this, we set $L_A=L_B=0$: 
@@
 That principle claims that two species in competition for the same resource cannot coexist~\cite{wangersky_lotka-volterra_1978}. 
-In previous studies it was shown that gLV equations with one bacterial state can be sucessfully applied to predict the microbial community dynamics, determine different types of the pairwise interspecies interactions~\cite{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
-In food studies the gLV was applied by Mounier \etal~\cite{mounier_microbial_2008} to study interactions of six bacteria species  with yeast omissin in cheese.
+In previous studies it was shown that \ac{glv} equations with one bacterial state can be sucessfully applied to predict the microbial community dynamics, determine different types of the pairwise interspecies interactions~\cite{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
+In food studies the \ac{glv} was applied by Mounier \etal~\cite{mounier_microbial_2008} to study interactions of six bacteria species  with yeast omissin in cheese.
 Cauchie \etal~\cite{cauchie_modeling_2020} applied Lotka-Volterra model to study the three-species system of the bacteria majorly present in minced pork meat.
@@
 \subsubsection{Species Inhibition}
-\cmt{\cite{lenski_coexistence_1986}: Modelling of the 2 sp. system with 1resource and 1 inhibitor.}
+\cmt{\cite{lenski_coexistence_1986}: Modeling of the 2 sp. system with 1resource and 1 inhibitor.}
 The simplest way to introduce inhibition of species A by species B is via a growth rate of species A depending on the abundance of species B. 
@@
 \end{eqnarray}
-The function $\mathcal{B}_{A/B}$ captures the inhibiting effects in an effective way. 
+The function $\mathcal{B}_{A/B}$ captures the inhibiting effects effectively. 
 It is straight forward to make the mechanism of inhibition more explicit by, e.g., introducing a waste pool. 
@@
 \subsection{Interspecies Cooperation}
-On the other hand, one can also observe some cooperative behaviours between species in case when it facilitates the competition with others.
+On the other hand, one can also observe some cooperative behaviors between species in case when it facilitates the competition with others.
 Quite often it happens between the species that are close relatives or clonal populations and such process is called kin selection \cite{west_social_2007}.
-Sometimes neighbours, so-called 'social cheaters', can benefit from a certain behaviour without expressing it by itself \cite{rainey_evolution_2003}.
+Sometimes neighbors, so-called 'social cheaters', can benefit from a certain behavior without expressing it by itself \cite{rainey_evolution_2003}.
 %
@@
 
-- \cite{emonet_agentcell_2005, gorochowski_bsim_2012, li_nufeb_2019, kreft_bacsim_1998}: examples of another toolboxes for bacterial communities modelling (ABM).\\
+- \cite{emonet_agentcell_2005, gorochowski_bsim_2012, li_nufeb_2019, kreft_bacsim_1998}: examples of another toolboxes for bacterial communities modeling (ABM).\\
 More about the modelling of the motile bacteria can be seen in works:~\cite{decoene_microscopic_2011, rosser_modelling_2014, sokolov_physical_2012, li_amplified_2008}.
@@
     Chemostaxis here was driven by the oxygen concentrations.
-    \item \cite{rosser_modelling_2014}: modelling of the bacterial reorintation, rotation.
-    \item \cite{sokolov_physical_2012, li_amplified_2008}: modelling of the bacteria using hydrodynamic interactions and Brownian motion.
+    \item \cite{rosser_modelling_2014}: modeling of the bacterial reorintation, rotation.
+    \item \cite{sokolov_physical_2012, li_amplified_2008}: modeling of the bacteria using hydrodynamic interactions and Brownian motion.
     Fluctuations in velocity and rotations.
     (\cite{li_amplified_2008}: Brownian motion gives right motion observed using microscopy).
-    \item \cite{farrell_mechanically_2013}: 2D modelling using Newtonian dynamics and only mechanical interactions + nutritient diffusion.
+    \item \cite{farrell_mechanically_2013}: 2D modeling using Newtonian dynamics and only mechanical interactions + nutritient diffusion.
     Branching parameter defines diff limits where diffusion dominant and limits growth, and where it's irrelevant.
@@
     \item \cite{farrell_mechanically_2013}: Showed the influence of mechanic interactions between bacteria in 2D system with bacteria consuming the nutritient.
-    The growth of bacterial colonies where nonmotile microorganisms replicate and push each other away as they grow.
+    The growth of bacterial colonies where non-motile microorganisms replicate and push each other away as they grow.
     Found a transition between two different growth regimes, controlled by the balance between growth and uptake of nutrient:\\

### 2023-12-06 pool_model.tex
Add descriprion to plots for 3 pool model and temp shift

@@
 In the described simplified version of the real-life system three possible bacteria states are mentioned corresponding to which process is happening: adaptation, growth or death.
-Motivated by this we divide a bacterial population into three pools (states) (\textbf{Figure \ref{fig:SchematicRep}}): 
+Motivated by this we divide a bacterial population into three pools (states) (Figure~\ref{fig:SchematicRep}): 
 \begin{itemize}
@@
 In this case $[L]=[G]=[N_t]=1$.
-For simplicity we rather make a transition to the dimensionless constant $\alpha:=\alpha_0 N_t$ and diminsioneless resource pool $R(t) = R_0 / N_t$ ($0 \leqslant R \leqslant 1$).
+For simplicity we rather make a transition to the dimensionless constant $\alpha:=\alpha_0 N_t$ and diminsionless resource pool $R(t) = R_0 / N_t$ ($0 \leqslant R \leqslant 1$).
 Here the full capacity of the resource pool $N_t$ was used as a scaling constant.
@@
 %
-Such model can be simplified by explointing that limitation of the resource pool also limits the total bacterial count of the system:
+Such model can be simplified by exploiting that limitation of the resource pool also limits the total bacterial count of the system:
 \begin{equation}
@@
          \lambda         & \alpha  & 0 \\
-         \mu & + \mu'& - \beta 
+         \mu &  \mu'& - \beta 
     \end{pmatrix} 
@@
 \end{figure}
+The resulting growth curves for the three pool case are presented in Figure~\ref{fig:3pool_resource_plots}.
+Here we can see that the value of the transition constant $\lambda$ determines how long the bacteria is in lag-phase.
+The larger is the rate the shorter is the lag-time.
+Moreover, by including the bacteria flow to the death pool we can diversify the shape of the growth curve allowing the decrease of the bacterial count.
 %
@@
 The relevant question here is whether cells in the pool $D$ still contribute to the limitation of the resources by, e.g., occupying space or consuming important nutrients. 
-If yes, the three pool model would be advantageous, if not the two pool model is sufficient with the advantages of having one parameter less.
-
+If yes, the three pool model would be advantageous, if not the two pool model is sufficient with the advantages of having one parameter less.\\
+%
 The dynamics of the three pool model can be captures by the coupled system of ordinary differential equations (the two pool model can be achieved by setting $D\equiv 0$ and replacing $\mu$ with $\beta$):
@@
 \end{eqnarray}
-
+%
+The ordinary differential equations for the pool model with environmental-shock based back-flow read:
+\begin{equation}
+    \begin{cases}
+    \dot{L} =\gamma(t) G - (\mu + \lambda) L\\
+    \dot{G} = \lambda L -\gamma(t) G + \alpha G\left(1-\frac{L+G+D}{N_t}\right)\\
+    \dot{D} = \mu  L - \beta D
+    \label{eq:ODE_tempshift}
+    \end{cases}
+\end{equation}
 \begin{figure}[t]
@@
         The parameters of the back-flow rate are $\Gamma=1$ and $\delta=5$.
-        The other model parameters are $\alpha=2.5$, $\mu'=0.5$, $\beta=5$, $\lambda=10^{-2}$, $\mu = 0$ fe each curve.
+        The other model parameters are $\alpha=2.5$, $\mu'=0.5$, $\beta=5$, $\lambda=10^{-2}$, $\mu = 0$ for each curve.
     }
@@
 \end{figure}
-The ordinary differential equations for the pool model with environmental-shock based back-flow read:
-\begin{equation}
-    \begin{cases}
-    \dot{L} =\gamma(t) G - (\mu + \lambda) L\\
-    \dot{G} = \lambda L -\gamma(t) G + \alpha G\left(1-\frac{L+G+D}{N_t}\right)\\
-    \dot{D} = \mu  L - \beta D
-    \label{eq:ODE_tempshift}
-    \end{cases}
-\end{equation}
+The effect of the temperature shift on the total bacterial count can be shown in Figure~\ref{fig:TempJump}.
+After the rapid temperature increase, the growth slows down till the bacteria that transitioned back to lag-phase go again to growth pool.
+If the shift happened rather early in time, when most of the bacteria were in lag-phase, the effect is not noticeable.
 \newpage
@@
 These survival mechanisms mostly aim to secure as much available nutrition which lead to competitive strategies or, if it is advantageous, cooperative ones~\cite{hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
-Further we show that the TPM can be readily extended to the systems with more than one bacteria and describe more complex relationships between species. 
+Further we show that the TPM can be readily extended to the systems with more than one bacterium and describe more complex relationships between species. 
 %
@@
 \subsection{Resource Competition}
-One on the ways to describe competition between bacteria populations is as a function of limiting resources ratio~\cite{tilman_resource_1977, smith_effects_2002}.
+One on the ways to describe competition between bacteria populations is as a function of limiting resource ratio~\cite{tilman_resource_1977, smith_effects_2002}.
 To describe such a situation we consider for simplicity only one common limiting resource pool. 

### 2023-12-06 pool_model.tex
Resource competition: rearrange, add Figures discussions

@@
     \begin{center}
-    \includegraphics[width=0.7\textwidth]{Figures/TPM_fig.pdf}
+    \includegraphics[width=0.6\textwidth]{Figures/TPM_fig.pdf}
     \caption{Schematic representation of exchange between three states or pools.}
@@
 \subsection{Back flow from Growth to Lag pool}
-Due to its life in rather unfavourable conditions bacteria developed different surviving mechanisms, one of which is adapting their metabolism and transitioning into dormant state.
+Due to its life in rather unfavorable conditions bacteria developed different surviving mechanisms, one of which is adapting their metabolism and transitioning into dormant state.
 By definition dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\cite{kaprelyants_dormancy_1993}. 
@@
 These survival mechanisms mostly aim to secure as much available nutrition which lead to competitive strategies or, if it is advantageous, cooperative ones~\cite{hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
-Further we show that the TPM can be readily extended to the systems with more than one bacterium and describe more complex relationships between species. 
+Further we show that the model can be readily extended to the systems with more than one bacterium and describe more complex relationships between species. 
 %
@@
 \begin{eqnarray}
-\label{Resource}
+\label{eq:Resource}
 R &=&1-\frac{L_A+G_A+L_B+G_B}{N_t}
 \end{eqnarray}
-Inserting Eq. \ref{Resource} into Eqs. \ref{eq:model_2sp_resource_comp} yields:
+Inserting Eq.~\ref{eq:Resource} into Eqs.~\ref{eq:model_2sp_resource_comp} yields:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-which has the structure of a \ac{glv}.
-In order to better see this, we set $L_A=L_B=0$: 
+which has the structure of a \ac{glv} expanded with different internal states for one species.
+As one can see in Figure~\ref{fig:2pool_resource_2sp} even without direct interaction between species the bacterial growth was inhibited due to the limitation of the resource.
+Here species $A$ has a longer lag-phase than $B$ but the exponential growth of $A$ is stronger than the one that bacteria $B$ experiences.
+As a result, species $A$ was able to suppress species $B$ consuming the larger amount of nutritient and reaching larger bacterial count at the end.
+\begin{figure}[H]
+    \begin{center}
+    \includegraphics[width=0.6\textwidth]{Figures/pool_model_2pools_resource_competition.pdf}
+    \caption{
+        The competition of the two bacterial species due to the common limiting resource calculated using Eq.~(\ref{eq:model_2sp_resource_comp}).
+        The red curve describes the growth of the bacteria $A$ ($N_A = \sum_{k} x_A^k = L_A+G_A$) with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
+        The blue curve describes the growth of the bacteria $B$ ($N_B = \sum_{k} x_B^k = L_B+G_B$) with model parameters $\lambda_B=5\cdot 10^{-2}$, $\alpha_B=2.5$.
+        The purple line shows the total bacterial count of the system  $N(t)=\sum_i\sum_k\mbx_i^k=N_A+N_B$.
+        The gray line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
+    }
+    \label{fig:2pool_resource_2sp}
+    \end{center}
+\end{figure}
+%
+In order to get the commonly used form of the \ac{glv}, we need to simply eliminate of the bacterial states by setting $L_A=L_B=0$: 
 \begin{eqnarray}
@@
 \end{eqnarray} 
-
-Equations (\ref{eq:LV_simple1}-\ref{eq:LV_simple2}) are basically competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competive excursion.
+%
+Equations (\ref{eq:LV_simple1}-\ref{eq:LV_simple2}) are basically competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competitive excursion.
 That principle claims that two species in competition for the same resource cannot coexist~\cite{wangersky_lotka-volterra_1978}. 
 In previous studies it was shown that \ac{glv} equations with one bacterial state can be sucessfully applied to predict the microbial community dynamics, determine different types of the pairwise interspecies interactions~\cite{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
-In food studies the \ac{glv} was applied by Mounier \etal~\cite{mounier_microbial_2008} to study interactions of six bacteria species  with yeast omissin in cheese.
-Cauchie \etal~\cite{cauchie_modeling_2020} applied Lotka-Volterra model to study the three-species system of the bacteria majorly present in minced pork meat.
-
-The pool model, in contrast to the classical Lotka-Volterra approach, allows for far more flexibility.
-
-\begin{figure}[H]
-    \begin{center}
-    \includegraphics[width=0.67\textwidth]{Figures/pool_model_2pools_resource_competition.pdf}
-    \caption{
-        The competition of the two bacterial species due to the common limiting resource calculated using Eq. (\ref{eq:model_2sp_resource_comp}).
-        The red curve describes the growth of the bacteria $A$ $N_A = \sum_{k} x_A^k = L_A+G_A$ with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
-        The blue curve describes the growth of the bacteria $B$ $N_B = \sum_{k} x_B^k = L_B+G_B$ with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
-        The purple line shows the total bacterial count of the system  $N(t)=\sum_i\sum_k\mbx_i^k=N_A+N_B$.
-        The gray line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
-    }
-    \label{fig:2pool_resource_2sp}
-    \end{center}
-\end{figure}
-As one can see in Figure~\ref{fig:2pool_resource_2sp} even without direct interaction between species the bacterial growth was inhibited due to the limitation of the resource.
-Here species $A$ was able to supress species $B$ due to the higher values of the growth constant $\alpha$.
+In food studies the \ac{glv} was applied by Mounier \etal~\cite{mounier_microbial_2008} to study interactions of six bacteria species influenced by yeast omissions in cheese.
+Cauchie \etal~\cite{cauchie_modeling_2020} applied Lotka-Volterra model to study the three-species system of the bacteria majorly present in minced pork meat.\\
+%
+However, in contrast to the classical Lotka-Volterra approach, the introduced pool model allows for far more flexibility in tuning properties of the bacterial interactions within and between species. 
 %
@@
 One of the examples of such mechanisms is a production of the antimicrobial compounds, inhibitors or toxins~\cite{wloch-salamon_effect_2008, chao_structured_1981}.
+On the other hand, some bacteria can produce public goods, the compounds that benefit growth of other species.
+In this way, can suppress the non-preferred species by helping their competitors~\cite{hibbing_bacterial_2010}.
 %
@@
 \end{equation}
-
+%
 As it can be seen in Figure~\ref{fig:2pool_2sp_toxin}, model $\mathcal{F}_1$ results in $I_1(t)=\kappa\int_{t_0}^tG(t')dt'$. 
@@
 This model assumes that inhibitor is only produced when the bacteria grow and divide.\\
-
+%
 \begin{figure}[H]
@@
     \begin{center}
-    \includegraphics[width=0.5\textwidth]{Figures/pool_model_1pool_inhib.pdf}
+    \includegraphics[width=0.6\textwidth]{Figures/pool_model_1pool_inhib.pdf}
     \caption{
@@
     \begin{center}
-    \includegraphics[width=0.67\textwidth]{Figures/pool_model_spatial.pdf}
+    \includegraphics[width=0.6\textwidth]{Figures/pool_model_spatial.pdf}
     \caption{

### 2023-12-06 pool_model.tex
Corrections to the toxin production section

@@
 The dynamics of the three pool model can be captures by the coupled system of ordinary differential equations (the two pool model can be achieved by setting $D\equiv 0$ and replacing $\mu$ with $\beta$):
-to be solved with the initial conditions $L(0)=n_0$, $G(0)=0$, and $D(0)=0$, where $n_0$ is the initial bacterial population. 
+to be solved with the initial conditions $L(0)=N_0$, $G(0)=0$, and $D(0)=0$, where $N_0$ is the initial bacterial population. 
 $N_t$ denotes the maximal size of the total bacterial population due to environmental limitations (\textbf{Figure \ref{fig:3pool_resource_plots}}). 
@@
     \begin{cases}
-        L(t) = n_0 e^{-(\mu+\lambda)t}\\
-        D(t) = n_0 \mu e^{-\beta t}\frac{1-e^{-(\mu+\lambda-\beta)t}}{\mu+\lambda-\beta}
+        L(t) = N_0 e^{-(\mu+\lambda)t}\\
+        D(t) = N_0 \mu e^{-\beta t}\frac{1-e^{-(\mu+\lambda-\beta)t}}{\mu+\lambda-\beta}
     \end{cases}.
@@
 However, for time dependent parameters, e.g., for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$.
-It is interesting to see what type of equations result for the total bacterial population $n=L+G+D$:
+It is interesting to see what type of equations result for the total bacterial population $N=L+G+D$:
 \begin{equation}
     \begin{cases}
-        \dot{G} = \lambda L + \alpha G\left(1-\frac{n}{N_t}\right)\\
-        \dot{n} = - \beta D + \alpha G\left(1-\frac{n}{N_t}\right)
+        \dot{G} = \lambda L + \alpha G\left(1-\frac{N}{N_t}\right)\\
+        \dot{N} = - \beta D + \alpha G\left(1-\frac{N}{N_t}\right)
     \end{cases}
 \end{equation}
-with the initial conditions $G(0)=0$ and $n(0)=n_0$. 
+with the initial conditions $G(0)=0$ and $N(0)=N_0$. 
 These equations have a clear interpretation, are derived from simple principles, and are different from the popular Barayni-Roberts and modified-Gompertz models. 
@@
 To find a substantial initial drop one needs $\beta \mu \gg \alpha \lambda$. 
-The lag time $t_L$ can be approximated by the time at which half of the initial population is in the growth phase, i.e., $n_0=2G(t_L)$. 
+The lag time $t_L$ can be approximated by the time at which half of the initial population is in the growth phase, i.e., $N_0=2G(t_L)$. 
 Assuming $L\ll N_t$ we can ignore the non-linear term in the equation for $G$, which gives rise to:
@@
 \subsubsection{Toxins Production}
-Now extend our model where bacteria not only consume nutritions from resource pool $R$ but one of the species also actively produce the inhibiting compounds $T$. 
-It is the accumulation of the bacteria waste which is responsible for the spoilage of food products. 
-In order to take this into account we add an equation for the waste production:
+To introduce active competitive mechanisms, we extend our model with one of the species producing the toxic compounds $T$ during its growth:
 \begin{eqnarray}
-    \dot{L} &=& - \lambda_f R L + \mathcal{B}(R)G\\
-    \dot{G} &=& \lambda_f R L +\alpha R G-\mathcal{B}(R)G\\
-    \dot{R} &=&-\frac{\alpha}{N_t} R G\\
     \dot{T} &=& \mathcal{F}(R,G).
 \end{eqnarray}
-The function $\mathcal{F}$ denotes the model for the antibiotic production rate. 
-We shall discuss two possible models:
-\begin{eqnarray}
-    \mathcal{F}_1(R,G)&=&\kappa G\\
-    \mathcal{F}_2(R,G)&=&\kappa\alpha R G=-\kappa N_t\dot{R}
-\end{eqnarray}
-We would like to test the following model:
+The function $\mathcal{F}$ denotes the model for the toxin production rate.
+It can be, for example, the accumulation of the bacteria waste which is responsible for the spoilage of food products and affecting another species in negative way.
+Consider the system for two species competing for resource $R$ with the production of the toxin $T$ by species $B$.
+Assume that the toxin causes the death of the species $A$ with rate $\mu$.
+Then the full \ac{ode} system is
 \begin{equation}
@@
         \dot{T} = \mathcal{F}_{1,2} (R, G_B) \\
-    \end{cases}
+    \end{cases}.
     \label{eq:model_2sp_toxin}
@@
 %
-As it can be seen in Figure~\ref{fig:2pool_2sp_toxin}, model $\mathcal{F}_1$ results in $I_1(t)=\kappa\int_{t_0}^tG(t')dt'$. 
-As long as $G>0$ toxin is produced, no matter whether the amount of bacteria is still growing. 
-Model $\mathcal{F}_2$ results in waste accumulation in steady state $I_2^s=\lim_{t\to\infty}I_1(t)=\kappa(N_t-n_0)$. 
-To see this we note $\dot{T}_2=\kappa(\dot{L}+\dot{G})$, which yields: $I_2(t)=\kappa(L(t)+G(t)-n_0)$ where we used $T_2(0)=G(0)=0$, $L(0)=n_0$. 
-This model assumes that inhibitor is only produced when the bacteria grow and divide.\\
+Where we consider two possible models of the toxin production:
+\begin{eqnarray}
+    \mathcal{F}_1(R,G)&=&\kappa G_B\\
+    \mathcal{F}_2(R,G)&=&\kappa\alpha_B R G_B.
+\end{eqnarray}
+Here $\kappa$ is the toxin production rate.
+The difference between the two toxin production models for resulting solution of the \acp{ode} is presented in Figure~\ref{fig:2pool_2sp_toxin}.
+%
 %
@@
 \end{figure}
+%
+Model $\mathcal{F}_1$ results in toxin concentration $T_1(t)=\kappa\int_{t_0}^tG(t')dt'$.
+That means that as long as there are bacteria in the growth phase $G_B>0$ toxin is produced, no matter whether the total bacterial count $N_B$ is still growing. 
+On the other hand, model $\mathcal{F}_2$ results in toxin accumulation in steady state $T_2^s=\lim_{t\to\infty}T_2(t)=\kappa(N_t-N_0)$. 
+To see this we note $\dot{T}_2=\kappa(\dot{L_B}+\dot{G_B})$, which yields: $T_2(t)=\kappa(L_B(t)+G_B(t)-N_{B0})$ where we used $T_2(0)=G_B(0)=0$, $L_B(0)=N_{B0}$. 
+This model assumes that toxin is only produced when the total bacterial grow and divide, hence, the total bacterial count $N_B$ increases.\\
 %
@@
 In the first case the inhibition is of minor importance and could be ignored. 
-In the latter case one can simplify the equations, given $n_A$ and $n_B$ are non-zero:
+In the latter case one can simplify the equations, given $N_A$ and $N_B$ are non-zero:
 \begin{eqnarray}

### 2023-12-06 pool_model.tex
Corrections: mutual inhibition od species

@@
 Consider for simplicity the system where only one bacteria species is present $N=1$.
-The bacterial growth curves usually consists of different consecutive phases~\cite{buchanan_when_1997}.
+The bacterial growth curves usually consist of different consecutive phases~\cite{buchanan_when_1997}.
 At the beginning no increase in bacterial count is present and so-called lag-phase is observed.
@@
 When the nutrition is exhausted by the growing bacteria, the stationary state of the growth curve occurs.
-In this state the cultures accumulate waste products of metabolitic activity that leads to cell death of the bacteria.
+In this state the cultures accumulate waste products of metabolic activity that leads to cell death of the bacteria.
 At the same time dying cells release the nutritions for survivors to continue growing.
@@
 Based on experimental observations, it was shown that the bacterial growth is strongly influenced by the limiting resource.
-In the system where all needed nutritients are in excess except for one the total growth is linear to the concentration of this liminting nutritient~\cite{hibbing_bacterial_2010, monod_growth_1949}.
+According to the works of Jacques Monod, in the system where all needed nutritients are in excess except for one the total growth is linear to the concentration of this liminting nutritient~\cite{monod_growth_1949}.
 The control of the population density with respect to the available nutritions is implemented using cell communication process called quorum sensing. 
@@
         The toxin $T$ causes the cell death of bacteria $A$ with rate constant $\mu = 0.2$.
-        The red curve describes the growth of the bacteria $A$ $N_A = \sum_{k} x_A^k = L_A+G_A$ with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
-        The blue curve describes the growth of the bacteria $B$ $N_B = \sum_{k} x_B^k = L_B+G_B$ with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
+        The red curve describes the growth of the bacteria $A$ ($N_A = \sum_{k} x_A^k = L_A+G_A$) with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
+        The blue curve describes the growth of the bacteria $B$ ($N_B = \sum_{k} x_B^k = L_B+G_B$) with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
         The gray line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
@@
 % ................................................................................................
-\subsubsection{Species Inhibition}
-\cmt{\cite{lenski_coexistence_1986}: Modeling of the 2 sp. system with 1resource and 1 inhibitor.}
-The simplest way to introduce inhibition of species A by species B is via a growth rate of species A depending on the abundance of species B. 
+\subsubsection{Mutual Inhibition}
+Another special case of the species competition is when the produced toxin does not cause the death of the bacteria but rather inhibit its growth.
+Such situation with 2 bacteria species, limiting resource and produced inhibitor was modeled, e.g., by Lenski and Hattingh~\cite{lenski_coexistence_1986}.
+In our case, we would like to look at rather simplified case reducing the number of variables of the system.
+The easiest way to introduce inhibition of species $A$ by species $B$ is via a growth rate of species $A$ depending on the abundance of species $B$.
 For sake of simplicity we ignore the lag pool and write for a mutual inhibition:
@@
 The function $\mathcal{B}_{A/B}$ captures the inhibiting effects effectively.
-It is straight forward to make the mechanism of inhibition more explicit by, e.g., introducing a waste pool. 
+It is straight forward to make the mechanism of inhibition more explicit by, e.g., introducing a toxin pool. 
 One possible choice for $\mathcal{B}$ is: 
@@
 \end{eqnarray}
-This corresponds to the case $W\sim G$, where $W$ is the waste (or whatever one chooses to call it) of species A affecting species B. 
-If the waste accumulates one needs to replace $G$ by $\int G dt$. Keeping it simple yields:
+This corresponds to the case $T\sim G$, where $T$ is the toxin produced. 
+If the toxin accumulates one needs to replace $G$ by $\int G dt$. Keeping it simple yields:
 \begin{eqnarray}
@@
 \end{eqnarray}
-We can separate three cases: 
+We can separate three cases (see Figure~\ref{fig:1pool_2sp_inhibit}): 
 \begin{eqnarray}
@@
     \begin{center}
-    \includegraphics[width=0.6\textwidth]{Figures/pool_model_1pool_inhib.pdf}
+    \includegraphics[width=0.5\textwidth]{Figures/pool_model_1pool_inhib.pdf}
     \caption{
-        The two species system where species inhibit each other defined with equations~(\ref{eq:2sp_1pool_inhibit_unitless1}-\ref{eq:2sp_1pool_inhibit_unitless2}).
+        The two-species system mutually inhibiting each other is defined with equations~(\ref{eq:2sp_1pool_inhibit_unitless1}-\ref{eq:2sp_1pool_inhibit_unitless2}).
         Here the red lines describes the evolution of the species $A$ and blue lines correspond to species $B$.

### 2023-12-06 pool_model.tex
Add citainions to existing models

@@
 \section{Introduction}
-
-Comprehensive reviews about competition/cooperation bacterial mechanisms: \cite{ghoul_ecology_2016, stubbendieck_bacterial_2016, hibbing_bacterial_2010, west_social_2007}
+Motivation: general, flexible approach for bacterial systems that allows easily modify the system depending on processes on different scales.
+Allows us to describe the processes within populations using different bacterial states as well as interactions between different populations using ODEs.\\
 %
 %
+Existing models in Predictive Microbiology (Primary Models)~\cite{perez-rodriguez_predictive_2012}:
+\begin{itemize}
+    \item An overview of primary models used in food science can be found in~\cite{van_boekel_kinetic_2008}.
+    \item Baranyi-Roberts Model~\cite{baranyi_modeling_1993, baranyi_dynamic_1994}
+    \item modified Gompertz
+    \item Buchanan et al.: three-phase linear model for lag phase, exponential growth phase, and stationary phase~\cite{buchanan_when_1997}
+    \item McKellar two-compartment model with bacteria in growing and non-growing state~\cite{mckellar_heterogeneous_1997}
+    \item Inactivation models
+\end{itemize}
+%
+Secondary models:
+\begin{itemize}
+    \item Polynomial models 
+    \item Square Root-Type Models, e.g., Ratkowski for temperature~\cite{ratkowsky_relationship_1982}
+    \item The Cardinal Parameter Model~\cite{zwietering_decision_1992}
+\end{itemize}
+%
+%
+Comprehensive reviews about competition/cooperation bacterial mechanisms: \cite{ghoul_ecology_2016, stubbendieck_bacterial_2016, hibbing_bacterial_2010, west_social_2007}\\
 %
@@
 % ================================================================================================
-\section{Motivating Example}
-
-...
-
-\cmtb{Maybe add here motivations about bacteria states: So toy system where we can eliminate states (sum and get F(n1, n2)).
-And describe it initially without states/1state. Get the difference.}
-
-\cmtb{Maybe Example from Seminar by Jens? 2 species,  1 resource, 1 waste production.}
-%
-%
-%
-%
-% ================================================================================================
-\section{General Mathemetical Formulation}
+\section{General Mathematical Formulation}
 Consider the system consisting of $N$ species in $M$ different states.
@@
 \end{equation}
-where  $\mbx \in \mathbb{R}^{N}  \otimes \mathbb{R}^{M}$ is a vector descibing bacterial count.
+where  $\mbx \in \mathbb{R}^{N}  \otimes \mathbb{R}^{M}$ is a vector describing bacterial count.
 In matrix form it can be written as
@@
 $\mbu$ is the vector of different inputs of the model, e.g., temperature, pressure etc.
-The vector $\mbm \in \mathbb{R}^{K}$ includes all variables corresponding to the microenvironment of the system which can be concentartion of the inhibitors or activators:
+The vector $\mbm \in \mathbb{R}^{K}$ includes all variables corresponding to the microenvironment of the system which can be concentration of the inhibitors or activators:
 \begin{equation}
@@
 \end{equation}
-Here function $\mbH$ is in general case also non-linear.
-
+Here function $\mbH$ is in general case also non-linear.\\
+%
 The advantage of the following formulation is its generality.
 Depending on the knowledge about the system as well as the needed level of accuracy one can define an arbitrary number of states $M$ and control the complexity of the model.
-Moreover, opposite to the already known Predictive Microbiology like Baranyi and Roberts one can easily change.
-
+Moreover, opposite to the already known Predictive Microbiology like Baranyi and Roberts one can easily change.\\
+%
 To calculate the bacterial count for one species $i$, we simply need to sum the tensor/vector $\mbx$ over all possible states:
@@
 \end{equation}
-In the following section we demonstate the application of this approach for different toy models describing the processes within and between bacteria communities.
-\newpage
+In the following section we demonstrate the application of this approach for different toy models describing the processes within and between bacteria communities.
 %
@@
 This phenomenon appears due to adaptation of the microbial community, for example, to the new environmental conditions.
-This so-called preparation to the exponential growth includes synthesys of the needed compounds, initiation of the transciption~\cite{rolfe_lag_2012} and poorly studied physiological and regulatory mechanisms~\cite{monod_growth_1949}
-After this, the succeeded in adaptation bacteria enter the exponential growth phase, where the process of cell division is directly happening \cmt{Add Ref}.
+This so-called preparation to the exponential growth includes synthesys of the needed compounds, initiation of the transciption~\cite{rolfe_lag_2012} and poorly studied physiological and regulatory mechanisms~\cite{monod_growth_1949}.
+After this, the succeeded in adaptation bacteria enter the exponential growth phase, where the process of cell division is directly happening.
 When the nutrition is exhausted by the growing bacteria, the stationary state of the growth curve occurs.
@@
 In the described simplified version of the real-life system three possible bacteria states are mentioned corresponding to which process is happening: adaptation, growth or death.
-Motivated by this we divide a bacterial population into three pools (states) (Figure~\ref{fig:SchematicRep}): 
+Motivated by this we divide a bacterial population into three pools (states) (~\ref{fig:SchematicRep}):
 \begin{itemize}
@@
 For now, we assume for simplicity that there is no back flow between growth state $G$ and lag-state $L$.
-The suggested model can be cast in the following reactions: \cmt{Use notations x ??}
+The suggested model can be cast in the following reactions:
 \begin{eqnarray}
@@
 Based on experimental observations, it was shown that the bacterial growth is strongly influenced by the limiting resource.
-According to the works of Jacques Monod, in the system where all needed nutritients are in excess except for one the total growth is linear to the concentration of this liminting nutritient~\cite{monod_growth_1949}.
-The control of the population density with respect to the available nutritions is implemented using cell communication process called quorum sensing. 
+According to the works of Jacques Monod, in the system where all needed nutritients are in excess except for one the total growth is linear to the concentration of this limiting nutritient~\cite{monod_growth_1949}.
+The control of the population density with respect to the available nutrition is implemented using cell communication process called quorum sensing.
 Using release of the extracellular signalling molecules (autoinducers) it is possible to synchronize the gene expression of the whole group~\cite{ng_bacterial_2009}.
@@
 In this case $[L]=[G]=[N_t]=1$.
-For simplicity we rather make a transition to the dimensionless constant $\alpha:=\alpha_0 N_t$ and diminsionless resource pool $R(t) = R_0 / N_t$ ($0 \leqslant R \leqslant 1$).
+For simplicity, we rather make a transition to the dimensionless constant $\alpha:=\alpha_0 N_t$ and dimensionless resource pool $R(t) = R_0 / N_t$ ($0 \leqslant R \leqslant 1$).
 Here the full capacity of the resource pool $N_t$ was used as a scaling constant.
@@
 to be solved with the initial conditions $L(0)=N_0$, $G(0)=0$, and $D(0)=0$, where $N_0$ is the initial bacterial population.
-$N_t$ denotes the maximal size of the total bacterial population due to environmental limitations (\textbf{Figure \ref{fig:3pool_resource_plots}}). 
+$N_t$ denotes the maximal size of the total bacterial population due to environmental limitations (Figure \ref{fig:3pool_resource_plots}).
 For $\mu'=0$ the equations for $L$ and $D$ can be readily integrated resulting in:
@@
 Moreover, the rapid shift in temperature is associated with the appearance of an additional lag phase~\cite{zwietering_modeling_1994}.
-To model it we propose a kind of visco-elastic Maxwell type of stress-strain relation and write down the ordinary differential equation for $\gamma$ to be:
+To model it we propose a kind of viscoelastic Maxwell type of stress-strain relation and write down the ordinary differential equation for $\gamma$ to be:
 \begin{eqnarray}
@@
 In order to get the commonly used form of the \ac{glv}, we need to simply eliminate of the bacterial states by setting $L_A=L_B=0$:
-\begin{eqnarray}
-    \dot{G}_A &=& \alpha_A G_A\left(1 - \frac{G_A}{N_t}\right) - \frac{\alpha_A}{N_t}G_AG_B
-    \label{eq:LV_simple1}\\
-    \dot{G}_B &=& \alpha_B G_B\left(1-\frac{G_B}{N_t}\right) -\frac{\alpha_B}{N_t}G_AG_B. 
-    \label{eq:LV_simple2}
-\end{eqnarray} 
+\begin{equation}
+    \begin{cases}
+    \dot{G}_A = \alpha_A G_A\left(1 - \frac{G_A}{N_t}\right) - \frac{\alpha_A}{N_t}G_AG_B\\
+    \dot{G}_B = \alpha_B G_B\left(1-\frac{G_B}{N_t}\right) -\frac{\alpha_B}{N_t}G_AG_B. 
+    \label{eq:LV_simple}
+    \end{cases}
+\end{equation}
+
 %
-Equations (\ref{eq:LV_simple1}-\ref{eq:LV_simple2}) are basically competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competitive excursion.
+Equations (\ref{eq:LV_simple}) are basically competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competitive excursion.
 That principle claims that two species in competition for the same resource cannot coexist~\cite{wangersky_lotka-volterra_1978}.
@@
 However, in contrast to the classical Lotka-Volterra approach, the introduced pool model allows for far more flexibility in tuning properties of the bacterial interactions within and between species.
-%
-\cmt{But LV for several bacteria states in literature? Didn't find!}
 %
@@
 To see this we note $\dot{T}_2=\kappa(\dot{L_B}+\dot{G_B})$, which yields: $T_2(t)=\kappa(L_B(t)+G_B(t)-N_{B0})$ where we used $T_2(0)=G_B(0)=0$, $L_B(0)=N_{B0}$.
-This model assumes that toxin is only produced when the total bacterial grow and divide, hence, the total bacterial count $N_B$ increases.\\
+This model assumes that toxin is only produced when the total bacterial grow and divide, hence, the total bacterial count $N_B$ increases.
 %
@@
 Rescaling time with $\tilde{\alpha}_A$ and defining the dimensionless parameter
-\begin{eqnarray}
-    \psi&=&\tilde{\alpha}_B/\tilde{\alpha}_A\\
-    \psi&=&\frac{K_B\alpha_B}{K_A\alpha_A}
-\end{eqnarray}
+\begin{equation}
+    \psi = \tilde{\alpha}_B/\tilde{\alpha}_A = \frac{K_B\alpha_B}{K_A\alpha_A}
+\end{equation}
 results in: 
-\begin{eqnarray}
-    \dot{G}_A &=&\frac{G_A}{G_B}\left(1 - G_A-G_B\right)
-    \label{eq:2sp_1pool_inhibit_unitless1}\\
-    \dot{G}_B &=&\psi\frac{G_B}{G_A}\left(1-G_A-G_B\right). 
-    \label{eq:2sp_1pool_inhibit_unitless2}
-\end{eqnarray}
+\begin{equation}
+    \begin{cases}
+    \dot{G}_A =\frac{G_A}{G_B}\left(1 - G_A-G_B\right)\\
+    \dot{G}_B =\psi\frac{G_B}{G_A}\left(1-G_A-G_B\right). 
+    \end{cases}
+    \label{eq:2sp_1pool_inhibit_unitless}
+\end{equation}
 We can separate three cases (see Figure~\ref{fig:1pool_2sp_inhibit}):
@@
     \caption{
-        The two-species system mutually inhibiting each other is defined with equations~(\ref{eq:2sp_1pool_inhibit_unitless1}-\ref{eq:2sp_1pool_inhibit_unitless2}).
+        The two-species system mutually inhibiting each other is defined with equations~(\ref{eq:2sp_1pool_inhibit_unitless}).
         Here the red lines describes the evolution of the species $A$ and blue lines correspond to species $B$.
@@
 coli} or \textit{B.
-subtilis} can experience the chemostatic motion of driven by oxygen concentrations~\cite{decoene_microscopic_2011}.
+subtilis} can experience the chemotactic motion of driven by oxygen concentrations~\cite{decoene_microscopic_2011}.
 The bacterial social interactions such as releasing the toxins or public goods can also significantly affect the spatial structure of the community~\cite{blanchard_bacterial_2015}.
@@
 \begin{itemize}
-    \item \cite{decoene_microscopic_2011}: 2D chemostatic self-propelled bacterial motion of the flagellated bactreia (e.g. E. coli, B. subtilis).
-    Chemostaxis here was driven by the oxygen concentrations.
+    \item \cite{decoene_microscopic_2011}: 2D chemotactic self-propelled bacterial motion of the flagellated bacteria (e.g. E. coli, B. subtilis).
+    Chemotaxis here was driven by the oxygen concentrations.
     \item \cite{rosser_modelling_2014}: modeling of the bacterial reorintation, rotation.
@@
 %
-To test the spatial limitations of the described above approach consider the system consisting from two bacterial species thet can be described bt two internal states: lag and growth states. 
+To test the spatial limitations of the described above approach consider the system consisting from two bacterial species that can be described by two internal states: lag and growth states. 
 Both of these species grow using limiting resource pool $R$ with different growth constants $\alpha$ and different lengths of lag phase determined by parameter $\lambda$.

### 2023-12-07 pool_model.tex
Add model cooperation, corrections to previous Figures

@@
         \dot{R} = -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B\\
-        \dot{T} = \mathcal{F}_{1,2} (R, G_B) \\
+        \dot{T}_{B}^{1,2} = \mathcal{F}^{1,2} (R, G_B) \\
     \end{cases}.
@@
 \begin{eqnarray}
-    \mathcal{F}_1(R,G)&=&\kappa G_B\\
-    \mathcal{F}_2(R,G)&=&\kappa\alpha_B R G_B.
+    \mathcal{F}^1(R,G)&=&\kappa G_B\\
+    \mathcal{F}^2(R,G)&=&\kappa\alpha_B R G_B.
 \end{eqnarray}
@@
     \caption{
-        The two species system competing for the common nutritient $R$ with production of the toxin $T$ (brown line) calculated using Eq. (\ref{eq:model_2sp_toxin}).
-        The growing bacteria $B$ produces toxin $T$ (a) constantly $\dot{T} = \mathcal{F}_1 = \kappa G_B$ and (b) only during growth process $\dot{T} = \mathcal{F}_2 = \kappa \alpha_B R G_B$ with 
+        The two species system competing for the common nutritient $R$ with production of the toxin $T_B$ (brown line) calculated using Eq.~(\ref{eq:model_2sp_toxin}).
+        The growing bacteria $B$ produces toxin $T_B$ (a) constantly $\dot{T}_B^1 = \mathcal{F}^1 = \kappa G_B$ or (b) only during growth process $\dot{T}_B^2 = \mathcal{F}^2 = \kappa \alpha_B R G_B$ with
         production rate $\kappa=0.2$.
-        The toxin $T$ causes the cell death of bacteria $A$ with rate constant $\mu = 0.2$.
+        The toxin $T_B$ causes the cell death of bacteria $A$ with rate constant $\mu = 0.2$.
         The red curve describes the growth of the bacteria $A$ ($N_A = \sum_{k} x_A^k = L_A+G_A$) with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
@@
 %
-Model $\mathcal{F}_1$ results in toxin concentration $T_1(t)=\kappa\int_{t_0}^tG(t')dt'$.
+Model $\mathcal{F}^1$ results in toxin concentration $T_B^1(t)=\kappa\int_{t_0}^tG(t')dt'$.
 That means that as long as there are bacteria in the growth phase $G_B>0$ toxin is produced, no matter whether the total bacterial count $N_B$ is still growing.
-On the other hand, model $\mathcal{F}_2$ results in toxin accumulation in steady state $T_2^s=\lim_{t\to\infty}T_2(t)=\kappa(N_t-N_0)$.
-To see this we note $\dot{T}_2=\kappa(\dot{L_B}+\dot{G_B})$, which yields: $T_2(t)=\kappa(L_B(t)+G_B(t)-N_{B0})$ where we used $T_2(0)=G_B(0)=0$, $L_B(0)=N_{B0}$.
+On the other hand, model $\mathcal{F}^2$ results in toxin accumulation in steady state $T_B^{2s}=\lim_{t\to\infty}T_B^2(t)=\kappa(N_t-N_0)$.
+To see this we note $\dot{T}_B^2=\kappa(\dot{L_B}+\dot{G_B})$, which yields: $T_B^2(t)=\kappa(L_B(t)+G_B(t)-N_{B0})$ where we used $T_B^2(0)=G_B(0)=0$, $L_B(0)=N_{B0}$.
 This model assumes that toxin is only produced when the total bacterial grow and divide, hence, the total bacterial count $N_B$ increases.
@@
 \begin{eqnarray}
+    \psi&=&1:\quad\text{A and B are equivalent}\\
     \psi&>&1:\quad\text{Species B growth faster and/or inhibits A stronger}\\
-    \psi&=&1:\quad\text{A and B are equivalent}\\
     \psi&<&1:\quad\text{Species A growth faster and/or inhibits B stronger}
@@
         Here the red lines describes the evolution of the species $A$ and blue lines correspond to species $B$.
-        The relation between inhibition strength of species $B$ and $A$ are defined by control parameter value (a) $\psi$ = 0.5, (b) $\psi$ = 1 or (c) $\psi$ = 2.
+        The relation between inhibition strength of species $B$ and $A$ are defined by control parameter value (a) $\psi$ = 1, (b) $\psi$ = 0.5 or (c) $\psi$ = 2.
     }
@@
 On the other hand, one can also observe some cooperative behaviors between species in case when it facilitates the competition with others.
+It happens if bacteria species as a result of their metabolism produce the compounds that are beneficial for the growth of each other, so-called public goods.
 Quite often it happens between the species that are close relatives or clonal populations and such process is called kin selection \cite{west_social_2007}.
-Sometimes neighbors, so-called 'social cheaters', can benefit from a certain behavior without expressing it by itself \cite{rainey_evolution_2003}.
+Furthermore, Sometimes neighbors can benefit from a certain behavior without expressing it by itself, so-called 'social cheaters'~\cite{rainey_evolution_2003}.
 %
-\cmt{\cite{rocha_base_2002}: Symbiosis host/bacteria bacterial genomes suffering reductive evolution due to this.
-Maybe it's connected with 'social cheaters'?}
-%
-\cmt{Production of public goods!}
 %
@@
 \begin{eqnarray}
-    \dot{G}_A &=& \mathcal{B}_A(W_B,R)G_A\\
-    \dot{G}_B &=& \mathcal{B}_B(W_A,R)G_B\\
-    \dot{R} &=&-\mathcal{B}_A(W_B,R)G_A-\mathcal{B}_B(W_A,R)G_B\\
-    \dot{W}_A &=& \mathcal{F}_A(R,G_A)\\
-    \dot{W}_B &=& \mathcal{F}_B(R,G_B).
+    \dot{G}_A &=& \mathcal{B}_A(P_B,R)G_A\\
+    \dot{G}_B &=& \mathcal{B}_B(P_A,R)G_B\\
+    \dot{R} &=&-\mathcal{B}_A(P_B,R)G_A-\mathcal{B}_B(P_A,R)G_B\\
+    \dot{P}_A &=& \mathcal{F}_A(R,G_A)\\
+    \dot{P}_B &=& \mathcal{F}_B(R,G_B).
 \end{eqnarray}
-$W_A$ and $W_B$ denote the substance produced by species A and B, respectively.
-The functions $\mathcal{B}_A(W_B,R)$ and $\mathcal{B}_B(W_A,R)$ capture the resource dependent growth.
-Setting $\mathcal{B}_A(W_B,R)=\alpha_A W_BR$, $\mathcal{B}_B(W_A,R)=\alpha_B W_AR$, $\mathcal{F}_A(R,G_A)=\kappa_A RG_A$, $\mathcal{F}_B(R,G_B)=\kappa_B RG_B$, rescaling $G_A$, $G_B$, and $W_B$ with $N_t$, $W_A$ with $\kappa_A/\alpha_A$, time with $\alpha_AN_t^2$, defining $\phi=\kappa_A\alpha_B/(\alpha_A^2N_t)$, and $\psi=\kappa_B/(\alpha_A N_t)$ results in:
-\begin{eqnarray}
-    \dot{G}_A &=& W_BG_A\left(1 - G_A-G_B\right)\\
-    \dot{G}_B &=& \phi W_AG_B\left(1 - G_A-G_B\right)\\
-    \dot{W}_A &=& G_A\left(1 - G_A-G_B\right)\\
-    \dot{W}_B &=&\psi G_B\left(1 - G_A-G_B\right).
-\end{eqnarray}
+$P_A$ and $P_B$ denote the substance produced by species A and B, respectively.
+The functions $\mathcal{B}_A(P_B,R)$ and $\mathcal{B}_B(P_A,R)$ capture the resource dependent growth.
+Setting $\mathcal{B}_A(P_B,R)=\alpha_A P_BR$, $\mathcal{B}_B(P_A,R)=\alpha_B P_A R$, $\mathcal{F}_A(R,G_A)=\kappa_A RG_A$, $\mathcal{F}_B(R,G_B)=\kappa_B RG_B$, rescaling $G_A$, $G_B$, and $P_B$ with $N_t$, $P_A$ with $\kappa_A/\alpha_A$, time with $\alpha_AN_t^2$, defining $\psi=\kappa_A\alpha_B/(\alpha_A^2N_t)$, and $\phi=\kappa_B/(\alpha_A N_t)$ results in:
+\begin{equation}
+    \begin{cases}
+    \dot{G}_A = P_B G_A\left(1 - G_A-G_B\right)\\
+    \dot{G}_B = \psi P_A G_B\left(1 - G_A-G_B\right)\\
+    \dot{P}_A = G_A\left(1 - G_A-G_B\right)\\
+    \dot{P}_B =\phi G_B\left(1 - G_A-G_B\right)
+    \end{cases}.
+    \label{eq:1pool_2sp_cooper}
+\end{equation}
+%
+The solution of the system~\ref{eq:1pool_2sp_cooper} is presented in Figure~\ref{fig:1pool_2sp_cooper}.
+Here we can see that if $\psi=1$ and $\phi=1$ (see Figure~\ref{fig:1pool_2sp_cooper}a) the production of the public goods of the two species is happening with the same rate as well as the growth of bacteria.
+Increasing the production rate of the public good produced by species $B$ (see Figure~\ref{fig:1pool_2sp_cooper}b) causes increasing the growth of species $A$ that profit from the presence of $B$.
+On the other hand, choosing $\psi=2$ (see Figure~\ref{fig:1pool_2sp_cooper}c), we increase the profit that species $B$ has from public goods $P_A$.
+That indirectly leads to larger amount of $P_B$ produced by species $B$.
+\begin{figure}[H]
+    \begin{center}
+    \includegraphics[width=0.5\textwidth]{Figures/pool_model_2sp_1pool_coop.pdf}
+    \caption{
+        The two-species system where each of the species release public goods $P$ is defined with equations~(\ref{eq:1pool_2sp_cooper}) and calculated for different parameters:
+        (a) $\psi=1, \phi=1$, (b) $\psi=1, \phi=2$ and (b) $\psi=2, \phi=1$.
+        Here the red lines describes the evolution of the species $G_A$ and blue lines correspond to species $G_B$.
+        Light brown curve corresponds to public good released by species $A$ ($P_A$) and brown curve shows public goods produced by species $B$ ($P_B$).
+    }
+    \label{fig:1pool_2sp_cooper}
+    \end{center}
+\end{figure}
 %

### 2023-12-08 pool_model.tex
add spatial figures

@@
 More about the modelling of the motile bacteria can be seen in works:~\cite{decoene_microscopic_2011, rosser_modelling_2014, sokolov_physical_2012, li_amplified_2008}.
-
+\begin{figure}
+    \includegraphics[width=0.5\textwidth]{Figures/snapshot_00000000.png}%
+    \includegraphics[width=0.5\textwidth]{Figures/snapshot_00006500.png}\\
+    \includegraphics[width=0.5\textwidth]{Figures/snapshot_00013000.png}%
+    \includegraphics[width=0.5\textwidth]{Figures/snapshot_00020000.png}%
+    \caption{}
+    \label{fig:spatial-snapshots}
+\end{figure}
+%
+%
+\begin{figure}
+    \centering
+    \includegraphics[width=0.6\textwidth]{Figures/cell_growth.pdf}
+    \caption{}
+    \label{fig:spatial-growth-curve}
+\end{figure}
+%
+%
+\begin{figure}
+    \centering
+    \includegraphics[width=0.6\textwidth]{Figures/lag_phase.pdf}
+    \caption{}
+    \label{fig:spatial-lag-phase}
+\end{figure}
+%
+%
+\begin{figure}
+    \centering
+    \includegraphics[width=0.6\textwidth]{Figures/entropy.pdf}
+    \caption{}
+    \label{fig:spatial-entropy}
+\end{figure}
+%
+%
 In general, bacterial communities in 2D can experience collective motion called swarming~\cite{wu_collective_2015}.

### 2023-12-20 pool_model.tex
adapt spatial figures
- add titles
- group two of them together
Corrections to plot names
make title smaller and add authors with affiliations
formatting
rename section; rename species from 1/2 to A/B; write some introductory material

@@
 
-\title{Pool Model Paper}
-\author{Polina Gaindrik}
+\title{Pool Model}
+\author{
+    Polina Gaindrik\textsuperscript{1},
+    Jonas Pleyer\textsuperscript{1}
+    and Christian Fleck\textsuperscript{1}
+}
+
+\date{
+    \footnotesize
+    \textsuperscript{\textbf{1}}Freiburg Center for Data Analysis and Modeling
+}
 
@@
 % ================================================================================================
-\section{Spatial Limitations}
-As described above, the proposed above model operates with the total number of the bacterial counts for different species and states in the system not accounting for their spatial distributions.
-Assuming that the bacteria is uniformly distributed we can simplify the problem to the system of the \acp{ode}.
-However, in reality spatial effects play significant role for population survival forcing some species develop such competitive mechanism as cell adhesion to surfaces in the areas rich with nutritients~\cite{htuson_bacteriasurface_2013}.
+\section{Spatial Effects}
+As described above, the proposed model operates with the total number of the bacterial counts for different species and states in the system not accounting for their spatial distributions.
+Assuming that the bacterial system is uniformly distributed (sometimes also called 'well-stirred') we can simplify the problem to a system of \acp{ode}.
+However, in reality spatial effects can play a significant role for population survival forcing some species develop competitive mechanisms such as cell adhesion to surfaces in the areas rich with nutritients~\cite{htuson_bacteriasurface_2013}.
 Another mechanism is the motility of bacteria, e.g., the flagellates bacteria as \textit{E.
@@
 The bacterial social interactions such as releasing the toxins or public goods can also significantly affect the spatial structure of the community~\cite{blanchard_bacterial_2015}.
-Therefore, it is rather important to establish the limitations for the model application.\\
-
+These effects can only be treated with a spatial model that is able to describe the observed effects.\\
+%
+%
 - \cite{emonet_agentcell_2005, gorochowski_bsim_2012, li_nufeb_2019, kreft_bacsim_1998}: examples of another toolboxes for bacterial communities modeling (ABM).\\
@@
 \begin{figure}
-    \includegraphics[width=0.5\textwidth]{Figures/snapshot_00000000.png}%
-    \includegraphics[width=0.5\textwidth]{Figures/snapshot_00006500.png}\\
-    \includegraphics[width=0.5\textwidth]{Figures/snapshot_00013000.png}%
-    \includegraphics[width=0.5\textwidth]{Figures/snapshot_00020000.png}%
+    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00000000.png}\hspace{0.02\textwidth}%
+    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00006500.png}\vspace{0.02\textwidth}\\
+    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00013000.png}\hspace{0.02\textwidth}%
+    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00020000.png}%
     \caption{}
@@
     \centering
-    \includegraphics[width=0.6\textwidth]{Figures/cell_growth.pdf}
+    \includegraphics[width=0.6\textwidth]{Figures/spatial_cell_growth.png}
     \caption{}
@@
 \begin{figure}
-    \centering
-    \includegraphics[width=0.6\textwidth]{Figures/lag_phase.pdf}
+    \includegraphics[width=0.5\textwidth]{Figures/lag_phase.png}%
+    \includegraphics[width=0.5\textwidth]{Figures/entropy.png}
+    % TODO difference to pool model
     \caption{}
-    \label{fig:spatial-lag-phase}
-\end{figure}
-%
-%
-\begin{figure}
-    \centering
-    \includegraphics[width=0.6\textwidth]{Figures/entropy.pdf}
-    \caption{}
-    \label{fig:spatial-entropy}
+    \label{fig:spatial-lag-phase-entropy}
 \end{figure}
@@
     \begin{cases}
-        \dot{L_1} = -\lambda_1 R L_1\\
-        \dot{G_1} = \lambda_1 R L_1 + \alpha_1 R G_1\\
-        \dot{L_2} = -\lambda_2 R L_2\\
-        \dot{G_2} = \lambda_2 R L_2 + \frac{\alpha_2}{1 + \mu_I I} R G_2\\
+        \dot{L_A} = -\lambda_A R L_A\\
+        \dot{G_A} = \lambda_A R L_A + \alpha_A R G_A\\
+        \dot{L_B} = -\lambda_B R L_B\\
+        \dot{G_B} = \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B\\
     \end{cases}.
@@
     \begin{cases}
-        \dot{R} = -\frac{\alpha_1}{N_t} R G_1-\frac{\alpha_2}{(1 + \mu_I I) N_t} R G_2 \\
-        \dot{I} = \kappa G_1
+        \dot{R} = -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \mu_I I) N_t} R G_B \\
+        \dot{I} = \kappa G_A
     \end{cases}.

### 2023-12-20 pool_model.tex
first iteration of abm description

@@
 %
-\begin{table}[H]
+%
+%
+% ------------------------------------------------------------------------------------------------
+\subsection{Constructing an Agent-Based Pool Model}
+We use \texttt{cellular\_raza}~\cite{Pleyer_cellular_raza_2023} to implement an \ac{abm}.
+\subsubsection{Cellular Reactions}
+To construct an \ac{abm} which resembles the given \ac{ode} system of equations~\eqref{eq:spatial_limit_F} and \eqref{eq:spatial_limit_H}, we consider multiple cellular reactions.
+The first is uptake of an extracellular resource $[R_e]$ into the cell $[R_i]$ with rate $u$ in \unit{\per\minute}.
+Nutrients are only taken up at the position of the cell $x_c$
+\begin{align}
+    \dot{[R_i]} &= u [R_e](x_c)\\
+    \dot{[R_e]}(x) &= -u \frac{V_c}{V_D} [R_e] \delta(x-x_c)
+\end{align}
+where $\delta(x-x_\text{cell})$ is the dirac delta-distribution.
+Note that $[R_e]=[R_e](x)$ explicitly depends on the spatial coordinate but we choose to omit its notation.
+In the second equation, we need to account for the ratio of cellular $V_c$ and domain volume $V_D$ to calculate the correct change.
+Further, we want to modify the cellular volume by growing it increasingly with rate $a$.
+Therefore, we need to specify how many resource molecules are turned into one volumetric units $\sigma$ in \unit{\micro\metre\cubed\per\mole}.
+The total amount of resource molecules present inside the cell is given by the current concentration multiplied by the cell-volume $V_c [R_i]$.
+With a rate $a$ and conversion $\sigma$, the cell turns these resource molecules into volume.
+\begin{align}
+    \dot{V_c} &= a\sigma [R_i] V_c\\
+    \dot{[R_i]} &= - a[R_i]
+\end{align}
+Furthermore, we let cells produce an inhibitor with production rate $b$ which inhibits the cells ability to uptake and thus use the resource by a factor of $1+\nu_I [I]$
+\begin{align}
+    \dot{[I]} &= b\frac{V_c}{V_D}\delta(x-x_c)\\
+    u &\longrightarrow \frac{u}{1+\nu_I [I]}.
+\end{align}
+If we combine all equations and consider multiple cells, we obtain the system
+\begin{align}
+    \dot{V_c} &= a\sigma [R_i] V_c\\
+    \dot{[R_i]} &= \frac{u}{1+\nu_I [I]} [R_e] - a[R_i]\\
+    \dot{[R_e]} &= -\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R_e] \delta(x-x_c)\\
+    \dot{[I]} &= b\frac{V_c}{V_D}\delta(x-x_c).
+    \label{eq:spatial-full-system-non-simplified}
+\end{align}
+In order to simplify this system, we can assume that the internal resources are constantly being depleted by the growth and the limiting process is the uptake of the resource.
+This means, we can effectively set $\dot{[R_i]}=0$.
+Therefore, we obtain the relation $\frac{u}{1+\nu_I [I]}[R_e] = a[R_i]$ which we can use to eliminate the variable monitoring internal resource concentrations $[R_i]$.
+Additionally, we require our external components to be freely diffusable with constants $D_R,D_I$.
+Plugging this into equation~\eqref{eq:spatial-full-system-non-simplified} yields.
+\begin{align}
+    \dot{V_c} &= \frac{u}{1+\nu_I [I]} \sigma [R]V_c\\
+    \dot{[R]} &= \sum\limits_c -\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c) + D_R\Delta [R]\\
+    \dot{[I]} &= \sum\limits_c b\frac{V_c}{V_D}\delta(x-x_c) + D_I\Delta [I].
+    \label{eq:spatial-full-system-simplified}
+\end{align}
+For the diffusion, we require Neumann (zero-flux) boundary conditions.
+Since, we will not explicitly model intracellular concentrations $[R_i]$, we will drop the index entirely, only considering the extracellular resource.
+%
+%
+% ................................................................................................
+\subsubsection{Physical Mechanics}
+We represent cells as spherical objects in 2D space.
+They can be represented by their position $x_c$ and radius $r_c$ which is of course related to their (2D) volume
+\begin{equation}
+    V_c = \pi r_c^2
+\end{equation}
+As the coupled reactions from equations~\eqref{eq:spatial-full-system-simplified} change $V_c$, we also automatically adapt the cell radius accordingly.
+Notice however, that how the radius is updated does not correspond to the cellular reactions defined previously but is a purely mechanical consideration.
+The dimensionality of the problem does not find its way into the cellular reactions but only manifests itself in these mechanical considerations.
+Cells are interacting via a point-like interaction potential.
+We can calculate the interaction strength via
+\begin{alignat}{3}
+    \vec{z} &= \vec{x_1} - \vec{x_2} \hspace{2em}&& r = ||\vec{z}||\\
+    \sigma &= \frac{r}{r_1+r_2} && q=0.2\\
+    \vec{F} &= \epsilon \frac{1+q}{q+\sigma} \theta(\sigma - 1) \frac{1}{r}\vec{z}
+\end{alignat}
+Cells are not sticky and thus only pushing on each other.
+No attractive force is present.
+This means that colonies are expanding as individual cells grow and push each other further apart.
+%
+\begin{figure}
     \centering
-    \begin{tabular}{ccc}
-        \specialrule{.1em}{.01em}{.05em} 
-        \textbf{Parameter} \hspace{3mm} & \textbf{Value} \hspace{3mm} & \textbf{Units}\\
-        \toprule
-        $\lambda_1$ & 0.008  & [time$^{-1}$]                 \\
-        $\lambda_2$ & 0.005  & [time$^{-1}$]                 \\
-        $\alpha_1$  & 4.0    & [time$^{-1}$]                 \\
-        $\alpha_2$  & 5.0    & [time$^{-1}$]                 \\
-        $\kappa$    & 0.1    & [(cell $\cdot$ time)$^{-1}$]  \\
-        $\mu_I$     & 0.1    & [cell$^{-1}$]                 \\
-        $N_t$       & $10^4$ & [cell]                        \\
-        \bottomrule
-    \end{tabular}
-    \caption{The list of the parameters of the system.}
-    \label{tab:spatial_limit_param}
-\end{table}
+    \includegraphics[width=0.6\textwidth]{Figures/interaciton_potential.png}
+    \caption{
+        Overall form of the interaction force between two cells.
+        A value $\sigma<1$ indicates that cells are overlapping which results in a positive force outwards.
+        Conversely, as long as cells are apart $\sigma>=1$, no force is acting.
+    }
+    \label{fig:spatial-interaction-potential}
+\end{figure}
+%
+%
+% ................................................................................................
+\subsubsection{Cell Cycle}
+Typically, cell cycles consist of multiple phases.
+Since our goal is to include only the necessary details of the underlying biological problem, we drastically simplify the description of the cell cycle.
+Cell growth is already being described the system of coupled \acp{ode} given in equations~\eqref{eq:spatial-full-system-simplified}.
+Thus we introduce an additional rule that splits apart the cell whenever a certain volume threshold is reached.
+The volume of the cell is divided equally between the new and old cell and both are assigned new random positions close to each other such that their radii do touch but not overlap.
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=0.4\textwidth]{Figures/snapshot-2-cells-before-division.png}
+    \includegraphics[width=0.4\textwidth]{Figures/snapshot-2-cells-after-division.png}
+    \caption{
+        A cell right before and after a division event.
+        The volume is equally distributed between each new cell and the new positions are picked random but at opposing sides on a circle around the original center of the single cell.
+    }
+    \label{fig:spatial-cell-division-example}
+    \end{center}
+\end{figure}
 %
@@
     \centering
-    \begin{tabular}{ccc}
-        \specialrule{.1em}{.01em}{.05em} 
-        \textbf{Variable} \hspace{3mm} & \textbf{Initial Value} \hspace{3mm} & \textbf{Units}\\
+    \begin{tabularx}{\textwidth}{@{}l *5{>{\centering\arraybackslash}X}@{}}
+        \multicolumn{3}{c}{\textbf{\acs{ode}}} & \multicolumn{3}{c}{\textbf{\acs{abm}}}\\
+        \cmidrule[1pt](l{\tabcolsep}r{\tabcolsep}){1-3}
+        \cmidrule[1pt](l{\tabcolsep}r{\tabcolsep}){4-6}
+         & \textbf{Symbol} & \textbf{Value} && \textbf{Symbol} & \textbf{Value}\\
         \toprule
-        $L_1$ & 5.0 & [cell]        \\
-        $G_1$ & 0.0 & [cell]        \\
-        $L_2$ & 5.0 & [cell]        \\
-        $G_2$ & 0.0 & [cell]        \\
-        $R$   & 1.0 & [cell/cell]   \\
-        $I$   & 0.0 & [cell]        \\
+        Rate $L_A\rightarrow G_A$   & $\lambda_A$   & $0.001$\unit{\per\minute}
+        &                           & $\lambda_A$   & $0.001$\unit{\per\minute}\\
+        Rate $L_B\rightarrow G_B$   & $\lambda_B$   & $0.0005$\unit{\per\minute}
+        &                           & $\lambda_B$   & $0.0005$\unit{\per\minute}\\
+        Production $G_A$            & $\alpha_A$    & ??\unit{\per\minute}
+        & Uptake                    & $u_A$         & $0.01$\unit{\per\minute}\\
+        Production $G_B$            & $\alpha_B$    & ??\unit{\per\minute}
+        & Uptake                    & $u_B$         & $0.01$\unit{\per\minute}\\
+        Production $I$              & $\kappa$      & $1.18\cdot10^{-5}$\unit{\per\minute}
+        &&&\\
+        Inhibition         & $\mu_I$       & $1.11\cdot 10^{-6}$\unit{\per\micro\metre\cubed}
+        &&&\\
+        Total volume     & $N_t$         & $1.80\cdot 10^4$\unit{\micro\metre\cubed}
+        &&&\\
+        \textbf{Initial Values}\\
+        \toprule
+        Volume of $L_A$             & $L_A$         & $35.34$\unit{\micro\metre\cubed}
+        & Cell Number               & $C_A$         & 5\\
+        Volume of $G_A$             & $G_A$         & $0.0$\unit{\micro\metre\cubed}
+        &                           &               & 0\\
+        Volume of $L_B$             & $L_B$         & $28.27$\unit{\micro\metre\cubed}
+        & Cell Number               & $C_B$         & 4\\
+        Volume of $G_B$             & $G_B$         & $0.0$\unit{\micro\metre\cubed}
+        &                           &               & 0\\
+        Relative nutrients          & $R$           & $1.0$
+        & Nutrients                 & $A$           & 2.0\unit{\mol\per\micro\metre\cubed}\\
+        Volume inhibitor            & $I$           & $0.0$\unit{\micro\metre\cubed}
+        & Inhibitor                 &               & $0.0$\unit{\mol\per\micro\metre\cubed}\\
         \bottomrule
-    \end{tabular}
+    \end{tabularx}
     \caption{The list of the initial values of the system.}
@@
 %
+%
+% ================================================================================================
+\beginsupplement
+\section*{Supplementary Material}
+\subsection{Deriving the Pool Model}
+Since the \ac{abm} can be considered more fundamental than the \ac{ode} model, we should be able to derive the latter one from the former.
+This can be achieved by starting with equations~\eqref{eq:spatial-full-system-simplified}.
+We integrate the spatial equations over the total domain starting with the resource concentration $[R]$.
+\begin{alignat}{7}
+    \int\limits_D\dot{[R]}\text{dVol} 
+    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c) &&+ D_R\Delta [R]\bigg]\text{dVol}\\
+    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c)\bigg]\text{dVol} &&+ \int\limits_D D_R\Delta [R]\text{dVol}\\
+    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c)\bigg]\text{dVol} &&+ \int\limits_{\partial D} D_R\vec{\nabla} [R]\vec{\text{dA}}\\
+    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c \int\limits_D \frac{1}{V_D} [R] \delta(x-x_c)\text{dVol}\\
+    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c [R]_{x=x_c}\\
+    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c [R]\\
+    &= &&&&-\frac{u}{1+\nu_I [I]V_D V_D^{-1}} V [R]\frac{V_D}{V_D}\\
+    \dot{R} &= &&&&-\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V\\
+\end{alignat}
+We can split up the integral into two parts.
+The second part can then be transformed into an integral over the flux of the resource along the boundary.
+Since we chose Neumann boundary conditions, this contribution vanishes identically.
+Afterwards, the first part can be evaluated by using the delta-distribution.
+Note that we additionally need to consume a $V_D^{-1}$ term since our volume is finite.
+We evaluate the resource concentration $[R]$ at the position of the individual cells and sum over all contributions.
+Now we assume homogeneity in the sense, that we can reduce $[R]$ to only a single variable instead of considering its spatial distribution.
+This means that every contribution does not depend on the position at which it was evaluated, ie. $[R]_{x=x_c}=[R]$.
+Finally, we can perform the sum to gain the total volume of all cells $V=\sum\limits_c V_c$.
+\begin{alignat}{7}
+    \int\limits_D\dot{[I]} &= \int\limits_D \bigg[&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) &&+ D_I\Delta [I]\bigg]\text{dVol}\\
+    &= \int\limits_D&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) \text{dVol} &&+ \int\limits_D D_I\Delta [I]\text{dVol}\\
+    &= \int\limits_D&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) \text{dVol} &&+ \int\limits_{\partial D} D_I\nabla [I]\vec{\text{dA}}\\
+    &= &&\sum\limits_c b V_c\int\limits_D \frac{1}{V_D} &&\delta(x-x_c) \text{dVol}\\
+    \dot{I} &= b V
+\end{alignat}
+Following the same principle as before, we can integrate the inhibitor equation.
+For the equation describing the growth of the cell volume, we simply add up all cells to obtain the total volume of all cells.
+\begin{alignat}{3}
+    \sum\limits_c\dot{V_c}
+    &= \sum\limits_c&&\frac{u}{1+\nu_I [I]} \sigma [R]V_c\\
+    &= &&\frac{u}{1+\nu_I [I]} \sigma [R]V\\
+    \dot{V} &= &&\frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V
+\end{alignat}
+We can thus summarize the equations obtained
+\begin{align}
+    \dot{R} &= -\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V\\
+    \dot{I} &= b V\\
+    \dot{V} &= \frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V.
+\end{align}
+It is clear that the original model is recovered this way.
+%
+%
+% ................................................................................................
+\subsection{Code Samples}
+The code samples show sample code from the simulation that calculate cellular interactions and cell cycle.
+\begin{listing}
+\begin{lstlisting}[language=rust]
+let z = ext_pos - own_pos;
+let r = z.norm();
+let sigma = r / (self.cell_radius + ext_radius);
+if sigma < 1.0 {
+    let q = 0.2;
+    let dir = z.normalize();
+    let modifier = (1.0 + q) / (q + sigma);
+    return self.potential_strength * dir * modifier;
+} else {
+    return Vector2::zero();
+}
+\end{lstlisting}
+\caption{
+    Sample code for calculation of interactions between two bacteria cells.
+}
+\label{listing:spatial-interaction-strength}
+\end{listing}
+%
+%
+\begin{listing}
+\begin{lstlisting}[language=rust]
+// Clone existing cell
+let mut c2 = c1.clone();
+
+let r = c1.interaction.cell_radius;
+
+// Make both cells smaller
+c1.interaction.cell_radius /= std::f64::consts::SQRT_2;
+c2.interaction.cell_radius /= std::f64::consts::SQRT_2;
+
+// Generate cellular splitting direction randomly
+let angle_1 = rng.gen_range(0.0..2.0 * std::f64::consts::PI);
+let dir_vec = nalgebra::Rotation2::new(angle_1) * Vector2::from([1.0, 0.0]);
+
+// Define new positions for cells
+// It is randomly chosen if the old cell is left or right
+let offset = dir_vec * r / std::f64::consts::SQRT_2;
+let old_pos = c1.pos();
+
+// Increase the remaining food in the bacteria by the fraction of volumes
+// from before and after the division. This exactly coincides with factor 2
+// since we half the volume of the cell
+c1.cellular_reactions.intracellular_concentrations[0] *= 2.0;
+c2.cellular_reactions.intracellular_concentrations[0] *= 2.0;
+
+c1.set_pos(&(old_pos + offset));
+c2.set_pos(&(old_pos - offset));
+\end{lstlisting}
+\caption{
+    Sample code for dividing one bacteria cell into two.
+}
+\end{listing}
+%
+%
 \end{document}

### 2023-12-20 pool_model.tex
more comments about code
tell the reader how to get full equations (with lag phase)

@@
 It is clear that the original model is recovered this way.
+To obtain the full equations, we additionally need to consider transition between states of lag-phase and growth phase not depicted here.
+The preceding equations all assume that cells are in growth phase.
+It is however straiforward, to setup these equations.
 %
@@
 The code samples show sample code from the simulation that calculate cellular interactions and cell cycle.
+Since backend of the simulation is written in Rust, so is the implementation of cellular aspects.
+Additionally bindings are generated such that the full power of the simulation tool can be accessed.
+This significantly improves handling and development time of new models.
 \begin{listing}

### 2023-12-20 pool_model.tex
make sure to specify that this is not really the total nutrients

@@
 To test the spatial limitations of the described above approach consider the system consisting from two bacterial species that can be described by two internal states: lag and growth states. 
-Both of these species grow using limiting resource pool $R$ with different growth constants $\alpha$ and different lengths of lag phase determined by parameter $\lambda$.
+Both of these species grow using limiting resource pool $R^*$ with different growth constants $\alpha$ and different lengths of lag phase determined by parameter $\lambda$.
+It has to be noted that the resource pool is normalized by its initial value, ie $R^* = R/R_0$.
 Moreover, the species 1 is able to release a certain antimicrobial component $I$ that shows inhibitory effect on species 2 (amensalism).
@@
     \begin{cases}
-        \dot{L_A} = -\lambda_A R L_A\\
-        \dot{G_A} = \lambda_A R L_A + \alpha_A R G_A\\
-        \dot{L_B} = -\lambda_B R L_B\\
-        \dot{G_B} = \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B\\
+        \dot{L_A} = -\lambda_A R^* L_A\\
+        \dot{G_A} = \lambda_A R^* L_A + \alpha_A R^* G_A\\
+        \dot{L_B} = -\lambda_B R^* L_B\\
+        \dot{G_B} = \lambda_B R^* L_B + \frac{\alpha_B}{1 + \mu_I I} R^* G_B\\
     \end{cases}.
@@
     \begin{cases}
-        \dot{R} = -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \mu_I I) N_t} R G_B \\
+        \dot{R^*} = -\frac{\alpha_A}{N_t} R^* G_A-\frac{\alpha_B}{(1 + \mu_I I) N_t} R^* G_B \\
         \dot{I} = \kappa G_A

### 2023-12-20 pool_model.tex
explain implementation of lag phase in abm

@@
 \end{figure}
+In addition to modeling division events, the cells have are subject to the lag-phase which delays their ability to proliferate.
+We model this by assigning a transition rate to each cell.
+For each time point of the simulation, we generate a boolean value from a random number generator with probability
+\begin{equation}
+    P(\text{transition}) = \Delta t \lambda \frac{[R]}{[R_0]}.
+\end{equation}
+If this boolean is true, we move the cell from the inactive state to the active state where it starts to consume nutrients and grow.
+It is clear that this is exactly identical to the \ac{ode} model.
 %

### 2023-12-20 pool_model.tex
begin section to compare abm with ode model

@@
 %
+%
+%
+% ------------------------------------------------------------------------------------------------
+\subsection{Comparing \acs{abm} and \acs{ode} Models}
+The \ac{abm} model is more fundamental than the \ac{ode} model in the sense that it incorporates more biological aspects and maps them closer to realistic biological mechanisms.
+We are thus able to derive the \ac{ode} pool model from its \ac{abm} version in the homogenenous case (see supplementary material~\ref{subsec:supplement-deriving-pool-odes}).
+By doing so, we find a mapping between parameters from the \ac{abm} to parameters used in the \ac{ode} description.
+\begin{align}
+    \kappa &= b\\
+    \alpha &= u\sigma\frac{1}{V_D}\\
+    \mu_I &= \nu_I \frac{1}{V_D}
+\end{align}
+\begin{figure}[H]
+    \begin{center}
+    \includegraphics[width=0.6\textwidth]{Figures/pool_model_spatial.pdf}
+    \caption{
+        The output of the model (\ref{eq:spatial_limit_F}, \ref{eq:spatial_limit_H}) for given parameters and initial valued presented in Tables \ref{tab:spatial_limit_param}, \ref{tab:spatial_limit_init}, respectively.
+        The observables of the system are the bacterial count of each species $L+G$ (blue and orange lines) and the total bacterial count of the system (green).
+    }
+    \label{fig:spatial_limit_plot1}
+    \end{center}
+\end{figure}
+%
 \begin{table}[H]
@@
 \end{table}
-%
-\begin{figure}[H]
-    \begin{center}
-    \includegraphics[width=0.6\textwidth]{Figures/pool_model_spatial.pdf}
-    \caption{
-        The output of the model (\ref{eq:spatial_limit_F}, \ref{eq:spatial_limit_H}) for given parameters and initial valued presented in Tables \ref{tab:spatial_limit_param}, \ref{tab:spatial_limit_init}, respectively.
-        The observables of the system are the bacterial count of each species $L+G$ (blue and orange lines) and the total bacterial count of the system (green).
-    }
-    \label{fig:spatial_limit_plot1}
-    \end{center}
-\end{figure}
 %

### 2023-12-20 pool_model.tex
label subsection in supplement
update graphs and parameters to fit new simulation data

@@
 More about the modelling of the motile bacteria can be seen in works:~\cite{decoene_microscopic_2011, rosser_modelling_2014, sokolov_physical_2012, li_amplified_2008}.
-\begin{figure}
-    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00000000.png}\hspace{0.02\textwidth}%
-    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00006500.png}\vspace{0.02\textwidth}\\
-    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00013000.png}\hspace{0.02\textwidth}%
-    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00020000.png}%
-    \caption{}
-    \label{fig:spatial-snapshots}
-\end{figure}
-%
-%
-\begin{figure}
-    \centering
-    \includegraphics[width=0.6\textwidth]{Figures/spatial_cell_growth.png}
-    \caption{}
-    \label{fig:spatial-growth-curve}
-\end{figure}
-%
-%
-\begin{figure}
-    \includegraphics[width=0.5\textwidth]{Figures/lag_phase.png}%
-    \includegraphics[width=0.5\textwidth]{Figures/entropy.png}
-    % TODO difference to pool model
-    \caption{}
-    \label{fig:spatial-lag-phase-entropy}
-\end{figure}
 %
@@
     \begin{cases}
-        \dot{R^*} = -\frac{\alpha_A}{N_t} R^* G_A-\frac{\alpha_B}{(1 + \mu_I I) N_t} R^* G_B \\
+        \dot{R^*} = -\frac{\alpha_A}{G_\text{max}} R^* G_A-\frac{\alpha_B}{(1 + \mu_I I) G_\text{max}} R^* G_B \\
         \dot{I} = \kappa G_A
@@
 \end{align}
+Note that we do not need to distinguish between the two species because we can set parameters individually on agent-level.
+This means, we can implement the most general version of the equations and by choosing appropriate parameters, set the species of the respective cell-agent.
 If we combine all equations and consider multiple cells, we obtain the system
@@
 By doing so, we find a mapping between parameters from the \ac{abm} to parameters used in the \ac{ode} description.
+The derived equations read
 \begin{align}
-    \kappa &= b\\
-    \alpha &= u\sigma\frac{1}{V_D}\\
-    \mu_I &= \nu_I \frac{1}{V_D}
+    \dot{R} &= -\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V\\
+    \dot{I} &= b V\\
+    \dot{V} &= \frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V.
 \end{align}
-\begin{figure}[H]
-    \begin{center}
-    \includegraphics[width=0.6\textwidth]{Figures/pool_model_spatial.pdf}
+and the \ac{ode} equations are
+\begin{align}
+    \dot{L_A} &= -\lambda_A R^* L_A\\
+    \dot{G_A} &= \lambda_A R^* L_A + \alpha_A R^* G_A\\
+    \dot{L_B} &= -\lambda_B R^* L_B\\
+    \dot{G_B} &= \lambda_B R^* L_B + \frac{\alpha_B}{1 + \mu_I I} R^* G_B\\
+    \dot{R^*} &= -\frac{\alpha_A}{G_\text{max}} R^* G_A-\frac{\alpha_B}{(1 + \mu_I I) G_\text{max}} R^* G_B \\
+    \dot{I} &= \kappa G_A
+\end{align}
+We compare the most general version and thus omit species $A$ for the moment.
+\begin{table}
+    \centering
+    \renewcommand{\arraystretch}{1.5}
+    \begin{tabularx}{\textwidth}{@{}l *3{>{\centering\arraybackslash}X}@{}}
+        & \textbf{\acs{ode}}
+        & \textbf{\acs{abm}}
+        & \textbf{Mappings}\\
+
+        \toprule
+
+        \multirow{2}{*}{Resource}
+        & \multirow{2}{*}{$\displaystyle \dot{R^*} = - \frac{\alpha}{(1 + \mu_I I) G_\text{max}} R^* G$}
+        & \multirow{2}{*}{$\displaystyle \dot{R} = -\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V$}
+        & $\displaystyle G_\text{max} = [R_0] V_D \sigma$\\
+
+        &&& $\displaystyle \mu_I = \frac{\nu_I}{V_D}$\\
+
+        Volume
+        & $\displaystyle \dot{G} = \frac{\alpha}{1 + \mu_I I} R^* G$
+        & $\displaystyle \dot{V} = \frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V$
+        & $\displaystyle \alpha = u[R_0]\sigma$\\
+
+        Inhibitor
+        & $\displaystyle \dot{I} = \kappa G_A$
+        & $\displaystyle \dot{I} = b V$
+        & $\displaystyle \kappa = b$\\
+        \bottomrule
+    \end{tabularx}
+\end{table}
+We can use these mappings to numerically compare both approaches.
+Table~\ref{tab:spatial_limit_init} shows a comparison of paramters of two simulations for the \ac{ode} and \ac{abm} model which have been mapped onto each other.
+Figure~\ref{fig:spatial-snapshots} shows individual snapshots of different 
+%
+\begin{figure}
+    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00000000.png}\hspace{0.02\textwidth}%
+    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00006500.png}\vspace{0.02\textwidth}\\
+    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00013000.png}\hspace{0.02\textwidth}%
+    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00020000.png}%
+    \caption{}
+    \label{fig:spatial-snapshots}
+\end{figure}
+%
+%
+\begin{figure}
+    \centering
+    \includegraphics[width=0.7\textwidth]{Figures/abm_ode_comparison.png}
     \caption{
-        The output of the model (\ref{eq:spatial_limit_F}, \ref{eq:spatial_limit_H}) for given parameters and initial valued presented in Tables \ref{tab:spatial_limit_param}, \ref{tab:spatial_limit_init}, respectively.
-        The observables of the system are the bacterial count of each species $L+G$ (blue and orange lines) and the total bacterial count of the system (green).
-    }
-    \label{fig:spatial_limit_plot1}
-    \end{center}
+        Comparison of the \ac{abm} and \ac{ode} output corresponding to the parameters presented in table~\ref{tab:spatial_limit_init}.
+        % TODO
+    }
+    \label{fig:spatial-growth-curve}
 \end{figure}
+%
 %
@@
         \toprule
-        Rate $L_A\rightarrow G_A$   & $\lambda_A$   & $0.001$\unit{\per\minute}
-        &                           & $\lambda_A$   & $0.001$\unit{\per\minute}\\
+        Rate $L_A\rightarrow G_A$   & $\lambda_A$   & $0.0025$\unit{\per\minute}
+        &                           & $\lambda_A$   & $0.008$\unit{\per\minute}\\
         Rate $L_B\rightarrow G_B$   & $\lambda_B$   & $0.0005$\unit{\per\minute}
         &                           & $\lambda_B$   & $0.0005$\unit{\per\minute}\\
-        Production $G_A$            & $\alpha_A$    & ??\unit{\per\minute}
+        Production $G_A$            & $\alpha_A$    & $0.002$\unit{\per\minute}
         & Uptake                    & $u_A$         & $0.01$\unit{\per\minute}\\
-        Production $G_B$            & $\alpha_B$    & ??\unit{\per\minute}
+        Production $G_B$            & $\alpha_B$    & $0.002$\unit{\per\minute}
         & Uptake                    & $u_B$         & $0.01$\unit{\per\minute}\\
-        Production $I$              & $\kappa$      & $1.18\cdot10^{-5}$\unit{\per\minute}
+        Production $I$              & $\kappa$      & $0.1$\unit{\per\minute}
         &&&\\
-        Inhibition         & $\mu_I$       & $1.11\cdot 10^{-6}$\unit{\per\micro\metre\cubed}
+        Inhibition         & $\mu_I$       & $1.0\cdot 10^{-7}$\unit{\per\micro\metre\cubed}
         &&&\\
-        Total volume     & $N_t$         & $1.80\cdot 10^4$\unit{\micro\metre\cubed}
+        Total volume     & $G_\text{max}$         & $2.0\cdot 10^5$\unit{\micro\metre\cubed}
         &&&\\
@@
         \toprule
-        Volume of $L_A$             & $L_A$         & $35.34$\unit{\micro\metre\cubed}
-        & Cell Number               & $C_A$         & 5\\
+        Volume of $L_A$             & $L_A$         & $127.2$\unit{\micro\metre\cubed}
+        & Cell Number               & $C_A$         & 18\\
         Volume of $G_A$             & $G_A$         & $0.0$\unit{\micro\metre\cubed}
         &                           &               & 0\\
-        Volume of $L_B$             & $L_B$         & $28.27$\unit{\micro\metre\cubed}
-        & Cell Number               & $C_B$         & 4\\
+        Volume of $L_B$             & $L_B$         & $127.23$\unit{\micro\metre\cubed}
+        & Cell Number               & $C_B$         & 18\\
         Volume of $G_B$             & $G_B$         & $0.0$\unit{\micro\metre\cubed}
@@
         Relative nutrients          & $R$           & $1.0$
-        & Nutrients                 & $A$           & 2.0\unit{\mol\per\micro\metre\cubed}\\
+        & Nutrients                 & $A$           & $1.0$\unit{\mol\per\micro\metre\cubed}\\
         Volume inhibitor            & $I$           & $0.0$\unit{\micro\metre\cubed}
@@
 \subsection{Deriving the Pool Model}
+\label{subsec:supplement-deriving-pool-odes}
 Since the \ac{abm} can be considered more fundamental than the \ac{ode} model, we should be able to derive the latter one from the former.

### 2023-12-20 pool_model.tex
new simulation run with corrected intracellular reactions
adapt parameters to new run
display more json files in supplement

@@
         Rate $L_A\rightarrow G_A$   & $\lambda_A$   & $0.0025$\unit{\per\minute}
-        &                           & $\lambda_A$   & $0.008$\unit{\per\minute}\\
+        &                           & $\lambda_A$   & $0.0025$\unit{\per\minute}\\
         Rate $L_B\rightarrow G_B$   & $\lambda_B$   & $0.0005$\unit{\per\minute}
@@
         Production $I$              & $\kappa$      & $0.1$\unit{\per\minute}
-        &&&\\
+        & Production                & $b$           & $0.1$\unit{\per\minute}\\
         Inhibition                  & $\mu_I$       & $1.0\cdot 10^{-7}$\unit{\per\micro\metre\cubed}
-        &&&\\
-        Total volume     & $G_\text{max}$         & $2.0\cdot 10^5$\unit{\micro\metre\cubed}
+        & Inhibition                & $\nu_I$       & $0.1$\unit{\micro\metre\cubed\per\mol}\\
+        Total volume                & $G_\text{max}$& $2.0\cdot 10^5$\unit{\micro\metre\cubed}
         &&&\\
@@
     \end{tabularx}
-    \caption{The list of the initial values of the system.}
+    \caption{
+        List paramters for both models.
+        Specifying all parameters here would requrie a much larger table.
+        For the full information, we refer to the supplementary material~\ref{sec:supplement-full-parameters}.
+    }
     \label{tab:spatial_limit_init}
@@
 %
+% ................................................................................................
+\subsection{Full List of Parameters}
+\label{sec:supplement-full-parameters}
+\begin{listing}
+\begin{lstlisting}[language=json]
+{
+    "diffusion_constants": [
+        15.0,
+        15.0
+    ],
+    "initial_concentrations": [
+        2.0,
+        0.0
+    ],
+    "size": 1000.0,
+    "n_voxels": null
+}
+\end{lstlisting}
+\caption{
+    The \texttt{domain.json} file specifying initial values of the domain.
+}
+\end{listing}
+%
+\begin{listing}
+\begin{lstlisting}[language=json]
+{
+    "random_seed": 1,
+    "n_times": 20001,
+    "dt": 0.5,
+    "t_start": 0.0,
+    "save_interval": 500,
+    "n_threads": 8,
+    "show_progressbar": true,
+    "save_path": "out/pool_model",
+    "save_add_date": true
+}
+\end{lstlisting}
+\caption{
+    The \texttt{meta\_params.json} file specifying all parameters not directly related to the simulation content.
+    These settings do not represent any model-specific properties but rather generic ones such as time-span or number of threads.
+}
+\end{listing}
+%
+\begin{listing}
+\begin{lstlisting}[language=json]
+{
+    "mechanics": {
+        "pos": [
+        142.85714285714286,
+        142.85714285714286
+        ],
+        "vel": [
+        0.0,
+        0.0
+        ],
+        "damping_constant": 0.5,
+        "mass": 7.704755982928968
+    },
+    "cycle": {
+        "volume_division_threshold": 14.137166941154069,
+        "lag_phase_transition_rate_1": 0.005,
+        "lag_phase_transition_rate_2": 0.0025
+    },
+    "cellular_reactions": {
+        "potential_strength": 0.5,
+        "food_to_volume_conversion": 0.1,
+        "lag_phase_active": true,
+        "species": "S1",
+        "cell_volume": 7.0685834705770345,
+        "uptake_rate": 0.01,
+        "inhibition_production_rate": 0.1,
+        "inhibition_coefficient": 0.1
+    },
+    "interactionextracellulargradient": null
+}
+\end{lstlisting}
+\caption{
+    The first entry of the \texttt{initial\_cells.json} file which defines every individual particle present in the simulation initially.
+}
+\end{listing}
+%
+%
 \end{document}

### 2023-12-20 pool_model.tex
nicely format table and include all parameters for the model
put everything from additional file into main file

@@
 ]{geometry}
-\input{additional_tex/packages}
+
+\usepackage[utf8]{inputenc}
+\usepackage[a4paper,
+            inner=2.5cm,
+            outer=3cm
+            ]{geometry}
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off, font=small]{caption}            
+\usepackage{acronym}
+\usepackage{graphicx}% Include figure files
+\usepackage{bm}% bold math
+\usepackage{color}
+\usepackage{hyperref}
+\usepackage[backend=biber,
+            sorting=none,
+%            citestyle=numeric-comp,
+%\            bibstyle=ieee, 
+            style=numeric-comp,
+            hyperref=true,
+            maxnames=2,
+            url=false,
+            isbn=false,
+            eprint=false,
+            date=year,
+            doi=false,
+            giveninits=true
+            ]{biblatex}
+\addbibresource{Bacteria Modelling.bib}
+\usepackage{xpatch}
+\xpatchbibmacro{journal+issuetitle}{%
+\usebibmacro{journal}%
+\setunit*{\addspace}%
+}{%
+\usebibmacro{journal}%
+\setunit*{\addcomma\space}%
+}{}{}
+
+\renewbibmacro*{name:andothers}{% Based on name:andothers from biblatex.def
+  \ifboolexpr{
+    test {\ifnumequal{\value{listcount}}{\value{liststop}}}
+    and
+    test \ifmorenames
+  }
+    {\ifnumgreater{\value{liststop}}{1}
+       {\finalandcomma}
+       {}%
+     \andothersdelim\bibstring[\emph]{andothers}}
+    {}}
+
+%\DeclareFieldFormat[article]{title}{#1}
+\DeclareFieldFormat[article]{volume}{\textbf{#1}}
+%\DeclareFieldFormat[book]{volume}{\textbf{#1}}
+
+\renewcommand\newunitpunct{\addcomma\space}
+\DefineBibliographyStrings{english}{%
+page = {},
+pages = {},
+}
+
+\renewbibmacro{in:}{\addcomma\addspace}
+
+\renewbibmacro*{volume+number+eid}{%
+\isdot
+\printfield{volume}%
+\setunit*{\addcomma\addspace}%
+\iffieldundef{number}{}{%
+\printtext{no\adddot\addnbthinspace}%
+}%
+\printfield{number}%
+\setunit{\addcomma\space}%
+\printfield{eid}}
+
+\setlength{\tabcolsep}{1.3pt}
+\usepackage{bm}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{mathtools}
+\usepackage{mathrsfs}
+\usepackage[T1]{fontenc}
+\usepackage{diagbox}
+\usepackage{here}
+\usepackage[dvipsnames]{xcolor}
+\usepackage{lipsum}
+\usepackage{multirow}
+\usepackage{ctable} % for \specialrule command
+\usepackage{array,booktabs}
+\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
+\usepackage{subcaption}
+\usepackage{float}
+\usepackage{tabularx}
+\usepackage{siunitx}
+\usepackage{listings}
+\usepackage{listings-rust}
+\lstset{
+    language=Rust,
+    showstringspaces=false
+}
+\usepackage{newfloat}
+\DeclareFloatingEnvironment[
+  name=Listing,
+  listname=List of Listings,
+  within=section,
+]{listing}
+\restylefloat{table}
+\numberwithin{equation}{section}
+
+\newcommand{\beginsupplement}{%
+  \setcounter{subsection}{0}
+  \renewcommand{\thesubsection}{S\arabic{subsection}}
+  \setcounter{table}{0}
+  \renewcommand{\thetable}{S\arabic{table}}%
+  \setcounter{figure}{0}
+  \renewcommand{\thefigure}{S\arabic{figure}}%
+  \setcounter{listing}{0}
+  \renewcommand{\thelisting}{S\arabic{listing}}
+}
+
+\newcommand{\mbP}{\textbf{P: }}
+\newcommand{\mbC}{\textbf{C: }}
+\newcommand{\mbx}{\mathbf{x}}
+\newcommand{\mbm}{\mathbf{m}}
+\newcommand{\mbp}{\mathbf{p}}
+\newcommand{\mbu}{\mathbf{u}}
+\newcommand{\mbF}{\mathbf{F}}
+\newcommand{\mbH}{\mathbf{H}}
+\newcommand{\cmtb}[1]{\textcolor{blue} {#1}}
+\newcommand{\cmt}[1]{\textcolor{red} {[#1]}}
+\newcommand{\etal}{{\textit{et al. }}}
+
+\newacro{ode}[ODE]{Ordinary Differential Equation}
+\newacro{glv}[gLV]{generalized Lotka-Volterra}
+\newacro{abm}[ABM]{Agent-Based Model}
+
+% Define new style for json listings
+% Copied from https://www.latex4technics.com/?note=3QTU
+\colorlet{punct}{red!60!black}
+\definecolor{background}{HTML}{EEEEEE}
+\definecolor{delim}{RGB}{20,105,176}
+\colorlet{numb}{magenta!60!black}
+
+\lstdefinelanguage{json}{
+    basicstyle=\normalfont\ttfamily,
+    numbers=left,
+    numberstyle=\scriptsize,
+    stepnumber=1,
+    numbersep=8pt,
+    showstringspaces=false,
+    breaklines=true,
+    frame=lines,
+    backgroundcolor=\color{background},
+    literate=
+     *{0}{{{\color{numb}0}}}{1}
+      {1}{{{\color{numb}1}}}{1}
+      {2}{{{\color{numb}2}}}{1}
+      {3}{{{\color{numb}3}}}{1}
+      {4}{{{\color{numb}4}}}{1}
+      {5}{{{\color{numb}5}}}{1}
+      {6}{{{\color{numb}6}}}{1}
+      {7}{{{\color{numb}7}}}{1}
+      {8}{{{\color{numb}8}}}{1}
+      {9}{{{\color{numb}9}}}{1}
+      {:}{{{\color{punct}{:}}}}{1}
+      {,}{{{\color{punct}{,}}}}{1}
+      {\{}{{{\color{delim}{\{}}}}{1}
+      {\}}{{{\color{delim}{\}}}}}{1}
+      {[}{{{\color{delim}{[}}}}{1}
+      {]}{{{\color{delim}{]}}}}{1},
+}
 
@@
 % ................................................................................................
-\subsubsection{Physical Mechanics}
+\subsubsection{Physical Mechanics \& Interactions}
 We represent cells as spherical objects in 2D space.
@@
 The dimensionality of the problem does not find its way into the cellular reactions but only manifests itself in these mechanical considerations.
+The underlying equation of motion is that of a standard damped particle
+\begin{equation}
+    m \ddot{\vec{x}} = -\eta\dot{\vec{x}} + \vec{F}
+\end{equation}
 Cells are interacting via a point-like interaction potential.
@@
     \centering
-    \begin{tabularx}{\textwidth}{@{}l *5{>{\centering\arraybackslash}X}@{}}
+    \begin{tabularx}{\textwidth}{@{}l *2{>{\centering\arraybackslash}X} l *2{>{\centering\arraybackslash}X}@{}}
         \multicolumn{3}{c}{\textbf{\acs{ode}}} & \multicolumn{3}{c}{\textbf{\acs{abm}}}\\
-        \cmidrule[1pt](l{\tabcolsep}r{\tabcolsep}){1-3}
-        \cmidrule[1pt](l{\tabcolsep}r{\tabcolsep}){4-6}
-         & \textbf{Symbol} & \textbf{Value} && \textbf{Symbol} & \textbf{Value}\\
+        \toprule
+        \multicolumn{6}{c}{\textbf{Cellular Reactions}}\\
+        \toprule
+        Production $G_A$            & $\alpha_A$    & $0.002$\unit{\per\minute}
+        & Uptake                    & $u_A$         & $0.01$\unit{\per\minute}\\
+        Production $G_B$            & $\alpha_B$    & $0.002$\unit{\per\minute}
+        & Uptake                    & $u_B$         & $0.01$\unit{\per\minute}\\
+        Production $I$              & $\kappa$      & $0.1$\unit{\per\minute}
+        & Production $I$            & $b$           & $0.1$\unit{\per\minute}\\
+        Inhibition                  & $\mu_I$       & $1.0\cdot 10^{-7}$\unit{\per\micro\metre\cubed}
+        &                           & $\nu_I$       & $0.1$\unit{\micro\metre\cubed\per\mol}\\
+        &&
+        & Conversion                & $\sigma$      & $0.1$\unit{\micro\metre\cubed\per\mol}\\
+        Total volume                & $G_\text{max}$& $2.0\cdot 10^5$\unit{\micro\metre\cubed}
+        &&&\\
+        \toprule
+        \multicolumn{6}{c}{\textbf{Physical Mechanics \& Interactions}}\\
+        \toprule
+        &&& Damping                 & $\eta$        & $0.5$\unit{\per\minute}\\
+        &&& Mass                    & $m$           & $7.70$\unit{\pico\gram}\\
+        &&& Potential               & $\epsilon$    & $0.5$\unit{\pico\gram\micro\metre\per\minute\squared}\\
+        \toprule
+        \multicolumn{6}{c}{\textbf{Cell Cycle}}\\
         \toprule
@@
         &                           & $\lambda_B$   & $0.0005$\unit{\per\minute}\\
-        Production $G_A$            & $\alpha_A$    & $0.002$\unit{\per\minute}
-        & Uptake                    & $u_A$         & $0.01$\unit{\per\minute}\\
-        Production $G_B$            & $\alpha_B$    & $0.002$\unit{\per\minute}
-        & Uptake                    & $u_B$         & $0.01$\unit{\per\minute}\\
-        Production $I$              & $\kappa$      & $0.1$\unit{\per\minute}
-        & Production                & $b$           & $0.1$\unit{\per\minute}\\
-        Inhibition                  & $\mu_I$       & $1.0\cdot 10^{-7}$\unit{\per\micro\metre\cubed}
-        & Inhibition                & $\nu_I$       & $0.1$\unit{\micro\metre\cubed\per\mol}\\
-        Total volume                & $G_\text{max}$& $2.0\cdot 10^5$\unit{\micro\metre\cubed}
-        &&&\\
-        \textbf{Initial Values}\\
+        &&& Division Vol.           & $V_t$         & $14.14$\unit{\micro\metre\cubed}\\
+        \toprule
+        \multicolumn{6}{c}{\textbf{Initial Values}}\\
         \toprule
@@
         Volume of $G_A$             & $G_A$         & $0.0$\unit{\micro\metre\cubed}
-        &                           &               & 0\\
+        & Cell Volume               & $V_{c,A}$     & $7.07$\unit{\micro\metre\cubed}\\
         Volume of $L_B$             & $L_B$         & $127.23$\unit{\micro\metre\cubed}
@@
         Volume of $G_B$             & $G_B$         & $0.0$\unit{\micro\metre\cubed}
-        &                           &               & 0\\
-        Relative nutrients          & $R$           & $1.0$
-        & Nutrients                 & $A$           & $1.0$\unit{\mol\per\micro\metre\cubed}\\
-        Volume inhibitor            & $I$           & $0.0$\unit{\micro\metre\cubed}
-        & Inhibitor                 &               & $0.0$\unit{\mol\per\micro\metre\cubed}\\
+        & Cell Volume               & $V_{c,B}$     & $7.07$\unit{\micro\metre\cubed}\\
+        Relative Nutrients          & $R^*$         & $1.0$
+        & Nutrients                 & $[R_0]$       & $1.0$\unit{\mol\per\micro\metre\cubed}\\
+        Inhibitor                   & $I_0$         & $0.0$\unit{\mol}
+        & Inhibitor                 & $[I_0]$       & $0.0$\unit{\mol\per\micro\metre\cubed}\\
         \bottomrule

### 2023-12-20 pool_model.tex
compare abm and ode for different setups

@@
 \begin{figure}
-    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00000000.png}\hspace{0.02\textwidth}%
-    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00006500.png}\vspace{0.02\textwidth}\\
-    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00013000.png}\hspace{0.02\textwidth}%
-    \includegraphics[width=0.49\textwidth]{Figures/snapshot_00020000.png}%
-    \caption{}
+    \includegraphics[width=0.49\textwidth]{Figures/abm-homogenous/snapshot_00006000.png}\hspace{0.02\textwidth}%
+    \includegraphics[width=0.49\textwidth]{Figures/abm-homogenous/snapshot_00012000.png}\vspace{0.02\textwidth}\\
+    \includegraphics[width=0.49\textwidth]{Figures/abm-homogenous/snapshot_00018000.png}\hspace{0.02\textwidth}%
+    \includegraphics[width=0.49\textwidth]{Figures/abm-homogenous/snapshot_00024000.png}%
+    \caption{
+        Simulation snapshots at $1.04$\unit{\day},$2.08$\unit{\day},$3.13$\unit{\day} and $4.17$\unit{\day} corresponding to the parameters given in table~table~\ref{tab:spatial_limit_init}.
+        Species $B$ is colored blue while species $A$ is colored in red.
+    }
     \label{fig:spatial-snapshots}
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{Figures/abm_ode_comparison.png}
-    \caption{
-        Comparison of the \ac{abm} and \ac{ode} output corresponding to the parameters presented in table~\ref{tab:spatial_limit_init}.
-        % TODO
+    \begin{subfigure}[c]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{Figures/abm-homogenous/abm_ode_comparison.png}
+        \subcaption{Homogenous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
+    \end{subfigure}%
+    \begin{subfigure}[c]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{Figures/abm-homogenous-low-diffusion/abm_ode_comparison.png}
+        \subcaption{Homogenous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
+    \end{subfigure}\\
+    \begin{subfigure}[c]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{Figures/abm-inhomogenous/abm_ode_comparison.png}%
+        \subcaption{Inhomogenous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
+    \end{subfigure}%
+    \begin{subfigure}[c]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{Figures/abm-inhomogenous-low-diffusion/abm_ode_comparison.png}%
+        \subcaption{Inhomogenous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
+    \end{subfigure}
+    \caption{
+        Comparison of the \ac{ode} and \ac{abm} model for different parameter configurations.
+        Snapshots of the spatial configurations of cells are included in the appendix~\ref{fig:supplement-spatial-initial-config-cells}.
     }
@@
 %
+% ................................................................................................
+\subsection{Homogenous and Inhomogenous Initial Cell Distributions}
+To clarify what we mean by homogenous and inhomogenous initial configurations, we include snapshots of the simulation.
+At this point, cells have already grown such that the sizes of cell-clumps are better visible by the naked eye.
+The snapshots were taken after $2.08$\unit{\day}.
+\begin{figure}
+    \centering
+    \begin{subfigure}[c]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{Figures/abm-homogenous/snapshot_00012000.png}
+        \subcaption{Homogenous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
+    \end{subfigure}%
+    \begin{subfigure}[c]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{Figures/abm-homogenous-low-diffusion/snapshot_00012000.png}
+        \subcaption{Homogenous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
+    \end{subfigure}\\
+    \begin{subfigure}[c]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{Figures/abm-inhomogenous/snapshot_00012000.png}%
+        \subcaption{Inhomogenous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
+    \end{subfigure}%
+    \begin{subfigure}[c]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{Figures/abm-inhomogenous-low-diffusion/snapshot_00012000.png}%
+        \subcaption{Inhomogenous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
+    \end{subfigure}
+    \caption{
+        Spatial configuration of cells corresponding to the pictures shown in figure~\ref{fig:spatial-growth-curve}.
+    }
+    \label{fig:supplement-spatial-initial-config-cells}
+\end{figure}
+%
+%
 \end{document}

### 2023-12-20 pool_model.tex
include values for the diffusion constants
describe figures better

@@
         Species $B$ is colored blue while species $A$ is colored in red.
+        The background shows the resource concentration $[R]$.
+        Darker colours indicate less present nutrients.
     }
@@
         Snapshots of the spatial configurations of cells are included in the appendix~\ref{fig:supplement-spatial-initial-config-cells}.
+        We see that in the highly-diffusive homogenous case of subfigure (a), the pool model is a good fit to the \ac{abm} model.
+        Heterogeneity and slower diffusion cause larger differences between the two models as can be seen in subfigures (b) and (c).
+        The combination of the two effects shows the largest difference between model predictions.
+        In subfigure (d), due to the heterogenous cell distribution (see also~\ref{fig:supplement-spatial-initial-config-cells}) and slow diffusion, overall inhibition of species $B$ does not play as large of a role.
+        We furthermore note that the total bacteria volume of the \ac{abm} model is always smaller than the \ac{ode} model.
+        This is expected, since processes such as diffusion and cells growing outward introduce another time component which is simply assumed to be instantanous in the \ac{ode} model.
+        The chosen diffusion constants $D_I=D_R=30.0$\unit{\micro\metre\squared\per\minute} are in the order of small molecules such as $NaCl$ while the lower values $D_I=D_R=5.0$\unit{\micro\metre\squared\per\minute} are around values of Glucose in water.
     }
@@
         &&&\\
+        &&
+        & Diffusion $R$             & $D_R$         & $30.0$\unit{\micro\metre\squared\per\minute}\\
+        &&
+        & Diffusion $I$             & $D_I$         & $30.0$\unit{\micro\metre\squared\per\minute}\\
+        % Note: this diffusion roughly corresponds to the one of ethanol in water
+        % (see https://de.wikipedia.org/wiki/Diffusionskoeffizient#In_Fl%C3%BCssigkeiten)
+        % 15.0 µm^2/min = 0.9 m^2/s
+        % Ethanol: 0.84 m^2/s
         \toprule
@@
 \end{table}
-%
-\cmtb{The ABM additional parameters $D_R$, $D_I$, $q$ (At which level of $R_i$ cell division happens) and more parameters describing bacterial division.}
 %

### 2023-12-20 pool_model.tex
comment out many citations; specify that we track volume and total amount of nutrients/inhibitor

@@
 %
-- \cite{emonet_agentcell_2005, gorochowski_bsim_2012, li_nufeb_2019, kreft_bacsim_1998}: examples of another toolboxes for bacterial communities modeling (ABM).\\
-More about the modelling of the motile bacteria can be seen in works:~\cite{decoene_microscopic_2011, rosser_modelling_2014, sokolov_physical_2012, li_amplified_2008}.
-%
-%
-In general, bacterial communities in 2D can experience collective motion called swarming~\cite{wu_collective_2015}.
-Some works modelled 2D motile bacteria:
-\begin{itemize}
-    \item \cite{decoene_microscopic_2011}: 2D chemotactic self-propelled bacterial motion of the flagellated bacteria (e.g. E. coli, B. subtilis).
-    Chemotaxis here was driven by the oxygen concentrations.
-    \item \cite{rosser_modelling_2014}: modeling of the bacterial reorintation, rotation.
-    \item \cite{sokolov_physical_2012, li_amplified_2008}: modeling of the bacteria using hydrodynamic interactions and Brownian motion.
-    Fluctuations in velocity and rotations.
-    (\cite{li_amplified_2008}: Brownian motion gives right motion observed using microscopy).
-    \item \cite{farrell_mechanically_2013}: 2D modeling using Newtonian dynamics and only mechanical interactions + nutritient diffusion.
-    Branching parameter defines diff limits where diffusion dominant and limits growth, and where it's irrelevant.
-    Mechanical pressure (due to bacteria-bacteria interactions) affects growth, can limit it.(!!!)
-\end{itemize}
-%
-For simulation results:
-\begin{itemize}
-    \item \cite{farrell_mechanically_2013}: Showed the influence of mechanic interactions between bacteria in 2D system with bacteria consuming the nutritient.
-    The growth of bacterial colonies where non-motile microorganisms replicate and push each other away as they grow.
-    Found a transition between two different growth regimes, controlled by the balance between growth and uptake of nutrient:\\
-    Exponential growth where diffusion of the nutritient does not play any role. Sublinear growth where growth is limited by nutritient diffusion.
-    (Structurally in 2D it's regimes: smooth propagating of bacterial front vs. branching.)
-    \item \cite{nagarajan_agent-based_2022-1}: review on the ABM for bacterial population.
-    The ABM represent a microbial colony as a group of discrete agents that are governed by a set of rules, where each agent can be an individual cell or a collection of cells.
-    Can use citation when talk about general processes we are accounting as growth, uptake of nutritient, production of inhibitor, e.g.,:
-    Physical interactions between agents are calculated using excluded volume interactions.
-    Chemical interactions as releasing inhibitors are described using concentrations and solved using PDEs.
-\end{itemize}
+% - \cite{emonet_agentcell_2005, gorochowski_bsim_2012, li_nufeb_2019, kreft_bacsim_1998}: examples of another toolboxes for bacterial communities modeling (ABM).\\
+% More about the modelling of the motile bacteria can be seen in works:~\cite{decoene_microscopic_2011, rosser_modelling_2014, sokolov_physical_2012, li_amplified_2008}.
+%
+%
+% In general, bacterial communities in 2D can experience collective motion called swarming~\cite{wu_collective_2015}.
+% Some works modelled 2D motile bacteria:
+% \begin{itemize}
+%     \item \cite{decoene_microscopic_2011}: 2D chemotactic self-propelled bacterial motion of the flagellated bacteria (e.g. E. coli, B. subtilis).
+%     Chemotaxis here was driven by the oxygen concentrations.
+%     \item \cite{rosser_modelling_2014}: modeling of the bacterial reorintation, rotation.
+%     \item \cite{sokolov_physical_2012, li_amplified_2008}: modeling of the bacteria using hydrodynamic interactions and Brownian motion.
+%     Fluctuations in velocity and rotations.
+%     (\cite{li_amplified_2008}: Brownian motion gives right motion observed using microscopy).
+%     \item \cite{farrell_mechanically_2013}: 2D modeling using Newtonian dynamics and only mechanical interactions + nutritient diffusion.
+%     Branching parameter defines diff limits where diffusion dominant and limits growth, and where it's irrelevant.
+%     Mechanical pressure (due to bacteria-bacteria interactions) affects growth, can limit it.(!!!)
+% \end{itemize}
+%
+% For simulation results:
+% \begin{itemize}
+%     \item \cite{farrell_mechanically_2013}: Showed the influence of mechanic interactions between bacteria in 2D system with bacteria consuming the nutritient.
+%     The growth of bacterial colonies where non-motile microorganisms replicate and push each other away as they grow.
+%     Found a transition between two different growth regimes, controlled by the balance between growth and uptake of nutrient:\\
+%     Exponential growth where diffusion of the nutritient does not play any role. Sublinear growth where growth is limited by nutritient diffusion.
+%     (Structurally in 2D it's regimes: smooth propagating of bacterial front vs. branching.)
+%     \item \cite{nagarajan_agent-based_2022-1}: review on the ABM for bacterial population.
+%     The ABM represent a microbial colony as a group of discrete agents that are governed by a set of rules, where each agent can be an individual cell or a collection of cells.
+%     Can use citation when talk about general processes we are accounting as growth, uptake of nutritient, production of inhibitor, e.g.,:
+%     Physical interactions between agents are calculated using excluded volume interactions.
+%     Chemical interactions as releasing inhibitors are described using concentrations and solved using PDEs.
+% \end{itemize}
 %
@@
 It has to be noted that the resource pool is normalized by its initial value, ie $R^* = R/R_0$.
-Moreover, the species 1 is able to release a certain antimicrobial component $I$ that shows inhibitory effect on species 2 (amensalism).
+Moreover, the species $A$ is able to release a certain antimicrobial component $I$ that shows inhibitory effect on species 2 (amensalism).
+We assume that $L,G$ are measured in terms of overall volume and $R,I$ are measured as total amount.
 The \acp{ode} the bacterial population can be written as:

### 2023-12-20 pool_model.tex
change description and shown example

@@
         \includegraphics[width=\textwidth]{Figures/abm-inhomogenous/abm_ode_comparison.png}%
-        \subcaption{Inhomogenous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
+        \subcaption{Heterogeneous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}%
     \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures/abm-inhomogenous-low-diffusion/abm_ode_comparison.png}%
-        \subcaption{Inhomogenous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
+        \includegraphics[width=\textwidth]{Figures/abm-random/abm_ode_comparison.png}%
+        \subcaption{Randomized initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}
@@
         The combination of the two effects shows the largest difference between model predictions.
-        In subfigure (d), due to the heterogenous cell distribution (see also~\ref{fig:supplement-spatial-initial-config-cells}) and slow diffusion, overall inhibition of species $B$ does not play as large of a role.
+        In subfigure (c), due to the heterogenous cell distribution (see also~\ref{fig:supplement-spatial-initial-config-cells}), overall inhibition of species $B$ is significantly reduced.
+        Subfigure (d) shows how a randomized homogenous initial distribution can lead to even more notable effects of inhibition.
         We furthermore note that the total bacteria volume of the \ac{abm} model is always smaller than the \ac{ode} model.
@@
         \includegraphics[width=\textwidth]{Figures/abm-inhomogenous/snapshot_00012000.png}%
-        \subcaption{Inhomogenous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
+        \subcaption{Heterogeneous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}%
     \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures/abm-inhomogenous-low-diffusion/snapshot_00012000.png}%
-        \subcaption{Inhomogenous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
+        \includegraphics[width=\textwidth]{Figures/abm-random/snapshot_00012000.png}%
+        \subcaption{Random initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}

### 2023-12-20 pool_model.tex
slightly correct description

@@
         List paramters for both models.
-        Specifying all parameters here would requrie a much larger table.
-        For the full information, we refer to the supplementary material~\ref{sec:supplement-full-parameters}.
+        To view, how these parameters are stored inside a cell-agent, we refer to the supplementary material~\ref{sec:supplement-full-parameters}.
     }

### 2025-08-21 pool_paper_plotting_code/digital_twin/__init__.py
plotting_code:upload python script used for plots generation

@@ -1,0 +1,3 @@
+from digital_twin.model import *
+from digital_twin.data_generation import *
+from digital_twin.plotting_templates import *

### 2025-08-21 pool_paper_plotting_code/digital_twin/data_generation.py

@@ -1,0 +1,62 @@
+#!/usr/bin/env python3
+import numpy as np
+import itertools as iter
+
+from digital_twin.model import model_ODE_solution
+
+
+def generate_insilico_data(model, times, param_arr, prob_arr, x0_arr, x0_prob_arr=None, const=[], obs_func=None,
+                           n_traj=5,  noise=.0, rel_noise=.0, add_name=''):
+    # Convert probabilities and param array to flatten format
+    param_arr_flatten, prob_arr_flatten = [], []
+    if len(param_arr) != 0:
+        for ind in iter.product(*[[i for i in range(np.shape(param_arr)[-1])] for _ in range(np.shape(param_arr)[0])]):  
+            param_arr_flatten.append([param_arr[j][ind[j]] for j in range (len(ind))])
+            prob_arr_flatten.append(prob_arr[*ind])
+    prob_arr_flatten = np.array(prob_arr_flatten)
+    data = []
+    for j in range (n_traj):
+        param_new = sample_parameter_from_distribution(param_arr_flatten, prob_arr_flatten, size=1)
+        if x0_prob_arr is None:
+            x0_new = [x0[0] for x0 in x0_arr]
+        else:
+            x0_new = [sample_parameter_from_distribution(x0, x0_prob, size=1) for x0, x0_prob in zip(x0_arr, x0_prob_arr)]
+        # Generate timepoints if not given:
+        if type(times) == tuple:
+            times_generated = generate_discrete_timepoints(*times)
+        else:
+            times_generated = times[j]
+        model_output = model_ODE_solution(model, times_generated, param_new, x0_new, const)
+        if obs_func is None:
+            observables = model_output + np.random.normal(0.0, rel_noise*model_output+noise, size=np.shape(model_output))
+        else:
+            observables = np.array(obs_func(model_output))
+            observables = observables + np.random.normal(0.0, rel_noise*observables+noise, size=np.shape(observables))
+        data.append({
+            'experiment_name': 'In-silico' + add_name,
+            'times': np.array(times_generated),
+            'obs_mean': observables,
+            'obs_std': np.abs(rel_noise * observables + noise),
+            'obs_whole': [[[tr] for tr in obs] for obs in observables],
+            'model_output': model_output,
+            'const' : const,
+            'other': [param_new, x0_new] # Save parameters and init conditions for generated curve
+            })
+    return data
+
+
[…]

### 2025-08-21 pool_paper_plotting_code/digital_twin/model.py

@@ -1,0 +1,32 @@
+#!/usr/bin/env python3
+
+import numpy as np
+from scipy.integrate import solve_ivp
+
+# Define the pool model ODEs (general case)
+def pool_model(t, x, param, x0, const):
+    (alpha, lambd, n_max, ) = param
+    (x01, ) = x0
+    (x1, ) = x
+    return [alpha * (x1 - x01 * np.exp(- lambd * t)) * (1 - x1 / n_max)]
+
+# Define the pool model ODEs for 2 species model with resource competition between species
+# With the waste production ofspecies 1, with waste-induced death of the species 2 
+def pool_model_2species_resource_waste(t, x, param, x0, const):
+    (L1, L2, G1, G2, R, W1,) = x
+    (alph1, alph2, lambd1, lambd2, chi1, muW2, Nt, ) = const
+    
+    return [
+        - lambd1 * R * L1,
+        - lambd2 * R * L2,
+          lambd1 * R * L1 + alph1 * R * G1,
+          lambd2 * R * L2 + alph2 * R * G2 - muW2/(chi1*Nt) * W1 * G2,
+        - alph1/Nt * R * G1 - alph2/Nt * R * G2,
+          chi1 * G1
+    ]
+
+
+# The model solution for given ODEs
+def model_ODE_solution(model, t, param, x0, const):
+    sol_model = solve_ivp(model, [0, t[-1]], x0, method='LSODA', t_eval=t, args=(param, x0, const), rtol=2e-4)
+    return sol_model.y

### 2025-08-21 pool_paper_plotting_code/digital_twin/plotting_templates.py

@@ -1,0 +1,70 @@
+#!/usr/bin/env python3
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+plt.rcParams["font.family"] = "serif"
+plt.rc('text', usetex=True)
+plt.rcParams['text.latex.preamble'] = r"\usepackage{bm} \usepackage{amsmath}"
+
+# Set some custom colors/scatter markers for plotting (mb add more or remove later)
+colors = [(0 / 255, 102 / 255, 217 / 255),
+          (90 / 255, 170 / 255, 90 / 255),
+          (215 / 255, 48 / 255, 39 / 255)]  
+gray = (128 / 255, 128 / 255, 128 / 255)
+gray2 = (160 / 255, 160 / 255, 160 / 255)
+scatter_marker = ['^', 'o', 'D', "s", "X", ".", "1", "2", "3", "4", "*", "P", "d", "+", "x"]
+
+
+def template_plot_measurements0(ax0, temp, data, c='b'):
+    n_plot = 0
+    for d in data: 
+        if d['const'] == temp:
+            for meas, std in zip(d['obs_mean'], d['obs_std']):
+                ax0.errorbar(d['times'], np.log10(meas), fmt=scatter_marker[n_plot],  yerr=np.abs(0.43*std/meas),
+                             markersize=7, color=c, label="T = {}".format(temp))
+            n_plot += 1
+
+
+def template_fig_for_many_temps(temp_plot, xlabel, ylabel, time_lim=[]):
+    fig, ax = plt.subplots(1, len(temp_plot), figsize=(4.8*len(temp_plot), 5), sharey=True)
+    fig.subplots_adjust(hspace=0.2, wspace=0.15)
+    for i, ax0 in enumerate(ax): 
+        ax0.tick_params(axis='x', which='major', labelsize=13, labelrotation=0)
+        if len(time_lim) == len(temp_plot):
+            ax0.set_xlim(-0.2, time_lim[i])
+        fig, ax0 = set_labels(fig, ax0, xlabel, ylabel)
+    return fig, ax
+
+
+def template_fig_1_temp(temp_plot, xlabel, ylabel, tcr, time_lim=[]):
+    fig, ax = plt.subplots()
+    if len(time_lim) != 0:
+        ax.set_xlim(-0.2, time_lim[0])
+        ticks_val = [int(4*j) for j in range (int(time_lim[0]*0.25)+1)]
[…]

### 2025-08-21 pool_paper_plotting_code/examples/main_paper.py

@@ -1,0 +1,480 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import os, sys
+sys.path.append(os.getcwd())
+from digital_twin import *
+
+
+def pool_model_3pools(t, x, param, x0, const):
+    (lambd, mu1, alpha, mu2, beta, ) = const
+    (L0, G0, D0, ) = x0
+    (L, G, D, ) = x
+        
+    return [
+        - (mu1 + lambd) * L,
+                  lambd * L + (alpha - mu2) * G,
+                    mu1 * L +           mu2 * G - beta * D
+        ]
+
+def pool_model_3pools_resource(t, x, param, x0, const):
+    (lambd, mu1, alpha, mu2, beta, N_t, ) = const
+    (L0, G0, D0, ) = x0
+    (L, G, D, ) = x
+        
+    return [
+        - (mu1 + lambd) * L,
+                  lambd * L + (alpha - alpha * ((L + G + D)/N_t) - mu2) * G,
+                    mu1 * L +                                       mu2 * G - beta * D
+        ]
+
+def pool_model_2pools_resource_lag(t, x, param, x0, const):
+    (lambd, alpha, N_t, ) = const
+    (L0, G0, R0, ) = x0
+    (L, G, R ) = x
+        
+    return [
+        - lambd * R * L,
+          lambd * R * L + alpha * R * G,
+                - (alpha / N_t) * R * G
+        ]
+
+def pool_model_2pools_resource_lag2(t, x, param, x0, const):
+    (lambd, alpha, N_t, ) = const
+    (L0, G0, ) = x0
+    (L, G, ) = x
+    r_term = 1 - (L+G)/N_t
+    return [
+        - lambd * r_term * L,
+          lambd * r_term * L + alpha * r_term * G,
+        ]
+
+def pool_model_dormant(t, x, param, x0, const):
+    pass
+
+def pool_model_tempstress(t, x, param, x0, const):
+    (lambd, mu1, alpha, mu2, beta, N_t, Gamma, delta, shift_cnd,) = const
+    (L0, G0, D0, ) = x0
+    (L, G, D, ) = x
+    
+    gamma = temp_stress(t, Gamma, delta, shift_cnd)
+    return [
+        - (mu1 + lambd) * L +                                             gamma * G,
[…]

### 2025-08-21 pool_paper_plotting_code/examples/pool_paper_spatial.py

@@ -1,0 +1,60 @@
+#!/usr/bin/env python3
+
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy.integrate import solve_ivp
+
+
+def model_solve(model, times, param, x0, const, obs_func=None):
+    model_output = model_ODE_solution(model, times, param, x0, const)
+    if obs_func is None:
+        observables = model_output
+    else:
+        observables = np.array(obs_func(model_output))
+    return times, observables
+
+
+# The model solution for given ODEs
+def model_ODE_solution(model, t, param, x0, const):
+    sol_model = solve_ivp(model, [0, t[-1]], x0, method='LSODA', t_eval=t, args=(param, x0, const), rtol=2e-4)
+    return sol_model.y
+
+def pool_model_spatial_limit(t, x, temp, x0, const):
+    (L1, G1, L2, G2, R, I,) = x
+    (lambd1, lambd2, alph1, alph2, chi, muI, N_t, ) = const
+    res = 1 - (L1+L2+G1+G2)/N_t
+    alpha2_inhib = alph2/(1 + muI*I)
+    return [
+        - lambd1 * R * L1,
+          lambd1 * R * L1 + alph1 * R * G1,
+        - lambd2 * R * L2,
+          lambd2 * R * L2 + alpha2_inhib * R * G2,
+        - (alph1/N_t) * R * G1 - (alpha2_inhib/  N_t) * R * G2,
+          chi * G1
+    ]
+
+def observable_2pool_2species(output):
+    return [output[0]+output[1], # NA
+            output[2]+output[3], # NB
+            output[0]+output[1]+output[2]+output[3]] # NA + NB
+
+if __name__ == "__main__":
+    # 9. Spatial Limitation:
+    Nt = 1e4
+    x0sp = [5., 0., 5., 0., 1., 0.]
+    constsp = [.008, .005, 4., 5., .1, .1, Nt] # (lambd1, lambd2, alph1, alph2, chi, muI, Nt, )
+    times, observabl = model_solve(pool_model_spatial_limit, np.linspace(0, 10, 100), [], x0sp, constsp,
+                             obs_func=observable_2pool_2species) 
+    
+    fig, ax = plt.subplots()
+    labels = ['Species 1', 'Species 2', 'Sp 1 + Sp 2']
+    for j, obs in enumerate(observabl):
[…]

### 2025-08-21 pool_model.tex
update citation for cellular_raza
Restructure: relocate parts with mathematical formalism to the supplement section

@@
 %
-% ================================================================================================
-\section{General Mathematical Formulation}
-Consider the system consisting of $N$ species in $M$ different states.
-In the general representation, the system can be described with the Ordinary Differential Equations (\acp{ode}) using a non-linear function $\mbF$:
-\begin{equation}
-   \dot{\mbx} = \mbF(\mbx, \mbm, \mbp, \mbu),
-   \label{eq:model_ODE}
-\end{equation}
-where  $\mbx \in \mathbb{R}^{N}  \otimes \mathbb{R}^{M}$ is a vector describing bacterial count.
-In matrix form it can be written as
-\begin{equation}
-    \mbx = \begin{pmatrix}
-        x^1_1  & \dots & x^1_N  \\
-        \vdots &       & \vdots \\
-        x^M_1  & \dots & x^M_N  \\
-            \end{pmatrix}
-    \label{eq:model_bact}
-\end{equation}
-with $x_{j}^{i}$ describing the bacterial counts for species $i$ in state $j$.
-$\mbu$ is the vector of different inputs of the model, e.g., temperature, pressure etc.
-The vector $\mbm \in \mathbb{R}^{K}$ includes all variables corresponding to the microenvironment of the system which can be concentration of the inhibitors or activators:
-\begin{equation}
-    \dot{\mbm} = \mbH (\mbx, \mbm, \mbp, \mbu).
-    \label{eq:model_microenv}
-\end{equation}
-Here function $\mbH$ is in general case also non-linear.\\
-%
-The advantage of the following formulation is its generality.
-Depending on the knowledge about the system as well as the needed level of accuracy one can define an arbitrary number of states $M$ and control the complexity of the model.
-Moreover, opposite to the already known Predictive Microbiology like Baranyi and Roberts one can easily change.\\
-%
-To calculate the bacterial count for one species $i$, we simply need to sum the tensor/vector $\mbx$ over all possible states:
-\begin{equation}
-    N_i = \sum_k \mbx_i^k.
-\end{equation}
-While the total bacterial count of all species of our system is
-\begin{equation}
-    N = \sum_i \sum_k \mbx_i^k.
-\end{equation}
-In the following section we demonstrate the application of this approach for different toy models describing the processes within and between bacteria communities.
-%
-%
-%
-%
+
 % ================================================================================================
@@
 In the described simplified version of the real-life system three possible bacteria states are mentioned corresponding to which process is happening: adaptation, growth or death.
-Motivated by this we divide a bacterial population into three pools (states) (~\ref{fig:SchematicRep}):
-\begin{itemize}
-    \item $x^1_1 (t) = L(t)$: the fraction of the population in the lag phase
-    \item $x^2_1 (t) = G(t)$: the fraction of the population growing and dividing
-    \item $x^3_1 (t) = D(t)$: the fraction of the population undergoing cell death
-\end{itemize}
-Here the concentrations of the pools are components of our variable vector $\mbx (t) = (L(t), G(t), D(t))^T$.
+Motivated by this we divide a bacterial population into three pools (states)~(\ref{fig:SchematicRep}) defining the fraction of the population in the lag phase $L(t)$, growing and dividing $G(t)$ and undergoing cell death $D(t)$.
 \begin{figure}[t]
@@
 \end{figure}
-To determine the function $\mbF(\mbx, \mbm, \mbp, \mbu)$ consider the flows between different states.
+To determine the system of equations describing the system dynamics consider the flows between different states.
 The pools $G$ and $D$ are unidirectionally connected with the pool $L$.
@@
     \dot{D} = \mu  L + \mu' G- \beta D  
-\end{cases}
-\end{equation}
-or in matrix form it shows the linear dependency
-\begin{equation}
-    \dot \mbx  = \begin{pmatrix}
-        -(\mu + \lambda) & 0             & 0      \\
-        \lambda          & \alpha - \mu' & 0      \\
-        \mu              & \mu'          & -\beta 
-    \end{pmatrix} \mbx .
-\end{equation}
-%
-%
-%
-% ------------------------------------------------------------------------------------------------
-\subsection{Resource Dependence}
+\end{cases}.
+\label{eq:ode_3pools}
+\end{equation}
+%
 Based on experimental observations, it was shown that the bacterial growth is strongly influenced by the limiting resource.
@@
 In this case the growth term can be simply written as $\alpha G R$ and the model is fully formulated in the following way:
-\begin{eqnarray}
-    \dot \mbx  &=& \mbF(\mbx, \mbm, \mbp, \mbu) = \begin{pmatrix}
-        -(\mu + \lambda) & 0               & 0      \\
-        \lambda          & \alpha R - \mu' & 0      \\
-        \mu              & \mu'            & -\beta 
-    \end{pmatrix} \mbx\\
-    \dot \mbm &=& \mbH (\mbx, \mbm, \mbp, \mbu) = -\frac{\alpha}{N_t} R G.
-\end{eqnarray}
+\begin{equation}
+    \begin{cases}
+        \dot{L} = -(\mu + \lambda) L\\
+        \dot{G} = \lambda L + \alpha R G-\mu' G\\
+        \dot{D} = \mu  L + \mu' G- \beta D\\
+        \dot{R} = - \frac{\alpha}{N_t} R G
+    \end{cases}.
+\label{eq:ode_3pool_resource} 
+\end{equation}
 %
@@
         \dot{D} = \mu  L + \mu' G- \beta D 
-        \label{eq:3pool_resource} 
     \end{cases}.
-\end{equation}
-Rewriting the equations above, we get the following non-linear system:
-\begin{equation}
-    \dot \mbx = \begin{pmatrix}
-        -(\mu + \lambda) & 0       & 0 \\
-         \lambda         & \alpha  & 0 \\
-         \mu &  \mu'& - \beta 
-    \end{pmatrix} 
-    \mbx + \frac{\alpha}{N_t} x^2 \sum_i^3 x^i \begin{pmatrix} 0 \\ 1 \\ 0  \end{pmatrix}.
-\end{equation}
-Here the first term is the linear part and the second term define the quadratic behavior of the system.
-\begin{figure}[t]
+\label{eq:ode_3pool_resource2} 
+\end{equation}
+\begin{figure}[h]
     \begin{center}
@@
     \caption{
-        The time dependence of the total bacterial count $N(t) = \sum_i \sum_k \mbx_i^k = L(t) + G(t) + D(t)$ is obtained as a sum of bacteria in all states.
-        All curves were calculated using the model (\ref{eq:3pool_resource}) with parameters $\alpha=2.5$, $\mu'=0.5$ and $\beta=5$.
+        The time dependence of the total bacterial count $N(t) = L(t) + G(t) + D(t)$ is obtained as a sum of bacteria in all states.
+        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\alpha=2.5$, $\mu'=0.5$ and $\beta=5$.
         The blue curve shows significant dip in behavior of the system where the death of the bacteria is present $\mu = 0.5$.
@@
 %
-\cmt{So now is it better to use flow to dormancy instead?}
-%
 We explore now the idea that under certain conditions, as sudden change in the environment, part of the population $G$ will enter a lag-phase, i.e., there is a back-flow from $G$ to $L$.
@@
 % ................................................................................................
-\subsubsection{Maxwell type of stress - strain relation}
+\paragraph{Maxwell type of stress - strain relation}
+
 One of the reasons of the dormant state appearance can be also the temperature shift~\cite{oliver_viable_1995}.
@@
     \label{eq:model_2sp_resource_comp}
-\end{equation}
-This system of equations defines the microenvironment $\mbm$ and bacteria state matrix $\mbx$ as
-\begin{equation}
-    \mbm = R, \hspace{1cm}
-    \mbx = \begin{pmatrix}
-        L_A & L_B \\
-        G_A & G_B 
-    \end{pmatrix}.
 \end{equation}
@@
         The competition of the two bacterial species due to the common limiting resource calculated using Eq.~(\ref{eq:model_2sp_resource_comp}).
-        The red curve describes the growth of the bacteria $A$ ($N_A = \sum_{k} x_A^k = L_A+G_A$) with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
-        The blue curve describes the growth of the bacteria $B$ ($N_B = \sum_{k} x_B^k = L_B+G_B$) with model parameters $\lambda_B=5\cdot 10^{-2}$, $\alpha_B=2.5$.
-        The purple line shows the total bacterial count of the system  $N(t)=\sum_i\sum_k\mbx_i^k=N_A+N_B$.
+        The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
+        The blue curve describes the growth of the bacteria $B$ ($N_B = L_B+G_B$) with model parameters $\lambda_B=5\cdot 10^{-2}$, $\alpha_B=2.5$.
+        The purple line shows the total bacterial count of the system  $N(t)=N_A+N_B$.
         The gray line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
@@
 That principle claims that two species in competition for the same resource cannot coexist~\cite{wangersky_lotka-volterra_1978}.
-In previous studies it was shown that \ac{glv} equations with one bacterial state can be sucessfully applied to predict the microbial community dynamics, determine different types of the pairwise interspecies interactions~\cite{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
+In previous studies it was shown that \ac{glv} equations with one bacterial state can be successfully applied to predict the microbial community dynamics, determine different types of the pairwise interspecies interactions~\cite{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
 In food studies the \ac{glv} was applied by Mounier \etal~\cite{mounier_microbial_2008} to study interactions of six bacteria species influenced by yeast omissions in cheese.
@@
 However, in contrast to the classical Lotka-Volterra approach, the introduced pool model allows for far more flexibility in tuning properties of the bacterial interactions within and between species.
+For further comparison of the derived model with the matrix formulation of the \ac{glv} we refer the reader to the supplement materials~\ref{ssec:supplement1}.
 %
@@
     \caption{
-        The two species system competing for the common nutritient $R$ with production of the toxin $T_B$ (brown line) calculated using Eq.~(\ref{eq:model_2sp_toxin}).
+        The two species system competing for the common nutrient $R$ with production of the toxin $T_B$ (brown line) calculated using Eq.~(\ref{eq:model_2sp_toxin}).
         The growing bacteria $B$ produces toxin $T_B$ (a) constantly $\dot{T}_B^1 = \mathcal{F}^1 = \kappa G_B$ or (b) only during growth process $\dot{T}_B^2 = \mathcal{F}^2 = \kappa \alpha_B R G_B$ with
@@
         The toxin $T_B$ causes the cell death of bacteria $A$ with rate constant $\mu = 0.2$.
-        The red curve describes the growth of the bacteria $A$ ($N_A = \sum_{k} x_A^k = L_A+G_A$) with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
-        The blue curve describes the growth of the bacteria $B$ ($N_B = \sum_{k} x_B^k = L_B+G_B$) with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
+        The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
+        The blue curve describes the growth of the bacteria $B$ ($N_B = L_B+G_B$) with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
         The gray line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
@@
 \subsection{Constructing an Agent-Based Pool Model}
-We use \texttt{cellular\_raza}~\cite{Pleyer_cellular_raza_2023} to implement an \ac{abm}.
+We use \texttt{cellular\_raza}~\cite{Pleyer2025} to implement an \ac{abm}.
+
 \subsubsection{Cellular Reactions}
@@
 % ================================================================================================
+\section{Discussion}
+% ------------------------------------------------------------------------------------------------
 \section{Conclusion}
@@
 \section*{Supplementary Material}
-\subsection{Deriving the Pool Model}
+% ================================================================================================
+\subsection{General Mathematical Formalism}  \label{ssec:supplement1}
+Consider the system consisting of $N$ species in $M$ different states.
+In the general representation, the system can be described with the Ordinary Differential Equations (\acp{ode}) using a non-linear function $\mbF$:
+\begin{equation}
+   \dot{\mbx} = \mbF(\mbx, \mbm, \mbp, \mbu),
+   \label{eq:model_ODE}
+\end{equation}
+where  $\mbx \in \mathbb{R}^{N}  \otimes \mathbb{R}^{M}$ is a vector describing bacterial count.
+In matrix form it can be written as
+\begin{equation}
+    \mbx = \begin{pmatrix}
+        x^1_1  & \dots & x^1_N  \\
+        \vdots &       & \vdots \\
+        x^M_1  & \dots & x^M_N  \\
+            \end{pmatrix}
+    \label{eq:model_bact}
+\end{equation}
+with $x_{j}^{i}$ describing the bacterial counts for species $i$ in state $j$.
+$\mbu$ is the vector of different inputs of the model, e.g., temperature, pressure etc.
+The vector $\mbm \in \mathbb{R}^{K}$ includes all variables corresponding to the microenvironment of the system which can be concentration of the inhibitors or activators:
+\begin{equation}
+    \dot{\mbm} = \mbH (\mbx, \mbm, \mbp, \mbu).
+    \label{eq:model_microenv}
+\end{equation}
+Here function $\mbH$ is in general case also non-linear.\\
+%
+The advantage of the following formulation is its generality.
+Depending on the knowledge about the system as well as the needed level of accuracy one can define an arbitrary number of states $M$ and control the complexity of the model.
+Moreover, opposite to the already known Predictive Microbiology like Baranyi and Roberts one can easily change.\\
+%
+To calculate the bacterial count for one species $i$, we simply need to sum the tensor/vector $\mbx$ over all possible states:
+\begin{equation}
+    N_i = \sum_k \mbx_i^k.
+\end{equation}
+While the total bacterial count of all species of our system is
+\begin{equation}
+    N = \sum_i \sum_k \mbx_i^k.
+\end{equation}
+%
+Using such matrix representation of the bacterial system, we can rewrite introduced above system in more compact form.
+Here the concentrations of the pools are components of our variable vector $\mbx (t) = (L(t), G(t), D(t))^T$.
+The three pool system described with equations~\ref{eq:ode_3pools} will be written as
+\begin{equation}
+    \dot \mbx  = A \mbx, \hspace{3mm} A = \begin{pmatrix}
+        -(\mu + \lambda) & 0             & 0      \\
+        \lambda          & \alpha - \mu' & 0      \\
+        \mu              & \mu'          & -\beta 
+    \end{pmatrix}.
+\label{eq:x_repres_3pools}
+\end{equation}
+%
+With introduction of the limiting resource the system dynamics presented by the \acp{ode}~\ref{eq:ode_3pool_resource} will be rewritten in the following non-linear system:
+\begin{eqnarray}
+    \dot \mbx  &=& \mbF(\mbx, \mbm, \mbp, \mbu) = \begin{pmatrix}
+        -(\mu + \lambda) & 0               & 0      \\
+        \lambda          & \alpha R - \mu' & 0      \\
+        \mu              & \mu'            & -\beta 
+    \end{pmatrix} \mbx\\
+    \dot \mbm &=& \mbH (\mbx, \mbm, \mbp, \mbu) = -\frac{\alpha}{N_t} R G.
+\label{eq:x_repres_3pools_resource}
+\end{eqnarray}
+\begin{equation}
+    \dot \mbx = \begin{pmatrix}
+        -(\mu + \lambda) & 0       & 0 \\
+         \lambda         & \alpha  & 0 \\
+         \mu &  \mu'& - \beta 
+    \end{pmatrix} 
+    \mbx + \frac{\alpha}{N_t} x^2 \sum_i^3 x^i \begin{pmatrix} 0 \\ 1 \\ 0  \end{pmatrix}.
+\label{eq:x_repres_3pools_resource2}
+\end{equation}
+Here the first term is the linear part and the second term define the quadratic behavior of the system.
+%
+\paragraph{Generalized Lotka-Volterra}
+
+This system of equations defines the microenvironment $\mbm$ and bacteria state matrix $\mbx$ as
+\begin{equation}
+    \mbm = R, \hspace{1cm}
+    \mbx = \begin{pmatrix}
+        L_A & L_B \\
+        G_A & G_B 
+    \end{pmatrix}.
+\end{equation}
+%
+
+\subsection{ABM Details} \label{ssec:supplement2}
+
+\subsubsection*{Deriving the Pool Model}
 \label{subsec:supplement-deriving-pool-odes}
@@
 % ................................................................................................
-\subsection{Code Samples}
+\subsubsection*{Code Samples}
 The code samples show sample code from the simulation that calculate cellular interactions and cell cycle.
@@
 % ................................................................................................
-\subsection{Full List of Parameters}
+\subsubsection*{Full List of Parameters}
 \label{sec:supplement-full-parameters}
@@
 % ................................................................................................
-\subsection{Homogenous and Inhomogenous Initial Cell Distributions}
+\subsubsection*{Homogenous and Inhomogenous Initial Cell Distributions}
 To clarify what we mean by homogenous and inhomogenous initial configurations, we include snapshots of the simulation.

### 2025-08-21 pool_model.tex
mutual inhibition: use I notation for inhibitor instead of T

@@
 \begin{eqnarray}
-    \mathcal{B}(G)&=&\frac{\alpha}{1+KG}. 
-\end{eqnarray}
-This corresponds to the case $T\sim G$, where $T$ is the toxin produced.
-If the toxin accumulates one needs to replace $G$ by $\int G dt$. Keeping it simple yields:
+    \mathcal{B}(G)&=&\frac{\alpha}{1+KI(G)},
+\end{eqnarray}
+where $I$ is the concentration of the inhibiting metabolic byproduct producing during bacterial growth.
 \begin{eqnarray}
@@
 \end{eqnarray}
+Note that if the toxin accumulates, i.e., $\dot I \sim G$,  one needs to replace $G$ by $\int G dt$.
 One can distinguish two limiting cases: {\it i}) $K_{A/B}\ll N_t$ and {\it ii}) $K_{A/B}\gg N_t$.
@@
 Quite often it happens between the species that are close relatives or clonal populations and such process is called kin selection \cite{west_social_2007}.
-Furthermore, Sometimes neighbors can benefit from a certain behavior without expressing it by itself, so-called 'social cheaters'~\cite{rainey_evolution_2003}.
+Furthermore, neighbors can also benefit from a certain behavior without expressing it by itself, so-called 'social cheaters'~\cite{rainey_evolution_2003}.
 %

### 2025-08-21 pool_model.tex
Corresc mistakes

@@
 This phenomenon appears due to adaptation of the microbial community, for example, to the new environmental conditions.
-This so-called preparation to the exponential growth includes synthesys of the needed compounds, initiation of the transciption~\cite{rolfe_lag_2012} and poorly studied physiological and regulatory mechanisms~\cite{monod_growth_1949}.
+This so-called preparation to the exponential growth includes synthesis of the needed compounds, initiation of the transcription~\cite{rolfe_lag_2012} and poorly studied physiological and regulatory mechanisms~\cite{monod_growth_1949}.
 After this, the succeeded in adaptation bacteria enter the exponential growth phase, where the process of cell division is directly happening.
@@
 In this state the cultures accumulate waste products of metabolic activity that leads to cell death of the bacteria.
-At the same time dying cells release the nutritions for survivors to continue growing.
+At the same time dying cells release the nutrients for survivors to continue growing.
 As a result the balance between the growing and dying processes leads to the constant with time bacterial count~\cite{navarro_llorens_stationary_2010}.
@@
 Based on experimental observations, it was shown that the bacterial growth is strongly influenced by the limiting resource.
-According to the works of Jacques Monod, in the system where all needed nutritients are in excess except for one the total growth is linear to the concentration of this limiting nutritient~\cite{monod_growth_1949}.
+According to the works of Jacques Monod, in the system where all needed nutrients are in excess except for one the total growth is linear to the concentration of this limiting nutrient~\cite{monod_growth_1949}.
 The control of the population density with respect to the available nutrition is implemented using cell communication process called quorum sensing.
-Using release of the extracellular signalling molecules (autoinducers) it is possible to synchronize the gene expression of the whole group~\cite{ng_bacterial_2009}.
+Using release of the extracellular signalling molecules (autoinducer molecules) it is possible to synchronize the gene expression of the whole group~\cite{ng_bacterial_2009}.
 We would not dive into the processes of quorum sensing and try to keep it as simple as possible.
@@
 By definition dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\cite{kaprelyants_dormancy_1993}.
-This survival strategy can appear in response to different external stresses, e.g., nutritient deficiency, temperature stress or extreme fluctuations of environment parameters.
+This survival strategy can appear in response to different external stresses, e.g., nutrient deficiency, temperature stress or extreme fluctuations of environment parameters.
 And after "difficult times" bacteria, also called viable but non-culturable (VBNC), can continue cell division~\cite{kell_viability_1998}.
@@
 Here species $A$ has a longer lag-phase than $B$ but the exponential growth of $A$ is stronger than the one that bacteria $B$ experiences.
-As a result, species $A$ was able to suppress species $B$ consuming the larger amount of nutritient and reaching larger bacterial count at the end.
+As a result, species $A$ was able to suppress species $B$ consuming the larger amount of nutrient and reaching larger bacterial count at the end.
 \begin{figure}[H]
@@
 Assuming that the bacterial system is uniformly distributed (sometimes also called 'well-stirred') we can simplify the problem to a system of \acp{ode}.
-However, in reality spatial effects can play a significant role for population survival forcing some species develop competitive mechanisms such as cell adhesion to surfaces in the areas rich with nutritients~\cite{htuson_bacteriasurface_2013}.
+However, in reality spatial effects can play a significant role for population survival forcing some species develop competitive mechanisms such as cell adhesion to surfaces in the areas rich with nutrients~\cite{htuson_bacteriasurface_2013}.
 Another mechanism is the motility of bacteria, e.g., the flagellates bacteria as \textit{E.
@@
 %     (\cite{li_amplified_2008}: Brownian motion gives right motion observed using microscopy).
-%     \item \cite{farrell_mechanically_2013}: 2D modeling using Newtonian dynamics and only mechanical interactions + nutritient diffusion.
+%     \item \cite{farrell_mechanically_2013}: 2D modeling using Newtonian dynamics and only mechanical interactions + nutrient diffusion.
 %     Branching parameter defines diff limits where diffusion dominant and limits growth, and where it's irrelevant.
@@
 % \begin{itemize}
-%     \item \cite{farrell_mechanically_2013}: Showed the influence of mechanic interactions between bacteria in 2D system with bacteria consuming the nutritient.
+%     \item \cite{farrell_mechanically_2013}: Showed the influence of mechanic interactions between bacteria in 2D system with bacteria consuming the nutrient.
 %     The growth of bacterial colonies where non-motile microorganisms replicate and push each other away as they grow.
 %     Found a transition between two different growth regimes, controlled by the balance between growth and uptake of nutrient:\\
-%     Exponential growth where diffusion of the nutritient does not play any role. Sublinear growth where growth is limited by nutritient diffusion.
+%     Exponential growth where diffusion of the nutrient does not play any role. Sublinear growth where growth is limited by nutrient diffusion.
 %     (Structurally in 2D it's regimes: smooth propagating of bacterial front vs. branching.)
@@
 %     The ABM represent a microbial colony as a group of discrete agents that are governed by a set of rules, where each agent can be an individual cell or a collection of cells.
-%     Can use citation when talk about general processes we are accounting as growth, uptake of nutritient, production of inhibitor, e.g.,:
+%     Can use citation when talk about general processes we are accounting as growth, uptake of nutrient, production of inhibitor, e.g.,:
 %     Physical interactions between agents are calculated using excluded volume interactions.
@@
 Both of these species grow using limiting resource pool $R^*$ with different growth constants $\alpha$ and different lengths of lag phase determined by parameter $\lambda$.
-It has to be noted that the resource pool is normalized by its initial value, ie $R^* = R/R_0$.
-Moreover, the species $A$ is able to release a certain antimicrobial component $I$ that shows inhibitory effect on species 2 (amensalism).
+It has to be noted that the resource pool is normalized by its initial value, i. e., $R^* = R/R_0$.
+Moreover, the species $A$ is able to release a certain antimicrobial component $I$ that shows inhibitory effect on species 2 with any harmful effect on itself (amensalism).
 We assume that $L,G$ are measured in terms of overall volume and $R,I$ are measured as total amount.

### 2025-08-21 pool_model.tex
Add labels to the equations

@@
     D &\stackrel{\beta}{\longrightarrow} & \emptyset
+\label{eq:reactions_3pools}
 \end{eqnarray}
@@
     G + R_0  \stackrel{\alpha_0}{\longrightarrow} 2G.
+\label{eq:reactions_resource}
 \end{equation}
@@
     R = 1 - \frac{L+G+D}{N_t}
+\label{eq:resource}
 \end{equation}
@@
     \end{cases}.
+\label{eq:lagtime_deriv1}
 \end{equation}
@@
     \end{cases}
+\label{eq:lagtime_deriv2}
 \end{equation}
@@
     1  &=& 2\lambda e^{\alpha t_L}\frac{1-e^{-(\mu+\lambda+\alpha)t_L}}{\mu+\lambda+\alpha}.
+\label{eq:lagtime_deriv3}
 \end{eqnarray}
@@
     t_L &\approx& -\frac{1}{\alpha}\ln\left(\frac{2\lambda}{\mu+\lambda+\alpha}\right).
+\label{eq:lagtime}
 \end{eqnarray}
@@
     D &\stackrel{\beta}{\longrightarrow} & \emptyset
+\label{eq:reactions_backlag}
 \end{eqnarray}
@@
     \dot{\gamma} &=& \Gamma \left |\dot{T}\right |-\delta \gamma.
+\label{eq:rate_backlag}
 \end{eqnarray}
@@
     \gamma(t) &=& \Gamma \int_0^t \left |\dot{T}(t^{\prime})\right |e^{-\delta (t-t^{\prime})}dt^{\prime}.
+\label{eq:rate_backlag2}
 \end{eqnarray}
@@
     \gamma(t) &=& \Gamma\delta\left |\left[T[t]-T[0]e^{-\delta t}-\delta \int_0^t T(t^{\prime})e^{-\delta (t-t^{\prime})}dt^{\prime}\right]\right |.
+\label{eq:rate_backlag3}
 \end{eqnarray}
@@
     \dot{D} = \mu  L - \beta D
-    \label{eq:ODE_tempshift}
     \end{cases}
+\label{eq:ode_tempshift_backlag}
 \end{equation}
@@
     \caption{
-        (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stress~(\ref{eq:ODE_tempshift}).
+        (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stress~(\ref{eq:ode_tempshift_backlag}).
         (b) The value of the corresponding back-flow rate $\gamma$ calculated according to (\ref{eq:gamma_tempshift}).
@@
     \dot{G}_B &=& \mathcal{B}_B(G_A) G_B\left(1-\frac{G_A+G_B}{N_t}\right).
+\label{eq:ode_inhib1}
 \end{eqnarray}
@@
     \mathcal{B}(G)&=&\frac{\alpha}{1+KI(G)},
+\label{eq:ode_inhib_B}
 \end{eqnarray}
@@
     \dot{G}_B &=& \frac{\alpha_B G_B}{1+K_AG_A}\left(1-\frac{G_A+G_B}{N_t}\right).
+\label{eq:ode_inhib2}
 \end{eqnarray}
@@
 \begin{eqnarray}
-\label{Mut_Inhib}
     \dot{G}_A &=&\tilde{\alpha}_A\frac{G_A}{G_B}\left(1 - G_A-G_B\right)\\
     \dot{G}_B &=& \tilde{\alpha}_B\frac{G_B}{G_A}\left(1-G_A-G_B\right).
+\label{eq:ode_inhib3}
 \end{eqnarray}
@@
     \psi = \tilde{\alpha}_B/\tilde{\alpha}_A = \frac{K_B\alpha_B}{K_A\alpha_A}
+\label{eq:ode_inhib_psi}
 \end{equation}
@@
    \dot{\mbx} = \mbF(\mbx, \mbm, \mbp, \mbu),
-   \label{eq:model_ODE}
+\label{eq:model_ODE_x}
 \end{equation}
@@
     N_i = \sum_k \mbx_i^k.
+\label{eq:N_pro_species}
 \end{equation}
@@
     N = \sum_i \sum_k \mbx_i^k.
+\label{eq:N_total}
 \end{equation}
@@
     \end{pmatrix}.
+\label{eq:gLV_x_m}
 \end{equation}

### 2025-08-22 pool_model.tex
correctly number supplement and list in TOC
formatting

@@
 \newcommand{\beginsupplement}{%
+  \renewcommand{\thesection}{}
   \setcounter{subsection}{0}
@@
 %
-
+%
 % ================================================================================================
@@
 %
-% ................................................................................................
 \paragraph{Maxwell type of stress - strain relation}
-
 One of the reasons of the dormant state appearance can be also the temperature shift~\cite{oliver_viable_1995}.
@@
 
-%
 Equations (\ref{eq:LV_simple}) are basically competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competitive excursion.
@@
 \end{figure}
-%
-%
-%
-%
+
 % ================================================================================================
@@
 \end{equation}
-%
-%
-%
+
 % ------------------------------------------------------------------------------------------------
@@
 \end{table}
-%
-%
+
 % ================================================================================================
@@
 \printbibliography
-%
-%
+
 % ================================================================================================
 \beginsupplement
-\section*{Supplementary Material}
+\section{Supplementary Material}
 % ================================================================================================
@@
 % ................................................................................................
-\subsubsection*{Code Samples}
+\subsubsection{Code Samples}
 The code samples show sample code from the simulation that calculate cellular interactions and cell cycle.
@@
 % ................................................................................................
-\subsubsection*{Full List of Parameters}
+\subsubsection{Full List of Parameters}
 \label{sec:supplement-full-parameters}
@@
 % ................................................................................................
-\subsubsection*{Homogenous and Inhomogenous Initial Cell Distributions}
+\subsubsection{Homogenous and Inhomogenous Initial Cell Distributions}
 To clarify what we mean by homogenous and inhomogenous initial configurations, we include snapshots of the simulation.

### 2025-08-26 pool_model.tex
rename bib file
Supplement1: math formalism, do some rewritings

@@
             ]{biblatex}
-\addbibresource{Bacteria Modelling.bib}
+\addbibresource{references.bib}
 \usepackage{xpatch}
@@
 Consider the system consisting of $N$ species in $M$ different states.
-In the general representation, the system can be described with the Ordinary Differential Equations (\acp{ode}) using a non-linear function $\mbF$:
+In the general representation, all discussed in the main part systems  can be described with the Ordinary Differential Equations (\acp{ode}) using a non-linear function $\mbF$:
 \begin{equation}
@@
 %
-Using such matrix representation of the bacterial system, we can rewrite introduced above system in more compact form.
-Here the concentrations of the pools are components of our variable vector $\mbx (t) = (L(t), G(t), D(t))^T$.
-The three pool system described with equations~\ref{eq:ode_3pools} will be written as
-\begin{equation}
-    \dot \mbx  = A \mbx, \hspace{3mm} A = \begin{pmatrix}
-        -(\mu + \lambda) & 0             & 0      \\
-        \lambda          & \alpha - \mu' & 0      \\
-        \mu              & \mu'          & -\beta 
-    \end{pmatrix}.
-\label{eq:x_repres_3pools}
-\end{equation}
-%
-With introduction of the limiting resource the system dynamics presented by the \acp{ode}~\ref{eq:ode_3pool_resource} will be rewritten in the following non-linear system:
+%
+%The three pool system described with equations~\ref{eq:ode_3pools} will be written as
+%\begin{equation}
+%    \dot \mbx  = A \mbx, \hspace{3mm} A = \begin{pmatrix}
+%        -(\mu + \lambda) & 0             & 0      \\
+%        \lambda          & \alpha - \mu' & 0      \\
+%        \mu              & \mu'          & -\beta 
+%    \end{pmatrix}.
+%\label{eq:x_repres_3pools}
+%\end{equation}
+%
+Using such matrix representation of the bacterial system, we can rewrite, for example, the three-pool model with limiting resource described with equations~\ref{eq:ode_3pool_resource} in a more compact form.
+The concentrations of the pools are then the components of our variable vector $\mbx (t) = (L(t), G(t), D(t))^T$.
+%
 \begin{eqnarray}
@@
 \end{eqnarray}
+Separating linear and non-linear parts in two terms yields
 \begin{equation}
@@
 
-This system of equations defines the microenvironment $\mbm$ and bacteria state matrix $\mbx$ as
+Likewise, the multispecies system with resource competition present described by equations~\ref{eq:model_2sp_resource_comp} can be represented in the same way by defining
 \begin{equation}
@@
 \end{equation}
+
+The simpler version of one state per species (Eq.~\ref{eq:LV_simple}) can be simply written in a general mathematical form as generalized Lotka-Volterra:
+\begin{equation}
+    \frac{d x_i}{d t} = x_i f_i (\mbx), \hspace{3mm} \mathbf{f} = \mathbf{r} + A \mbx.
+\label{eq:gLV}
+\end{equation}
+
 %

### 2025-08-28 pool_model.tex
write introductory part for spatial effects

@@
 \section{Spatial Effects}
-As described above, the proposed model operates with the total number of the bacterial counts for different species and states in the system not accounting for their spatial distributions.
-Assuming that the bacterial system is uniformly distributed (sometimes also called 'well-stirred') we can simplify the problem to a system of \acp{ode}.
-However, in reality spatial effects can play a significant role for population survival forcing some species develop competitive mechanisms such as cell adhesion to surfaces in the areas rich with nutrients~\cite{htuson_bacteriasurface_2013}.
-Another mechanism is the motility of bacteria, e.g., the flagellates bacteria as \textit{E.
-coli} or \textit{B.
-subtilis} can experience the chemotactic motion of driven by oxygen concentrations~\cite{decoene_microscopic_2011}.
-The bacterial social interactions such as releasing the toxins or public goods can also significantly affect the spatial structure of the community~\cite{blanchard_bacterial_2015}.
-These effects can only be treated with a spatial model that is able to describe the observed effects.\\
+As described above, the proposed model operates with the total number of the bacterial counts for
+different species and states in the system, not accounting for their spatial distributions.
+Assuming that the bacterial system is uniformly distributed (sometimes also called 'well-stirred')
+we expect to be able to simplify the problem and map it to a system of \acp{ode}.
+However, in reality spatial effects can play a significant role for population survival forcing some
+species to develop competitive mechanisms such as cell adhesion to surfaces in the areas rich with
+nutritients~\cite{htuson_bacteriasurface_2013}.
+Another mechanism is the motility of bacteria, e.g., the flagellates bacteria as \textit{E. coli} or
+\textit{B. subtilis} can experience the chemotactic motion of driven by oxygen
+concentrations~\cite{decoene_microscopic_2011}.
+The bacterial social interactions such as releasing the toxins or public goods can also
+significantly affect the spatial structure of the community~\cite{blanchard_bacterial_2015}.
+These effects can only be treated with a spatial model that is able to describe the observed
+effects.\\
 %
@@
 %
-To test the spatial limitations of the described above approach consider the system consisting from two bacterial species that can be described by two internal states: lag and growth states. 
-Both of these species grow using limiting resource pool $R^*$ with different growth constants $\alpha$ and different lengths of lag phase determined by parameter $\lambda$.
-It has to be noted that the resource pool is normalized by its initial value, i. e., $R^* = R/R_0$.
-Moreover, the species $A$ is able to release a certain antimicrobial component $I$ that shows inhibitory effect on species 2 with any harmful effect on itself (amensalism).
-We assume that $L,G$ are measured in terms of overall volume and $R,I$ are measured as total amount.
+To test the spatial limitations of the described above approach consider the system consisting from
+two bacterial species that can be described by two internal states: lag and growth states.
+Both of these species grow using the same resource pool $R$ but with different growth constants
+$\alpha$ and different lengths of lag phase $\lambda$.
+Moreover, the species $A$ is able to release a certain antimicrobial component $I$ that shows an
+inhibitory effect on species 2 (amensalism).
+We assume that $L,G$ are measured in terms of the overall volume and $R,I$ are measured as total
+amount.
 The \acp{ode} the bacterial population can be written as:

### 2025-08-28 pool_model.tex
rename variable
write short description of abm formulation

@@
     \begin{cases}
-        \dot{L_A} = -\lambda_A R^* L_A\\
-        \dot{G_A} = \lambda_A R^* L_A + \alpha_A R^* G_A\\
-        \dot{L_B} = -\lambda_B R^* L_B\\
-        \dot{G_B} = \lambda_B R^* L_B + \frac{\alpha_B}{1 + \mu_I I} R^* G_B\\
+        \dot{L_A} = -\lambda_A R L_A\\
+        \dot{G_A} = \lambda_A R L_A + \alpha_A R G_A\\
+        \dot{L_B} = -\lambda_B R L_B\\
+        \dot{G_B} = \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B\\
     \end{cases}.
@@
     \begin{cases}
-        \dot{R^*} = -\frac{\alpha_A}{G_\text{max}} R^* G_A-\frac{\alpha_B}{(1 + \mu_I I) G_\text{max}} R^* G_B \\
+        \dot{R} = -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \mu_I I) N_t} R G_B \\
         \dot{I} = \kappa G_A
@@
 \subsection{Constructing an Agent-Based Pool Model}
-We use \texttt{cellular\_raza}~\cite{Pleyer2025} to implement an \ac{abm}.
+The previously considered \ac{ode} formulation makes extensive use of the assumption that the
+bacteria are evenly distributed and all communications and interactions between bacteria or
+nutrients occurr on time scales much faster than the remaining processes.
+However, for heterogenous systems or nutrients with low diffusability, this assumption might not be
+valid.\\
+In order to construct a model which can encapsulate diffusion of nutrients and heterogeneity of
+the bacteria or nutrients, we have to take into consideration the spatial component of our setup.
+\acp{abm} force researchers to think in terms of cellular processes, thus bridging the gap from a
+collective (bulk) model to an individual-based (single-cell) description.
+The following sections will demonstrate how a given \ac{ode} system can be translated into a
+custom-built \ac{abm} which has only a few more overall parameters.
+To construct our \ac{abm}, we use \texttt{cellular\_raza}~\cite{Pleyer2025} which was designed to
+provide maximal flexibility for researchers in order to construct minimalist models.
+Without the ability to define and control the precise behavior of our cell-agents, the following
+approaches would not be possible.
 
@@
 \begin{align}
-    \dot{L_A} &= -\lambda_A R^* L_A\\
-    \dot{G_A} &= \lambda_A R^* L_A + \alpha_A R^* G_A\\
-    \dot{L_B} &= -\lambda_B R^* L_B\\
-    \dot{G_B} &= \lambda_B R^* L_B + \frac{\alpha_B}{1 + \mu_I I} R^* G_B\\
-    \dot{R^*} &= -\frac{\alpha_A}{G_\text{max}} R^* G_A-\frac{\alpha_B}{(1 + \mu_I I) G_\text{max}} R^* G_B \\
+    \dot{L_A} &= -\lambda_A R L_A\\
+    \dot{G_A} &= \lambda_A R L_A + \alpha_A R G_A\\
+    \dot{L_B} &= -\lambda_B R L_B\\
+    \dot{G_B} &= \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B\\
+    \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \mu_I I) N_t} R G_B \\
     \dot{I} &= \kappa G_A
@@
         \multirow{2}{*}{Resource}
-        & \multirow{2}{*}{$\displaystyle \dot{R^*} = - \frac{\alpha}{(1 + \mu_I I) G_\text{max}} R^* G$}
+        & \multirow{2}{*}{$\displaystyle \dot{R} = - \frac{\alpha}{(1 + \mu_I I) N_t} R G$}
         & \multirow{2}{*}{$\displaystyle \dot{R} = -\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V$}
-        & $\displaystyle G_\text{max} = [R_0] V_D \sigma$\\
+        & $\displaystyle N_t = [R_0] V_D \sigma$\\
 
@@
         Volume
-        & $\displaystyle \dot{G} = \frac{\alpha}{1 + \mu_I I} R^* G$
+        & $\displaystyle \dot{G} = \frac{\alpha}{1 + \mu_I I} R G$
         & $\displaystyle \dot{V} = \frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V$
@@
         & Conversion                & $\sigma$      & $0.1$\unit{\micro\metre\cubed\per\mol}\\
-        Total volume                & $G_\text{max}$& $2.0\cdot 10^5$\unit{\micro\metre\cubed}
+        Total volume                & $N_t$& $2.0\cdot 10^5$\unit{\micro\metre\cubed}
         &&&\\
@@
         & Cell Volume               & $V_{c,B}$     & $7.07$\unit{\micro\metre\cubed}\\
-        Relative Nutrients          & $R^*$         & $1.0$
+        Relative Nutrients          & $R$         & $1.0$
         & Nutrients                 & $[R_0]$       & $1.0$\unit{\mol\per\micro\metre\cubed}\\

### 2025-08-28 pool_model.tex
partially rewrite introductory part again

@@
 Assuming that the bacterial system is uniformly distributed (sometimes also called 'well-stirred')
-we expect to be able to simplify the problem and map it to a system of \acp{ode}.
+we expect the \ac{ode} description to be valid.
 However, in reality spatial effects can play a significant role for population survival forcing some
@@
 effects.\\
+In order to construct a model which can encapsulate diffusion of nutrients and heterogeneity of
+the bacteria or nutrients, we have to take into consideration the spatial component of our setup.
+\acp{abm} force researchers to think in terms of cellular processes, thus bridging the gap from a
+collective (bulk) model to an individual-based (single-cell) description.
+We consider the system consisting from two bacterial species that can be described by two internal
+states: lag and growth states.
+Both of these species grow using the same resource pool $R$ but with different growth constants
+$\alpha$ and different lengths of lag phase $\lambda$.
+Moreover, the species $A$ is able to release a certain antimicrobial component $I$ that shows an
+inhibitory effect on species 2 (amensalism).
+We assume that $L,G$ are measured in terms of the overall volume and $R,I$ are measured as total
+amount.
+The \acp{ode} the bacterial population can be written as:
 %
@@
 %
-To test the spatial limitations of the described above approach consider the system consisting from
-two bacterial species that can be described by two internal states: lag and growth states.
-Both of these species grow using the same resource pool $R$ but with different growth constants
-$\alpha$ and different lengths of lag phase $\lambda$.
-Moreover, the species $A$ is able to release a certain antimicrobial component $I$ that shows an
-inhibitory effect on species 2 (amensalism).
-We assume that $L,G$ are measured in terms of the overall volume and $R,I$ are measured as total
-amount.
-The \acp{ode} the bacterial population can be written as:
 \begin{equation}
@@
 \subsection{Constructing an Agent-Based Pool Model}
-The previously considered \ac{ode} formulation makes extensive use of the assumption that the
-bacteria are evenly distributed and all communications and interactions between bacteria or
-nutrients occurr on time scales much faster than the remaining processes.
-However, for heterogenous systems or nutrients with low diffusability, this assumption might not be
-valid.\\
-In order to construct a model which can encapsulate diffusion of nutrients and heterogeneity of
-the bacteria or nutrients, we have to take into consideration the spatial component of our setup.
-\acp{abm} force researchers to think in terms of cellular processes, thus bridging the gap from a
-collective (bulk) model to an individual-based (single-cell) description.
 The following sections will demonstrate how a given \ac{ode} system can be translated into a

### 2025-08-28 pool_model.tex
make caption nicer
comment out fiture; no need

@@
 %
-\begin{figure}
-    \centering
-    \includegraphics[width=0.6\textwidth]{Figures/interaciton_potential.png}
-    \caption{
-        Overall form of the interaction force between two cells.
-        A value $\sigma<1$ indicates that cells are overlapping which results in a positive force outwards.
-        Conversely, as long as cells are apart $\sigma>=1$, no force is acting.
-    }
-    \label{fig:spatial-interaction-potential}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=0.6\textwidth]{Figures/interaciton_potential.png}
+%     \caption{
+%         Overall form of the interaction force between two cells.
+%         A value $\sigma<1$ indicates that cells are overlapping which results in a positive force outwards.
+%         Conversely, as long as cells are apart $\sigma>=1$, no force is acting.
+%     }
+%     \label{fig:spatial-interaction-potential}
+% \end{figure}
 %
@@
     \caption{
-        List paramters for both models.
-        To view, how these parameters are stored inside a cell-agent, we refer to the supplementary material~\ref{sec:supplement-full-parameters}.
+        Comparison of the parameters for the \ac{ode} model and \ac{abm}.
+        To view, how these parameters are stored inside a cell-agent, we refer to the supplementary
+        material~\ref{sec:supplement-full-parameters}.
     }

### 2025-08-28 pool_model.tex
enhance some figures
add notes for discussion

@@
     \caption{
-        A cell right before and after a division event.
+        (A) A cell right before and (B) after a division event.
         The volume is equally distributed between each new cell and the new positions are picked random but at opposing sides on a circle around the original center of the single cell.
     }
-    \label{fig:spatial-cell-division-example}
     \end{center}
@@
 We compare the most general version and thus omit species $A$ for the moment.
-\begin{table}
+\begin{table}[h]
     \centering
@@
 \section{Discussion}
+
+How well does the pool model do to encapsulate all the different growth models
+\begin{itemize}
+    \item All 6 models can be represented by pool model
+    \item Baranyi-Roberts results can be reproduced, but not identical; requires some work
+    \item what are some general limitations within this framework? how can they be circumvented?
+\end{itemize}
+
+\noindent
+How justified was the construction of the \ac{abm}?
+\begin{itemize}
+    \item Discuss simplified physical interaction model
+    \item Heterogeneity plays a role; as we have shown
+    \item why did we not simply use a PDE?; because we have to think in terms of cellular processes
+        and this ensures that our model is actually realistic
+\end{itemize}
+
+\noindent
+How can we cover more effects of bacteria?
+\begin{itemize}
+    \item More states for metabolic processes
+    \item More states for more species
+    \item Other cooperiation/inhibition effects
+    \item external influences
+\end{itemize}
+
+\noindent
+What are limitations of our approach?
+\begin{itemize}
+    \item spatial stuff; heterogeneity (because \ac{ode})
+    \item larger models require more parameters, harder to estimate
+    \item only simple models have analytical solutions
+\end{itemize}
+
 % ------------------------------------------------------------------------------------------------

### 2025-08-28 pool_model.tex
move table of parameters to supplement

@@
 %
+% ================================================================================================
+\section{Discussion}
+
+How well does the pool model do to encapsulate all the different growth models
+\begin{itemize}
+    \item All 6 models can be represented by pool model
+    \item Baranyi-Roberts results can be reproduced, but not identical; requires some work
+    \item what are some general limitations within this framework? how can they be circumvented?
+\end{itemize}
+
+\noindent
+How justified was the construction of the \ac{abm}?
+\begin{itemize}
+    \item Discuss simplified physical interaction model
+    \item Heterogeneity plays a role; as we have shown
+    \item why did we not simply use a PDE?; because we have to think in terms of cellular processes
+        and this ensures that our model is actually realistic
+\end{itemize}
+
+\noindent
+How can we cover more effects of bacteria?
+\begin{itemize}
+    \item More states for metabolic processes
+    \item More states for more species
+    \item Other cooperiation/inhibition effects
+    \item external influences
+\end{itemize}
+
+\noindent
+What are limitations of our approach?
+\begin{itemize}
+    \item spatial stuff; heterogeneity (because \ac{ode})
+    \item larger models require more parameters, harder to estimate
+    \item only simple models have analytical solutions
+\end{itemize}
+
+% ------------------------------------------------------------------------------------------------
+\section{Conclusion}
+\newpage
+\printbibliography
+
+% ================================================================================================
+\beginsupplement
+\section{Supplementary Material}
+% ================================================================================================
+\subsection{General Mathematical Formalism}  \label{ssec:supplement1}
+Consider the system consisting of $N$ species in $M$ different states.
+In the general representation, all discussed in the main part systems  can be described with the Ordinary Differential Equations (\acp{ode}) using a non-linear function $\mbF$:
+\begin{equation}
+   \dot{\mbx} = \mbF(\mbx, \mbm, \mbp, \mbu),
+\label{eq:model_ODE_x}
+\end{equation}
+where  $\mbx \in \mathbb{R}^{N}  \otimes \mathbb{R}^{M}$ is a vector describing bacterial count.
+In matrix form it can be written as
+\begin{equation}
+    \mbx = \begin{pmatrix}
+        x^1_1  & \dots & x^1_N  \\
+        \vdots &       & \vdots \\
+        x^M_1  & \dots & x^M_N  \\
+            \end{pmatrix}
+    \label{eq:model_bact}
+\end{equation}
+with $x_{j}^{i}$ describing the bacterial counts for species $i$ in state $j$.
+$\mbu$ is the vector of different inputs of the model, e.g., temperature, pressure etc.
+The vector $\mbm \in \mathbb{R}^{K}$ includes all variables corresponding to the microenvironment of the system which can be concentration of the inhibitors or activators:
+\begin{equation}
+    \dot{\mbm} = \mbH (\mbx, \mbm, \mbp, \mbu).
+    \label{eq:model_microenv}
+\end{equation}
+Here function $\mbH$ is in general case also non-linear.\\
+%
+The advantage of the following formulation is its generality.
+Depending on the knowledge about the system as well as the needed level of accuracy one can define an arbitrary number of states $M$ and control the complexity of the model.
+Moreover, opposite to the already known Predictive Microbiology like Baranyi and Roberts one can easily change.\\
+%
+To calculate the bacterial count for one species $i$, we simply need to sum the tensor/vector $\mbx$ over all possible states:
+\begin{equation}
+    N_i = \sum_k \mbx_i^k.
+\label{eq:N_pro_species}
+\end{equation}
+While the total bacterial count of all species of our system is
+\begin{equation}
+    N = \sum_i \sum_k \mbx_i^k.
+\label{eq:N_total}
+\end{equation}
+%
+%
+%The three pool system described with equations~\ref{eq:ode_3pools} will be written as
+%\begin{equation}
+%    \dot \mbx  = A \mbx, \hspace{3mm} A = \begin{pmatrix}
+%        -(\mu + \lambda) & 0             & 0      \\
+%        \lambda          & \alpha - \mu' & 0      \\
+%        \mu              & \mu'          & -\beta 
+%    \end{pmatrix}.
+%\label{eq:x_repres_3pools}
+%\end{equation}
+%
+Using such matrix representation of the bacterial system, we can rewrite, for example, the three-pool model with limiting resource described with equations~\ref{eq:ode_3pool_resource} in a more compact form.
+The concentrations of the pools are then the components of our variable vector $\mbx (t) = (L(t), G(t), D(t))^T$.
+%
+\begin{eqnarray}
+    \dot \mbx  &=& \mbF(\mbx, \mbm, \mbp, \mbu) = \begin{pmatrix}
+        -(\mu + \lambda) & 0               & 0      \\
+        \lambda          & \alpha R - \mu' & 0      \\
+        \mu              & \mu'            & -\beta 
+    \end{pmatrix} \mbx\\
+    \dot \mbm &=& \mbH (\mbx, \mbm, \mbp, \mbu) = -\frac{\alpha}{N_t} R G.
+\label{eq:x_repres_3pools_resource}
+\end{eqnarray}
+Separating linear and non-linear parts in two terms yields
+\begin{equation}
+    \dot \mbx = \begin{pmatrix}
+        -(\mu + \lambda) & 0       & 0 \\
+         \lambda         & \alpha  & 0 \\
+         \mu &  \mu'& - \beta 
+    \end{pmatrix} 
+    \mbx + \frac{\alpha}{N_t} x^2 \sum_i^3 x^i \begin{pmatrix} 0 \\ 1 \\ 0  \end{pmatrix}.
+\label{eq:x_repres_3pools_resource2}
+\end{equation}
+Here the first term is the linear part and the second term define the quadratic behavior of the system.
+%
+\paragraph{Generalized Lotka-Volterra}
+
+Likewise, the multispecies system with resource competition present described by equations~\ref{eq:model_2sp_resource_comp} can be represented in the same way by defining
+\begin{equation}
+    \mbm = R, \hspace{1cm}
+    \mbx = \begin{pmatrix}
+        L_A & L_B \\
+        G_A & G_B 
+    \end{pmatrix}.
+\label{eq:gLV_x_m}
+\end{equation}
+
+The simpler version of one state per species (Eq.~\ref{eq:LV_simple}) can be simply written in a general mathematical form as generalized Lotka-Volterra:
+\begin{equation}
+    \frac{d x_i}{d t} = x_i f_i (\mbx), \hspace{3mm} \mathbf{f} = \mathbf{r} + A \mbx.
+\label{eq:gLV}
+\end{equation}
+
+%
+
+\subsection{ABM Details} \label{ssec:supplement2}
+
+\subsubsection*{Deriving the Pool Model}
+\label{subsec:supplement-deriving-pool-odes}
+Since the \ac{abm} can be considered more fundamental than the \ac{ode} model, we should be able to derive the latter one from the former.
+This can be achieved by starting with equations~\eqref{eq:spatial-full-system-simplified}.
+We integrate the spatial equations over the total domain starting with the resource concentration $[R]$.
+\begin{alignat}{7}
+    \int\limits_D\dot{[R]}\text{dVol} 
+    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c) &&+ D_R\Delta [R]\bigg]\text{dVol}\\
+    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c)\bigg]\text{dVol} &&+ \int\limits_D D_R\Delta [R]\text{dVol}\\
+    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c)\bigg]\text{dVol} &&+ \int\limits_{\partial D} D_R\vec{\nabla} [R]\vec{\text{dA}}\\
+    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c \int\limits_D \frac{1}{V_D} [R] \delta(x-x_c)\text{dVol}\\
+    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c [R]_{x=x_c}\\
+    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c [R]\\
+    &= &&&&-\frac{u}{1+\nu_I [I]V_D V_D^{-1}} V [R]\frac{V_D}{V_D}\\
+    \dot{R} &= &&&&-\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V\\
+\end{alignat}
+We can split up the integral into two parts.
+The second part can then be transformed into an integral over the flux of the resource along the boundary.
+Since we chose Neumann boundary conditions, this contribution vanishes identically.
+Afterwards, the first part can be evaluated by using the delta-distribution.
+Note that we additionally need to consume a $V_D^{-1}$ term since our volume is finite.
+We evaluate the resource concentration $[R]$ at the position of the individual cells and sum over all contributions.
+Now we assume homogeneity in the sense, that we can reduce $[R]$ to only a single variable instead of considering its spatial distribution.
+This means that every contribution does not depend on the position at which it was evaluated, ie. $[R]_{x=x_c}=[R]$.
+Finally, we can perform the sum to gain the total volume of all cells $V=\sum\limits_c V_c$.
+\begin{alignat}{7}
+    \int\limits_D\dot{[I]} &= \int\limits_D \bigg[&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) &&+ D_I\Delta [I]\bigg]\text{dVol}\\
+    &= \int\limits_D&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) \text{dVol} &&+ \int\limits_D D_I\Delta [I]\text{dVol}\\
+    &= \int\limits_D&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) \text{dVol} &&+ \int\limits_{\partial D} D_I\nabla [I]\vec{\text{dA}}\\
+    &= &&\sum\limits_c b V_c\int\limits_D \frac{1}{V_D} &&\delta(x-x_c) \text{dVol}\\
+    \dot{I} &= b V
+\end{alignat}
+Following the same principle as before, we can integrate the inhibitor equation.
+For the equation describing the growth of the cell volume, we simply add up all cells to obtain the total volume of all cells.
+\begin{alignat}{3}
+    \sum\limits_c\dot{V_c}
+    &= \sum\limits_c&&\frac{u}{1+\nu_I [I]} \sigma [R]V_c\\
+    &= &&\frac{u}{1+\nu_I [I]} \sigma [R]V\\
+    \dot{V} &= &&\frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V
+\end{alignat}
+We can thus summarize the equations obtained
+\begin{align}
+    \dot{R} &= -\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V\\
+    \dot{I} &= b V\\
+    \dot{V} &= \frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V.
+\end{align}
+It is clear that the original model is recovered this way.
+To obtain the full equations, we additionally need to consider transition between states of lag-phase and growth phase not depicted here.
+The preceding equations all assume that cells are in growth phase.
+It is however straiforward, to setup these equations.
+%
+%
+% ................................................................................................
+\subsubsection{Code Samples}
+The code samples show sample code from the simulation that calculate cellular interactions and cell cycle.
+Since backend of the simulation is written in Rust, so is the implementation of cellular aspects.
+Additionally bindings are generated such that the full power of the simulation tool can be accessed.
+This significantly improves handling and development time of new models.
+\begin{listing}
+\begin{lstlisting}[language=rust]
+let z = ext_pos - own_pos;
+let r = z.norm();
+let sigma = r / (self.cell_radius + ext_radius);
+if sigma < 1.0 {
+    let q = 0.2;
+    let dir = z.normalize();
+    let modifier = (1.0 + q) / (q + sigma);
+    return self.potential_strength * dir * modifier;
+} else {
+    return Vector2::zero();
+}
+\end{lstlisting}
+\caption{
+    Sample code for calculation of interactions between two bacteria cells.
+}
+\label{listing:spatial-interaction-strength}
+\end{listing}
+%
+%
+\begin{listing}
+\begin{lstlisting}[language=rust]
+// Clone existing cell
+let mut c2 = c1.clone();
+
+let r = c1.interaction.cell_radius;
+
+// Make both cells smaller
+c1.interaction.cell_radius /= std::f64::consts::SQRT_2;
+c2.interaction.cell_radius /= std::f64::consts::SQRT_2;
+
+// Generate cellular splitting direction randomly
+let angle_1 = rng.gen_range(0.0..2.0 * std::f64::consts::PI);
+let dir_vec = nalgebra::Rotation2::new(angle_1) * Vector2::from([1.0, 0.0]);
+
+// Define new positions for cells
+// It is randomly chosen if the old cell is left or right
+let offset = dir_vec * r / std::f64::consts::SQRT_2;
+let old_pos = c1.pos();
+
+// Increase the remaining food in the bacteria by the fraction of volumes
+// from before and after the division. This exactly coincides with factor 2
+// since we half the volume of the cell
+c1.cellular_reactions.intracellular_concentrations[0] *= 2.0;
+c2.cellular_reactions.intracellular_concentrations[0] *= 2.0;
+
+c1.set_pos(&(old_pos + offset));
+c2.set_pos(&(old_pos - offset));
+\end{lstlisting}
+\caption{
+    Sample code for dividing one bacteria cell into two.
+}
+\end{listing}
+%
+%
+% ................................................................................................
+\subsubsection{Full List of Parameters}
+\label{sec:supplement-full-parameters}
+
 \begin{table}[H]
@@
         Comparison of the parameters for the \ac{ode} model and \ac{abm}.
-        To view, how these parameters are stored inside a cell-agent, we refer to the supplementary
-        material~\ref{sec:supplement-full-parameters}.
+        % To view, how these parameters are stored inside a cell-agent, we refer to the supplementary
+        % material~\ref{sec:supplement-full-parameters}.
     }
@@
 
-% ================================================================================================
-\section{Discussion}
-
-How well does the pool model do to encapsulate all the different growth models
-\begin{itemize}
-    \item All 6 models can be represented by pool model
-    \item Baranyi-Roberts results can be reproduced, but not identical; requires some work
-    \item what are some general limitations within this framework? how can they be circumvented?
-\end{itemize}
-
-\noindent
-How justified was the construction of the \ac{abm}?
-\begin{itemize}
-    \item Discuss simplified physical interaction model
-    \item Heterogeneity plays a role; as we have shown
-    \item why did we not simply use a PDE?; because we have to think in terms of cellular processes
-        and this ensures that our model is actually realistic
-\end{itemize}
-
-\noindent
-How can we cover more effects of bacteria?
-\begin{itemize}
-    \item More states for metabolic processes
-    \item More states for more species
-    \item Other cooperiation/inhibition effects
-    \item external influences
-\end{itemize}
-
-\noindent
-What are limitations of our approach?
-\begin{itemize}
-    \item spatial stuff; heterogeneity (because \ac{ode})
-    \item larger models require more parameters, harder to estimate
-    \item only simple models have analytical solutions
-\end{itemize}
-
-% ------------------------------------------------------------------------------------------------
-\section{Conclusion}
-\newpage
-\printbibliography
-
-% ================================================================================================
-\beginsupplement
-\section{Supplementary Material}
-% ================================================================================================
-\subsection{General Mathematical Formalism}  \label{ssec:supplement1}
-Consider the system consisting of $N$ species in $M$ different states.
-In the general representation, all discussed in the main part systems  can be described with the Ordinary Differential Equations (\acp{ode}) using a non-linear function $\mbF$:
-\begin{equation}
-   \dot{\mbx} = \mbF(\mbx, \mbm, \mbp, \mbu),
-\label{eq:model_ODE_x}
-\end{equation}
-where  $\mbx \in \mathbb{R}^{N}  \otimes \mathbb{R}^{M}$ is a vector describing bacterial count.
-In matrix form it can be written as
-\begin{equation}
-    \mbx = \begin{pmatrix}
-        x^1_1  & \dots & x^1_N  \\
-        \vdots &       & \vdots \\
-        x^M_1  & \dots & x^M_N  \\
-            \end{pmatrix}
-    \label{eq:model_bact}
-\end{equation}
-with $x_{j}^{i}$ describing the bacterial counts for species $i$ in state $j$.
-$\mbu$ is the vector of different inputs of the model, e.g., temperature, pressure etc.
-The vector $\mbm \in \mathbb{R}^{K}$ includes all variables corresponding to the microenvironment of the system which can be concentration of the inhibitors or activators:
-\begin{equation}
-    \dot{\mbm} = \mbH (\mbx, \mbm, \mbp, \mbu).
-    \label{eq:model_microenv}
-\end{equation}
-Here function $\mbH$ is in general case also non-linear.\\
-%
-The advantage of the following formulation is its generality.
-Depending on the knowledge about the system as well as the needed level of accuracy one can define an arbitrary number of states $M$ and control the complexity of the model.
-Moreover, opposite to the already known Predictive Microbiology like Baranyi and Roberts one can easily change.\\
-%
-To calculate the bacterial count for one species $i$, we simply need to sum the tensor/vector $\mbx$ over all possible states:
-\begin{equation}
-    N_i = \sum_k \mbx_i^k.
-\label{eq:N_pro_species}
-\end{equation}
-While the total bacterial count of all species of our system is
-\begin{equation}
-    N = \sum_i \sum_k \mbx_i^k.
-\label{eq:N_total}
-\end{equation}
-%
-%
-%The three pool system described with equations~\ref{eq:ode_3pools} will be written as
-%\begin{equation}
-%    \dot \mbx  = A \mbx, \hspace{3mm} A = \begin{pmatrix}
-%        -(\mu + \lambda) & 0             & 0      \\
-%        \lambda          & \alpha - \mu' & 0      \\
-%        \mu              & \mu'          & -\beta 
-%    \end{pmatrix}.
-%\label{eq:x_repres_3pools}
-%\end{equation}
-%
-Using such matrix representation of the bacterial system, we can rewrite, for example, the three-pool model with limiting resource described with equations~\ref{eq:ode_3pool_resource} in a more compact form.
-The concentrations of the pools are then the components of our variable vector $\mbx (t) = (L(t), G(t), D(t))^T$.
-%
-\begin{eqnarray}
-    \dot \mbx  &=& \mbF(\mbx, \mbm, \mbp, \mbu) = \begin{pmatrix}
-        -(\mu + \lambda) & 0               & 0      \\
-        \lambda          & \alpha R - \mu' & 0      \\
-        \mu              & \mu'            & -\beta 
-    \end{pmatrix} \mbx\\
-    \dot \mbm &=& \mbH (\mbx, \mbm, \mbp, \mbu) = -\frac{\alpha}{N_t} R G.
-\label{eq:x_repres_3pools_resource}
-\end{eqnarray}
-Separating linear and non-linear parts in two terms yields
-\begin{equation}
-    \dot \mbx = \begin{pmatrix}
-        -(\mu + \lambda) & 0       & 0 \\
-         \lambda         & \alpha  & 0 \\
-         \mu &  \mu'& - \beta 
-    \end{pmatrix} 
-    \mbx + \frac{\alpha}{N_t} x^2 \sum_i^3 x^i \begin{pmatrix} 0 \\ 1 \\ 0  \end{pmatrix}.
-\label{eq:x_repres_3pools_resource2}
-\end{equation}
-Here the first term is the linear part and the second term define the quadratic behavior of the system.
-%
-\paragraph{Generalized Lotka-Volterra}
-
-Likewise, the multispecies system with resource competition present described by equations~\ref{eq:model_2sp_resource_comp} can be represented in the same way by defining
-\begin{equation}
-    \mbm = R, \hspace{1cm}
-    \mbx = \begin{pmatrix}
-        L_A & L_B \\
-        G_A & G_B 
-    \end{pmatrix}.
-\label{eq:gLV_x_m}
-\end{equation}
-
-The simpler version of one state per species (Eq.~\ref{eq:LV_simple}) can be simply written in a general mathematical form as generalized Lotka-Volterra:
-\begin{equation}
-    \frac{d x_i}{d t} = x_i f_i (\mbx), \hspace{3mm} \mathbf{f} = \mathbf{r} + A \mbx.
-\label{eq:gLV}
-\end{equation}
-
-%
-
-\subsection{ABM Details} \label{ssec:supplement2}
-
-\subsubsection*{Deriving the Pool Model}
-\label{subsec:supplement-deriving-pool-odes}
-Since the \ac{abm} can be considered more fundamental than the \ac{ode} model, we should be able to derive the latter one from the former.
-This can be achieved by starting with equations~\eqref{eq:spatial-full-system-simplified}.
-We integrate the spatial equations over the total domain starting with the resource concentration $[R]$.
-\begin{alignat}{7}
-    \int\limits_D\dot{[R]}\text{dVol} 
-    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c) &&+ D_R\Delta [R]\bigg]\text{dVol}\\
-    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c)\bigg]\text{dVol} &&+ \int\limits_D D_R\Delta [R]\text{dVol}\\
-    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c)\bigg]\text{dVol} &&+ \int\limits_{\partial D} D_R\vec{\nabla} [R]\vec{\text{dA}}\\
-    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c \int\limits_D \frac{1}{V_D} [R] \delta(x-x_c)\text{dVol}\\
-    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c [R]_{x=x_c}\\
-    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c [R]\\
-    &= &&&&-\frac{u}{1+\nu_I [I]V_D V_D^{-1}} V [R]\frac{V_D}{V_D}\\
-    \dot{R} &= &&&&-\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V\\
-\end{alignat}
-We can split up the integral into two parts.
-The second part can then be transformed into an integral over the flux of the resource along the boundary.
-Since we chose Neumann boundary conditions, this contribution vanishes identically.
-Afterwards, the first part can be evaluated by using the delta-distribution.
-Note that we additionally need to consume a $V_D^{-1}$ term since our volume is finite.
-We evaluate the resource concentration $[R]$ at the position of the individual cells and sum over all contributions.
-Now we assume homogeneity in the sense, that we can reduce $[R]$ to only a single variable instead of considering its spatial distribution.
-This means that every contribution does not depend on the position at which it was evaluated, ie. $[R]_{x=x_c}=[R]$.
-Finally, we can perform the sum to gain the total volume of all cells $V=\sum\limits_c V_c$.
-\begin{alignat}{7}
-    \int\limits_D\dot{[I]} &= \int\limits_D \bigg[&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) &&+ D_I\Delta [I]\bigg]\text{dVol}\\
-    &= \int\limits_D&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) \text{dVol} &&+ \int\limits_D D_I\Delta [I]\text{dVol}\\
-    &= \int\limits_D&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) \text{dVol} &&+ \int\limits_{\partial D} D_I\nabla [I]\vec{\text{dA}}\\
-    &= &&\sum\limits_c b V_c\int\limits_D \frac{1}{V_D} &&\delta(x-x_c) \text{dVol}\\
-    \dot{I} &= b V
-\end{alignat}
-Following the same principle as before, we can integrate the inhibitor equation.
-For the equation describing the growth of the cell volume, we simply add up all cells to obtain the total volume of all cells.
-\begin{alignat}{3}
-    \sum\limits_c\dot{V_c}
-    &= \sum\limits_c&&\frac{u}{1+\nu_I [I]} \sigma [R]V_c\\
-    &= &&\frac{u}{1+\nu_I [I]} \sigma [R]V\\
-    \dot{V} &= &&\frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V
-\end{alignat}
-We can thus summarize the equations obtained
-\begin{align}
-    \dot{R} &= -\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V\\
-    \dot{I} &= b V\\
-    \dot{V} &= \frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V.
-\end{align}
-It is clear that the original model is recovered this way.
-To obtain the full equations, we additionally need to consider transition between states of lag-phase and growth phase not depicted here.
-The preceding equations all assume that cells are in growth phase.
-It is however straiforward, to setup these equations.
-%
-%
-% ................................................................................................
-\subsubsection{Code Samples}
-The code samples show sample code from the simulation that calculate cellular interactions and cell cycle.
-Since backend of the simulation is written in Rust, so is the implementation of cellular aspects.
-Additionally bindings are generated such that the full power of the simulation tool can be accessed.
-This significantly improves handling and development time of new models.
-\begin{listing}
-\begin{lstlisting}[language=rust]
-let z = ext_pos - own_pos;
-let r = z.norm();
-let sigma = r / (self.cell_radius + ext_radius);
-if sigma < 1.0 {
-    let q = 0.2;
-    let dir = z.normalize();
-    let modifier = (1.0 + q) / (q + sigma);
-    return self.potential_strength * dir * modifier;
-} else {
-    return Vector2::zero();
-}
-\end{lstlisting}
-\caption{
-    Sample code for calculation of interactions between two bacteria cells.
-}
-\label{listing:spatial-interaction-strength}
-\end{listing}
-%
-%
-\begin{listing}
-\begin{lstlisting}[language=rust]
-// Clone existing cell
-let mut c2 = c1.clone();
-
-let r = c1.interaction.cell_radius;
-
-// Make both cells smaller
-c1.interaction.cell_radius /= std::f64::consts::SQRT_2;
-c2.interaction.cell_radius /= std::f64::consts::SQRT_2;
-
-// Generate cellular splitting direction randomly
-let angle_1 = rng.gen_range(0.0..2.0 * std::f64::consts::PI);
-let dir_vec = nalgebra::Rotation2::new(angle_1) * Vector2::from([1.0, 0.0]);
-
-// Define new positions for cells
-// It is randomly chosen if the old cell is left or right
-let offset = dir_vec * r / std::f64::consts::SQRT_2;
-let old_pos = c1.pos();
-
-// Increase the remaining food in the bacteria by the fraction of volumes
-// from before and after the division. This exactly coincides with factor 2
-// since we half the volume of the cell
-c1.cellular_reactions.intracellular_concentrations[0] *= 2.0;
-c2.cellular_reactions.intracellular_concentrations[0] *= 2.0;
-
-c1.set_pos(&(old_pos + offset));
-c2.set_pos(&(old_pos - offset));
-\end{lstlisting}
-\caption{
-    Sample code for dividing one bacteria cell into two.
-}
-\end{listing}
-%
-%
-% ................................................................................................
-\subsubsection{Full List of Parameters}
-\label{sec:supplement-full-parameters}
 \begin{listing}

### 2025-08-28 pool_model.tex
add more notes

@@
     \item what are some general limitations within this framework? how can they be circumvented?
+    \item why is the pool model "better" than the other \ac{ode} models? (why did we do this?);
+        generalization of many models; better explainability;
+    \item link to supplement; we can generalize this and prove theorems for the abstract,
+        generalized model; this allows us to discuss many cases at once
 \end{itemize}
@@
 \section{Conclusion}
+\noindent
+\begin{itemize}
+    \item New theoretical framework
+    \item consistent with \ac{ode} models (Baranyi-Roberts)
+    \item Better explainability
+    \item Include effects such as inhibition, toxins in understandable manner
+    \item showed spatial limitations of \ac{ode} model
+    \item \ac{abm} model allowed modeling of cellular processes; ensure that pool model is derivable
+        from the individual behvaiour of cellular processes
+\end{itemize}
+
 \newpage
@@
 \end{equation}
+
+\textbf{!Christian extend this!}
 

### 2025-08-28 pool_model.tex
more outlook comment
format introduction

@@
 Motivation: general, flexible approach for bacterial systems that allows easily modify the system depending on processes on different scales.
-Allows us to describe the processes within populations using different bacterial states as well as interactions between different populations using ODEs.\\
-%
-%
+Allows us to describe the processes within populations using different bacterial states as well as interactions between different populations using ODEs.
+
+\paragraph{Bacteria Consortia and processes (growth, lag phase, inhibition, etc.)}
+\begin{itemize}
+    \item 
+\end{itemize}
+
+\noindent
+\paragraph{ODE models (predictive microbiology, existing, overview, most popular)}
 Existing models in Predictive Microbiology (Primary Models)~\cite{perez-rodriguez_predictive_2012}:
@@
 Comprehensive reviews about competition/cooperation bacterial mechanisms: \cite{ghoul_ecology_2016, stubbendieck_bacterial_2016, hibbing_bacterial_2010, west_social_2007}\\
+
+\noindent
+\paragraph{Spatial effects in cosortia (adhesion, migration, chemotaxis)}
+\begin{itemize}
+    \item 
+\end{itemize}
+
+\noindent
+\paragraph{Models for spatial effects; PDEs, \acp{abm}, others}
 %
@@
         from the individual behvaiour of cellular processes
+    \item outlook: generalized mathematical formulation; extend inhibition model (because it is a
+        dimensional reduction of existing models)
 \end{itemize}

### 2025-08-28 pool_model.tex
reorder package imports

@@
 \documentclass[10pt,A4paper]{article}
+\usepackage[utf8]{inputenc}
+
+\setlength{\tabcolsep}{1.3pt}
+\usepackage{acronym}
+\usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage[
-    a4paper,
-    inner=2.5cm,
-    outer=3cm
-]{geometry}
-
-\usepackage[utf8]{inputenc}
-\usepackage[a4paper,
-            inner=2.5cm,
-            outer=3cm
-            ]{geometry}
-\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off, font=small]{caption}            
-\usepackage{acronym}
-\usepackage{graphicx}% Include figure files
-\usepackage{bm}% bold math
-\usepackage{color}
-\usepackage{hyperref}
-\usepackage[backend=biber,
+    backend=biber,
     sorting=none,
-%            citestyle=numeric-comp,
-%\            bibstyle=ieee, 
     style=numeric-comp,
@@
 ]{biblatex}
-\addbibresource{references.bib}
-\usepackage{xpatch}
-\xpatchbibmacro{journal+issuetitle}{%
-\usebibmacro{journal}%
-\setunit*{\addspace}%
-}{%
-\usebibmacro{journal}%
-\setunit*{\addcomma\space}%
-}{}{}
-
-\renewbibmacro*{name:andothers}{% Based on name:andothers from biblatex.def
-  \ifboolexpr{
-    test {\ifnumequal{\value{listcount}}{\value{liststop}}}
-    and
-    test \ifmorenames
-  }
-    {\ifnumgreater{\value{liststop}}{1}
-       {\finalandcomma}
-       {}%
-     \andothersdelim\bibstring[\emph]{andothers}}
-    {}}
-
-%\DeclareFieldFormat[article]{title}{#1}
-\DeclareFieldFormat[article]{volume}{\textbf{#1}}
-%\DeclareFieldFormat[book]{volume}{\textbf{#1}}
-
-\renewcommand\newunitpunct{\addcomma\space}
-\DefineBibliographyStrings{english}{%
-page = {},
-pages = {},
-}
-
-\renewbibmacro{in:}{\addcomma\addspace}
-
-\renewbibmacro*{volume+number+eid}{%
-\isdot
-\printfield{volume}%
-\setunit*{\addcomma\addspace}%
-\iffieldundef{number}{}{%
-\printtext{no\adddot\addnbthinspace}%
-}%
-\printfield{number}%
-\setunit{\addcomma\space}%
-\printfield{eid}}
-
-\setlength{\tabcolsep}{1.3pt}
+\usepackage{array}
 \usepackage{bm}
-\usepackage{amsmath}
-\usepackage{amssymb}
+\usepackage{booktabs}
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off, font=small]{caption}
+\usepackage{color}
+\usepackage{ctable} % for \specialrule command
+\usepackage{diagbox}
+\usepackage{float}
+\usepackage[T1]{fontenc}
+\usepackage[a4paper,inner=2.5cm,outer=3cm]{geometry}
+\usepackage{graphicx}% Include figure files
+\usepackage{here}
+\usepackage{hyperref}
+\usepackage{lipsum}
+\usepackage{listings}
+\usepackage{listings-rust}
 \usepackage{mathtools}
 \usepackage{mathrsfs}
-\usepackage[T1]{fontenc}
-\usepackage{diagbox}
-\usepackage{here}
+\usepackage{multirow}
+\usepackage{newfloat}
+\usepackage{siunitx}
+\usepackage{subcaption}
+\usepackage{tabularx}
 \usepackage[dvipsnames]{xcolor}
-\usepackage{lipsum}
-\usepackage{multirow}
-\usepackage{ctable} % for \specialrule command
-\usepackage{array,booktabs}
+\usepackage{xpatch}
+
+% Define new column type with multirow
 \newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
-\usepackage{subcaption}
-\usepackage{float}
-\usepackage{tabularx}
-\usepackage{siunitx}
-\usepackage{listings}
-\usepackage{listings-rust}
+
+% Add bibliography
+\addbibresource{references.bib}
+
+% Define Rust lisintgs
 \lstset{
@@
 }
-\usepackage{newfloat}
 \DeclareFloatingEnvironment[
@@
 ]{listing}
+
 \restylefloat{table}
+
+% Define equation numbering
 \numberwithin{equation}{section}

### 2025-08-28 pool_model.tex
adapt to twocolumn layout

@@
 
+\twocolumn
+
 \newpage
@@
     \begin{center}
-    \includegraphics[width=0.6\textwidth]{Figures/TPM_fig.pdf}
+    \includegraphics[width=\columnwidth]{Figures/TPM_fig.pdf}
     \caption{Schematic representation of exchange between three states or pools.}
@@
     D &\stackrel{\beta}{\longrightarrow} & \emptyset
-\label{eq:reactions_3pools}
 \end{eqnarray}
 Then the \acp{ode} describing such a model can be written as
-\begin{equation}
-\begin{cases}
-    \dot{L} = -(\mu + \lambda) L\\
-    \dot{G} = \lambda L + \alpha G - \mu' G\\
-    \dot{D} = \mu  L + \mu' G- \beta D  
-\end{cases}.
-\label{eq:ode_3pools}
-\end{equation}
+\begin{align}\begin{split}
+    \dot{L} &= -(\mu + \lambda) L\\
+    \dot{G} &= \lambda L + \alpha G - \mu' G\\
+    \dot{D} &= \mu  L + \mu' G- \beta D  
+\end{split}\end{align}.
 %
@@
 In this case the growth term can be simply written as $\alpha G R$ and the model is fully formulated in the following way:
-\begin{equation}
-    \begin{cases}
-        \dot{L} = -(\mu + \lambda) L\\
-        \dot{G} = \lambda L + \alpha R G-\mu' G\\
-        \dot{D} = \mu  L + \mu' G- \beta D\\
-        \dot{R} = - \frac{\alpha}{N_t} R G
-    \end{cases}.
+\begin{align}
+    \begin{split}
+        \dot{L} &= -(\mu + \lambda) L\\
+        \dot{G} &= \lambda L + \alpha R G-\mu' G\\
+        \dot{D} &= \mu  L + \mu' G- \beta D\\
+        \dot{R} &= - \frac{\alpha}{N_t} R G
+    \end{split}.
 \label{eq:ode_3pool_resource} 
-\end{equation}
+\end{align}
 %
@@
     R = 1 - \frac{L+G+D}{N_t}
-\label{eq:resource}
 \end{equation}
 and, hence, the full system can be written as
-\begin{equation}
-    \begin{cases}
-        \dot{L} = -(\mu + \lambda) L\\
-        \dot{G} = \lambda L + \alpha G\left(1-\frac{L+G+D}{N_t}\right)-\mu' G\\
-        \dot{D} = \mu  L + \mu' G- \beta D 
-    \end{cases}.
+\begin{align}
+    \begin{split}
+        \dot{L} &= -(\mu + \lambda) L\\
+        \dot{G} &= \lambda L + \alpha G\left(1-\frac{L+G+D}{N_t}\right)-\mu' G\\
+        \dot{D} &= \mu  L + \mu' G- \beta D 
+    \end{split}.
 \label{eq:ode_3pool_resource2} 
-\end{equation}
+\end{align}
 \begin{figure}[h]
     \begin{center}
-    \includegraphics[width=0.6\textwidth]{Figures/pool_model_3pools_resource.pdf}
+    \includegraphics[width=\columnwidth]{Figures/pool_model_3pools_resource.pdf}
     \caption{
@@
 For $\mu'=0$ the equations for $L$ and $D$ can be readily integrated resulting in:
-\begin{equation}
-    \begin{cases}
-        L(t) = N_0 e^{-(\mu+\lambda)t}\\
-        D(t) = N_0 \mu e^{-\beta t}\frac{1-e^{-(\mu+\lambda-\beta)t}}{\mu+\lambda-\beta}
-    \end{cases}.
-\label{eq:lagtime_deriv1}
-\end{equation}
+\begin{align}
+    \begin{split}
+        L(t) &= N_0 e^{-(\mu+\lambda)t}\\
+        D(t) &= N_0 \mu e^{-\beta t}\frac{1-e^{-(\mu+\lambda-\beta)t}}{\mu+\lambda-\beta}
+    \end{split}.
+\end{align}
 For constant parameters $\mu$, $\lambda$, and $\beta$ the integrals for $L$ and $D$ can be put into the differential equation for $G$.
@@
 It is interesting to see what type of equations result for the total bacterial population $N=L+G+D$:
-\begin{equation}
-    \begin{cases}
-        \dot{G} = \lambda L + \alpha G\left(1-\frac{N}{N_t}\right)\\
-        \dot{N} = - \beta D + \alpha G\left(1-\frac{N}{N_t}\right)
-    \end{cases}
-\label{eq:lagtime_deriv2}
-\end{equation}
+\begin{align}
+    \begin{split}
+        \dot{G} &= \lambda L + \alpha G\left(1-\frac{N}{N_t}\right)\\
+        \dot{N} &= - \beta D + \alpha G\left(1-\frac{N}{N_t}\right)
+    \end{split}
+\end{align}
 with the initial conditions $G(0)=0$ and $N(0)=N_0$. 
@@
     1  &=& 2\lambda e^{\alpha t_L}\frac{1-e^{-(\mu+\lambda+\alpha)t_L}}{\mu+\lambda+\alpha}.
-\label{eq:lagtime_deriv3}
 \end{eqnarray}
@@
     t_L &\approx& -\frac{1}{\alpha}\ln\left(\frac{2\lambda}{\mu+\lambda+\alpha}\right).
-\label{eq:lagtime}
 \end{eqnarray}
@@
 The ordinary differential equations for the pool model with environmental-shock based back-flow read:
-\begin{equation}
-    \begin{cases}
+\begin{align}
+    \begin{split}
     \dot{L} =\gamma(t) G - (\mu + \lambda) L\\
@@
     \dot{D} = \mu  L - \beta D
-    \end{cases}
+    \end{split}
 \label{eq:ode_tempshift_backlag}
-\end{equation}
+\end{align}
 \begin{figure}[t]
     \begin{center}
-    \includegraphics[width=0.6\textwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}
+    \includegraphics[width=\columnwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}
     \caption{
@@
 %
-%
-%
 % ================================================================================================
@@
 %
-%
-%
 % ------------------------------------------------------------------------------------------------
@@
 The simplest pool model for bacterial species $A$ and $B$ encompassing lag-phases, growth-phases and a common resource pool is given by (no back-flow to the lag-pool):
-\begin{equation}
-    \begin{cases}
+\begin{align}
+    \begin{split}
         \dot{L}_A = - \lambda_A R L_A\\
@@
         \dot{R} =-\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B
-    \end{cases}
+    \end{split}
     \label{eq:model_2sp_resource_comp}
-\end{equation}
+\end{align}
 We again rescaled the resource pool with $N_t$ and find
@@
 \begin{eqnarray*}
-    \dot{L}_A &=& - \lambda_A  L_A + \frac{\lambda_A}{N_t}\left(L_A+G_A+L_B+G_B\right)L_A\\
-    \dot{L}_B &=& - \lambda_B L_B + \frac{\lambda_B}{N_t}\left(L_A+G_A+L_B+G_B\right)L_B \\
-    \dot{G}_A &=&  \lambda_A  L_A - \frac{\lambda_A}{N_t}\left(L_A+G_A+L_B+G_B\right)L_A +\alpha_A G_A - \frac{\alpha_A}{N_t}\left(L_A+G_A+L_B+G_B\right)G_A\\
-    \dot{G}_B &=&  \lambda_B L_B - \frac{\lambda_B}{N_t}\left(L_A+G_A+L_B+G_B\right)L_B  +\alpha_B G_B -\frac{\alpha_B}{N_t}\left(L_A+G_A+L_B+G_B\right)G_B
+    \dot{L}_A &=& - \lambda_A  L_A + \frac{\lambda_A}{N_t}L_A\left<LG\right>\\
+    \dot{L}_B &=& - \lambda_B L_B + \frac{\lambda_B}{N_t}L_B \left<LG\right>\\
+    \dot{G}_A &=&  \lambda_A  L_A + \alpha_A G_A - \frac{\lambda_A L_A + \alpha_A G_A}{N_t} \left<LG\right>\\
+    \dot{G}_B &=&  \lambda_B L_B - \alpha_B G_B \frac{\lambda_B L_B + \alpha_B G_B}{N_t}\left<LG\right>
 \end{eqnarray*}
-which has the structure of a \ac{glv} expanded with different internal states for one species.
+where we used the short notation $\left<LG\right>=(L_A+G_A+L_B+G_B)$.
+These equations have the structure of a \ac{glv} expanded with different internal states for one species.
 As one can see in Figure~\ref{fig:2pool_resource_2sp} even without direct interaction between species the bacterial growth was inhibited due to the limitation of the resource.
@@
     \begin{center}
-    \includegraphics[width=0.6\textwidth]{Figures/pool_model_2pools_resource_competition.pdf}
+    \includegraphics[width=\columnwidth]{Figures/pool_model_2pools_resource_competition.pdf}
     \caption{
@@
 In order to get the commonly used form of the \ac{glv}, we need to simply eliminate of the bacterial states by setting $L_A=L_B=0$:
-\begin{equation}
-    \begin{cases}
-    \dot{G}_A = \alpha_A G_A\left(1 - \frac{G_A}{N_t}\right) - \frac{\alpha_A}{N_t}G_AG_B\\
-    \dot{G}_B = \alpha_B G_B\left(1-\frac{G_B}{N_t}\right) -\frac{\alpha_B}{N_t}G_AG_B. 
+\begin{align}
+    \begin{split}
+        \dot{G}_A &= \alpha_A G_A\left(1 - \frac{G_A}{N_t}\right) - \frac{\alpha_A}{N_t}G_AG_B\\
+        \dot{G}_B &= \alpha_B G_B\left(1-\frac{G_B}{N_t}\right) -\frac{\alpha_B}{N_t}G_AG_B. 
     \label{eq:LV_simple}
-    \end{cases}
-\end{equation}
+    \end{split}
+\end{align}
 
@@
 To introduce active competitive mechanisms, we extend our model with one of the species producing the toxic compounds $T$ during its growth:
-\begin{eqnarray}
-    \dot{T} &=& \mathcal{F}(R,G).
-\end{eqnarray}
+\begin{equation}
+    \dot{T} = \mathcal{F}(R,G).
+\end{equation}
 The function $\mathcal{F}$ denotes the model for the toxin production rate.
@@
 Then the full \ac{ode} system is
-\begin{equation}
-    \begin{cases}
-        \dot{L}_A = - \lambda_A R L_A\\
-        \dot{L}_B = - \lambda_B R L_B \\
-        \dot{G}_A = \lambda_A R L_A + \alpha_A R G_A - \frac{\mu}{\kappa N_t} T G_A\\
-        \dot{G}_B = \lambda_B R L_B + \alpha_B R G_B\\ 
-        \dot{R} = -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B\\
-        \dot{T}_{B}^{1,2} = \mathcal{F}^{1,2} (R, G_B) \\
-    \end{cases}.
+\begin{align}
+    \begin{split}
+        \dot{L}_A &= - \lambda_A R L_A\\
+        \dot{L}_B &= - \lambda_B R L_B \\
+        \dot{G}_A &= \lambda_A R L_A + \alpha_A R G_A - \frac{\mu}{\kappa N_t} T G_A\\
+        \dot{G}_B &= \lambda_B R L_B + \alpha_B R G_B\\ 
+        \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B\\
+        \dot{T}_{B}^{1,2} &= \mathcal{F}^{1,2} (R, G_B) \\
+    \end{split}.
     \label{eq:model_2sp_toxin}
-\end{equation}
+\end{align}
 %
 Where we consider two possible models of the toxin production:
-\begin{eqnarray}
-    \mathcal{F}^1(R,G)&=&\kappa G_B\\
-    \mathcal{F}^2(R,G)&=&\kappa\alpha_B R G_B.
-\end{eqnarray}
+\begin{align}
+    \mathcal{F}^1(R,G)&=\kappa G_B\\
+    \mathcal{F}^2(R,G)&=\kappa\alpha_B R G_B.
+\end{align}
 Here $\kappa$ is the toxin production rate.
@@
     \begin{center}
-    \includegraphics[width=1.\textwidth]{Figures/pool_model_2pools_toxin.pdf}
+    \includegraphics[width=\columnwidth]{Figures/pool_model_2pools_toxin.pdf}
     \caption{
@@
 results in: 
-\begin{equation}
-    \begin{cases}
-    \dot{G}_A =\frac{G_A}{G_B}\left(1 - G_A-G_B\right)\\
-    \dot{G}_B =\psi\frac{G_B}{G_A}\left(1-G_A-G_B\right). 
-    \end{cases}
+\begin{align}
+    \begin{split}
+        \dot{G}_A &=\frac{G_A}{G_B}\left(1 - G_A-G_B\right)\\
+        \dot{G}_B &=\psi\frac{G_B}{G_A}\left(1-G_A-G_B\right). 
+    \end{split}
 \label{eq:2sp_1pool_inhibit_unitless}
-\end{equation}
-We can separate three cases (see Figure~\ref{fig:1pool_2sp_inhibit}):
+\end{align}
+We can separate three cases (which are displayed in Figure~\ref{fig:1pool_2sp_inhibit}):
 \begin{eqnarray}
     \psi&=&1:\quad\text{A and B are equivalent}\\
-    \psi&>&1:\quad\text{Species B growth faster and/or inhibits A stronger}\\
-    \psi&<&1:\quad\text{Species A growth faster and/or inhibits B stronger}
+    \psi&>&1:\quad\text{B grows faster/inhibits A}\\
+    \psi&<&1:\quad\text{A grows faster/inhibits B}
 \end{eqnarray}
@@
     \begin{center}
-    \includegraphics[width=0.5\textwidth]{Figures/pool_model_1pool_inhib.pdf}
+    % TODO make this into 2x2 grid for plots and insert eqnarray from before as well
+    \includegraphics[width=\columnwidth]{Figures/pool_model_1pool_inhib.pdf}
     \caption{
@@
     \begin{center}
-    \includegraphics[width=0.5\textwidth]{Figures/pool_model_2sp_1pool_coop.pdf}
+    \includegraphics[width=\columnwidth]{Figures/pool_model_2sp_1pool_coop.pdf}
     \caption{
@@
     \begin{center}
-    \includegraphics[width=0.4\textwidth]{Figures/snapshot-2-cells-before-division.png}
-    \includegraphics[width=0.4\textwidth]{Figures/snapshot-2-cells-after-division.png}
+    \includegraphics[width=0.49\columnwidth]{Figures/snapshot-2-cells-before-division.png}%
+    \hspace{0.01\columnwidth}%
+    \includegraphics[width=0.49\columnwidth]{Figures/snapshot-2-cells-after-division.png}
     \caption{
@@
 We compare the most general version and thus omit species $A$ for the moment.
-\begin{table}[h]
+\begin{table*}[h]
     \centering
     \renewcommand{\arraystretch}{1.5}
-    \begin{tabularx}{\textwidth}{@{}l *3{>{\centering\arraybackslash}X}@{}}
+    \begin{tabularx}{\columnwidth}{@{}l *3{>{\centering\arraybackslash}X}@{}}
         & \textbf{\acs{ode}}
@@
     \end{tabularx}
-\end{table}
+\end{table*}
 We can use these mappings to numerically compare both approaches.
@@
 \begin{figure}
-    \includegraphics[width=0.49\textwidth]{Figures/abm-homogenous/snapshot_00006000.png}\hspace{0.02\textwidth}%
-    \includegraphics[width=0.49\textwidth]{Figures/abm-homogenous/snapshot_00012000.png}\vspace{0.02\textwidth}\\
-    \includegraphics[width=0.49\textwidth]{Figures/abm-homogenous/snapshot_00018000.png}\hspace{0.02\textwidth}%
-    \includegraphics[width=0.49\textwidth]{Figures/abm-homogenous/snapshot_00024000.png}%
+    \includegraphics[width=0.49\columnwidth]{Figures/abm-homogenous/snapshot_00006000.png}\hspace{0.01\columnwidth}%
+    \includegraphics[width=0.49\columnwidth]{Figures/abm-homogenous/snapshot_00012000.png}\vspace{0.01\columnwidth}\\
+    \includegraphics[width=0.49\columnwidth]{Figures/abm-homogenous/snapshot_00018000.png}\hspace{0.01\columnwidth}%
+    \includegraphics[width=0.49\columnwidth]{Figures/abm-homogenous/snapshot_00024000.png}%
     \caption{
@@
     \centering
-    \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures/abm-homogenous/abm_ode_comparison.png}
-        \subcaption{Homogenous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
+    \begin{subfigure}[c]{0.5\columnwidth}
+        \includegraphics[width=\columnwidth]{Figures/abm-homogenous/abm_ode_comparison.png}
+        \subcaption{Homogenous initial cell distribution, $D=30.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}%
-    \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures/abm-homogenous-low-diffusion/abm_ode_comparison.png}
-        \subcaption{Homogenous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
+    \begin{subfigure}[c]{0.5\columnwidth}
+        \includegraphics[width=\columnwidth]{Figures/abm-homogenous-low-diffusion/abm_ode_comparison.png}
+        \subcaption{Homogenous initial cell distribution, $D=5.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}\\
-    \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures/abm-inhomogenous/abm_ode_comparison.png}%
-        \subcaption{Heterogeneous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
+    \begin{subfigure}[c]{0.5\columnwidth}
+        \includegraphics[width=\columnwidth]{Figures/abm-inhomogenous/abm_ode_comparison.png}%
+        \subcaption{Heterogeneous initial cell distribution, $D=30.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}%
-    \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures/abm-random/abm_ode_comparison.png}%
-        \subcaption{Randomized initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
+    \begin{subfigure}[c]{0.5\columnwidth}
+        \includegraphics[width=\columnwidth]{Figures/abm-random/abm_ode_comparison.png}%
+        \subcaption{Randomized initial cell distribution, $D=5.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}
@@
         Snapshots of the spatial configurations of cells are included in the appendix~\ref{fig:supplement-spatial-initial-config-cells}.
+        The diffusion constants are chosen $D_R=D_I=D$.
         We see that in the highly-diffusive homogenous case of subfigure (a), the pool model is a good fit to the \ac{abm} model.
@@
 \end{itemize}
+
+\onecolumn
 

### 2025-08-28 pool_model.tex
more cleanup; more formatting for twocolumn layout

@@
 \usepackage[T1]{fontenc}
-\usepackage[a4paper,inner=2.5cm,outer=3cm]{geometry}
+\usepackage[a4paper,inner=2cm,outer=2cm]{geometry}
 \usepackage{graphicx}% Include figure files
@@
     \begin{center}
-    \includegraphics[width=\columnwidth]{Figures/TPM_fig.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures/TPM_fig.pdf}
     \caption{Schematic representation of exchange between three states or pools.}
@@
     \begin{center}
-    \includegraphics[width=\columnwidth]{Figures/pool_model_3pools_resource.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_3pools_resource.pdf}
     \caption{
@@
     D &\stackrel{\beta}{\longrightarrow} & \emptyset
-\label{eq:reactions_backlag}
 \end{eqnarray}
@@
 To model it we propose a kind of viscoelastic Maxwell type of stress-strain relation and write down the ordinary differential equation for $\gamma$ to be:
-\begin{eqnarray}
-    \dot{\gamma} &=& \Gamma \left |\dot{T}\right |-\delta \gamma.
-\label{eq:rate_backlag}
-\end{eqnarray}
+\begin{equation}
+    \dot{\gamma} = \Gamma |\dot{T}|-\delta \gamma.
+\end{equation}
 We assumed that the direction of the temperature change does not matter, i.e., a change from $T=2$ to $T=12$ has the same effect as a change from $T=14$ to $T=4$.
@@
 Integration of this equation yields (with $\gamma(0)=0$):
-\begin{eqnarray}
-    \gamma(t) &=& \Gamma \int_0^t \left |\dot{T}(t^{\prime})\right |e^{-\delta (t-t^{\prime})}dt^{\prime}.
-\label{eq:rate_backlag2}
-\end{eqnarray}
+\begin{equation}
+    \gamma(t) = \Gamma \int_0^t |\dot{T}(t^{\prime})|e^{-\delta (t-t^{\prime})}dt^{\prime}.
+\end{equation}
 After some manipulation of this equation we find:
-\begin{eqnarray}
-    \gamma(t) &=& \Gamma\delta\left |\left[T[t]-T[0]e^{-\delta t}-\delta \int_0^t T(t^{\prime})e^{-\delta (t-t^{\prime})}dt^{\prime}\right]\right |.
-\label{eq:rate_backlag3}
-\end{eqnarray}
+\begin{equation}
+    \gamma(t) = \Gamma\delta\left |T[t]-T[0]e^{-\delta t}-\delta
+        \int\limits_0^t T(t^{\prime})e^{-\delta (t-t^{\prime})}dt^{\prime}\right |.
+\end{equation}
 This representation of the rate $\gamma(t)$ has the advantage that it only involves $T(t)$ and does not require the calculation of the derivative of $T$.
@@
     \begin{split}
-    \dot{L} =\gamma(t) G - (\mu + \lambda) L\\
-    \dot{G} = \lambda L -\gamma(t) G + \alpha G\left(1-\frac{L+G+D}{N_t}\right)\\
-    \dot{D} = \mu  L - \beta D
+        \dot{L} &=\gamma(t) G - (\mu + \lambda) L\\
+        \dot{G} &= \lambda L -\gamma(t) G + \alpha G\left(1-\frac{L+G+D}{N_t}\right)\\
+        \dot{D} &= \mu  L - \beta D
     \end{split}
@@
     \begin{center}
-    \includegraphics[width=\columnwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}
     \caption{
@@
 If the shift happened rather early in time, when most of the bacteria were in lag-phase, the effect is not noticeable.
-\newpage
 %
@@
     \begin{split}
-        \dot{L}_A = - \lambda_A R L_A\\
-        \dot{L}_B = - \lambda_B R L_B \\
-        \dot{G}_A = \lambda_A R L_A +\alpha_A R G_A\\
-        \dot{G}_B = \lambda_B R L_B +\alpha_B R G_B\\
-        \dot{R} =-\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B
+        \dot{L}_A &= - \lambda_A R L_A\\
+        \dot{L}_B &= - \lambda_B R L_B \\
+        \dot{G}_A &= \lambda_A R L_A +\alpha_A R G_A\\
+        \dot{G}_B &= \lambda_B R L_B +\alpha_B R G_B\\
+        \dot{R} &=-\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B
     \end{split}
@@
 We again rescaled the resource pool with $N_t$ and find
-\begin{eqnarray}
+\begin{equation}
+    R =1-\frac{L_A+G_A+L_B+G_B}{N_t}
     \label{eq:Resource}
-R &=&1-\frac{L_A+G_A+L_B+G_B}{N_t}
-\end{eqnarray}
+\end{equation}
 Inserting Eq.~\ref{eq:Resource} into Eqs.~\ref{eq:model_2sp_resource_comp} yields:
-\begin{eqnarray*}
-    \dot{L}_A &=& - \lambda_A  L_A + \frac{\lambda_A}{N_t}L_A\left<LG\right>\\
-    \dot{L}_B &=& - \lambda_B L_B + \frac{\lambda_B}{N_t}L_B \left<LG\right>\\
-    \dot{G}_A &=&  \lambda_A  L_A + \alpha_A G_A - \frac{\lambda_A L_A + \alpha_A G_A}{N_t} \left<LG\right>\\
-    \dot{G}_B &=&  \lambda_B L_B - \alpha_B G_B \frac{\lambda_B L_B + \alpha_B G_B}{N_t}\left<LG\right>
-\end{eqnarray*}
+\begin{align}
+    \dot{L}_A &= -\lambda_A  L_A + \frac{\lambda_A}{N_t}L_A\left<LG\right>\\
+    \dot{L}_B &= -\lambda_B L_B + \frac{\lambda_B}{N_t}L_B \left<LG\right>\\
+    \dot{G}_A &=  \lambda_A  L_A + \alpha_A G_A - \frac{\lambda_A L_A + \alpha_A G_A}{N_t} \left<LG\right>\\
+    \dot{G}_B &=  \lambda_B L_B - \alpha_B G_B \frac{\lambda_B L_B + \alpha_B G_B}{N_t}\left<LG\right>
+\end{align}
 where we used the short notation $\left<LG\right>=(L_A+G_A+L_B+G_B)$.
@@
     \begin{center}
-    \includegraphics[width=\columnwidth]{Figures/pool_model_2pools_resource_competition.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_2pools_resource_competition.pdf}
     \caption{
@@
 %
-\begin{figure}[H]
+\begin{figure*}[t!]
     \begin{center}
-    \includegraphics[width=\columnwidth]{Figures/pool_model_2pools_toxin.pdf}
+    \includegraphics[width=2.0\columnwidth]{Figures/pool_model_2pools_toxin.pdf}
     \caption{
@@
     \end{center}
-\end{figure}
+\end{figure*}
 %
@@
     % TODO make this into 2x2 grid for plots and insert eqnarray from before as well
-    \includegraphics[width=\columnwidth]{Figures/pool_model_1pool_inhib.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_1pool_inhib.pdf}
     \caption{
@@
     \begin{center}
-    \includegraphics[width=\columnwidth]{Figures/pool_model_2sp_1pool_coop.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_2sp_1pool_coop.pdf}
     \caption{
@@
     \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=\columnwidth]{Figures/abm-homogenous/abm_ode_comparison.png}
+        \includegraphics[width=0.9\columnwidth]{Figures/abm-homogenous/abm_ode_comparison.png}
         \subcaption{Homogenous initial cell distribution, $D=30.0$\unit{\micro\metre\squared\per\minute}}
@@
     \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=\columnwidth]{Figures/abm-homogenous-low-diffusion/abm_ode_comparison.png}
+        \includegraphics[width=0.9\columnwidth]{Figures/abm-homogenous-low-diffusion/abm_ode_comparison.png}
         \subcaption{Homogenous initial cell distribution, $D=5.0$\unit{\micro\metre\squared\per\minute}}
@@
     \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=\columnwidth]{Figures/abm-inhomogenous/abm_ode_comparison.png}%
+        \includegraphics[width=0.9\columnwidth]{Figures/abm-inhomogenous/abm_ode_comparison.png}%
         \subcaption{Heterogeneous initial cell distribution, $D=30.0$\unit{\micro\metre\squared\per\minute}}
@@
     \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=\columnwidth]{Figures/abm-random/abm_ode_comparison.png}%
+        \includegraphics[width=0.9\columnwidth]{Figures/abm-random/abm_ode_comparison.png}%
         \subcaption{Randomized initial cell distribution, $D=5.0$\unit{\micro\metre\squared\per\minute}}

### 2025-08-28 pool_paper_plotting_code/digital_twin/plotting_templates.py
plotting: relocale all important general model parameters to plotting_templates

@@ -3,68 +3,41 @@
 import matplotlib.pyplot as plt
 import numpy as np
 
+colors_all = {
+        'R': '#808080',
+        'N_A':'#D06062',
+        'N_B': '#4E89B1',
+        'N':'#7E57A5',
+        'T':'#99582A',
+        'T_A':'#c79758',
+        'N_lambd_1e-2_mu_0':'#E2B100',
+        'N_lambd_1e-3_mu_0':'#386641',
+        'N_lambd_1e-3_mu_0_5':'#0982A4',
+        'N_wo_tempshift':'#679E48',
+        'N_tempshift_10':'#ED733E',
+        'N_tempshift_10_5_15':'#C3568A',
+    }
+
+figsize_default = (6.5, 4.0)
+figsize_default2subpl = (13, 4.0)
+
 plt.rcParams["font.family"] = "serif"
 plt.rc('text', usetex=True)
 plt.rcParams['text.latex.preamble'] = r"\usepackage{bm} \usepackage{amsmath}"
 
-# Set some custom colors/scatter markers for plotting (mb add more or remove later)
-colors = [(0 / 255, 102 / 255, 217 / 255),
-          (90 / 255, 170 / 255, 90 / 255),
-          (215 / 255, 48 / 255, 39 / 255)]  
-gray = (128 / 255, 128 / 255, 128 / 255)
-gray2 = (160 / 255, 160 / 255, 160 / 255)
-scatter_marker = ['^', 'o', 'D', "s", "X", ".", "1", "2", "3", "4", "*", "P", "d", "+", "x"]
+plt.rcParams['legend.fontsize'] = 15.
+plt.rcParams['legend.framealpha'] = 0.
+plt.rcParams['legend.handlelength'] = 1.8
+plt.rcParams['axes.prop_cycle'] = plt.cycler(linewidth=[2.5])
+plt.rcParams['font.size'] = 15
+plt.rc('xtick', labelsize=13)
+plt.rc('ytick', labelsize=13)
 
-
-def template_plot_measurements0(ax0, temp, data, c='b'):
-    n_plot = 0
-    for d in data: 
-        if d['const'] == temp:
-            for meas, std in zip(d['obs_mean'], d['obs_std']):
-                ax0.errorbar(d['times'], np.log10(meas), fmt=scatter_marker[n_plot],  yerr=np.abs(0.43*std/meas),
-                             markersize=7, color=c, label="T = {}".format(temp))
-            n_plot += 1
-
-
[…]

### 2025-08-28 pool_paper_plotting_code/examples/main_paper.py

@@ -179,141 +179,55 @@ def observable_2pool_2species_resource_toxin(output):
             output[4],
             output[5]]
 
+
 if __name__ == "__main__":
-    plt.rcParams['legend.fontsize'] = 25.
-    plt.rcParams['legend.framealpha'] = 0.
-
-    plt.rcParams['axes.prop_cycle'] = plt.cycler(linewidth=[2.5])
-
-    grey = '#808080' #(128 / 255, 128 / 255, 128 / 255)
-
-    color_palette_2sp = ['#D06062', '#4E89B1', '#7E57A5', grey] # NA, NB, NA+NB, R
-    color_palette_2sp_toxin = ['#D06062', '#4E89B1', grey, '#99582A'] # NA, NB, R, T
-    color_palette_2sp_coop = ['#D06062', '#4E89B1', '#c79758', '#99582A'] # NA, NB, R, TA, TB
-    color_palette_2sp_only = ['#D06062', '#4E89B1']
-    color_palette_1sp = ['#386641', '#679E48', '#E2B100', '#0982A4'] 
-    color_palette_1sp_tempshift = ['#679E48', '#ED733E', '#C3568A']
-
-
-    #colorsp = np.zeros((6, 3))
-    #colorsp[0, :] = (5,48,97)
-    #colorsp[1, :] = (50,147,200)
-    #colorsp[2, :] = (90,170,90)
-    #colorsp[3, :] = (250,140,70)
-    #colorsp[4, :] = (215,48,39)
-    #colorsp[5, :] = (103, 0, 31 )
-    #colorsp = colorsp / 255
-
-
-    #colors2 = np.zeros((7, 3))
-    #colors2[0] = (228,26,28)
-    #colors2[1] = (55,126,184)
-    #colors2[2] = (77,175,74)
-    #colors2[3] = (152,78,163)
-    #colors2[4] = (255,127,0)
-    #colors2[5] = (166,86,40)
-    #colors2 = colors2 / 255
-
-    figsize_default = (6.5, 4.0)
-    figsize_default2subpl = (13, 4.0)
-
-
-    # 1. 3 pool model generate and plot
-
-    # 2. 3 pool models with resource pool 
-    # 3. Resource with lag-phase
-
-    '''
-    x02 = [[1e3], [0.], [0.]]
-    
-    lambd_vals = [1e-1, 1e-3, 1e-5]
-    fig, ax = plt.subplots()
[…]

### 2025-08-28 pool_paper_plotting_code/examples/pool_paper_spatial.py

@@ -4,6 +4,10 @@ import matplotlib.pyplot as plt
 import numpy as np
 from scipy.integrate import solve_ivp
 
+import os, sys
+sys.path.append(os.getcwd())
+from digital_twin import *
+
 
 def model_solve(model, times, param, x0, const, obs_func=None):
     model_output = model_ODE_solution(model, times, param, x0, const)
@@ -50,11 +54,11 @@ if __name__ == "__main__":
     labels = ['Species 1', 'Species 2', 'Sp 1 + Sp 2']
     for j, obs in enumerate(observabl):
         ax.plot(times, obs, label=labels[j])#, color=colorsp[i+2])
-    ax.set_xlabel('Time, [days]', fontsize=15)
+    ax.set_xlabel('Time, [days]')
     ax.tick_params(labelsize=13)
-    ax.set_ylabel('Total Bacterial Count', fontsize=15)
+    ax.set_ylabel('Total Bacterial Count')
     #ax.set_yscale('log')
     ax.set_xlim(0., 10.)
-    ax.legend(fontsize=15)
+    ax.legend()
     plt.savefig(f'pool_model_spatial.pdf', bbox_inches='tight')
     plt.close(fig) 

### 2025-08-28 pool_model.tex
formatting of margins
insert blank abstract on new page
new caption

--- 2025-08-28 pool_model.tex
@@
 \setlength{\tabcolsep}{1.3pt}
+\setlength{\columnsep}{0.4cm}
 \usepackage{acronym}
@@
 \usepackage[T1]{fontenc}
-\usepackage[a4paper,inner=2cm,outer=2cm]{geometry}
+\usepackage[a4paper,inner=2cm,outer=2cm,top=2.5cm,bottom=2.5cm]{geometry}
 \usepackage{graphicx}% Include figure files
@@
 
+\begin{abstract}
+    TODO
+\end{abstract}
+
+\newpage
 \tableofcontents
@@
     G + R_0  \stackrel{\alpha_0}{\longrightarrow} 2G.
-\label{eq:reactions_resource}
 \end{equation}
@@
 \end{align}
-\begin{figure}[h]
+%
+%
+\begin{figure}
     \begin{center}
@@
 Assuming $L\ll N_t$ we can ignore the non-linear term in the equation for $G$, which gives rise to:
-\begin{eqnarray}
-    1&=& 2\int_0^{t_L} \lambda L(t)e^{\alpha(t_L-t)}dt\\
-    1  &=& 2\lambda e^{\alpha t_L}\frac{1-e^{-(\mu+\lambda+\alpha)t_L}}{\mu+\lambda+\alpha}.
-\end{eqnarray}
+\begin{align}
+    1&= 2\int_0^{t_L} \lambda L(t)e^{\alpha(t_L-t)}dt\\
+    1  &= 2\lambda e^{\alpha t_L}\frac{1-e^{-(\mu+\lambda+\alpha)t_L}}{\mu+\lambda+\alpha}.
+\end{align}
 For $\alpha > \mu+\lambda$ $t_L$ can be approximated by:
-\begin{eqnarray}
-    t_L &\approx& -\frac{1}{\alpha}\ln\left(\frac{2\lambda}{\mu+\lambda+\alpha}\right).
-\end{eqnarray}
+\begin{align}
+    t_L &\approx -\frac{1}{\alpha}\ln\left(\frac{2\lambda}{\mu+\lambda+\alpha}\right).
+\end{align}
 The lag time $t_L$ diverges very slowly, logarithmically with $\lambda\to 0$.
@@
 As a result, species $A$ was able to suppress species $B$ consuming the larger amount of nutrient and reaching larger bacterial count at the end.
-\begin{figure}[H]
+\begin{figure}
     \begin{center}
@@
 %
-\begin{figure*}[t!]
+\begin{figure}
     \begin{center}
@@
 For sake of simplicity we ignore the lag pool and write for a mutual inhibition:
-\begin{eqnarray}
-    \dot{G}_A &=& \mathcal{B}_A(G_B)G_A\left(1 - \frac{G_A+G_B}{N_t}\right)\\
-    \dot{G}_B &=& \mathcal{B}_B(G_A) G_B\left(1-\frac{G_A+G_B}{N_t}\right).
-\label{eq:ode_inhib1}
-\end{eqnarray}
+\begin{align}
+    \dot{G}_A &= \mathcal{B}_A(G_B)G_A\left(1 - \frac{G_A+G_B}{N_t}\right)\\
+    \dot{G}_B &= \mathcal{B}_B(G_A) G_B\left(1-\frac{G_A+G_B}{N_t}\right).
+\end{align}
 The function $\mathcal{B}_{A/B}$ captures the inhibiting effects effectively.
@@
 One possible choice for $\mathcal{B}$ is:
-\begin{eqnarray}
-    \mathcal{B}(G)&=&\frac{\alpha}{1+KI(G)},
-\label{eq:ode_inhib_B}
-\end{eqnarray}
+\begin{equation}
+    \mathcal{B}(G) = \frac{\alpha}{1+KI(G)},
+\end{equation}
 where $I$ is the concentration of the inhibiting metabolic byproduct producing during bacterial growth.
-\begin{eqnarray}
-    \dot{G}_A &=& \frac{\alpha_A G_A}{1+K_BG_B}\left(1 - \frac{G_A+G_B}{N_t}\right)\\
-    \dot{G}_B &=& \frac{\alpha_B G_B}{1+K_AG_A}\left(1-\frac{G_A+G_B}{N_t}\right).
-\label{eq:ode_inhib2}
-\end{eqnarray}
+\begin{align}\begin{split}
+    \dot{G}_A &= \frac{\alpha_A G_A}{1+K_BG_B}\left(1 - \frac{G_A+G_B}{N_t}\right)\\
+    \dot{G}_B &= \frac{\alpha_B G_B}{1+K_AG_A}\left(1-\frac{G_A+G_B}{N_t}\right).
+\end{split}\end{align}
 Note that if the toxin accumulates, i.e., $\dot I \sim G$,  one needs to replace $G$ by $\int G dt$.
@@
 In the latter case one can simplify the equations, given $N_A$ and $N_B$ are non-zero:
-\begin{eqnarray}
-    \dot{G}_A &=&\tilde{\alpha}_A\frac{G_A}{G_B}\left(1 - G_A-G_B\right)\\
-    \dot{G}_B &=& \tilde{\alpha}_B\frac{G_B}{G_A}\left(1-G_A-G_B\right).
-\label{eq:ode_inhib3}
-\end{eqnarray}
+\begin{align}\begin{split}
+    \dot{G}_A &=\tilde{\alpha}_A\frac{G_A}{G_B}\left(1 - G_A-G_B\right)\\
+    \dot{G}_B &= \tilde{\alpha}_B\frac{G_B}{G_A}\left(1 - G_A-G_B\right).
+\end{split}\end{align}
 We introduced the rescaled growth rates $\tilde{\alpha}_A=\alpha_A/(N_tK_B)$ and $\tilde{\alpha}_B=\alpha_B/(N_tK_A)$. Note that $[\tilde{\alpha}]=\mathrm{s}^{-1}$.
@@
     \psi = \tilde{\alpha}_B/\tilde{\alpha}_A = \frac{K_B\alpha_B}{K_A\alpha_A}
-\label{eq:ode_inhib_psi}
 \end{equation}
@@
 We can separate three cases (which are displayed in Figure~\ref{fig:1pool_2sp_inhibit}):
-\begin{eqnarray}
-    \psi&=&1:\quad\text{A and B are equivalent}\\
-    \psi&>&1:\quad\text{B grows faster/inhibits A}\\
-    \psi&<&1:\quad\text{A grows faster/inhibits B}
-\end{eqnarray}
-%
-\begin{figure}[H]
+\begin{align}
+    \psi&=1:\quad\text{A and B are equivalent}\\
+    \psi&>1:\quad\text{B grows faster/inhibits A}\\
+    \psi&<1:\quad\text{A grows faster/inhibits B}
+\end{align}
+%
+\begin{figure}
     \begin{center}
-    % TODO make this into 2x2 grid for plots and insert eqnarray from before as well
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_1pool_inhib.pdf}
+    % TODO make this into 2x2 grid for plots and insert cases for psi from before as well
+    \includegraphics[width=\columnwidth]{Figures/pool_model_1pool_inhib.pdf}
     \caption{
@@
 For sake of clarity we omit again the lag pool and write:
-\begin{eqnarray}
-    \dot{G}_A &=& \mathcal{B}_A(P_B,R)G_A\\
-    \dot{G}_B &=& \mathcal{B}_B(P_A,R)G_B\\
-    \dot{R} &=&-\mathcal{B}_A(P_B,R)G_A-\mathcal{B}_B(P_A,R)G_B\\
-    \dot{P}_A &=& \mathcal{F}_A(R,G_A)\\
-    \dot{P}_B &=& \mathcal{F}_B(R,G_B).
-\end{eqnarray}
+\begin{align}
+    \dot{G}_A &= \mathcal{B}_A(P_B,R)G_A\\
+    \dot{G}_B &= \mathcal{B}_B(P_A,R)G_B\\
+    \dot{R} &=-\mathcal{B}_A(P_B,R)G_A-\mathcal{B}_B(P_A,R)G_B\\
+    \dot{P}_A &= \mathcal{F}_A(R,G_A)\\
+    \dot{P}_B &= \mathcal{F}_B(R,G_B).
+\end{align}
 $P_A$ and $P_B$ denote the substance produced by species A and B, respectively.
@@
 Setting $\mathcal{B}_A(P_B,R)=\alpha_A P_BR$, $\mathcal{B}_B(P_A,R)=\alpha_B P_A R$, $\mathcal{F}_A(R,G_A)=\kappa_A RG_A$, $\mathcal{F}_B(R,G_B)=\kappa_B RG_B$, rescaling $G_A$, $G_B$, and $P_B$ with $N_t$, $P_A$ with $\kappa_A/\alpha_A$, time with $\alpha_AN_t^2$, defining $\psi=\kappa_A\alpha_B/(\alpha_A^2N_t)$, and $\phi=\kappa_B/(\alpha_A N_t)$ results in:
-\begin{equation}
-    \begin{cases}
-    \dot{G}_A = P_B G_A\left(1 - G_A-G_B\right)\\
-    \dot{G}_B = \psi P_A G_B\left(1 - G_A-G_B\right)\\
-    \dot{P}_A = G_A\left(1 - G_A-G_B\right)\\
-    \dot{P}_B =\phi G_B\left(1 - G_A-G_B\right)
-    \end{cases}.
+\begin{align}
+    \dot{G}_A &= P_B G_A\left(1 - G_A-G_B\right)\\
+    \dot{G}_B &= \psi P_A G_B\left(1 - G_A-G_B\right)\\
+    \dot{P}_A &= G_A\left(1 - G_A-G_B\right)\\
+    \dot{P}_B &=\phi G_B\left(1 - G_A-G_B\right).
     \label{eq:1pool_2sp_cooper}
-\end{equation}
+\end{align}
 %
@@
 That indirectly leads to larger amount of $P_B$ produced by species $B$.
-\begin{figure}[H]
+\begin{figure}
     \begin{center}
@@
 %
-\begin{equation}
-    \begin{cases}
-        \dot{L_A} = -\lambda_A R L_A\\
+\begin{align}
+    \dot{L_A} &= -\lambda_A R L_A\\
     \dot{G_A} = \lambda_A R L_A + \alpha_A R G_A\\
     \dot{L_B} = -\lambda_B R L_B\\
-        \dot{G_B} = \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B\\
-    \end{cases}.
+    \dot{G_B} = \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B
     \label{eq:spatial_limit_F}
-\end{equation}
+\end{align}
 %
 The microenvironment of our system contains the resource and the inhibitor molecules:
-\begin{equation}
-    \begin{cases}
-        \dot{R} = -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \mu_I I) N_t} R G_B \\
-        \dot{I} = \kappa G_A
-    \end{cases}.
+\begin{align}
+    \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \mu_I I) N_t} R G_B \\
+    \dot{I} &= \kappa G_A
     \label{eq:spatial_limit_H}
-\end{equation}
+\end{align}
 
@@
         (A) A cell right before and (B) after a division event.
-        The volume is equally distributed between each new cell and the new positions are picked random but at opposing sides on a circle around the original center of the single cell.
+        The volume is equally distributed between the two cells and the new positions are picked
+        random but at opposing sides on a circle around the original center of the single cell.
     }
@@
 We compare the most general version and thus omit species $A$ for the moment.
-\begin{table*}[h]
+\begin{table*}
     \centering
     \renewcommand{\arraystretch}{1.5}
-    \begin{tabularx}{\columnwidth}{@{}l *3{>{\centering\arraybackslash}X}@{}}
+    \begin{tabularx}{0.8\textwidth}{@{}l *3{>{\centering\arraybackslash}X}@{}}
         & \textbf{\acs{ode}}
@@
     \end{tabularx}
+    \caption{Mappings between \ac{ode} and \ac{abm} parameters.}
 \end{table*}
@@
 % ================================================================================================
+\newpage
 \beginsupplement
@@
 %
-\begin{eqnarray}
-    \dot \mbx  &=& \mbF(\mbx, \mbm, \mbp, \mbu) = \begin{pmatrix}
+\begin{align}
+    \dot \mbx  &= \mbF(\mbx, \mbm, \mbp, \mbu) = \begin{pmatrix}
         -(\mu + \lambda) & 0               & 0      \\
@@
     \end{pmatrix} \mbx\\
-    \dot \mbm &=& \mbH (\mbx, \mbm, \mbp, \mbu) = -\frac{\alpha}{N_t} R G.
-\label{eq:x_repres_3pools_resource}
-\end{eqnarray}
+    \dot \mbm &= \mbH (\mbx, \mbm, \mbp, \mbu) = -\frac{\alpha}{N_t} R G.
+\end{align}
 Separating linear and non-linear parts in two terms yields
@@
 
-\begin{table}[H]
+\begin{table}
     \centering

### 2025-08-28 pool_model.tex
formatting
simplify caption; move to main text

@@
         The background shows the resource concentration $[R]$.
-        Darker colours indicate less present nutrients.
+        Darker colours indicate that less nutrients are present.
     }
@@
 %
-\begin{figure}
-    \centering
-    \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=0.9\columnwidth]{Figures/abm-homogenous/abm_ode_comparison.png}
-        \subcaption{Homogenous initial cell distribution, $D=30.0$\unit{\micro\metre\squared\per\minute}}
-    \end{subfigure}%
-    \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=0.9\columnwidth]{Figures/abm-homogenous-low-diffusion/abm_ode_comparison.png}
-        \subcaption{Homogenous initial cell distribution, $D=5.0$\unit{\micro\metre\squared\per\minute}}
-    \end{subfigure}\\
-    \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=0.9\columnwidth]{Figures/abm-inhomogenous/abm_ode_comparison.png}%
-        \subcaption{Heterogeneous initial cell distribution, $D=30.0$\unit{\micro\metre\squared\per\minute}}
-    \end{subfigure}%
-    \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=0.9\columnwidth]{Figures/abm-random/abm_ode_comparison.png}%
-        \subcaption{Randomized initial cell distribution, $D=5.0$\unit{\micro\metre\squared\per\minute}}
-    \end{subfigure}
-    \caption{
-        Comparison of the \ac{ode} and \ac{abm} model for different parameter configurations.
-        Snapshots of the spatial configurations of cells are included in the appendix~\ref{fig:supplement-spatial-initial-config-cells}.
-        The diffusion constants are chosen $D_R=D_I=D$.
 We see that in the highly-diffusive homogenous case of subfigure (a), the pool model is a good fit to the \ac{abm} model.
@@
 The chosen diffusion constants $D_I=D_R=30.0$\unit{\micro\metre\squared\per\minute} are in the order of small molecules such as $NaCl$ while the lower values $D_I=D_R=5.0$\unit{\micro\metre\squared\per\minute} are around values of Glucose in water.
+\begin{figure}
+    \centering
+    \begin{subfigure}[c]{0.5\columnwidth}
+        \includegraphics[width=\columnwidth]{Figures/abm-homogenous/abm_ode_comparison.png}
+        \subcaption{Homogenous distribution, $D=30.0$\unit{\micro\metre\squared\per\minute}}
+    \end{subfigure}%
+    \begin{subfigure}[c]{0.5\columnwidth}
+        \includegraphics[width=\columnwidth]{Figures/abm-homogenous-low-diffusion/abm_ode_comparison.png}
+        \subcaption{Homogenous distribution, $D=5.0$\unit{\micro\metre\squared\per\minute}}
+    \end{subfigure}\\
+    \begin{subfigure}[c]{0.5\columnwidth}
+        \includegraphics[width=\columnwidth]{Figures/abm-inhomogenous/abm_ode_comparison.png}%
+        \subcaption{Heterogeneous distribution, $D=30.0$\unit{\micro\metre\squared\per\minute}}
+    \end{subfigure}%
+    \begin{subfigure}[c]{0.5\columnwidth}
+        \includegraphics[width=\columnwidth]{Figures/abm-random/abm_ode_comparison.png}%
+        \subcaption{Randomized distribution, $D=5.0$\unit{\micro\metre\squared\per\minute}}
+    \end{subfigure}
+    \caption{
+        Comparison of the \ac{ode} and \ac{abm} model for different parameter configurations.
+        Snapshots of the spatial configurations of cells are included in the appendix~\ref{fig:supplement-spatial-initial-config-cells}.
+        The diffusion constants are chosen such that $D_R=D_I=D$.
     }

### 2025-08-28 pool_model.tex
fix merge
fit both images in column
change title
incluide notes from christian

@@
 
-\title{Pool Model}
+\title{Pool Model - A unified Mathematical\\ Modelling Framework for Food Science}
 \author{
@@
 \begin{abstract}
-    TODO
+    \textbf{Notes from Christian}
+    Different empirical and inactivation models applied in the estimation
+    of parameters and modelling of phenomena in food microbiology, gen-
+    eralized Lotka Voltera approach used to describe the interaction of mi-
+    croorganisms, mechanistic models utilized to provide more insight into
+    the underlying dynamic can all be unified as variations of a new model
+    known as three pool model. Compared against the aforementioned
+    methods which are applicable in specific applications with some limi-
+    tations, the present model turns out to be less complex with general
+    applicability, low number of parameters and preferably to explain the
+    mechanism of the phenomenon of interest. To illustrate the generality
+    of the model, three different examples are reported.
 \end{abstract}
@@
     \begin{center}
-    \includegraphics[width=2.0\columnwidth]{Figures/pool_model_2pools_toxin.pdf}
+    \includegraphics[width=\columnwidth]{Figures/pool_model_2pools_toxin.pdf}
     \caption{
@@
     \end{center}
-\end{figure*}
+\end{figure}
 %

### 2025-08-29 pool_model.tex
formatting
Do separate files for subplots

@@
 
-\begin{abstract}
+\begin{abstract}\noindent
     \textbf{Notes from Christian}
@@
     \begin{center}
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}
+    %\includegraphics[width=0.9\columnwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_3pools_resource_tempshift1.pdf}\\
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_3pools_resource_tempshift2.pdf}
     \caption{
@@
     \begin{center}
-    \includegraphics[width=\columnwidth]{Figures/pool_model_2pools_toxin.pdf}
+    %\includegraphics[width=\columnwidth]{Figures/pool_model_2pools_toxin.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_2pools_toxin1.pdf}\\
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_2pools_toxin2.pdf}
     \caption{
@@
     % TODO make this into 2x2 grid for plots and insert cases for psi from before as well
-    \includegraphics[width=\columnwidth]{Figures/pool_model_1pool_inhib.pdf}
+    %\includegraphics[width=\columnwidth]{Figures/pool_model_1pool_inhib.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_1pool_inhib1.pdf}\\
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_1pool_inhib2.pdf}\\
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_1pool_inhib3.pdf}
     \caption{
@@
     \begin{center}
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_2sp_1pool_coop.pdf}
+    %\includegraphics[width=0.9\columnwidth]{Figures/pool_model_2sp_1pool_coop.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_2sp_1pool_coop1.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_2sp_1pool_coop2.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_2sp_1pool_coop3.pdf}
     \caption{

### 2025-08-29 pool_paper_plotting_code/digital_twin/plotting_templates.py
plotting code: add plotting of subplots separately

@@ -19,8 +19,10 @@ colors_all = {
     }
 
 figsize_default = (6.5, 4.0)
+figsize_default_small = (6.5, 2.0)
 figsize_default2subpl = (13, 4.0)
 
+plt.rcParams['figure.dpi'] = 400
 plt.rcParams["font.family"] = "serif"
 plt.rc('text', usetex=True)
 plt.rcParams['text.latex.preamble'] = r"\usepackage{bm} \usepackage{amsmath}"

### 2025-08-29 pool_paper_plotting_code/examples/main_paper.py

@@ -220,7 +220,7 @@ if __name__ == "__main__":
     const6 = [1e-2, 0., 2.5, .5, 5., Nt, Gamma, delta, ((1, 10), (3, 5), (7, 15))] # (lambd, mu1, alpha, mu2, beta, N_t, Gamma, delta, shift_cnd)
     consts_Tempshift = [const1, const5, const6]
     data_Tempshift = [data_3pool_res[0]] + [generate_insilico_data(pool_model_tempstress, [t], [], [[]], x01,
-                                            const=const, obs_func=observable_3pool, n_traj=1) for const in consts_Tempshift]
+                                            const=const, obs_func=observable_3pool, n_traj=1) for const in consts_Tempshift[1:]]
 
     labels = [r'$\Delta T = 0$', r'$\Delta T = 10$',  r'$\Delta T = 10, 5, 15$']
     color_palette_1sp_tempshift = [colors_all['N_wo_tempshift'], colors_all['N_tempshift_10'], colors_all['N_tempshift_10_5_15']]
@@ -244,6 +244,29 @@ if __name__ == "__main__":
     plt.savefig(f'out/pdfs/pool_model_3pools_resource_tempshift.pdf', bbox_inches='tight')
     plt.close(fig)
 
+    fig, ax = plt.subplots(1, 1, figsize=figsize_default)
+    for i, data in enumerate(data_Tempshift):
+        for d in data:
+            ax.plot(d['times'], d['obs_mean'][0], color=color_palette_1sp_tempshift[i], label=labels[i])
+    fig, ax = set_labels(fig, ax, r'', r'Bacterial Count, $N$')
+    ax.set_yscale('log')
+    ax.set_xlim(0.7, 8.5)
+    ax.set_ylim(6e0, 1.1e4)
+    ax.legend()
+    ax.text(8., 9.9, r'(a)')
+    plt.savefig(f'out/pdfs/pool_model_3pools_resource_tempshift1.pdf', bbox_inches='tight')
+    plt.close(fig)
+
+    fig, ax = plt.subplots(1, 1, figsize=figsize_default_small)
+    ax.plot(t, y, color=color_palette_1sp_tempshift[i-1])
+    y2 = [temp_stress(tt, *const6[-3:]) for tt in t]
+    ax.plot(t, y2, color=color_palette_1sp_tempshift[i])
[…]

### 2025-08-30 pool_paper_plotting_code/examples/main_paper.py
fix imports with path

@@ -1,8 +1,11 @@
+# fmt: off
 #!/usr/bin/env python3
-
 import numpy as np
-import os, sys
-sys.path.append(os.getcwd())
+import os
+import sys
+import matplotlib.pyplot as plt
+sys.path.append(f"{os.getcwd()}")
+sys.path.append(f"{os.getcwd()}/pool_paper_plotting_code")
 from digital_twin import *
 
 

### 2025-08-30 pool_paper_plotting_code/examples/main_paper.py
make into normal string instead of f-string

@@ -211,7 +211,7 @@ if __name__ == "__main__":
     ax.set_xlim(0., 9.0)
     ax.set_ylim(3.3e0, 1.1e4)
     ax.legend()
-    plt.savefig(f'out/pdfs/pool_model_3pools_resource.pdf', bbox_inches='tight')
+    plt.savefig('out/pdfs/pool_model_3pools_resource.pdf', bbox_inches='tight')
     plt.close(fig)
 
 #################################### Temperature shift ##########################################
@@ -244,7 +244,7 @@ if __name__ == "__main__":
     ax[0].legend()
     ax[0].text(8., 9.9, r'(a)')
     ax[1].text(8., 1.5, r'(b)')
-    plt.savefig(f'out/pdfs/pool_model_3pools_resource_tempshift.pdf', bbox_inches='tight')
+    plt.savefig('out/pdfs/pool_model_3pools_resource_tempshift.pdf', bbox_inches='tight')
     plt.close(fig)
 
     fig, ax = plt.subplots(1, 1, figsize=figsize_default)
@@ -257,7 +257,7 @@ if __name__ == "__main__":
     ax.set_ylim(6e0, 1.1e4)
     ax.legend()
     ax.text(8., 9.9, r'(a)')
-    plt.savefig(f'out/pdfs/pool_model_3pools_resource_tempshift1.pdf', bbox_inches='tight')
+    plt.savefig('out/pdfs/pool_model_3pools_resource_tempshift1.pdf', bbox_inches='tight')
     plt.close(fig)
 
     fig, ax = plt.subplots(1, 1, figsize=figsize_default_small)
@@ -267,7 +267,7 @@ if __name__ == "__main__":
     fig, ax = set_labels(fig, ax, r'Time, $t$', r'Rate, $\gamma$')
     ax.set_xlim(0.7, 8.5)
     ax.text(8., 1.5, r'(b)')
-    plt.savefig(f'out/pdfs/pool_model_3pools_resource_tempshift2.pdf', bbox_inches='tight')
+    plt.savefig('out/pdfs/pool_model_3pools_resource_tempshift2.pdf', bbox_inches='tight')
     plt.close(fig)
 
 #################################### Resource competition ################################################3
@@ -290,7 +290,7 @@ if __name__ == "__main__":
     ax.set_xlim(0., 8.0)
[…]

### 2025-08-30 pool_paper_plotting_code/examples/main_paper.py
remove whitespace

@@ -13,7 +13,7 @@ def pool_model_3pools(t, x, param, x0, const):
     (lambd, mu1, alpha, mu2, beta, ) = const
     (L0, G0, D0, ) = x0
     (L, G, D, ) = x
-        
+
     return [
         - (mu1 + lambd) * L,
                   lambd * L + (alpha - mu2) * G,
@@ -24,7 +24,7 @@ def pool_model_3pools_resource(t, x, param, x0, const):
     (lambd, mu1, alpha, mu2, beta, N_t, ) = const
     (L0, G0, D0, ) = x0
     (L, G, D, ) = x
-        
+
     return [
         - (mu1 + lambd) * L,
                   lambd * L + (alpha - alpha * ((L + G + D)/N_t) - mu2) * G,
@@ -35,7 +35,7 @@ def pool_model_2pools_resource_lag(t, x, param, x0, const):
     (lambd, alpha, N_t, ) = const
     (L0, G0, R0, ) = x0
     (L, G, R ) = x
-        
+
     return [
         - lambd * R * L,
           lambd * R * L + alpha * R * G,
@@ -59,7 +59,7 @@ def pool_model_tempstress(t, x, param, x0, const):
     (lambd, mu1, alpha, mu2, beta, N_t, Gamma, delta, shift_cnd,) = const
     (L0, G0, D0, ) = x0
     (L, G, D, ) = x
-    
+
     gamma = temp_stress(t, Gamma, delta, shift_cnd)
     return [
         - (mu1 + lambd) * L +                                             gamma * G,
@@ -79,7 +79,7 @@ def pool_model_resource_comp(t, x, param, x0, const):
     (lambd_A, alpha_A, lambd_B, alpha_B, N_t, ) = const
     (L_A0, G_A0, L_B0, G_B0, R0, ) = x0
     (L_A, G_A, L_B, G_B, R, ) = x
-        
+
     return [
         - lambd_A * R * L_A,
           lambd_A * R * L_A + alpha_A * R * G_A,
@@ -92,7 +92,7 @@ def pool_model_2sp_comp_toxin1(t, x, param, x0, const):
     (lambd_A, alpha_A, lambd_B, alpha_B, N_t, k, mu,) = const
     (L_A0, G_A0, L_B0, G_B0, R0, T0) = x0
     (L_A, G_A, L_B, G_B, R, T) = x
-        
+
     return [
         - lambd_A * R * L_A,
[…]

### 2025-08-30 pool_paper_plotting_code/examples/main_paper.py
simplify and fix imports

@@ -4,7 +4,7 @@ import numpy as np
 import os
 import sys
 import matplotlib.pyplot as plt
-sys.path.append(f"{os.getcwd()}")
+sys.path.append(os.getcwd())
 sys.path.append(f"{os.getcwd()}/pool_paper_plotting_code")
 from digital_twin import *
 

### 2025-08-30 pool_paper_plotting_code/examples/pool_paper_spatial.py

@@ -1,11 +1,14 @@
+# fmt: off
 #!/usr/bin/env python3
 
 import matplotlib.pyplot as plt
 import numpy as np
 from scipy.integrate import solve_ivp
 
-import os, sys
+import os
+import sys
 sys.path.append(os.getcwd())
+sys.path.append(f"{os.getcwd()}/pool_paper_plotting_code")
 from digital_twin import *
 
 

### 2025-08-30 pool_paper_plotting_code/examples/pool_paper_spatial.py
fix whitespace

@@ -52,7 +52,7 @@ if __name__ == "__main__":
     constsp = [.008, .005, 4., 5., .1, .1, Nt] # (lambd1, lambd2, alph1, alph2, chi, muI, Nt, )
     times, observabl = model_solve(pool_model_spatial_limit, np.linspace(0, 10, 100), [], x0sp, constsp,
                              obs_func=observable_2pool_2species) 
-    
+
     fig, ax = plt.subplots()
     labels = ['Species 1', 'Species 2', 'Sp 1 + Sp 2']
     for j, obs in enumerate(observabl):
@@ -63,5 +63,5 @@ if __name__ == "__main__":
     #ax.set_yscale('log')
     ax.set_xlim(0., 10.)
     ax.legend()
-    plt.savefig(f'pool_model_spatial.pdf', bbox_inches='tight')
-    plt.close(fig) 
+    plt.savefig('pool_model_spatial.pdf', bbox_inches='tight')
+    plt.close(fig)

### 2025-08-30 pool_paper_plotting_code/examples/main_paper.py
change save path: directly save to paper dir

@@ -211,7 +211,7 @@ if __name__ == "__main__":
     ax.set_xlim(0., 9.0)
     ax.set_ylim(3.3e0, 1.1e4)
     ax.legend()
-    plt.savefig('out/pdfs/pool_model_3pools_resource.pdf', bbox_inches='tight')
+    plt.savefig('paper/Figures/pool_model_3pools_resource.pdf', bbox_inches='tight')
     plt.close(fig)
 
 #################################### Temperature shift ##########################################
@@ -244,7 +244,7 @@ if __name__ == "__main__":
     ax[0].legend()
     ax[0].text(8., 9.9, r'(a)')
     ax[1].text(8., 1.5, r'(b)')
-    plt.savefig('out/pdfs/pool_model_3pools_resource_tempshift.pdf', bbox_inches='tight')
+    plt.savefig('paper/Figures/pool_model_3pools_resource_tempshift.pdf', bbox_inches='tight')
     plt.close(fig)
 
     fig, ax = plt.subplots(1, 1, figsize=figsize_default)
@@ -257,7 +257,7 @@ if __name__ == "__main__":
     ax.set_ylim(6e0, 1.1e4)
     ax.legend()
     ax.text(8., 9.9, r'(a)')
-    plt.savefig('out/pdfs/pool_model_3pools_resource_tempshift1.pdf', bbox_inches='tight')
+    plt.savefig('paper/Figures/pool_model_3pools_resource_tempshift1.pdf', bbox_inches='tight')
     plt.close(fig)
 
     fig, ax = plt.subplots(1, 1, figsize=figsize_default_small)
@@ -267,7 +267,7 @@ if __name__ == "__main__":
     fig, ax = set_labels(fig, ax, r'Time, $t$', r'Rate, $\gamma$')
     ax.set_xlim(0.7, 8.5)
     ax.text(8., 1.5, r'(b)')
-    plt.savefig('out/pdfs/pool_model_3pools_resource_tempshift2.pdf', bbox_inches='tight')
+    plt.savefig('paper/Figures/pool_model_3pools_resource_tempshift2.pdf', bbox_inches='tight')
     plt.close(fig)
 
 #################################### Resource competition ################################################3
@@ -290,7 +290,7 @@ if __name__ == "__main__":
     ax.set_xlim(0., 8.0)
[…]

### 2025-08-30 cr_pool/__init__.py
add source code for abm simulation

@@ -1,0 +1,2 @@
+from .simulation import *
+from .analysis import *

### 2025-08-30 cr_pool/simulation.py

@@ -1,0 +1,427 @@
+import os
+import json
+from typing import Any
+import pandas as pd
+from pathlib import Path
+from .cr_pool import *
+import multiprocessing as mp
+import numpy as np
+import scipy as sp
+from types import SimpleNamespace
+import matplotlib.pyplot as plt
+import tqdm
+import matplotlib
+import gc
+
+
+def get_last_output_path(name="pool_model") -> Path:
+    return Path("out") / name / sorted(os.listdir(Path("out") / name))[-1]
+
+
+def get_simulation_settings(output_path) -> tuple[Any, Any, Any]:
+    f_domain = open(output_path / "domain.json")
+    f_initial_cells = open(output_path / "initial_cells.json")
+    f_meta_params = open(output_path / "meta_params.json")
+
+    domain = json.load(f_domain, object_hook=lambda d: SimpleNamespace(**d))
+    initial_cells = json.load(
+        f_initial_cells, object_hook=lambda d: SimpleNamespace(**d)
+    )
+    meta_params = json.load(f_meta_params, object_hook=lambda d: SimpleNamespace(**d))
+    return domain, initial_cells, meta_params
+
+
+def _combine_batches(run_directory):
+    # Opens all batches in a given directory and stores
+    # them in one unified big list
+    combined_batch = []
+    for batch_file in os.listdir(run_directory):
+        f = open(run_directory / batch_file)
+        b = json.load(f)["data"]
+        combined_batch.extend(b)
+    return combined_batch
+
+
+def _convert_entries(df, element_path):
+    if element_path == "cell_storage":
+        df["identifier.Division"] = df["identifier.Division"].apply(lambda x: tuple(x))
+        df["element.identifier.Division"] = df["element.identifier.Division"].apply(
+            lambda x: tuple(x)
+        )
+        df["element.cell.mechanics.pos"] = df["element.cell.mechanics.pos"].apply(
+            lambda x: np.array(x)
+        )
+        df["element.cell.mechanics.vel"] = df["element.cell.mechanics.vel"].apply(
+            lambda x: np.array(x)
+        )
+        df["element.cell.interactionextracellulargradient"] = df[
+            "element.cell.interactionextracellulargradient"
+        ].apply(lambda x: np.array(x))
+
+    if element_path == "voxel_storage":
[…]

### 2025-08-30 src/bacterial_properties.rs

@@ -1,0 +1,419 @@
+use cellular_raza::concepts::reactions_old::*;
+use cellular_raza::prelude::*;
+
+use nalgebra::Vector2;
+use num::Zero;
+use serde::{Deserialize, Serialize};
+
+use pyo3::prelude::*;
+
+pub const NUMBER_OF_REACTION_COMPONENTS: usize = 2;
+pub type ReactionVector = nalgebra::SVector<f64, NUMBER_OF_REACTION_COMPONENTS>;
+
+#[derive(CellAgent, Clone, Debug, Deserialize, Serialize)]
+#[pyclass(get_all, set_all)]
+pub struct Bacteria {
+    #[Mechanics]
+    pub mechanics: NewtonDamped2D,
+    #[Cycle]
+    pub cycle: BacteriaCycle,
+    #[Interaction]
+    pub cellular_reactions: BacteriaReactions,
+    #[ExtracellularGradient]
+    pub interactionextracellulargradient: GradientSensing,
+}
+
+impl CellularReactions<f64, ReactionVector> for Bacteria {
+    fn get_intracellular(&self) -> f64 {
+        self.cellular_reactions.get_intracellular()
+    }
+
+    fn set_intracellular(&mut self, concentration_vector: f64) {
+        self.cellular_reactions
+            .set_intracellular(concentration_vector)
+    }
+
+    fn calculate_intra_and_extracellular_reaction_increment(
+        &self,
+        internal_concentration_vector: &f64,
+        external_concentration_vector: &ReactionVector,
+    ) -> Result<(f64, ReactionVector), CalcError> {
+        self.cellular_reactions
+            .calculate_intra_and_extracellular_reaction_increment(
+                internal_concentration_vector,
+                external_concentration_vector,
+            )
+    }
+}
+
+#[derive(Clone, Debug, Deserialize, Serialize)]
+#[pyclass(get_all, set_all)]
+pub struct BacteriaTemplate {
+    pub mechanics: Py<NewtonDamped2D>,
+    pub cycle: Py<BacteriaCycle>,
+    pub cellular_reactions: Py<BacteriaReactions>,
+    pub interactionextracellulargradient: Py<GradientSensing>,
+}
+
+fn bacteria_default_volume() -> f64 {
+    std::f64::consts::PI * 1.5_f64.powi(2)
+}
+
+fn bacteria_default_newton_damped() -> NewtonDamped2D {
+    NewtonDamped2D::new(
+        [0.0; 2],                         // pos
[…]

### 2025-08-30 src/lib.rs

@@ -1,0 +1,25 @@
+mod bacterial_properties;
+mod simulation;
+
+use bacterial_properties::*;
+use simulation::*;
+
+use pyo3::prelude::*;
+
+#[pymodule]
+fn cr_pool(m: &Bound<'_, PyModule>) -> PyResult<()> {
+    m.add_function(wrap_pyfunction!(run_simulation, m)?)?;
+    m.add_function(wrap_pyfunction!(generate_cells, m)?)?;
+
+    m.add_class::<Bacteria>()?;
+    m.add_class::<BacteriaTemplate>()?;
+    m.add_class::<Species>()?;
+    m.add_class::<cellular_raza::prelude::NewtonDamped2D>()?;
+    m.add_class::<BacteriaCycle>()?;
+    m.add_class::<BacteriaReactions>()?;
+
+    m.add_class::<MetaParams>()?;
+    m.add_class::<Domain>()?;
+
+    Ok(())
+}

### 2025-08-30 src/simulation.rs

@@ -1,0 +1,308 @@
+use cellular_raza::building_blocks::cartesian_cuboid_n_old::*;
+use cellular_raza::prelude::cpu_os_threads::*;
+
+use num::Zero;
+
+use rand::{Rng, SeedableRng};
+use rand_chacha::ChaCha8Rng;
+
+use serde::{Deserialize, Serialize};
+
+use pyo3::prelude::*;
+
+use crate::bacterial_properties::*;
+
+#[derive(Clone, Debug, Deserialize, Serialize)]
+#[pyclass(get_all, set_all)]
+pub struct Domain {
+    pub diffusion_constants: [f64; NUMBER_OF_REACTION_COMPONENTS],
+    pub initial_concentrations: [f64; NUMBER_OF_REACTION_COMPONENTS],
+
+    pub size: f64,
+    pub n_voxels: Option<usize>,
+}
+
+#[pymethods]
+impl Domain {
+    fn __repr__(&self) -> String {
+        format!("{self:#?}")
+    }
+
+    #[new]
+    #[pyo3(signature = (
+        diffusion_constants=[10.0, 5.0],
+        initial_concentrations=[2.0, 0.0],
+        size=200.0,
+        n_voxels=None
+    ))]
+    fn new(
+        diffusion_constants: [f64; NUMBER_OF_REACTION_COMPONENTS],
+        initial_concentrations: [f64; NUMBER_OF_REACTION_COMPONENTS],
+        size: f64,
+        n_voxels: Option<usize>,
+    ) -> Self {
+        Self {
+            diffusion_constants,
+            initial_concentrations,
+            size,
+            n_voxels,
+        }
+    }
+
+    /* #[staticmethod]
+    fn default() -> Self {
+        Self {
+            diffusion_constants: [2.0, 0.2],
+            initial_concentrations: [1.0, 0.0],
+
+            size: 250.0,
+            n_voxels: None,
+        }
+    }*/
+}
+
+#[derive(Clone, Debug, Deserialize, Serialize)]
+#[pyclass(get_all, set_all)]
+pub struct MetaParams {
+    // General Settings
+    pub random_seed: u64,
+
+    // SIMULATION FLOW SETTINGS
+    pub n_times: usize,
+    pub dt: f64,
+    pub t_start: f64,
+    pub save_interval: usize,
+    pub n_threads: usize,
+    pub show_progressbar: bool,
+    pub save_path: String,
+    pub save_add_date: bool,
+}
+
[…]

### 2025-08-30 cr_pool/simulation.py
add docstrings for functions

@@ -1,3 +1,30 @@
+"""
+
+.. list-table:: Recurring Arguments
+    :header-rows: 1
+    :widths: 15 10 75
+
+    * - Name
+      - Type
+      - Description
+    * - **output_path**
+      - :class:`Path`
+      - Path of the stored results.
+    * - **iteration**
+      - :class:`int`
+      - Iteration count at which to obtain results.
+    * - **element_path**
+      - :class:`str`
+      - Identifier which should be \"cell_storage\" or \"voxel_storage\"
+    * - **threads**
+      - :class:`int`
+      - Number of threads to use simultaneously.
+    * - **iteration**
+      - :class:`int`
+      - Iteration count at which to obtain results.
+
+"""
+
 import os
 import json
 from typing import Any
@@ -15,10 +42,27 @@ import gc
 
 
 def get_last_output_path(name="pool_model") -> Path:
+    """
+    Returns the path of the last numerical result.
+
+    Args:
+        name(str): Name of the subfolder in which to look for.
+    Returns:
+        Path: Storage Path of the last simulation result
+    """
     return Path("out") / name / sorted(os.listdir(Path("out") / name))[-1]
 
 
 def get_simulation_settings(output_path) -> tuple[Any, Any, Any]:
+    """
+    Obtain simulation settings for a given output path.
+
+    Args:
+        output_path(Path): Path of the stored results.
+    Returns:
+        (Any, Any, Any): A tuple containing json results of the domain, initial cells and meta
+        parameters.
+    """
     f_domain = open(output_path / "domain.json")
     f_initial_cells = open(output_path / "initial_cells.json")
     f_meta_params = open(output_path / "meta_params.json")
@@ -101,6 +145,16 @@ def _convert_entries(df, element_path):
 def get_elements_at_iter(
     output_path: Path, iteration, element_path="cell_storage"
 ) -> pd.DataFrame:
+    """
+    Helper function to obtain information about cells or the domain at a given iteration point.
+
+    Args:
+        output_path(Path): Path of the stored results.
+        iteration(int): Iteration count at which to obtain results.
+        element_path(str): Identifier which should be \"cell_storage\" or \"voxel_storage\"
+    Returns:
+        pd.DataFrame: DataFrame containing all information
+    """
     dir = Path(output_path) / element_path / "json"
     run_directory = None
     for x in os.listdir(dir):
@@ -116,6 +170,15 @@ def get_elements_at_iter(
 
 
[…]

### 2025-08-30 docs/source/conf.py
add documentation code with sphinx

@@ -1,0 +1,58 @@
+# Configuration file for the Sphinx documentation builder.
+#
+# For the full list of built-in configuration values, see the documentation:
+# https://www.sphinx-doc.org/en/master/usage/configuration.html
+
+# -- Project information -----------------------------------------------------
+# https://www.sphinx-doc.org/en/master/usage/configuration.html#project-information
+
+import os
+import sys
+
+sys.path.append(os.path.abspath("../../"))
+import cr_pool
+
+project = "cr_pool"
+copyright = "2024, Jonas Pleyer"
+author = "Jonas Pleyer"
+release = "0.2.0"
+
+# -- General configuration ---------------------------------------------------
+# https://www.sphinx-doc.org/en/master/usage/configuration.html#general-configuration
+
+extensions = [
+    "myst_parser",
+    "sphinx.ext.napoleon",
+    "sphinx.ext.duration",
+    "sphinx.ext.doctest",
+    "sphinx.ext.autodoc",
+    "sphinx.ext.autosummary",
+    "sphinx.ext.mathjax",
+    "sphinx.ext.autosectionlabel",
+    "sphinxcontrib.bibtex",
+    "sphinx_subfigure",
+]
+
+source_suffix = {
+    ".rst": "restructuredtext",
+    ".txt": "markdown",
+    ".md": "markdown",
+}
+
+autosummary_generate = True
+add_module_names = False
+autodoc_typehints_format = "short"
+
+templates_path = ["_templates"]
+exclude_patterns = []
+
+
+# -- Options for HTML output -------------------------------------------------
+# https://www.sphinx-doc.org/en/master/usage/configuration.html#options-for-html-output
+
+html_theme = "sphinx_rtd_theme"
+html_static_path = ["_static"]
+
+bibtex_bibfiles = ["references.bib"]
+bibtex_default_style = "unsrt"
+bibtex_reference_style = "label"

### 2025-09-02 pool_paper_plotting_code/examples/main_paper.py
f-string for plots naming

@@ -395,7 +395,7 @@ if __name__ == "__main__":
         fig, ax = set_labels(fig, ax, r'Time, $t$', r'$N / N_t$')
         ax.text(*text_coord[j], text[j])
         ax.set_ylim(0.1, y_max_lim[j])
-        plt.savefig('paper/Figures/pool_model_1pool_inhib{j+1}.pdf', bbox_inches='tight')
+        plt.savefig(f'paper/Figures/pool_model_1pool_inhib{j+1}.pdf', bbox_inches='tight')
         plt.close(fig)
 
 
@@ -448,7 +448,7 @@ if __name__ == "__main__":
         ax.set_ylim(0.0, ymax_lim[k])
         ax.text(*text_coord[k], text[k])
         ax.legend(loc='upper left')
-        plt.savefig('paper/Figures/pool_model_2sp_1pool_coop{k+1}.pdf', bbox_inches='tight')
+        plt.savefig(f'paper/Figures/pool_model_2sp_1pool_coop{k+1}.pdf', bbox_inches='tight')
         plt.close(fig)
 
 

### 2025-09-03 cr_pool/analysis.py
abstract analysis functions from jupyter notebook into separate module

@@ -1,0 +1,166 @@
+import numpy as np
+import multiprocessing as mp
+import pandas as pd
+import tqdm
+
+from .simulation import (
+    get_all_iterations,
+    get_elements_at_iter,
+    calculate_entropy,
+)
+
+# Define units for time
+UNIT_HOUR = 60
+UNIT_DAY = 24 * UNIT_HOUR
+
+
+def print_cell_properties(output_path):
+    # Cellular Properties and their names in dataframe
+    iter_0_particles = get_elements_at_iter(output_path, 0)
+    print("Name".ljust(62), "Type")
+    print("⎯" * 80)
+    for ty, col in zip(
+        iter_0_particles.dtypes,
+        iter_0_particles.columns,
+    ):
+        print(
+            "{}".format(col).ljust(62),
+            "{}".format(ty),
+        )
+
+
+def print_domain_properties(output_path):
+    # Voxel Properties and their names in dataframe
+    iter_0_voxels = get_elements_at_iter(
+        output_path,
+        0,
+        element_path="voxel_storage",
+    )
+    print("Name".ljust(62), "Type")
+    print("⎯" * 80)
+    for ty, col in zip(
+        iter_0_voxels.dtypes,
+        iter_0_voxels.columns,
+    ):
+        print(
+            "{}".format(col).ljust(62),
+            "{}".format(ty),
+        )
+
+
+def analyze_voxel_data(output_path, iteration, meta_params):
+    # Calculate total extracellular concentrations
+    entry = get_elements_at_iter(
+        output_path,
+        iteration,
+        element_path="voxel_storage",
+    )
+    mi = np.array([x for x in entry["element.voxel.min"]])
+    ma = np.array([x for x in entry["element.voxel.max"]])
+    diff = np.abs(ma - mi)
+    conc = np.array([c for c in entry["element.voxel.extracellular_concentrations"]])
+    volume = np.prod(diff, axis=1)
+    conc_total = np.sum(volume[:, None] * conc, axis=0)
+
+    return {
+        "iteration": iteration,
+        "time": iteration * meta_params.dt / UNIT_DAY,
+        "nutrients_conc_total": conc_total[0],
+        "inhib_total": conc_total[1],
+    }
+
+
+def _analyze_voxel_data_helper(args):
+    return analyze_voxel_data(*args)
+
+
+def analyze_cell_data(output_path, iteration, meta_params):
+    entry = get_elements_at_iter(output_path, iteration)
+
+    # Calculate number of cells for species
[…]

### 2025-09-03 cr_pool/ode_models.py
include ODE models and solver functions

@@ -1,0 +1,196 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import scipy as sp
+from dataclasses import dataclass
+
+UNIT_MINUTE = 1
+UNIT_HOUR = 60 * UNIT_MINUTE
+UNIT_DAY = 24 * UNIT_HOUR
+
+
+def observable_2pool_2species(out):
+    output = out.y
+    return np.array(
+        [
+            output[0] + output[1],  # NA
+            output[2] + output[3],  # NB
+            output[0] + output[1] + output[2] + output[3],  # NA + NB
+        ]
+    )
+
+
+@dataclass
+class ODEParameters:
+    # TODO adjust these parameters
+    lag_phase_transition_rate_1: float = 0.0025
+    lag_phase_transition_rate_2: float = 0.005
+    production_rate_1: float = 0.002
+    production_rate_2: float = 0.002
+    inhibitor_production: float = 0.1
+    inhibition_constant: float = 1e-7
+    total_cell_volume: int = 2e5
+
+    @property
+    def parameters(self):
+        return (
+            self.lag_phase_transition_rate_1,  # lambd1,
+            self.lag_phase_transition_rate_2,  # lambd2,
+            self.production_rate_1,  # alpha1,
+            self.production_rate_2,  # alpha2,
+            self.inhibitor_production,  # kappa,
+            self.inhibition_constant,  # muI,
+            self.total_cell_volume,  # N_t
+        )
+
+
+@dataclass
+class ODEInitialValues:
+    initial_total_volume_lag_phase_1: int = 127.2
+    initial_total_volume_growth_phase_1: int = 0
+    initial_total_volume_lag_phase_2: int = 127.2
+    initial_total_volume_growth_phase_2: int = 0
+    # NEVER TOUCH THIS VALUE (IF YOU WANT TO GET TOTAL_CELL_VOLUME)!
+    initial_nutrients: float = 1.0
+    initial_inhibitor: float = 0
+
+    @property
+    def initial_values(self):
+        return np.array(
+            [
+                self.initial_total_volume_lag_phase_1,
+                self.initial_total_volume_growth_phase_1,
+                self.initial_total_volume_lag_phase_2,
+                self.initial_total_volume_growth_phase_2,
+                self.initial_nutrients,
+                self.initial_inhibitor,
+            ]
+        )
+
+
+@dataclass
[…]

### 2025-09-03 cr_pool/plotting.py
abstract plotting functionality from analysis.ipynb script
- also add templates from polina

@@ -1,0 +1,594 @@
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import numpy as np
+import tqdm
+import multiprocessing as mp
+import gc
+import os
+from pathlib import Path
+
+from .analysis import UNIT_DAY, UNIT_HOUR
+from .ode_models import ODEModel, observable_2pool_2species
+from .plotting_templates import COLORS_ALL, FIGSIZE_DEFAULT
+from .cr_pool import Domain, BacteriaTemplate, MetaParams
+from .simulation import (
+    get_simulation_settings,
+    get_all_iterations,
+    get_elements_at_iter,
+    calculate_spatial_density,
+)
+
+CA = COLORS_ALL["N_A"]
+CB = COLORS_ALL["N_B"]
+CN = COLORS_ALL["N"]
+
+
+def plot_growth_curve(data_cells, output_path):
+    # Growth Curve
+    fig, ax1 = plt.subplots(figsize=FIGSIZE_DEFAULT)
+
+    data_cells.plot(x="time", y="bacteria_count_1", ax=ax1, label="Species 1", color=CA, alpha=0.5)
+    data_cells.plot(x="time", y="bacteria_count_2", ax=ax1, label="Species 2", color=CB, alpha=0.5)
+    data_cells.plot(
+        x="time", y="bacteria_count_total", ax=ax1, label="Combined", color=CN, alpha=0.5
+    )
+    ax1.set_ylabel("Bacteria Count")
+    ax1.set_xlabel("Time [days]")
+
+    ax2 = ax1.twinx()
+    data_cells.plot(x="time", y="bacteria_volume_1", ax=ax2, label="Species 1", color=CA)
+    data_cells.plot(x="time", y="bacteria_volume_2", ax=ax2, label="Species 2", color=CB)
+    data_cells.plot(x="time", y="bacteria_volume_total", ax=ax2, label="Combined", color=CN)
+    ax2.set_ylabel("Combined Bacteria Volume")
+
+    ax1.legend()
+    fig.tight_layout()
+    fig.savefig(f"{output_path}/cell_growth.png")
+    fig.savefig(f"{output_path}/cell_growth.pdf")
+    plt.show(fig)
+
+
+def plot_nutrients(data_voxels, output_path):
+    # Nutrients
+    fig, ax = plt.subplots(figsize=FIGSIZE_DEFAULT)
+    data_voxels.plot(
[…]

### 2025-09-03 cr_pool/plotting_templates.py

@@ -1,0 +1,38 @@
+import matplotlib.pyplot as plt
+
+COLORS_ALL = {
+    "R": "#808080",
+    "N_A": "#D06062",
+    "N_B": "#4E89B1",
+    "N": "#7E57A5",
+    "T": "#99582A",
+    "T_A": "#c79758",
+    "N_lambd_1e-2_mu_0": "#E2B100",
+    "N_lambd_1e-3_mu_0": "#386641",
+    "N_lambd_1e-3_mu_0_5": "#0982A4",
+    "N_wo_tempshift": "#679E48",
+    "N_tempshift_10": "#ED733E",
+    "N_tempshift_10_5_15": "#C3568A",
+}
+
+FIGSIZE_DEFAULT = (6.5, 4.0)
+FIGSIZE_DEFAULT_SMALL = (6.5, 2.0)
+FIGSIZE_DEFAULT_2SUBPL = (13, 4)
+
+
+def load_style():
+    plt.rcParams["figure.dpi"] = 400
+    plt.rcParams["font.family"] = "serif"
+    plt.rc("text", usetex=True)
+    plt.rcParams["text.latex.preamble"] = r"\usepackage{bm} \usepackage{amsmath}"
+
+    plt.rcParams["legend.fontsize"] = 15.0
+    plt.rcParams["legend.framealpha"] = 0.0
+    plt.rcParams["legend.handlelength"] = 1.8
+    plt.rcParams["axes.prop_cycle"] = plt.cycler(linewidth=[2.5])
+    plt.rcParams["axes.labelsize"] = 15
+    plt.rcParams["xtick.labelsize"] = 13
+    plt.rcParams["ytick.labelsize"] = 13
+    plt.rcParams["font.size"] = 15
+    plt.rc("xtick", labelsize=13)
+    plt.rc("ytick", labelsize=13)

### 2025-09-03 cr_pool/simulation.py
add optional prefix

@@ -41,7 +41,7 @@ import matplotlib
 import gc
 
 
-def get_last_output_path(name="pool_model") -> Path:
+def get_last_output_path(name="pool_model", prefix="out") -> Path:
     """
     Returns the path of the last numerical result.
 
@@ -50,7 +50,7 @@ def get_last_output_path(name="pool_model") -> Path:
     Returns:
         Path: Storage Path of the last simulation result
     """
-    return Path("out") / name / sorted(os.listdir(Path("out") / name))[-1]
+    return Path(prefix) / name / sorted(os.listdir(Path(prefix) / name))[-1]
 
 
 def get_simulation_settings(output_path) -> tuple[Any, Any, Any]:

### 2025-09-03 cr_pool/simulation.py
cleanup imports

@@ -35,10 +35,7 @@ import multiprocessing as mp
 import numpy as np
 import scipy as sp
 from types import SimpleNamespace
-import matplotlib.pyplot as plt
 import tqdm
-import matplotlib
-import gc
 
 
 def get_last_output_path(name="pool_model", prefix="out") -> Path:

### 2025-09-03 cr_pool/simulation.py
formatting and cleanup

@@ -65,9 +65,7 @@ def get_simulation_settings(output_path) -> tuple[Any, Any, Any]:
     f_meta_params = open(output_path / "meta_params.json")
 
     domain = json.load(f_domain, object_hook=lambda d: SimpleNamespace(**d))
-    initial_cells = json.load(
-        f_initial_cells, object_hook=lambda d: SimpleNamespace(**d)
-    )
+    initial_cells = json.load(f_initial_cells, object_hook=lambda d: SimpleNamespace(**d))
     meta_params = json.load(f_meta_params, object_hook=lambda d: SimpleNamespace(**d))
     return domain, initial_cells, meta_params
 
@@ -103,28 +101,24 @@ def _convert_entries(df, element_path):
         df["element.index"] = df["element.index"].apply(lambda x: np.array(x))
         df["element.voxel.min"] = df["element.voxel.min"].apply(lambda x: np.array(x))
         df["element.voxel.max"] = df["element.voxel.max"].apply(lambda x: np.array(x))
-        df["element.voxel.middle"] = df["element.voxel.middle"].apply(
-            lambda x: np.array(x)
-        )
+        df["element.voxel.middle"] = df["element.voxel.middle"].apply(lambda x: np.array(x))
         df["element.voxel.dx"] = df["element.voxel.dx"].apply(lambda x: np.array(x))
-        df["element.voxel.index"] = df["element.voxel.index"].apply(
-            lambda x: np.array(x)
-        )
+        df["element.voxel.index"] = df["element.voxel.index"].apply(lambda x: np.array(x))
         df["element.voxel.extracellular_concentrations"] = df[
             "element.voxel.extracellular_concentrations"
         ].apply(lambda x: np.array(x))
         df["element.voxel.extracellular_gradient"] = df[
             "element.voxel.extracellular_gradient"
         ].apply(lambda x: np.array(x))
-        df["element.voxel.diffusion_constant"] = df[
-            "element.voxel.diffusion_constant"
-        ].apply(lambda x: np.array(x))
+        df["element.voxel.diffusion_constant"] = df["element.voxel.diffusion_constant"].apply(
+            lambda x: np.array(x)
+        )
         df["element.voxel.production_rate"] = df["element.voxel.production_rate"].apply(
             lambda x: np.array(x)
         )
-        df["element.voxel.degradation_rate"] = df[
[…]

### 2025-09-03 cr_pool/plotting.py
plot under correct name

@@ -56,8 +56,8 @@ def plot_nutrients(data_voxels, output_path):
     )
     ax.set_ylabel("Total Nutrients")
     ax.set_xlabel("Time [days]")
-    fig.savefig(f"{output_path}/cell_growth.png")
-    fig.savefig(f"{output_path}/cell_growth.pdf")
+    fig.savefig(f"{output_path}/nutrients.png")
+    fig.savefig(f"{output_path}/nutrients.pdf")
 
 
 def abm_to_ode(

### 2025-09-03 cr_pool/__init__.py
import all modules

@@ -1,2 +1,6 @@
-from .simulation import *
 from .analysis import *
+from .cr_pool import *
+from .ode_models import *
+from .plotting import *
+from .plotting_templates import *
+from .simulation import *

### 2025-09-03 scripts/run_sim.py
script to run simulations

@@ -1,0 +1,111 @@
+import numpy as np
+import scipy as sp
+
+import cr_pool as crp
+
+
+def calculate_lattice_points(x_min, x_max, n_agents):
+    # Calculate the lowest number possible such that
+    # we can create a almost even lattice based on that
+    m = np.ceil(np.sqrt(n_agents)) + 1
+    dx = (x_max - x_min) / m
+
+    x, y = np.mgrid[x_min + dx : x_max : dx, x_min + dx : x_max : dx]
+    if m % 2 == 0:
+        z = np.vstack([x.reshape(-1), y.reshape(-1)]).T
+    else:
+        for i in range(len(y)):
+            y[i] = np.roll(y[i], -(i % 2))
+        z = np.vstack([x.reshape(-1), y.reshape(-1)]).T
+    return z
+
+
+def calculate_index_distributions(n_cells_1, n_cells_2, n_positions, homogenous=True):
+    if homogenous:
+        indices = np.arange(n_positions)
+
+        ind_even = indices[::2]
+        ind_uneven = indices[1::2]
+
+        ind1_even = ind_even[:n_cells_1]
+        ind1_uneven = ind_uneven[: n_cells_1 - len(ind1_even)]
+
+        ind1 = np.sort(np.hstack([ind1_even, ind1_uneven]))
+        ind2 = np.sort(np.setdiff1d(indices, ind1))[:n_cells_2]
+
+        return ind1, ind2
+    else:
+        indices = np.arange(n_positions)
+        ind1 = indices[:n_cells_1]
+        ind2 = indices[n_cells_1:]
+        return ind1, ind2
+
+
+def generate_cells(n_cells_1, n_cells_2, domain, randomness=0.0, pad=0.15, seed=0):
+    """
+    n_cells_1: int
+    n_cells_2: int
+    uniformity: float
+        Floating point number between 0.0 and 1.0
+    """
+    # Fix numpy random seed
+    rng = np.random.default_rng(seed)
+
+    # Get the domain size
+    d_min = pad * domain.size
+    d_max = (1 - pad) * domain.size
+    r = np.clip(randomness, 0, 1)
+
+    positions_random = d_min + rng.random((n_cells_1 + n_cells_2, 2)) * (d_max - d_min)
+
+    positions_lattice = calculate_lattice_points(
+        0.0, domain.size, n_cells_1 + n_cells_2
+    )
+    positions = positions_random * r + (1 - r) * positions_lattice
[…]

### 2025-09-03 pool_paper_plotting_code/examples/pool_paper_spatial.py
move to new place

@@ -1,67 +1,0 @@
-# fmt: off
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import numpy as np
-from scipy.integrate import solve_ivp
-
-import os
-import sys
-sys.path.append(os.getcwd())
-sys.path.append(f"{os.getcwd()}/pool_paper_plotting_code")
-from digital_twin import *
-
-
-def model_solve(model, times, param, x0, const, obs_func=None):
-    model_output = model_ODE_solution(model, times, param, x0, const)
-    if obs_func is None:
-        observables = model_output
-    else:
-        observables = np.array(obs_func(model_output))
-    return times, observables
-
-
-# The model solution for given ODEs
-def model_ODE_solution(model, t, param, x0, const):
-    sol_model = solve_ivp(model, [0, t[-1]], x0, method='LSODA', t_eval=t, args=(param, x0, const), rtol=2e-4)
-    return sol_model.y
-
-def pool_model_spatial_limit(t, x, temp, x0, const):
-    (L1, G1, L2, G2, R, I,) = x
-    (lambd1, lambd2, alph1, alph2, chi, muI, N_t, ) = const
-    res = 1 - (L1+L2+G1+G2)/N_t
-    alpha2_inhib = alph2/(1 + muI*I)
-    return [
-        - lambd1 * R * L1,
-          lambd1 * R * L1 + alph1 * R * G1,
-        - lambd2 * R * L2,
-          lambd2 * R * L2 + alpha2_inhib * R * G2,
-        - (alph1/N_t) * R * G1 - (alpha2_inhib/  N_t) * R * G2,
-          chi * G1
-    ]
-
-def observable_2pool_2species(output):
-    return [output[0]+output[1], # NA
-            output[2]+output[3], # NB
-            output[0]+output[1]+output[2]+output[3]] # NA + NB
-
-if __name__ == "__main__":
-    # 9. Spatial Limitation:
-    Nt = 1e4
-    x0sp = [5., 0., 5., 0., 1., 0.]
-    constsp = [.008, .005, 4., 5., .1, .1, Nt] # (lambd1, lambd2, alph1, alph2, chi, muI, Nt, )
-    times, observabl = model_solve(pool_model_spatial_limit, np.linspace(0, 10, 100), [], x0sp, constsp,
-                             obs_func=observable_2pool_2species) 
-
-    fig, ax = plt.subplots()
[…]

### 2025-09-03 scripts/pool_paper_spatial.py

@@ -1,0 +1,164 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import scipy as sp
+from dataclasses import dataclass
+
+
+UNIT_MINUTE = 1
+UNIT_HOUR = 60 * UNIT_MINUTE
+UNIT_DAY = 24 * UNIT_HOUR
+
+
+def observable_2pool_2species(out):
+    output = out.y
+    return np.array(
+        [
+            output[0] + output[1],  # NA
+            output[2] + output[3],  # NB
+            output[0] + output[1] + output[2] + output[3],  # NA + NB
+        ]
+    )
+
+
+@dataclass
+class ODEParameters:
+    # TODO adjust these parameters
+    lag_phase_transition_rate_1: float = 0.0025
+    lag_phase_transition_rate_2: float = 0.005
+    production_rate_1: float = 0.002
+    production_rate_2: float = 0.002
+    inhibitor_production: float = 0.1
+    inhibition_constant: float = 1e-7
+    total_cell_volume: int = 2e5
+
+    @property
+    def parameters(self):
+        return (
+            self.lag_phase_transition_rate_1,  # lambd1,
+            self.lag_phase_transition_rate_2,  # lambd2,
+            self.production_rate_1,  # alpha1,
+            self.production_rate_2,  # alpha2,
+            self.inhibitor_production,  # kappa,
+            self.inhibition_constant,  # muI,
+            self.total_cell_volume,  # N_t
+        )
+
+
+@dataclass
+class ODEInitialValues:
+    initial_total_volume_lag_phase_1: int = 127.2
+    initial_total_volume_growth_phase_1: int = 0
+    initial_total_volume_lag_phase_2: int = 127.2
+    initial_total_volume_growth_phase_2: int = 0
+    # NEVER TOUCH THIS VALUE (IF YOU WANT TO GET TOTAL_CELL_VOLUME)!
+    initial_nutrients: float = 1.0
+    initial_inhibitor: float = 0
+
+    @property
+    def initial_values(self):
+        return np.array(
+            [
+                self.initial_total_volume_lag_phase_1,
+                self.initial_total_volume_growth_phase_1,
+                self.initial_total_volume_lag_phase_2,
+                self.initial_total_volume_growth_phase_2,
+                self.initial_nutrients,
+                self.initial_inhibitor,
+            ]
+        )
+
+
+@dataclass
+class ODEMetaParameters:
[…]

### 2025-09-03 paper/pool_model.tex
add bullet points for introduction
Co-authored-by: NAME <polinagayndrik@gmail.com>

--- 2025-08-29 pool_model.tex
@@
 \section{Introduction}
+
+\paragraph{General Introduction to Bacterial Consortia}
+\begin{itemize}
+    \item Complex system; many players (cell-types, interactions)
+    \item processes on different scales (population, cellular, molecular, gene-level)
+    \item Food Safety, Production
+    \item Healthcare: Drug development
+    \item Environmental Sustainabiliy
+\end{itemize}
 Motivation: general, flexible approach for bacterial systems that allows easily modify the system depending on processes on different scales.
@@
 
-\paragraph{Bacteria Consortia and processes (growth, lag phase, inhibition, etc.)}
+\noindent
+\paragraph{Experimental/Biological evidence/phenomena}
 \begin{itemize}
-    \item 
+    \item Lag Phase, Competition, Cooperation, Limiting Resource
+    \item Comprehensive reviews about competition/cooperation bacterial mechanisms: \cite{ghoul_ecology_2016, stubbendieck_bacterial_2016, hibbing_bacterial_2010, west_social_2007}
+    \item Spatial Effects: adhesion, migration, chemotaxis, diffusion
+    \item Quorum-sensing
+    \item Biofilm formation \cite{Jin2020}
 \end{itemize}
@@
     \item Baranyi-Roberts Model~\cite{baranyi_modeling_1993, baranyi_dynamic_1994}
-    \item modified Gompertz
+    \item modified Gompertz~\cite{Gompertz1825-wi,Gibson1987}
     \item Buchanan et al.: three-phase linear model for lag phase, exponential growth phase, and stationary phase~\cite{buchanan_when_1997}
@@
 \end{itemize}
-%
-%
-Comprehensive reviews about competition/cooperation bacterial mechanisms: \cite{ghoul_ecology_2016, stubbendieck_bacterial_2016, hibbing_bacterial_2010, west_social_2007}\\
-
-\noindent
-\paragraph{Spatial effects in cosortia (adhesion, migration, chemotaxis)}
-\begin{itemize}
-    \item 
-\end{itemize}
 
@@
 \paragraph{Models for spatial effects; PDEs, \acp{abm}, others}
+\begin{itemize}
+    \item BSim \cite{Gorochowski2012,Dang2020}
+    \item Bacterial Branching patterns \cite{Kawasaki1997,Matsushita1998}
+\end{itemize}
+
+\noindent
+\paragraph{What is the problem/what is the question? How do we address this?}
+
 %

### 2025-09-03 paper/pool_model.tex
write first part of introduction

@@
 
-\paragraph{General Introduction to Bacterial Consortia}
-\begin{itemize}
-    \item Complex system; many players (cell-types, interactions)
-    \item processes on different scales (population, cellular, molecular, gene-level)
-    \item Food Safety, Production
-    \item Healthcare: Drug development
-    \item Environmental Sustainabiliy
-\end{itemize}
-Motivation: general, flexible approach for bacterial systems that allows easily modify the system depending on processes on different scales.
-Allows us to describe the processes within populations using different bacterial states as well as interactions between different populations using ODEs.
+% \paragraph{General Introduction to Bacterial Consortia}
+% \begin{itemize}
+%     \item Complex system; many players (cell-types, interactions)
+%     \item processes on different scales (population, cellular, molecular, gene-level)
+%     \item Food Safety, Production
+%     \item Healthcare: Drug development
+%     \item Environmental Sustainabiliy
+% \end{itemize}
+% Motivation: general, flexible approach for bacterial systems that allows easily modify the system depending on processes on different scales.
+% Allows us to describe the processes within populations using different bacterial states as well as interactions between different populations using ODEs.
+
+Bacterial consortia are highly complex systems which consist of multiple distinct cell types.
+The underlying processes are interconnected with each other and occur on all scales, ranging from
+population-based dynamics of colony growth over molecular processes such as metabolism to gene
+expression.
+Bacterial communities are ubiquitous and even able to survive under extreme conditions, making them
+virtually omnipresent around any living system.
+They play an especially important role within the context of food safety and production
+being a key contributor to the spoilage of food products as well as fermentation processes.
+Furthermore effects such as biofilm formation can hinder drug treatment by shielding the bacterial
+colony from antibiotic substances or immune response of the organism.
+The natural bacetrial processes are a crucial part of our planets ecosystem, being responsible for
+tasks such as nutrient cycling, decomposition of organic matter and oxygen production.
+It is of great interest to understand the dynamics of these various processes.
 

### 2025-09-03 paper/pool_model.tex
write next two sections of introduction

@@
 
+% \noindent
+% \paragraph{Experimental/Biological evidence/phenomena}
+% \begin{itemize}
+%     \item Lag Phase, Competition, Cooperation, Limiting Resource
+%     \item Comprehensive reviews about competition/cooperation bacterial mechanisms: \cite{ghoul_ecology_2016, stubbendieck_bacterial_2016, hibbing_bacterial_2010, west_social_2007}
+%     \item Spatial Effects: adhesion, migration, chemotaxis, diffusion
+%     \item Quorum-sensing
+%     \item Biofilm formation \cite{Jin2020}
+% \end{itemize}
+
 \noindent
-\paragraph{Experimental/Biological evidence/phenomena}
-\begin{itemize}
-    \item Lag Phase, Competition, Cooperation, Limiting Resource
-    \item Comprehensive reviews about competition/cooperation bacterial mechanisms: \cite{ghoul_ecology_2016, stubbendieck_bacterial_2016, hibbing_bacterial_2010, west_social_2007}
-    \item Spatial Effects: adhesion, migration, chemotaxis, diffusion
-    \item Quorum-sensing
-    \item Biofilm formation \cite{Jin2020}
-\end{itemize}
+In order to achieve a comprehensive description of the target system, it is important to incorporate
+all relevant biologial phenomena arising in these bacterial systems.
+During their reproductive process, bacteria undergo multiple phases, starting in a lag phase which
+is necessary to adapt to new environments and initially impedes the overall growth.
+Additionally, bacteria are able to cooperate e.g. via complementary metabolic processes or inhibit
+each other e.g. by means of waste production as a metabilic byproduct.
+A limited supply of nutrients will also affect the overall reproductive capacity of the colony.
+Bacteria have thus devised various survival strategies to ensure a sufficient access to nutrients.
+Effects on a single-cell level such as adhesion, migration and chemotaxis can lead to complex
+phenomena on a population-based level which can aid in the goal of survival.
+Cell signaling e.g. in the form of quorum-sensing can lead to collective emergent phenomena such as
+biofilm formation which play an important role for many mature microbial community.
+
+% \noindent
+% \paragraph{ODE models (predictive microbiology, existing, overview, most popular)}
+% Existing models in Predictive Microbiology (Primary Models)~\cite{perez-rodriguez_predictive_2012}:
+% \begin{itemize}
+%     \item An overview of primary models used in food science can be found in~\cite{van_boekel_kinetic_2008}.
+%     \item Baranyi-Roberts Model~\cite{baranyi_modeling_1993, baranyi_dynamic_1994}
+%     \item modified Gompertz~\cite{Gompertz1825-wi,Gibson1987}
+%     \item Buchanan et al.: three-phase linear model for lag phase, exponential growth phase, and stationary phase~\cite{buchanan_when_1997}
+%     \item McKellar two-compartment model with bacteria in growing and non-growing state~\cite{mckellar_heterogeneous_1997}
+%     \item Inactivation models
+% \end{itemize}
+% %
+% Secondary models:
+% \begin{itemize}
+%     \item Polynomial models 
+%     \item Square Root-Type Models, e.g., Ratkowski for temperature~\cite{ratkowsky_relationship_1982}
+%     \item The Cardinal Parameter Model~\cite{zwietering_decision_1992}
+% \end{itemize}
 
 \noindent
-\paragraph{ODE models (predictive microbiology, existing, overview, most popular)}
-Existing models in Predictive Microbiology (Primary Models)~\cite{perez-rodriguez_predictive_2012}:
-\begin{itemize}
-    \item An overview of primary models used in food science can be found in~\cite{van_boekel_kinetic_2008}.
-    \item Baranyi-Roberts Model~\cite{baranyi_modeling_1993, baranyi_dynamic_1994}
-    \item modified Gompertz~\cite{Gompertz1825-wi,Gibson1987}
-    \item Buchanan et al.: three-phase linear model for lag phase, exponential growth phase, and stationary phase~\cite{buchanan_when_1997}
-    \item McKellar two-compartment model with bacteria in growing and non-growing state~\cite{mckellar_heterogeneous_1997}
-    \item Inactivation models
-\end{itemize}
-%
-Secondary models:
-\begin{itemize}
-    \item Polynomial models 
-    \item Square Root-Type Models, e.g., Ratkowski for temperature~\cite{ratkowsky_relationship_1982}
-    \item The Cardinal Parameter Model~\cite{zwietering_decision_1992}
-\end{itemize}
+The first step in the construction of any mathematical model is to determine the required level of
+detail with respect to the description scale and consequently which effects to include.
+The bacterial growth curves let us describe the dynamics of the overall abundance of various
+species.
+The field of predictive microbiology has developed a variety of ODE-based models to capture
+bacterial growth dynamics.
+Various \acp{ode} \textit{primary models} describe the bacterial growth curve under contant
+environmental conditions.
+Widely used examples include the modified Gompertz model~\cite{Gompertz1825-wi}, Baranyi-Roberts
+model~\cite{baranyi_modeling_1993, baranyi_dynamic_1994} and MCKellar two-compartment model~\cite{buchanan_when_1997} which distinguishes between growing and non-growing bacteria.
+To take into account changing environmental conditions, so called \textit{secondary models}
+formulate the model parameters as a function of external quantities such as temperature, pH or
+humidity.
+These can be simple polynomial models, square root-type models~\cite{ratkowsky_relationship_1982} or
+more complex relations such as the cardinal parameter model~\cite{zwietering_decision_1992}.
+The advantage of the \ac{ode} formulation lies in the predictive power and simplicitiy of usage.
+However due to their coarse-grained nature, they lack the ability to describe heterogeneity and
+other spatial effects.
 

### 2025-09-03 paper/pool_model.tex
preliminary finish of introduction

@@
 \newacro{ode}[ODE]{Ordinary Differential Equation}
+\newacro{pde}[PDE]{Partial Differential Equation}
 \newacro{glv}[gLV]{generalized Lotka-Volterra}
@@
 
+% \paragraph{Models for spatial effects; PDEs, \acp{abm}, others}
+% \begin{itemize}
+%     \item BSim \cite{Gorochowski2012,Dang2020}
+%     \item Bacterial Branching patterns \cite{Kawasaki1997,Matsushita1998}
+% \end{itemize}
+
 \noindent
-\paragraph{Models for spatial effects; PDEs, \acp{abm}, others}
-\begin{itemize}
-    \item BSim \cite{Gorochowski2012,Dang2020}
-    \item Bacterial Branching patterns \cite{Kawasaki1997,Matsushita1998}
-\end{itemize}
+% To incorporate spatial heterogeneity or other effects other modeling approaches have been devised.
+% \acp{pde} follow a similar pattern as \acp{ode} but also distribute the described variables through
+% space.
+When dealing with spatial effects such as heterogeneity, we are required to use more extensive
+modeling formulations such as \acp{pde} or \acp{abm}.
+The former represent an intermediate step between the population-based \ac{ode} models and the
+individually-based \acp{abm} and can be effective, especially when dealing with large crowds of
+cells that can be approximated as a continuum.
+On the other hand, \acp{abm} excel at describing the individual behaviour of cells which can then
+lead to more complex collective phenomena.
+Tools such as BSim~\cite{Gorochowski2012,Dang2020} or \texttt{cellular\_raza}~\cite{Pleyer2025} are
+capable to describe large amounts of cells, performing various intracellular and extracellular
+processes.
+The latter has been used to adapt \ac{pde}-based models into an \ac{abm} formulation which describe
+bacterial branching patterns~\cite{Kawasaki1997,Matsushita1998}.
+
+% \paragraph{What is the problem/what is the question? How do we address this?}
 
 \noindent
-\paragraph{What is the problem/what is the question? How do we address this?}
-
-%
-%
-%
+In this work, we introduce a new \ac{ode}-based modeling framework which describes different
+bacterial species, their metabolic states and external quantities such as nutrients as individual
+pools.
+The dynamics of bacterial growth can be understood in terms of transitions between the different
+bacterial states i.e. pools.
+Our model also allows us to capture interactions between these bacterials states and the external
+microenvironment.
+Compared against the aforementioned methods, this approach turns out to be less complex with general
+applicability, low number of parameters and greater explainability for the underlying biological
+reality.
+To illustrate the generality of the model, multiple examples are reported.
+We analyze the spacial limitations of an interacting two-species model and show that it can be fully
+understood on an individual-based level in terms of single-cell processes.
+Furthermore, we asses how spatial heterogeneity can impact the shape of the growth curves.
+
 % ================================================================================================

### 2025-09-03 paper/pool_model.tex
Add citations to introduction and correct some mistakes

@@
 being a key contributor to the spoilage of food products as well as fermentation processes.
-Furthermore effects such as biofilm formation can hinder drug treatment by shielding the bacterial
+Furthermore, effects such as biofilm formation can hinder drug treatment by shielding the bacterial
 colony from antibiotic substances or immune response of the organism.
-The natural bacetrial processes are a crucial part of our planets ecosystem, being responsible for
+The natural bacterial processes are a crucial part of our planet's ecosystem, being responsible for
 tasks such as nutrient cycling, decomposition of organic matter and oxygen production.
@@
 In order to achieve a comprehensive description of the target system, it is important to incorporate
-all relevant biologial phenomena arising in these bacterial systems.
-During their reproductive process, bacteria undergo multiple phases, starting in a lag phase which
-is necessary to adapt to new environments and initially impedes the overall growth.
+all relevant biological phenomena arising in these bacterial systems.
+During their reproductive process, bacteria undergo multiple phases~\cite{buchanan_when_1997}, starting in a lag phase which
+is necessary to adapt to new environments and initially impedes the overall growth~\cite{rolfe_lag_2012, monod_growth_1949}.
 Additionally, bacteria are able to cooperate e.g. via complementary metabolic processes or inhibit
-each other e.g. by means of waste production as a metabilic byproduct.
-A limited supply of nutrients will also affect the overall reproductive capacity of the colony.
-Bacteria have thus devised various survival strategies to ensure a sufficient access to nutrients.
-Effects on a single-cell level such as adhesion, migration and chemotaxis can lead to complex
+each other e.g. by means of waste production as a metabolic byproduct~\cite{west_social_2007, rainey_evolution_2003, hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
+A limited supply of nutrients will also affect the overall reproductive capacity of the colony~\cite{monod_growth_1949}.
+Bacteria have thus devised various survival strategies to ensure sufficient access to nutrients.
+Effects on a single-cell level such as adhesion~\cite{htuson_bacteriasurface_2013}, migration~\cite{decoene_microscopic_2011} and chemotaxis can lead to complex
 phenomena on a population-based level which can aid in the goal of survival.
 Cell signaling e.g. in the form of quorum-sensing can lead to collective emergent phenomena such as
-biofilm formation which play an important role for many mature microbial community.
+biofilm formation which play an important role for many mature microbial communities~\cite{ng_bacterial_2009}.
 
@@
 The field of predictive microbiology has developed a variety of ODE-based models to capture
-bacterial growth dynamics.
-Various \acp{ode} \textit{primary models} describe the bacterial growth curve under contant
+bacterial growth dynamics~\cite{perez-rodriguez_predictive_2012}.
+Various \acp{ode} \textit{primary models} describe the bacterial growth curve under constant
 environmental conditions.
 Widely used examples include the modified Gompertz model~\cite{Gompertz1825-wi}, Baranyi-Roberts
-model~\cite{baranyi_modeling_1993, baranyi_dynamic_1994} and MCKellar two-compartment model~\cite{buchanan_when_1997} which distinguishes between growing and non-growing bacteria.
-To take into account changing environmental conditions, so called \textit{secondary models}
+model~\cite{baranyi_modeling_1993, baranyi_dynamic_1994} and McKellar two-compartment model~\cite{buchanan_when_1997} which distinguishes between growing and non-growing bacteria.
+To take into account changing environmental conditions, so-called \textit{secondary models}
 formulate the model parameters as a function of external quantities such as temperature, pH or
@@
 more complex relations such as the cardinal parameter model~\cite{zwietering_decision_1992}.
-The advantage of the \ac{ode} formulation lies in the predictive power and simplicitiy of usage.
-However due to their coarse-grained nature, they lack the ability to describe heterogeneity and
+The advantage of the \ac{ode} formulation lies in the predictive power and simplicity of usage.
+However, due to their coarse-grained nature, they lack the ability to describe heterogeneity and
 other spatial effects.
@@
 bacterial states i.e. pools.
-Our model also allows us to capture interactions between these bacterials states and the external
+Our model also allows us to capture interactions between these bacterial states and the external
 microenvironment.
@@
 To illustrate the generality of the model, multiple examples are reported.
-We analyze the spacial limitations of an interacting two-species model and show that it can be fully
+We analyze the spatial limitations of an interacting two-species model and show that it can be fully
 understood on an individual-based level in terms of single-cell processes.
-Furthermore, we asses how spatial heterogeneity can impact the shape of the growth curves.
+Furthermore, we assess how spatial heterogeneity can impact the shape of the growth curves.
 
@@
 Then $[\alpha_0]=V/s$, $[N_t]=1/V$, and $[\alpha]=1/s$.
-For homogeneous densities it is equivalent to consider the total amount, i.e., multiplying $L$ and $G$ by the volume.
+For homogenous densities it is equivalent to consider the total amount, i.e., multiplying $L$ and $G$ by the volume.
 In this case $[L]=[G]=[N_t]=1$.
@@
     \dot{L_A} &= -\lambda_A R L_A\\
-    \dot{G_A} = \lambda_A R L_A + \alpha_A R G_A\\
-    \dot{L_B} = -\lambda_B R L_B\\
-    \dot{G_B} = \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B
+    \dot{G_A} &= \lambda_A R L_A + \alpha_A R G_A\\
+    \dot{L_B} &= -\lambda_B R L_B\\
+    \dot{G_B} &= \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B
     \label{eq:spatial_limit_F}
@@
 \end{align}
-where $\delta(x-x_\text{cell})$ is the dirac delta-distribution.
+where $\delta(x-x_\text{cell})$ is the Dirac delta-distribution.
 Note that $[R_e]=[R_e](x)$ explicitly depends on the spatial coordinate but we choose to omit its notation.
@@
 \end{align}
-Furthermore, we let cells produce an inhibitor with production rate $b$ which inhibits the cells ability to uptake and thus use the resource by a factor of $1+\nu_I [I]$
+Furthermore, we let cells produce an inhibitor with production rate $b$ which inhibits the cell ability to uptake and thus use the resource by a factor of $1+\nu_I [I]$
 \begin{align}
@@
 Cell growth is already being described the system of coupled \acp{ode} given in equations~\eqref{eq:spatial-full-system-simplified}.
-Thus we introduce an additional rule that splits apart the cell whenever a certain volume threshold is reached.
+Thus, we introduce an additional rule that splits apart the cell whenever a certain volume threshold is reached.
 The volume of the cell is divided equally between the new and old cell and both are assigned new random positions close to each other such that their radii do touch but not overlap.
@@
 \end{figure}
-In addition to modeling division events, the cells have are subject to the lag-phase which delays their ability to proliferate.
+In addition to modeling division events, the cells are subject to the lag-phase which delays their ability to proliferate.
 We model this by assigning a transition rate to each cell.
@@
 The \ac{abm} model is more fundamental than the \ac{ode} model in the sense that it incorporates more biological aspects and maps them closer to realistic biological mechanisms.
-We are thus able to derive the \ac{ode} pool model from its \ac{abm} version in the homogenenous case (see supplementary material~\ref{subsec:supplement-deriving-pool-odes}).
+We are thus able to derive the \ac{ode} pool model from its \ac{abm} version in the homogenous case (see supplementary material~\ref{subsec:supplement-deriving-pool-odes}).
 By doing so, we find a mapping between parameters from the \ac{abm} to parameters used in the \ac{ode} description.
@@
 We can use these mappings to numerically compare both approaches.
-Table~\ref{tab:spatial_limit_init} shows a comparison of paramters of two simulations for the \ac{ode} and \ac{abm} model which have been mapped onto each other.
+Table~\ref{tab:spatial_limit_init} shows a comparison of parameters of two simulations for the \ac{ode} and \ac{abm} model which have been mapped onto each other.
 Figure~\ref{fig:spatial-snapshots} shows individual snapshots of different 
@@
         The background shows the resource concentration $[R]$.
-        Darker colours indicate that less nutrients are present.
+        Darker colors indicate that less nutrients are present.
     }

### 2025-09-03 cr_pool/analysis.py
only calculate entropy optionally

@@ -74,7 +74,7 @@ def _analyze_voxel_data_helper(args):
     return analyze_voxel_data(*args)
 
 
-def analyze_cell_data(output_path, iteration, meta_params):
+def analyze_cell_data(output_path, iteration, meta_params, get_entropy=False):
     entry = get_elements_at_iter(output_path, iteration)
 
     # Calculate number of cells for species
@@ -98,11 +98,8 @@ def analyze_cell_data(output_path, iteration, meta_params):
     volume_2 = np.sum(volume[entry["element.cell.cellular_reactions.species"] == "S2"])
     total_cell_volume = np.sum(volume)
 
-    # Calculate the entropy at this step
-    entropy = calculate_entropy(output_path, iteration)
-
     # Return combined results
-    return {
+    res = {
         "iteration": iteration,
         "time": iteration * meta_params.dt / UNIT_DAY,
         "bacteria_count_1": bacteria_count_1,
@@ -114,8 +111,13 @@ def analyze_cell_data(output_path, iteration, meta_params):
         "bacteria_volume_1": volume_1,
         "bacteria_volume_2": volume_2,
         "bacteria_volume_total": total_cell_volume,
-        "entropy": entropy,
     }
+
+    # Calculate the entropy at this step
+    if get_entropy:
+        res["entropy"] = calculate_entropy(output_path, iteration)
+
+    return res
 
 
 def _analyze_cell_data_helper(args):

### 2025-09-03 src/simulation.rs
simplify mappings

@@ -122,7 +122,7 @@ impl MetaParams {
 }
 
 fn save_initial_state(
-    path: &std::path::PathBuf,
+    path: std::path::PathBuf,
     cells: &Vec<BacteriaTemplate>,
     domain: &Domain,
     meta_params: &MetaParams,
@@ -143,19 +143,19 @@ fn save_initial_state(
     let writer_domain = std::io::BufWriter::new(f_domain);
     let writer_meta_params = std::io::BufWriter::new(f_meta_params);
 
-    serde_json::to_writer_pretty(writer_cells, &cells).or_else(|e| {
-        Err(pyo3::PyErr::new::<pyo3::exceptions::PyValueError, _>(
-            format!("serde_json error in writing simulation settings to file: {e}"),
+    serde_json::to_writer_pretty(writer_cells, &cells).map_err(|e| {
+        pyo3::PyErr::new::<pyo3::exceptions::PyValueError, _>(format!(
+            "serde_json error in writing simulation settings to file: {e}"
         ))
     })?;
-    serde_json::to_writer_pretty(writer_domain, &domain).or_else(|e| {
-        Err(pyo3::PyErr::new::<pyo3::exceptions::PyValueError, _>(
-            format!("serde_json error in writing simulation settings to file: {e}"),
+    serde_json::to_writer_pretty(writer_domain, &domain).map_err(|e| {
+        pyo3::PyErr::new::<pyo3::exceptions::PyValueError, _>(format!(
+            "serde_json error in writing simulation settings to file: {e}"
         ))
     })?;
-    serde_json::to_writer_pretty(writer_meta_params, &meta_params).or_else(|e| {
-        Err(pyo3::PyErr::new::<pyo3::exceptions::PyValueError, _>(
-            format!("serde_json error in writing simulation settings to file: {e}"),
+    serde_json::to_writer_pretty(writer_meta_params, &meta_params).map_err(|e| {
+        pyo3::PyErr::new::<pyo3::exceptions::PyValueError, _>(format!(
+            "serde_json error in writing simulation settings to file: {e}"
         ))
     })?;
 
@@ -235,9 +235,9 @@ pub fn run_simulation(
             )
         }
     }
-    .or_else(|e| {
-        Err(pyo3::PyErr::new::<pyo3::exceptions::PyValueError, _>(
-            format!("Rust error in construction of simulation domain: {e}"),
+    .map_err(|e| {
[…]

### 2025-09-03 cr_pool/analysis.py
only load entropy optionally
- this speeds up loading of the results tremendously

@@ -48,7 +48,7 @@ def print_domain_properties(output_path):
         )
 
 
-def analyze_voxel_data(output_path, iteration, meta_params):
+def analyze_voxel_data(output_path, iteration, meta_params, get_entropy):
     # Calculate total extracellular concentrations
     entry = get_elements_at_iter(
         output_path,
@@ -74,7 +74,7 @@ def _analyze_voxel_data_helper(args):
     return analyze_voxel_data(*args)
 
 
-def analyze_cell_data(output_path, iteration, meta_params, get_entropy=False):
+def analyze_cell_data(output_path, iteration, meta_params, get_entropy):
     entry = get_elements_at_iter(output_path, iteration)
 
     # Calculate number of cells for species
@@ -124,14 +124,17 @@ def _analyze_cell_data_helper(args):
     return analyze_cell_data(*args)
 
 
-def analyze_all_cell_voxel_data(output_path, meta_params, n_threads=None):
+def analyze_all_cell_voxel_data(output_path, meta_params, n_threads=None, get_entropy=False):
     # Construct pool to process results in parallel
     if n_threads is None:
         n_threads = mp.cpu_count()
     pool = mp.Pool(n_threads)
 
     # Gather args and use pool to get already pre-processed results
-    args = [(output_path, iteration, meta_params) for iteration in get_all_iterations(output_path)]
+    args = [
+        (output_path, iteration, meta_params, get_entropy)
+        for iteration in get_all_iterations(output_path)
+    ]
     print("Loading Cell Data")
     data_cells = (
         pd.DataFrame(

### 2025-09-03 cr_pool/plotting.py
cleanup

@@ -45,7 +45,7 @@ def plot_growth_curve(data_cells, output_path):
     fig.tight_layout()
     fig.savefig(f"{output_path}/cell_growth.png")
     fig.savefig(f"{output_path}/cell_growth.pdf")
-    plt.show(fig)
+    plt.close(fig)
 
 
 def plot_nutrients(data_voxels, output_path):
@@ -150,7 +150,6 @@ def plot_comparisons(data_cells, output_path):
     settings = get_simulation_settings(output_path)
 
     model = abm_to_ode(*settings)
-    print(model)
     res = model.solve_ode_raw()
     obs = observable_2pool_2species(res)
 
@@ -177,7 +176,7 @@ def plot_comparisons(data_cells, output_path):
     fig.savefig(output_path / "abm_ode_comparison.png")
     fig.savefig(output_path / "abm_ode_comparison.pdf")
     fig.savefig(output_path / "abm_ode_comparison.eps")
-    plt.show(fig)
+    plt.close(fig)
 
 
 def plot_lag_phase(data_cells, res, output_path):
@@ -228,7 +227,7 @@ def plot_lag_phase(data_cells, res, output_path):
     fig.tight_layout()
     fig.savefig(f"{output_path}/lag_phase.png")
     fig.savefig(f"{output_path}/lag_phase.pdf")
-    plt.show(fig)
+    plt.close(fig)
 
 
 def plot_entropy(data_cells, output_path):
@@ -245,7 +244,7 @@ def plot_entropy(data_cells, output_path):
     fig.tight_layout()
     fig.savefig(output_path / "entropy.png")
     fig.savefig(output_path / "entropy.pdf")
-    plt.show(fig)
+    plt.close(fig)
 
 
 def _determine_image_save_path(output_path, iteration, fmt="png"):

### 2025-09-03 cr_pool/simulation.py
move generation of cell data to cr_pool

@@ -265,3 +265,86 @@ def calculate_entropy(output_path, iteration) -> np.ndarray:
 
     # y = z1/(z1+z2)
     return np.array([z1, z2])
+
+
+def calculate_lattice_points(x_min, x_max, n_agents):
+    # Calculate the lowest number possible such that
+    # we can create a almost even lattice based on that
+    m = np.ceil(np.sqrt(n_agents)) + 1
+    dx = (x_max - x_min) / m
+
+    x, y = np.mgrid[x_min + dx : x_max : dx, x_min + dx : x_max : dx]
+    if m % 2 == 0:
+        z = np.vstack([x.reshape(-1), y.reshape(-1)]).T
+    else:
+        for i in range(len(y)):
+            y[i] = np.roll(y[i], -(i % 2))
+        z = np.vstack([x.reshape(-1), y.reshape(-1)]).T
+    return z
+
+
+def calculate_index_distributions(n_cells_1, n_cells_2, n_positions, homogenous=True):
+    if homogenous:
+        indices = np.arange(n_positions)
+
+        ind_even = indices[::2]
+        ind_uneven = indices[1::2]
+
+        ind1_even = ind_even[:n_cells_1]
+        ind1_uneven = ind_uneven[: n_cells_1 - len(ind1_even)]
+
+        ind1 = np.sort(np.hstack([ind1_even, ind1_uneven]))
+        ind2 = np.sort(np.setdiff1d(indices, ind1))[:n_cells_2]
+
+        return ind1, ind2
+    else:
+        indices = np.arange(n_positions)
+        ind1 = indices[:n_cells_1]
+        ind2 = indices[n_cells_1:]
+        return ind1, ind2
+
+
+def generate_cells(
+    n_cells_1, n_cells_2, domain, randomness=0.0, pad=0.15, seed=0, homogenous=False
+):
+    """
+    n_cells_1: int
+    n_cells_2: int
+    uniformity: float
+        Floating point number between 0.0 and 1.0
+    """
+    # Fix numpy random seed
+    rng = np.random.default_rng(seed)
+
+    # Get the domain size
+    d_min = pad * domain.size
+    d_max = (1 - pad) * domain.size
+    r = np.clip(randomness, 0, 1)
+
+    positions_random = d_min + rng.random((n_cells_1 + n_cells_2, 2)) * (d_max - d_min)
+
+    positions_lattice = calculate_lattice_points(0.0, domain.size, n_cells_1 + n_cells_2)
[…]

### 2025-09-03 scripts/run_sim.py

@@ -4,91 +4,6 @@ import scipy as sp
 import cr_pool as crp
 
 
-def calculate_lattice_points(x_min, x_max, n_agents):
-    # Calculate the lowest number possible such that
-    # we can create a almost even lattice based on that
-    m = np.ceil(np.sqrt(n_agents)) + 1
-    dx = (x_max - x_min) / m
-
-    x, y = np.mgrid[x_min + dx : x_max : dx, x_min + dx : x_max : dx]
-    if m % 2 == 0:
-        z = np.vstack([x.reshape(-1), y.reshape(-1)]).T
-    else:
-        for i in range(len(y)):
-            y[i] = np.roll(y[i], -(i % 2))
-        z = np.vstack([x.reshape(-1), y.reshape(-1)]).T
-    return z
-
-
-def calculate_index_distributions(n_cells_1, n_cells_2, n_positions, homogenous=True):
-    if homogenous:
-        indices = np.arange(n_positions)
-
-        ind_even = indices[::2]
-        ind_uneven = indices[1::2]
-
-        ind1_even = ind_even[:n_cells_1]
-        ind1_uneven = ind_uneven[: n_cells_1 - len(ind1_even)]
-
-        ind1 = np.sort(np.hstack([ind1_even, ind1_uneven]))
-        ind2 = np.sort(np.setdiff1d(indices, ind1))[:n_cells_2]
-
-        return ind1, ind2
-    else:
-        indices = np.arange(n_positions)
-        ind1 = indices[:n_cells_1]
-        ind2 = indices[n_cells_1:]
-        return ind1, ind2
-
-
-def generate_cells(n_cells_1, n_cells_2, domain, randomness=0.0, pad=0.15, seed=0):
-    """
-    n_cells_1: int
-    n_cells_2: int
-    uniformity: float
-        Floating point number between 0.0 and 1.0
-    """
-    # Fix numpy random seed
-    rng = np.random.default_rng(seed)
-
-    # Get the domain size
-    d_min = pad * domain.size
-    d_max = (1 - pad) * domain.size
-    r = np.clip(randomness, 0, 1)
-
-    positions_random = d_min + rng.random((n_cells_1 + n_cells_2, 2)) * (d_max - d_min)
-
-    positions_lattice = calculate_lattice_points(
-        0.0, domain.size, n_cells_1 + n_cells_2
-    )
-    positions = positions_random * r + (1 - r) * positions_lattice
[…]

### 2025-09-03 paper/pool_model.tex
remove listings from appendix

@@
 \end{table}
-
-\begin{listing}
-\begin{lstlisting}[language=json]
-{
-    "diffusion_constants": [
-        15.0,
-        15.0
-    ],
-    "initial_concentrations": [
-        2.0,
-        0.0
-    ],
-    "size": 1000.0,
-    "n_voxels": null
-}
-\end{lstlisting}
-\caption{
-    The \texttt{domain.json} file specifying initial values of the domain.
-}
-\end{listing}
-%
-\begin{listing}
-\begin{lstlisting}[language=json]
-{
-    "random_seed": 1,
-    "n_times": 20001,
-    "dt": 0.5,
-    "t_start": 0.0,
-    "save_interval": 500,
-    "n_threads": 8,
-    "show_progressbar": true,
-    "save_path": "out/pool_model",
-    "save_add_date": true
-}
-\end{lstlisting}
-\caption{
-    The \texttt{meta\_params.json} file specifying all parameters not directly related to the simulation content.
-    These settings do not represent any model-specific properties but rather generic ones such as time-span or number of threads.
-}
-\end{listing}
-%
-\begin{listing}
-\begin{lstlisting}[language=json]
-{
-    "mechanics": {
-        "pos": [
-        142.85714285714286,
-        142.85714285714286
-        ],
-        "vel": [
-        0.0,
-        0.0
-        ],
-        "damping_constant": 0.5,
-        "mass": 7.704755982928968
-    },
-    "cycle": {
-        "volume_division_threshold": 14.137166941154069,
-        "lag_phase_transition_rate_1": 0.005,
-        "lag_phase_transition_rate_2": 0.0025
-    },
-    "cellular_reactions": {
-        "potential_strength": 0.5,
-        "food_to_volume_conversion": 0.1,
-        "lag_phase_active": true,
-        "species": "S1",
-        "cell_volume": 7.0685834705770345,
-        "uptake_rate": 0.01,
-        "inhibition_production_rate": 0.1,
-        "inhibition_coefficient": 0.1
-    },
-    "interactionextracellulargradient": null
-}
-\end{lstlisting}
-\caption{
-    The first entry of the \texttt{initial\_cells.json} file which defines every individual particle present in the simulation initially.
-}
-\end{listing}
-%
 %

### 2025-09-03 paper/pool_model.tex
write abstract

@@
 \begin{abstract}\noindent
-    \textbf{Notes from Christian}
-    Different empirical and inactivation models applied in the estimation
-    of parameters and modelling of phenomena in food microbiology, gen-
-    eralized Lotka Voltera approach used to describe the interaction of mi-
-    croorganisms, mechanistic models utilized to provide more insight into
-    the underlying dynamic can all be unified as variations of a new model
-    known as three pool model. Compared against the aforementioned
-    methods which are applicable in specific applications with some limi-
-    tations, the present model turns out to be less complex with general
-    applicability, low number of parameters and preferably to explain the
-    mechanism of the phenomenon of interest. To illustrate the generality
-    of the model, three different examples are reported.
+    Bacterial consortia are complex systems which are often comprised of multiple distinct species.
+    Their dynamics on a population-based level often result from interactions that lead to
+    inter-species cooperation or competition.
+    Various empirical and inactivation models have been conceived and applied to problems related to
+    food microbiology.
+    However, the utilized models live on the popluation level, omitting any spatial effects such as
+    heterogeneity or diffusion.
+    In this work we present a novel generalized framework, called "pool model" which uses multiple
+    distinct pools that can hold bacterial states of distinct species or extracellular nutrient
+    abundance.
+    It allows us to describe these systems mechanistically as variations of an underlying, more
+    fundamental model.
+    To illustrate the usefulness of the model, multiple different examples are reported.
+    Finally, we construct an \acl{abm} for a given pool model, showcasing that the resulting model
+    arises naturally from the individual behaviour of many cellular agents.
+    We further apply this \ac{abm} to study the role of spatial effects on the dynamcis of bacterial
+    growth.
+    % \textbf{Notes from Christian}
+    % Different empirical and inactivation models applied in the estimation
+    % of parameters and modelling of phenomena in food microbiology, gen-
+    % eralized Lotka Voltera approach used to describe the interaction of mi-
+    % croorganisms, mechanistic models utilized to provide more insight into
+    % the underlying dynamic can all be unified as variations of a new model
+    % known as three pool model. Compared against the aforementioned
+    % methods which are applicable in specific applications with some limi-
+    % tations, the present model turns out to be less complex with general
+    % applicability, low number of parameters and preferably to explain the
+    % mechanism of the phenomenon of interest. To illustrate the generality
+    % of the model, three different examples are reported.
 \end{abstract}
+
+\noindent
+\textbf{Keywords:}
+Bacterial Consortia, Competition and Cooperation, Mechanistic Modeling, Spatial Effects
 

### 2025-09-04 paper/pool_model.tex
add data availability statement
add TODO to conclusion
fix typo
load onimage.sty for overlays of images
set indent to zero and remove commands

@@
 \usepackage{color}
-\usepackage{ctable} % for \specialrule command
 \usepackage{diagbox}
@@
 \usepackage{newfloat}
+\usepackage{onimage}
 \usepackage{siunitx}
@@
 \usepackage{tabularx}
+\usepackage{tikz}
 \usepackage[dvipsnames]{xcolor}
 \usepackage{xpatch}
+
+% Must be loaded after tikz
+\usepackage{ctable}
 
@@
 \restylefloat{table}
+
+% Remove indent across the whole document
+\setlength\parindent{0pt}
 
@@
 
-\begin{abstract}\noindent
+\begin{abstract}
     Bacterial consortia are complex systems which are often comprised of multiple distinct species.
@@
 
-\noindent
 \textbf{Keywords:}
@@
 
-% \noindent
 % \paragraph{Experimental/Biological evidence/phenomena}
@@
 
-\noindent
 In order to achieve a comprehensive description of the target system, it is important to incorporate
@@
 
-% \noindent
 % \paragraph{ODE models (predictive microbiology, existing, overview, most popular)}
@@
 
-\noindent
 The first step in the construction of any mathematical model is to determine the required level of
@@
 
-\noindent
 % To incorporate spatial heterogeneity or other effects other modeling approaches have been devised.
@@
 
-\noindent
 In this work, we introduce a new \ac{ode}-based modeling framework which describes different
@@
 
-\noindent
 How justified was the construction of the \ac{abm}?
@@
 
-\noindent
 How can we cover more effects of bacteria?
@@
 
-\noindent
 What are limitations of our approach?
@@
 \section{Conclusion}
-\noindent
 \begin{itemize}
@@
         dimensional reduction of existing models)
+    \item Provide code in a pypi package as well (TODO)
 \end{itemize}
@@
 \newpage
+\section*{Data Availability \& Reproducibility}
+
+The data and simulation code which are presented in this study are publicly available in
+\url{https://github.com/polinagaindrik/pool_model_paper}.
+
 \printbibliography

### 2025-09-04 paper/pool_model.tex
remvoe marker
remove todo
adapt formulations of first section

@@
 \subsection{Three-Pool Model}
-Consider for simplicity the system where only one bacteria species is present $N=1$.
-The bacterial growth curves usually consist of different consecutive phases~\cite{buchanan_when_1997}.
-At the beginning no increase in bacterial count is present and so-called lag-phase is observed.
-This phenomenon appears due to adaptation of the microbial community, for example, to the new environmental conditions.
-This so-called preparation to the exponential growth includes synthesis of the needed compounds, initiation of the transcription~\cite{rolfe_lag_2012} and poorly studied physiological and regulatory mechanisms~\cite{monod_growth_1949}.
-After this, the succeeded in adaptation bacteria enter the exponential growth phase, where the process of cell division is directly happening.
-When the nutrition is exhausted by the growing bacteria, the stationary state of the growth curve occurs.
-In this state the cultures accumulate waste products of metabolic activity that leads to cell death of the bacteria.
+We begin our discussion by considering a simplistic system where only one bacteria species is present $N=1$.
+Bacterial growth undergoes different consecutive phases~\cite{buchanan_when_1997}, starting with the so called lag-phase, during which no increase in the bacterial count occurs.
+This phenomenon appears due to adaptations of the microbial community, for example, to the new environmental conditions.
+This so-called preparation before the exponential growth includes synthesis of the needed compounds, initiation of the transcription~\cite{rolfe_lag_2012} and poorly studied physiological and regulatory mechanisms~\cite{monod_growth_1949}.
+After this, the bacteria which have succeeded in this adaptation stage enter the exponential growth phase, where the process of cell division is predominant.
+When the levels of nutrition are exhausted by the growing bacteria, the stationary state of the growth curve occurs.
+In this state the cultures accumulate waste products of metabolic activity that lead to cell death in the bacterial colony.
 At the same time dying cells release the nutrients for survivors to continue growing.
-As a result the balance between the growing and dying processes leads to the constant with time bacterial count~\cite{navarro_llorens_stationary_2010}.
-In the described simplified version of the real-life system three possible bacteria states are mentioned corresponding to which process is happening: adaptation, growth or death.
-Motivated by this we divide a bacterial population into three pools (states)~(\ref{fig:SchematicRep}) defining the fraction of the population in the lag phase $L(t)$, growing and dividing $G(t)$ and undergoing cell death $D(t)$.
+As a result the balance between the growing and dying processes leads to a constant bacterial count~\cite{navarro_llorens_stationary_2010}.
+In the described simplified version of the real-life system three possible bacterial states are mentioned, corresponding to the underlying process: adaptation, growth or death.
+Motivated by this we divide a bacterial population into three pools (states)~(\ref{fig:SchematicRep}), defining the fraction of the population in the lag phase $L(t)$, growing and dividing $G(t)$ and undergoing cell death $D(t)$.
 \begin{figure}[t]
@@
 \end{figure}
-To determine the system of equations describing the system dynamics consider the flows between different states.
+To determine the system of equations which describe the systems dynamics, we consider the flows between the different states.
 The pools $G$ and $D$ are unidirectionally connected with the pool $L$.
-For now, we assume for simplicity that there is no back flow between growth state $G$ and lag-state $L$.
+For now, we assume for that there is no back flow between growth state $G$ and lag-state $L$.
 The suggested model can be cast in the following reactions:
@@
     \dot{D} &= \mu  L + \mu' G- \beta D  
-\end{split}\end{align}.
+\end{split}\end{align}
+with the initial conditions $L(0)=N_0$, $G(0)=0$, and $D(0)=0$, where $N_0$ is the initial bacterial population.
 %
 Based on experimental observations, it was shown that the bacterial growth is strongly influenced by the limiting resource.
-According to the works of Jacques Monod, in the system where all needed nutrients are in excess except for one the total growth is linear to the concentration of this limiting nutrient~\cite{monod_growth_1949}.
-The control of the population density with respect to the available nutrition is implemented using cell communication process called quorum sensing.
-Using release of the extracellular signalling molecules (autoinducer molecules) it is possible to synchronize the gene expression of the whole group~\cite{ng_bacterial_2009}.
-We would not dive into the processes of quorum sensing and try to keep it as simple as possible.
-According to this, we aim to include a resource pool which is used by the growing bacteria using the second order kinetic reaction:
+According to the works of Jacques Monod, in the system where all needed nutrients are in excess except for one, the total growth is linear to the concentration of this limiting nutrient resource~\cite{monod_growth_1949}.
+The control of the population density with respect to the available nutrition is implemented using cell communication processes called quorum sensing.
+By releasing extracellular signalling molecules (autoinducer molecules) it is possible for the bacteria to synchronize the gene expression of the whole group~\cite{ng_bacterial_2009}.
+We will not dive into the processes of quorum sensing and try to keep it as simple as possible.
+We aim to include a resource pool which is consumed by the growing bacteria using the second order kinetic reaction:
 \begin{equation}
@@
 Then $[\alpha_0]=V/s$, $[N_t]=1/V$, and $[\alpha]=1/s$.
+$N_t$ denotes the maximal size of the total bacterial population due to environmental limitations (Figure \ref{fig:3pool_resource_plots}).
 For homogenous densities it is equivalent to consider the total amount, i.e., multiplying $L$ and $G$ by the volume.
 In this case $[L]=[G]=[N_t]=1$.
-For simplicity, we rather make a transition to the dimensionless constant $\alpha:=\alpha_0 N_t$ and dimensionless resource pool $R(t) = R_0 / N_t$ ($0 \leqslant R \leqslant 1$).
+For simplicity, we make a transition to the dimensionless constant $\alpha:=\alpha_0 N_t$ and dimensionless resource pool $R(t) = R_0 / N_t$ ($0 \leqslant R \leqslant 1$).
 Here the full capacity of the resource pool $N_t$ was used as a scaling constant.
@@
 %
-Such model can be simplified by exploiting that limitation of the resource pool also limits the total bacterial count of the system:
+Such a model can be simplified by exploiting that any limitation of the resource pool also leads to a limit of the total bacterial count
 \begin{equation}
@@
         All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\alpha=2.5$, $\mu'=0.5$ and $\beta=5$.
-        The blue curve shows significant dip in behavior of the system where the death of the bacteria is present $\mu = 0.5$.
-        Other curves show the case of no bacteria dying $\mu=0$ and represent different length of lag phases:
-        short $\lambda=10^{-2}$ (dark green), intermediate $\lambda=10^{-3}$ (light green) and long one $\lambda=10^{-4}$ (yellow).
+        The blue curve shows a significant dip in the behavior of the system where the death of the bacteria is present $\mu = 0.5$.
+        Other curves show the case of no bacteria dying $\mu=0$ and represent different lengths of lag phases: short $\lambda=10^{-2}$ (dark green), intermediate $\lambda=10^{-3}$ (light green) and long one $\lambda=10^{-4}$ (yellow).
     }
@@
 The resulting growth curves for the three pool case are presented in Figure~\ref{fig:3pool_resource_plots}.
-Here we can see that the value of the transition constant $\lambda$ determines how long the bacteria is in lag-phase.
-The larger is the rate the shorter is the lag-time.
-Moreover, by including the bacteria flow to the death pool we can diversify the shape of the growth curve allowing the decrease of the bacterial count.
-%
+Here we can see that the value of the transition constant $\lambda$ determines how long the bacteria stays in its lag-phase.
+The larger this rate, the shorter is the lag-time.
+Moreover, by including the flow of living bacteria to the death pool, we can model a greater variety of effects by allowing a decrease in the bacterial count.
 %
@@
 \subsection{Lag-time Estimation}
-Note that one could further simplify this model to a two pool or state model by omitting the pool $D$ and removing cells with rate $\beta$ directly from pool $L$.
+Note that the previously introduced model could be further simplified to a two pool or state model by omitting the pool $D$ and removing cells with rate $\beta$ directly from pool $L$.
 The relevant question here is whether cells in the pool $D$ still contribute to the limitation of the resources by, e.g., occupying space or consuming important nutrients.
@@
 %
-The dynamics of the three pool model can be captures by the coupled system of ordinary differential equations (the two pool model can be achieved by setting $D\equiv 0$ and replacing $\mu$ with $\beta$):
-to be solved with the initial conditions $L(0)=N_0$, $G(0)=0$, and $D(0)=0$, where $N_0$ is the initial bacterial population.
-$N_t$ denotes the maximal size of the total bacterial population due to environmental limitations (Figure \ref{fig:3pool_resource_plots}).
+% The dynamics of the three pool model can be captured by the coupled system of ordinary differential
+The dynamics of the three pool model is captured by equations~\ref{eq:ode_3pool_resource2} which can be used to derive the two pool model by setting $D\equiv 0$ and replacing $\mu$ with $\beta$).
 For $\mu'=0$ the equations for $L$ and $D$ can be readily integrated resulting in:
@@
 \end{align}
-For constant parameters $\mu$, $\lambda$, and $\beta$ the integrals for $L$ and $D$ can be put into the differential equation for $G$.
-However, for time dependent parameters, e.g., for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$.
+For constant parameters $\mu$, $\lambda$, and $\beta$ the integrals for $L$ and $D$ can be inserted into the differential equation for $G$.
+However, in the case of time dependent parameters, e.g., for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$.
 It is interesting to see what type of equations result for the total bacterial population $N=L+G+D$:
@@
 \end{align}
-with the initial conditions $G(0)=0$ and $N(0)=N_0$. 
-These equations have a clear interpretation, are derived from simple principles, and are different from the popular Barayni-Roberts and modified-Gompertz models.
+with initial conditions $G(0)=0$ and $N(0)=N_0$.
+These equations have a clear interpretation, are derived from first principles, and are not identical to the popular Barayni-Roberts and modified-Gompertz models.
 A notable difference is that the growth rate $\alpha$ is time independent.
@@
 It follows immediately that $\lim_{t\to\infty} G(t) = N_t$.
-The equation exhibit a lag phase and an initial drop due to cell death.
+The equation exhibits a lag phase and an initial drop due to cell death.
 To find a substantial initial drop one needs $\beta \mu \gg \alpha \lambda$.
 The lag time $t_L$ can be approximated by the time at which half of the initial population is in the growth phase, i.e., $N_0=2G(t_L)$.
-Assuming $L\ll N_t$ we can ignore the non-linear term in the equation for $G$, which gives rise to:
+Assuming $L\ll N_t$, we can ignore the non-linear term in the equation for $G$, which gives rise to:
 \begin{align}
@@
 \end{align}
-For $\alpha > \mu+\lambda$ $t_L$ can be approximated by:
+For $\alpha > \mu+\lambda$  we can approximate $t_L$ by:
 \begin{align}
@@
 \subsection{Back flow from Growth to Lag pool}
-Due to its life in rather unfavorable conditions bacteria developed different surviving mechanisms, one of which is adapting their metabolism and transitioning into dormant state.
-By definition dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\cite{kaprelyants_dormancy_1993}.
+Bacteria which have lived under rather unfavorable conditions have developed different surviving mechanisms, one of which is adapting their metabolism and transitioning into dormant state.
+Dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\cite{kaprelyants_dormancy_1993}.
 This survival strategy can appear in response to different external stresses, e.g., nutrient deficiency, temperature stress or extreme fluctuations of environment parameters.
 And after "difficult times" bacteria, also called viable but non-culturable (VBNC), can continue cell division~\cite{kell_viability_1998}.
-The models including transition of bacteria to the dormant state were proposed and studied by Bär \etal~\cite{bar_modelling_2002} and Jones and Lennon~\cite{jones_dormancy_2010}.
-%
-We explore now the idea that under certain conditions, as sudden change in the environment, part of the population $G$ will enter a lag-phase, i.e., there is a back-flow from $G$ to $L$.
+Multiple models which include transitions of bacteria to the dormant state were proposed by Bär \etal~\cite{bar_modelling_2002} and Jones with Lennon~\cite{jones_dormancy_2010}.
+%
+We explore now the idea that under certain conditions, a sudden change in the environment can lead to parts of the population $G$ entering a lag-phase.
+Within the framework of our pool model this means, that there is a back-flow from $G$ to $L$.
 \begin{eqnarray}
@@
 \paragraph{Maxwell type of stress - strain relation}
-One of the reasons of the dormant state appearance can be also the temperature shift~\cite{oliver_viable_1995}.
-Moreover, the rapid shift in temperature is associated with the appearance of an additional lag phase~\cite{zwietering_modeling_1994}.
+One of the reasons that bacteria enter the dormant state can be due to a shift in temperature~\cite{oliver_viable_1995}.
+Moreover, a rapid shift in temperature is associated with the appearance of an additional lag phase~\cite{zwietering_modeling_1994}.
 To model it we propose a kind of viscoelastic Maxwell type of stress-strain relation and write down the ordinary differential equation for $\gamma$ to be:
@@
 \section{Interaction between species}
-As it was mentioned above, bacteria usually coexist in large multispecies interacting communities whose growth is vastly determined by the developed survival mechanisms.
-These survival mechanisms mostly aim to secure as much available nutrition which lead to competitive strategies or, if it is advantageous, cooperative ones~\cite{hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
+As mentioned above, bacteria usually coexist in large interacting communities of multiple species whose growth is vastly determined by the developed survival mechanisms.
+These mechanisms mostly aim to secure as much available nutrition which lead to competitive strategies or, if it is advantageous, cooperative ones~\cite{hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
 Further we show that the model can be readily extended to the systems with more than one bacterium and describe more complex relationships between species.
@@
     \begin{center}
-    % TODO make this into 2x2 grid for plots and insert cases for psi from before as well
-    %\includegraphics[width=\columnwidth]{Figures/pool_model_1pool_inhib.pdf}
     \includegraphics[width=0.9\columnwidth]{Figures/pool_model_1pool_inhib1.pdf}\\
@@
 
-\textbf{!Christian extend this!}
-
-%
-
 \subsection{ABM Details} \label{ssec:supplement2}
-
 \subsubsection*{Deriving the Pool Model}

### 2025-09-04 paper/pool_model.tex
reformulate section 3

@@
 \subsection{Resource Competition}
-One on the ways to describe competition between bacteria populations is as a function of limiting resource ratio~\cite{tilman_resource_1977, smith_effects_2002}.
-To describe such a situation we consider for simplicity only one common limiting resource pool.
+One of the ways to describe competition between bacterial populations is as a function of the limiting resource ratio~\cite{tilman_resource_1977, smith_effects_2002}.
+To this end we consider for simplicity only one common limiting resource pool.
 The simplest pool model for bacterial species $A$ and $B$ encompassing lag-phases, growth-phases and a common resource pool is given by (no back-flow to the lag-pool):
@@
 where we used the short notation $\left<LG\right>=(L_A+G_A+L_B+G_B)$.
-These equations have the structure of a \ac{glv} expanded with different internal states for one species.
+These equations have the structure of a \ac{glv} model expanded with different internal states for one species.
 As one can see in Figure~\ref{fig:2pool_resource_2sp} even without direct interaction between species the bacterial growth was inhibited due to the limitation of the resource.
-Here species $A$ has a longer lag-phase than $B$ but the exponential growth of $A$ is stronger than the one that bacteria $B$ experiences.
-As a result, species $A$ was able to suppress species $B$ consuming the larger amount of nutrient and reaching larger bacterial count at the end.
+Here species $A$ has a longer lag-phase than $B$ but the exponential growth of $A$ is stronger compared to which is experienced by bacteria $B$.
+As a result, species $A$ was able to suppress species $B$, consuminl the larger amount of nutrients and reaching a larger bacterial count at the end.
 \begin{figure}
@@
 \end{align}
-
-Equations (\ref{eq:LV_simple}) are basically competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competitive excursion.
-That principle claims that two species in competition for the same resource cannot coexist~\cite{wangersky_lotka-volterra_1978}.
+%
+Equations (\ref{eq:LV_simple}) can be understood as competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competitive excursion.
+That principle claims that two species which compete for the same resource cannot coexist~\cite{wangersky_lotka-volterra_1978}.
 In previous studies it was shown that \ac{glv} equations with one bacterial state can be successfully applied to predict the microbial community dynamics, determine different types of the pairwise interspecies interactions~\cite{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
@@
 \subsection{Interspecies Competition}
-However, except the inhibition due to resource competition, bacteria developed different others active competition mechanisms.
-One of the examples of such mechanisms is a production of the antimicrobial compounds, inhibitors or toxins~\cite{wloch-salamon_effect_2008, chao_structured_1981}.
+Mutual inhibition between different species can not only result from competition for a shared resource for which bacteria have developed various other active competition mechanisms.
+One of the examples of such a mechanisms is the production of antimicrobial compounds, inhibitors or toxins~\cite{wloch-salamon_effect_2008, chao_structured_1981}.
 On the other hand, some bacteria can produce public goods, the compounds that benefit growth of other species.
-In this way, can suppress the non-preferred species by helping their competitors~\cite{hibbing_bacterial_2010}.
+In this way, they can suppress the non-preferred species by helping their competitors~\cite{hibbing_bacterial_2010}.
 %
@@
 \subsubsection{Toxins Production}
-To introduce active competitive mechanisms, we extend our model with one of the species producing the toxic compounds $T$ during its growth:
+One direct form of inhibition is the production of toxins.
+To introduce active competitive mechanisms, we extend our model with one of the species, producing the toxic compounds $T$ during its growth:
 \begin{equation}
@@
 %
-Where we consider two possible models of the toxin production:
+where we consider two possible models of the toxin production
 \begin{align}
@@
 Here $\kappa$ is the toxin production rate.
-The difference between the two toxin production models for resulting solution of the \acp{ode} is presented in Figure~\ref{fig:2pool_2sp_toxin}.
+The difference in the solutions of the given \acp{ode} between the two toxin production models is presented in Figure~\ref{fig:2pool_2sp_toxin}.
 %
@@
 Model $\mathcal{F}^1$ results in toxin concentration $T_B^1(t)=\kappa\int_{t_0}^tG(t')dt'$.
-That means that as long as there are bacteria in the growth phase $G_B>0$ toxin is produced, no matter whether the total bacterial count $N_B$ is still growing.
+That means that as long as there are bacteria in the growth phase $G_B>0$, toxin is produced, no matter whether the total bacterial count $N_B$ is still growing.
 On the other hand, model $\mathcal{F}^2$ results in toxin accumulation in steady state $T_B^{2s}=\lim_{t\to\infty}T_B^2(t)=\kappa(N_t-N_0)$.
@@
 Another special case of the species competition is when the produced toxin does not cause the death of the bacteria but rather inhibit its growth.
-Such situation with 2 bacteria species, limiting resource and produced inhibitor was modeled, e.g., by Lenski and Hattingh~\cite{lenski_coexistence_1986}.
-In our case, we would like to look at rather simplified case reducing the number of variables of the system.
+Such situation with 2 bacteria species, a shared resource and produced inhibitor was modeled, e.g., by Lenski and Hattingh~\cite{lenski_coexistence_1986}.
+In our case, we would like to look at a rather simplified case reducing the number of variables of the system.
 The easiest way to introduce inhibition of species $A$ by species $B$ is via a growth rate of species $A$ depending on the abundance of species $B$.
@@
 One can distinguish two limiting cases: {\it i}) $K_{A/B}\ll N_t$ and {\it ii}) $K_{A/B}\gg N_t$.
-In the first case the inhibition is of minor importance and could be ignored.
-In the latter case one can simplify the equations, given $N_A$ and $N_B$ are non-zero:
+In the first case the inhibition is of minor importance and could be ignored while in the latter case one can simplify the equations, given $N_A$ and $N_B$ are non-zero:
 \begin{align}\begin{split}
@@
 \end{split}\end{align}
-We introduced the rescaled growth rates $\tilde{\alpha}_A=\alpha_A/(N_tK_B)$ and $\tilde{\alpha}_B=\alpha_B/(N_tK_A)$. Note that $[\tilde{\alpha}]=\mathrm{s}^{-1}$.
-Rescaling time with $\tilde{\alpha}_A$ and defining the dimensionless parameter
+We introduced the rescaled growth rates $\tilde{\alpha}_A=\alpha_A/(N_tK_B)$ and $\tilde{\alpha}_B=\alpha_B/(N_tK_A)$.
+Note that $[\tilde{\alpha}]=\mathrm{s}^{-1}$.
+We rescale the time with $\tilde{\alpha}_A$ and define the dimensionless parameter
 \begin{equation}
@@
 \end{equation}
-results in: 
+which results in
 \begin{align}
@@
 \end{align}
-We can separate three cases (which are displayed in Figure~\ref{fig:1pool_2sp_inhibit}):
+We can further separate three cases
 \begin{align}
@@
 \end{align}
+which are displayed in Figure~\ref{fig:1pool_2sp_inhibit}.
 %
@@
 \subsection{Interspecies Cooperation}
-On the other hand, one can also observe some cooperative behaviors between species in case when it facilitates the competition with others.
-It happens if bacteria species as a result of their metabolism produce the compounds that are beneficial for the growth of each other, so-called public goods.
-Quite often it happens between the species that are close relatives or clonal populations and such process is called kin selection \cite{west_social_2007}.
-Furthermore, neighbors can also benefit from a certain behavior without expressing it by itself, so-called 'social cheaters'~\cite{rainey_evolution_2003}.
-%
-%
-The benefit, or mutual activation, of the two species from the presence of the other can be simply introduced by mathematical functions.
-For sake of clarity we omit again the lag pool and write:
+On the other hand, one can also observe some cooperative behaviors between species in cases when it facilitates the competition with others.
+This phenomenon occurs when bacterial metabolism produces compounds that are beneficial for the growth of each other, so-called public goods.
+It is most often observed between species which are close relatives or clonal populations.
+This process is called kin selection~\cite{west_social_2007}.
+Furthermore, neighbors can also benefit from a behavior which does not require any action on an expression level, so-called 'social cheaters'~\cite{rainey_evolution_2003}.
+This mutual activation benefits all involved species and can by easily included in our mathematical framework.
+For the sake of clarity we omit again the lag pool and write
 \begin{align}
@@
     \dot{P}_A &= \mathcal{F}_A(R,G_A)\\
-    \dot{P}_B &= \mathcal{F}_B(R,G_B).
-\end{align}
-$P_A$ and $P_B$ denote the substance produced by species A and B, respectively.
+    \dot{P}_B &= \mathcal{F}_B(R,G_B)
+\end{align}
+where $P_A$ and $P_B$ denote the substance produced by species A and B, respectively.
 The functions $\mathcal{B}_A(P_B,R)$ and $\mathcal{B}_B(P_A,R)$ capture the resource dependent growth.
-Setting $\mathcal{B}_A(P_B,R)=\alpha_A P_BR$, $\mathcal{B}_B(P_A,R)=\alpha_B P_A R$, $\mathcal{F}_A(R,G_A)=\kappa_A RG_A$, $\mathcal{F}_B(R,G_B)=\kappa_B RG_B$, rescaling $G_A$, $G_B$, and $P_B$ with $N_t$, $P_A$ with $\kappa_A/\alpha_A$, time with $\alpha_AN_t^2$, defining $\psi=\kappa_A\alpha_B/(\alpha_A^2N_t)$, and $\phi=\kappa_B/(\alpha_A N_t)$ results in:
+Setting $\mathcal{B}_A(P_B,R)=\alpha_A P_BR$, $\mathcal{B}_B(P_A,R)=\alpha_B P_A R$, $\mathcal{F}_A(R,G_A)=\kappa_A RG_A$, $\mathcal{F}_B(R,G_B)=\kappa_B RG_B$, rescaling $G_A$, $G_B$, and $P_B$ with $N_t$, $P_A$ with $\kappa_A/\alpha_A$, time with $\alpha_AN_t^2$, defining $\psi=\kappa_A\alpha_B/(\alpha_A^2N_t)$, and $\phi=\kappa_B/(\alpha_A N_t)$ results in
 \begin{align}
@@
 The solution of the system~\ref{eq:1pool_2sp_cooper} is presented in Figure~\ref{fig:1pool_2sp_cooper}.
-Here we can see that if $\psi=1$ and $\phi=1$ (see Figure~\ref{fig:1pool_2sp_cooper}a) the production of the public goods of the two species is happening with the same rate as well as the growth of bacteria.
-Increasing the production rate of the public good produced by species $B$ (see Figure~\ref{fig:1pool_2sp_cooper}b) causes increasing the growth of species $A$ that profit from the presence of $B$.
+Here we can see that if $\psi=1$ and $\phi=1$ (see Figure~\ref{fig:1pool_2sp_cooper}a) the production of the public goods of the two species is happening with the same rate as the growth of bacteria.
+Increasing the production rate of the public good (see Figure~\ref{fig:1pool_2sp_cooper}b), leads to an increase in the growth of species $A$ that profit from the presence of $B$.
 On the other hand, choosing $\psi=2$ (see Figure~\ref{fig:1pool_2sp_cooper}c), we increase the profit that species $B$ has from public goods $P_A$.

### 2025-09-04 paper/pool_model.tex
reformulate; add more citations; fix some mistakes

@@
 Cell signaling e.g. in the form of quorum-sensing can lead to collective emergent phenomena such as
-biofilm formation which play an important role for many mature microbial communities~\cite{ng_bacterial_2009}.
+biofilm formation which play an important role for many mature microbial communities~\cite{Jin2020,ng_bacterial_2009}.
 
@@
 environmental conditions.
-Widely used examples include the modified Gompertz model~\cite{Gompertz1825-wi}, Baranyi-Roberts
+Widely used examples include the modified Gompertz model~\cite{Gompertz1825-wi,Gibson1987}, Baranyi-Roberts
 model~\cite{baranyi_modeling_1993, baranyi_dynamic_1994} and McKellar two-compartment model~\cite{buchanan_when_1997} which distinguishes between growing and non-growing bacteria.
@@
 lead to more complex collective phenomena.
-Tools such as BSim~\cite{Gorochowski2012,Dang2020} or \texttt{cellular\_raza}~\cite{Pleyer2025} are
-capable to describe large amounts of cells, performing various intracellular and extracellular
+Tools such as BSim~\cite{Gorochowski2012,Dang2020}, AgentCell~\cite{emonet_agentcell_2005}, BacSim~\cite{kreft_bacsim_1998} or \texttt{cellular\_raza}~\cite{Pleyer2025} are
+capable to describe large amounts of bacteria, performing various intracellular and extracellular
 processes.
@@
 understood on an individual-based level in terms of single-cell processes.
-Furthermore, we assess how spatial heterogeneity can impact the shape of the growth curves.
+Furthermore, we assess how spatial heterogeneity~\cite{mckellar_heterogeneous_1997} can impact the shape of the growth curves.
 
@@
 \subsection{Resource Competition}
-One of the ways to describe competition between bacterial populations is as a function of the limiting resource ratio~\cite{tilman_resource_1977, smith_effects_2002}.
+One of the ways to describe competition between bacterial populations is as a function of the limiting resource ratio~\cite{tilman_resource_1977, smith_effects_2002,ghoul_ecology_2016}.
 To this end we consider for simplicity only one common limiting resource pool.
@@
 nutritients~\cite{htuson_bacteriasurface_2013}.
-Another mechanism is the motility of bacteria, e.g., the flagellates bacteria as \textit{E. coli} or
-\textit{B. subtilis} can experience the chemotactic motion of driven by oxygen
+Another mechanism is the motility of bacteria, e.g., the flagellates \textit{E. coli} or
+\textit{B. subtilis} which are capable of chemotactic motion of driven by oxygen
 concentrations~\cite{decoene_microscopic_2011}.
-The bacterial social interactions such as releasing the toxins or public goods can also
+The bacterial social interactions such as releas of toxins or public goods can also
 significantly affect the spatial structure of the community~\cite{blanchard_bacterial_2015}.
@@
 effects.\\
-In order to construct a model which can encapsulate diffusion of nutrients and heterogeneity of
-the bacteria or nutrients, we have to take into consideration the spatial component of our setup.
-\acp{abm} force researchers to think in terms of cellular processes, thus bridging the gap from a
+Compared to other modeling techniques such as \acp{pde}, \acp{abm}~\cite{nagarajan_agent-based_2022-1} force researchers to think in
+terms of cellular processes, thus bridging the gap from a
 collective (bulk) model to an individual-based (single-cell) description.
-We consider the system consisting from two bacterial species that can be described by two internal
-states: lag and growth states.
+Our goal is to provide an \ac{abm} which can be used to derive the given \ac{ode} model as a
+coarse-grained case.
+In this way, we can show that the bulk-based \ac{ode} model is directly related to the mechanistic
+single-cell behaviour of the bacteria.
+To this end, we consider the system consisting from two bacterial species that can be described by two internal states: lag and growth states.
 Both of these species grow using the same resource pool $R$ but with different growth constants
@@
 custom-built \ac{abm} which has only a few more overall parameters.
-To construct our \ac{abm}, we use \texttt{cellular\_raza}~\cite{Pleyer2025} which was designed to
-provide maximal flexibility for researchers in order to construct minimalist models.
-Without the ability to define and control the precise behavior of our cell-agents, the following
-approaches would not be possible.
+This procedure is a blueprint which can be followed to construct \acp{abm} for similar
+\ac{ode}-based systems.\\
+To construct our \ac{abm}, we use \texttt{cellular\_raza}~\cite{Pleyer2025}.
+It allows us to define our bacterial agents from a clean slate, without having to rely and inherit
+existing work.
+With this flexibility, researchers are able to construct minimal models using only as many parameters which are necessary.
+This ability, to precisely define and control the exact behavior of our cell-agents, enables the following parts.
 
@@
 Further, we want to modify the cellular volume by growing it increasingly with rate $a$.
-Therefore, we need to specify how many resource molecules are turned into one volumetric units $\sigma$ in \unit{\micro\metre\cubed\per\mole}.
+Therefore, we need to specify how many resource molecules are turned into volumetric units $\sigma$ in \unit{\micro\metre\cubed\per\mole}.
 The total amount of resource molecules present inside the cell is given by the current concentration multiplied by the cell-volume $V_c [R_i]$.
@@
 \end{align}
-Furthermore, we let cells produce an inhibitor with production rate $b$ which inhibits the cell ability to uptake and thus use the resource by a factor of $1+\nu_I [I]$
+Furthermore, we let cells produce an inhibitor with production rate $b$ which inhibits the cell's ability to uptake and thus use the resource by a factor of $1+\nu_I [I]$
 \begin{align}
@@
 \end{align}
-Note that we do not need to distinguish between the two species because we can set parameters individually on agent-level.
-This means, we can implement the most general version of the equations and by choosing appropriate parameters, set the species of the respective cell-agent.
+% Note that we do not need to distinguish between the two species because we can set parameters individually on agent-level.
+To distinguish between the two species, it suffices to set parameters for every individual agent accordingly.
+This means, we can implement the most general version of the equations and by choosing appropriate parameters, this defines the species of the respective cell-agent.
 If we combine all equations and consider multiple cells, we obtain the system
@@
 We represent cells as spherical objects in 2D space.
-They can be represented by their position $x_c$ and radius $r_c$ which is of course related to their (2D) volume
-\begin{equation}
-    V_c = \pi r_c^2
-\end{equation}
+They can be represented by their position $x_c$ and radius $r_c$ which is of course related to their (2D) volume $V_c = \pi r_c^2$.
 As the coupled reactions from equations~\eqref{eq:spatial-full-system-simplified} change $V_c$, we also automatically adapt the cell radius accordingly.
@@
 \end{alignat}
-Cells are not sticky and thus only pushing on each other.
-No attractive force is present.
-This means that colonies are expanding as individual cells grow and push each other further apart.
+We assume that the cells are not sticky and consider only a repulsive interaction potential, leaving out any possible attractive forces~\cite{Garrett2008}.
+These mechanisms will result in the overall expansion of the colony as individual cells push each further other apart.
 %
@@
 \subsubsection{Cell Cycle}
-Typically, cell cycles consist of multiple phases.
+Typically the cycle of a cell consist of multiple phases~\cite{Cooper2000_fn}.
 Since our goal is to include only the necessary details of the underlying biological problem, we drastically simplify the description of the cell cycle.
 Cell growth is already being described the system of coupled \acp{ode} given in equations~\eqref{eq:spatial-full-system-simplified}.
-Thus, we introduce an additional rule that splits apart the cell whenever a certain volume threshold is reached.
-The volume of the cell is divided equally between the new and old cell and both are assigned new random positions close to each other such that their radii do touch but not overlap.
+Thus, we introduce an additional rule that performs the division process when a certain volume threshold is reached.
+The volume of the cell is divided equally between the new and old cell and both are assigned new random positions close to each other such that their radii do touch but without any overlap.
 \begin{figure}
     \begin{center}
-    \includegraphics[width=0.49\columnwidth]{Figures/snapshot-2-cells-before-division.png}%
+    \begin{tikzonimage}[width=0.49\columnwidth]
+        {Figures/snapshot-2-cells-before-division.png}%
+        \node at (0.01, 0.99)[anchor=north west]{\textbf{A}};
+    \end{tikzonimage}%
     \hspace{0.01\columnwidth}%
-    \includegraphics[width=0.49\columnwidth]{Figures/snapshot-2-cells-after-division.png}
+    \begin{tikzonimage}[width=0.49\columnwidth]
+        {Figures/snapshot-2-cells-after-division.png}
+        \node at (0.01, 0.99)[anchor=north west]{\textbf{B}};
+    \end{tikzonimage}
     \caption{
@@
 If this boolean is true, we move the cell from the inactive state to the active state where it starts to consume nutrients and grow.
-It is clear that this is exactly identical to the \ac{ode} model.
-%
+The concatenation of many such stochastic processes on a population-based level converges identically to the lag-phase behaviour of the \ac{ode} model.
 %
@@
 \subsection{Comparing \acs{abm} and \acs{ode} Models}
-The \ac{abm} model is more fundamental than the \ac{ode} model in the sense that it incorporates more biological aspects and maps them closer to realistic biological mechanisms.
-We are thus able to derive the \ac{ode} pool model from its \ac{abm} version in the homogenous case (see supplementary material~\ref{subsec:supplement-deriving-pool-odes}).
+The \ac{abm} model should be considered to be the more fundamental model compared to the \ac{ode} model in the sense that it incorporates more biological aspects and maps them closer to the actual behaviour of the bacteria.
+We are thus able to derive the \ac{ode} pool model from its \ac{abm} version via a process of homogenization (see supplementary material~\ref{subsec:supplement-deriving-pool-odes}).
 By doing so, we find a mapping between parameters from the \ac{abm} to parameters used in the \ac{ode} description.
@@
 Table~\ref{tab:spatial_limit_init} shows a comparison of parameters of two simulations for the \ac{ode} and \ac{abm} model which have been mapped onto each other.
-Figure~\ref{fig:spatial-snapshots} shows individual snapshots of different 
+Figure~\ref{fig:spatial-snapshots} shows snapshots of a numerical simulation of the described system.
+The intial bacteria have been ordered on a grid such that only every diagonal neighbor is of the same species, thus resulting in a more homogenous initial distribution.
 %
 \begin{figure}
-    \includegraphics[width=0.49\columnwidth]{Figures/abm-homogenous/snapshot_00006000.png}\hspace{0.01\columnwidth}%
-    \includegraphics[width=0.49\columnwidth]{Figures/abm-homogenous/snapshot_00012000.png}\vspace{0.01\columnwidth}\\
-    \includegraphics[width=0.49\columnwidth]{Figures/abm-homogenous/snapshot_00018000.png}\hspace{0.01\columnwidth}%
-    \includegraphics[width=0.49\columnwidth]{Figures/abm-homogenous/snapshot_00024000.png}%
+    \begin{tikzonimage}[width=0.49\columnwidth]
+        {Figures/abm-homogenous/snapshot_00006000.png}
+        \node at (0.01, 0.99)[anchor=north west]{\textbf{A}};
+    \end{tikzonimage}%
+    \hspace{0.01\columnwidth}%
+    \begin{tikzonimage}[width=0.49\columnwidth]
+        {Figures/abm-homogenous/snapshot_00012000.png}
+        \node at (0.01, 0.99)[anchor=north west]{\textbf{B}};
+    \end{tikzonimage}\\
+    \begin{tikzonimage}[width=0.49\columnwidth]
+        {Figures/abm-homogenous/snapshot_00018000.png}
+        \node at (0.01, 0.99)[anchor=north west, white]{\textbf{C}};
+    \end{tikzonimage}%
+    \hspace{0.01\columnwidth}%
+    \begin{tikzonimage}[width=0.49\columnwidth]
+        {Figures/abm-homogenous/snapshot_00024000.png}
+        \node at (0.01, 0.99)[anchor=north west, white]{\textbf{D}};
+    \end{tikzonimage}%
     \caption{
@@
 %
+Figure~\ref{fig:spatial-growth-curve} shows multiple results for various configurations of the \ac{abm}, differing only in their initial spatial configuration and diffusivity.
 We see that in the highly-diffusive homogenous case of subfigure (a), the pool model is a good fit to the \ac{abm} model.
@@
 Subfigure (d) shows how a randomized homogenous initial distribution can lead to even more notable effects of inhibition.
-We furthermore note that the total bacteria volume of the \ac{abm} model is always smaller than the \ac{ode} model.
+We furthermore note that the total bacteria volume of the \ac{abm} model is always smaller or equal compared to the \ac{ode} model.
 This is expected, since processes such as diffusion and cells growing outward introduce another time component which is simply assumed to be instantanous in the \ac{ode} model.
@@
         \includegraphics[width=\columnwidth]{Figures/abm-homogenous/abm_ode_comparison.png}
-        \subcaption{Homogenous distribution, $D=30.0$\unit{\micro\metre\squared\per\minute}}
+        \subcaption{Hom. $D=30.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}%
@@
         \includegraphics[width=\columnwidth]{Figures/abm-homogenous-low-diffusion/abm_ode_comparison.png}
-        \subcaption{Homogenous distribution, $D=5.0$\unit{\micro\metre\squared\per\minute}}
+        \subcaption{Hom. $D=5.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}\\
@@
         \includegraphics[width=\columnwidth]{Figures/abm-inhomogenous/abm_ode_comparison.png}%
-        \subcaption{Heterogeneous distribution, $D=30.0$\unit{\micro\metre\squared\per\minute}}
+        \subcaption{Het. $D=30.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}%
@@
         \includegraphics[width=\columnwidth]{Figures/abm-random/abm_ode_comparison.png}%
-        \subcaption{Randomized distribution, $D=5.0$\unit{\micro\metre\squared\per\minute}}
+        \subcaption{Rand. $D=5.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}
     \caption{
-        Comparison of the \ac{ode} and \ac{abm} model for different parameter configurations.
+        Comparison of the \ac{ode} and \ac{abm} model for different configurations, showing
+        homogenous (Hom.), heterogenous (Het.) or randomized (Rand.) initial configurations.
+        The diffusion constants are chosen such that $D_R=D_I=D$.
         Snapshots of the spatial configurations of cells are included in the appendix~\ref{fig:supplement-spatial-initial-config-cells}.
-        The diffusion constants are chosen such that $D_R=D_I=D$.
     }
@@
 \end{figure}
+These results show how variations in properties that can not be described by simple \ac{ode} models can lead to large deviations in the growth curves of the species.
+Furthermore, it was shown that the utilized pool model can be derived from first principles by performing a homogenization of the mechanistic bacterial properties.
 %

### 2025-09-04 scripts/paper_spatial_plots.py
add scripts to produce comparison plots from ODE and ABM

@@ -1,0 +1,77 @@
+import cr_pool as crp
+from pathlib import Path
+import argparse
+
+# Diffusion constants, randomness, homogenous
+CONFIGS = [
+    # (30, 0, True),
+    # (5, 0, True),
+    # (30, 0, False),
+    (30, 0.25, True),
+    (30, 0.20, True),
+    (30, 0.15, True),
+    (30, 0.10, True),
+    (5, 0.25, True),
+    (5, 0.20, True),
+    (5, 0.15, True),
+    (5, 0.10, True),
+]
+
+
+def calcualte_results(diffusion_constant, randomness, homogenous):
+    # Domain Settings
+    domain = crp.Domain()
+    domain.size = 1_000
+    # domain.diffusion_constants = [5.0, 5.0]
+    domain.diffusion_constants = [diffusion_constant] * 2
+
+    # Meta Parameters
+    meta_params = crp.MetaParams()
+    meta_params.save_interval = 1_000
+    meta_params.n_times = 40_001
+    meta_params.dt = 0.25
+    meta_params.n_threads = 8
+
+    cells = crp.generate_cells(18, 18, domain, randomness, homogenous=homogenous)
+
+    output_path = crp.run_simulation(
+        cells,
+        domain,
+        meta_params,
+    )
+
+    return output_path
+
+
+if __name__ == "__main__":
+    parser = argparse.ArgumentParser(
+        prog="crp_spatial_plots", description="Plots results needed for the paper publication"
+    )
+    parser.add_argument("paths", type=str, nargs="*")
+    parser.add_argument("--save-snapshots", action="store_true")
+    args = parser.parse_args()
+
+    paths = args.paths
+
+    if len(paths) == 0:
+        paths = []
+        for diffusion_constant, randomness, homogenous in CONFIGS:
+            output_path = calcualte_results(diffusion_constant, randomness, homogenous)
+            paths.append(output_path)
+
+    paths = [Path(p) for p in paths]
+
+    if args.save_snapshots:
+        for output_path in paths:
+            picked_iter = crp.get_all_iterations(output_path)[20]
+            crp.save_snapshot(output_path, picked_iter)
+        # for output_path in paths:
+        #     crp.save_all_snapshots(Path(output_path))
+
+    crp.load_style()
+    for output_path in paths:
+        domain, initial_cells, meta_params = crp.get_simulation_settings(output_path)
[…]

### 2025-09-04 src/bacterial_properties.rs
implement PartialEq for all types

@@ -10,7 +10,7 @@ use pyo3::prelude::*;
 pub const NUMBER_OF_REACTION_COMPONENTS: usize = 2;
 pub type ReactionVector = nalgebra::SVector<f64, NUMBER_OF_REACTION_COMPONENTS>;
 
-#[derive(CellAgent, Clone, Debug, Deserialize, Serialize)]
+#[derive(CellAgent, Clone, Debug, Deserialize, Serialize, PartialEq)]
 #[pyclass(get_all, set_all)]
 pub struct Bacteria {
     #[Mechanics]
@@ -179,7 +179,7 @@ impl Interaction<Vector2<f64>, Vector2<f64>, Vector2<f64>, f64> for BacteriaReac
     }
 }
 
-#[derive(Serialize, Deserialize, Debug, Clone)]
+#[derive(Serialize, Deserialize, Debug, Clone, PartialEq)]
 #[pyclass(get_all, set_all)]
 pub struct BacteriaCycle {
     /// Threshold for the volume when the cell should divide
@@ -276,14 +276,14 @@ impl Cycle<Bacteria> for BacteriaCycle {
     }
 }
 
-#[derive(Clone, Debug, Deserialize, Serialize)]
+#[derive(Clone, Debug, Deserialize, Serialize, PartialEq)]
 #[pyclass]
 pub enum Species {
     S1,
     S2,
 }
 
-#[derive(Serialize, Deserialize, Clone, Debug)]
+#[derive(Serialize, Deserialize, Clone, Debug, PartialEq)]
 #[pyclass(get_all, set_all)]
 pub struct BacteriaReactions {
     pub potential_strength: f64,
@@ -400,7 +400,7 @@ impl Volume for Bacteria {
     }
 }
 
-#[derive(Clone, Debug, Serialize, Deserialize)]
+#[derive(Clone, Debug, Serialize, Deserialize, PartialEq)]
 #[pyclass]
 pub struct GradientSensing;
 

### 2025-09-04 src/simulation.rs

@@ -12,7 +12,7 @@ use pyo3::prelude::*;
 
 use crate::bacterial_properties::*;
 
-#[derive(Clone, Debug, Deserialize, Serialize)]
+#[derive(Clone, Debug, Deserialize, Serialize, PartialEq)]
 #[pyclass(get_all, set_all)]
 pub struct Domain {
     pub diffusion_constants: [f64; NUMBER_OF_REACTION_COMPONENTS],
@@ -61,7 +61,7 @@ impl Domain {
     }*/
 }
 
-#[derive(Clone, Debug, Deserialize, Serialize)]
+#[derive(Clone, Debug, Deserialize, Serialize, PartialEq)]
 #[pyclass(get_all, set_all)]
 pub struct MetaParams {
     // General Settings

### 2025-09-04 src/bacterial_properties.rs
add custom functino to compare BacteriaTemplate with each other

@@ -55,6 +55,25 @@ pub struct BacteriaTemplate {
     pub interactionextracellulargradient: Py<GradientSensing>,
 }
 
+impl BacteriaTemplate {
+    pub(crate) fn py_eq(&self, py: Python, other: &Self) -> PyResult<bool> {
+        let BacteriaTemplate {
+            mechanics,
+            cycle,
+            cellular_reactions,
+            interactionextracellulargradient,
+        } = self;
+        let e1 = mechanics.extract::<NewtonDamped2D>(py)? == other.mechanics.extract(py)?;
+        let e2 = cycle.extract::<BacteriaCycle>(py)? == other.cycle.extract(py)?;
+        let e3 = cellular_reactions.extract::<BacteriaReactions>(py)?
+            == other.cellular_reactions.extract(py)?;
+        let e4 = interactionextracellulargradient.extract::<GradientSensing>(py)?
+            == other.interactionextracellulargradient.extract(py)?;
+
+        Ok(e1 && e2 && e3 && e4)
+    }
+}
+
 fn bacteria_default_volume() -> f64 {
     std::f64::consts::PI * 1.5_f64.powi(2)
 }

### 2025-09-04 src/simulation.rs
cleanup

@@ -128,12 +128,9 @@ fn save_initial_state(
     meta_params: &MetaParams,
 ) -> PyResult<()> {
     // Also save the SimulationSettings into the same folder
-    let mut save_path_cells = path.clone();
-    let mut save_path_domain = path.clone();
-    let mut save_path_meta_params = path.clone();
-    save_path_cells.push("initial_cells.json");
-    save_path_domain.push("domain.json");
-    save_path_meta_params.push("meta_params.json");
+    let save_path_cells = path.join("initial_cells.json");
+    let save_path_domain = path.join("domain.json");
+    let save_path_meta_params = path.join("meta_params.json");
 
     let f_cells = std::fs::File::create(save_path_cells)?;
     let f_domain = std::fs::File::create(save_path_domain)?;
@@ -269,7 +266,6 @@ pub fn run_simulation(
         .location(&meta_params.save_path)
         .add_date(meta_params.save_add_date)
         .init();
-    // storage.export_formats = vec![ExportOptions::Vtk];
 
     let simulation_setup = create_simulation_setup!(
         Domain: simulation_domain,

### 2025-09-04 src/lib.rs
new public function which can load existing results instead of calculating them newly

@@ -9,6 +9,7 @@ use pyo3::prelude::*;
 #[pymodule]
 fn cr_pool(m: &Bound<'_, PyModule>) -> PyResult<()> {
     m.add_function(wrap_pyfunction!(run_simulation, m)?)?;
+    m.add_function(wrap_pyfunction!(run_or_load_simulation, m)?)?;
     m.add_function(wrap_pyfunction!(generate_cells, m)?)?;
 
     m.add_class::<Bacteria>()?;

### 2025-09-04 src/simulation.rs

@@ -8,7 +8,7 @@ use rand_chacha::ChaCha8Rng;
 
 use serde::{Deserialize, Serialize};
 
-use pyo3::prelude::*;
+use pyo3::{exceptions::PyValueError, prelude::*};
 
 use crate::bacterial_properties::*;
 
@@ -301,4 +301,66 @@ pub fn run_simulation(
     let simulation_result = supervisor.run_full_sim().unwrap();
 
     Ok(simulation_result.storage.get_full_path())
+}
+
+fn try_compare_settings(
+    py: Python,
+    path: &std::path::Path,
+    cells: &[BacteriaTemplate],
+    domain: &Domain,
+    meta_params: &MetaParams,
+) -> PyResult<bool> {
+    let save_path_cells = path.join("initial_cells.json");
+    let save_path_domain = path.join("domain.json");
+    let save_path_meta_params = path.join("meta_params.json");
+
+    let cells_read: Vec<BacteriaTemplate> = {
+        let f = std::fs::File::open(save_path_cells)?;
+        serde_json::from_reader(f).map_err(|e| PyValueError::new_err(format!("{e}")))?
+    };
+    let domain_read: Domain = {
+        let f = std::fs::File::open(save_path_domain)?;
+        serde_json::from_reader(f).map_err(|e| PyValueError::new_err(format!("{e}")))?
+    };
+    let meta_params_read: MetaParams = {
+        let f = std::fs::File::open(save_path_meta_params)?;
+        serde_json::from_reader(f).map_err(|e| PyValueError::new_err(format!("{e}")))?
+    };
+
+    // Compare all cells
+    let cell_comps = cells
+        .iter()
+        .zip(cells_read.iter())
+        .map(|(c1, c2)| c1.py_eq(py, c2))
+        .collect::<PyResult<Vec<_>>>()?
+        .into_iter()
+        .all(|x| x);
+
+    let e1 = cells.len() == cells_read.len() && cell_comps;
+    let e2 = &domain_read == domain;
+    let e3 = &meta_params_read == meta_params;
+
+    Ok(e1 && e2 && e3)
+}
+
+#[pyfunction]
+pub fn run_or_load_simulation(
+    py: Python,
+    cells: Vec<BacteriaTemplate>,
+    domain: Domain,
+    meta_params: MetaParams,
+) -> PyResult<std::path::PathBuf> {
+    let paths = glob::glob(&format!("{}/*", meta_params.save_path))
[…]

### 2025-09-04 scripts/paper_spatial_plots.py
use new function in favour of plain run_simulation

@@ -34,7 +34,7 @@ def calcualte_results(diffusion_constant, randomness, homogenous):
 
     cells = crp.generate_cells(18, 18, domain, randomness, homogenous=homogenous)
 
-    output_path = crp.run_simulation(
+    output_path = crp.run_or_load_simulation(
         cells,
         domain,
         meta_params,

### 2025-09-04 scripts/run_sim.py

@@ -19,7 +19,7 @@ if __name__ == "__main__":
 
     cells = crp.generate_cells(18, 18, domain, randomness=0.0)
 
-    output_path = crp.run_simulation(
+    output_path = crp.run_or_load_simulation(
         cells,
         domain,
         meta_params,

### 2025-09-04 src/simulation.rs
fixup! new public function which can load existing results instead of calculating them newly

@@ -361,6 +361,5 @@ pub fn run_or_load_simulation(
         }
     }
 
-    todo!();
     run_simulation(py, cells, domain, meta_params)
 }

### 2025-09-05 paper/pool_model.tex
write discussion
- also remove bullet points

@@
 
-How well does the pool model do to encapsulate all the different growth models
-\begin{itemize}
-    \item All 6 models can be represented by pool model
-    \item Baranyi-Roberts results can be reproduced, but not identical; requires some work
-    \item what are some general limitations within this framework? how can they be circumvented?
-    \item why is the pool model "better" than the other \ac{ode} models? (why did we do this?);
-        generalization of many models; better explainability;
-    \item link to supplement; we can generalize this and prove theorems for the abstract,
-        generalized model; this allows us to discuss many cases at once
-\end{itemize}
-
-How justified was the construction of the \ac{abm}?
-\begin{itemize}
-    \item Discuss simplified physical interaction model
-    \item Heterogeneity plays a role; as we have shown
-    \item why did we not simply use a PDE?; because we have to think in terms of cellular processes
-        and this ensures that our model is actually realistic
-\end{itemize}
-
-How can we cover more effects of bacteria?
-\begin{itemize}
-    \item More states for metabolic processes
-    \item More states for more species
-    \item Other cooperiation/inhibition effects
-    \item external influences
-\end{itemize}
-
-What are limitations of our approach?
-\begin{itemize}
-    \item spatial stuff; heterogeneity (because \ac{ode})
-    \item larger models require more parameters, harder to estimate
-    \item only simple models have analytical solutions
-\end{itemize}
+% How well does the pool model do to encapsulate all the different growth models
+% \begin{itemize}
+%     \item All 6 models can be represented by pool model
+%     \item Baranyi-Roberts results can be reproduced, but not identical; requires some work
+%     \item what are some general limitations within this framework? how can they be circumvented?
+%     \item why is the pool model "better" than the other \ac{ode} models? (why did we do this?);
+%         generalization of many models; better explainability;
+%     \item link to supplement; we can generalize this and prove theorems for the abstract,
+%         generalized model; this allows us to discuss many cases at once
+% \end{itemize}
+%
+% How can we cover more effects of bacteria?
+% \begin{itemize}
+%     \item More states for metabolic processes
+%     \item More states for more species
+%     \item Other cooperiation/inhibition effects
+%     \item external influences
+% \end{itemize}
+%
+% How justified was the construction of the \ac{abm}?
+% \begin{itemize}
+%     \item Discuss simplified physical interaction model
+%     \item Heterogeneity plays a role; as we have shown
+%     \item why did we not simply use a PDE?; because we have to think in terms of cellular processes
+%         and this ensures that our model is actually realistic
+% \end{itemize}
+%
+% What are limitations of our approach?
+% \begin{itemize}
+%     \item spatial stuff; heterogeneity (because \ac{ode})
+%     \item larger models require more parameters, harder to estimate
+%     \item only simple models have analytical solutions
+% \end{itemize}
+
+The pool model presented here was used to describe the dynamics of 6 systems containing established models of bacterial growth.
+Comparing our approach with the Baranyi-Roberts model which is widely used in food science, it is possible to obtain results which provide growth curves of very similar shape.
+However, their analytical form is not completely identical.
+The inclusion of additional metabolic processes or larger number of cell-species can be achieved by extending the list of possible states and defining transitions between them.
+In addition we have shown how to treat interspecies interactions in the form of cooperation and competition.
+This framework provides a starting point for a mathematically more abstract formulation which is thus able to cover a much wider class of systems.
+By constructing an \ac{abm}, we have exemplified how close the devised model is to the individual biological processes that occur within the bacteria.
+We have investigated the effect of spatial effects in the form of varying diffusion parameters and different initial spatial distributions of bacteria.
+Our findings reinforce that the \ac{ode} model is a good simplification in the highly-diffusive regime within a mostly homogenous spatial bacterial distribution.
+However, deviations to these assumptions can lead to non-negligible effects, influencing the growth of the colony.
 

### 2025-09-05 paper/pool_model.tex
exit before finishing Conclusion

@@
 \end{itemize}
+
+We introduced the pool model as a novel framework to unify existing approaches and allows researchers to follow a mechanistic approach during model development.
+It is able to capture the effects of many established models and provides a lens which is useful to generalize existing systems and results in better overall explainability.
+
+\textbf{CHRISTIAN HELP!}
 

### 2025-09-05 paper/pool_model.tex
finish discussion and conclusion for now

@@
 In addition we have shown how to treat interspecies interactions in the form of cooperation and competition.
-This framework provides a starting point for a mathematically more abstract formulation which is thus able to cover a much wider class of systems.
 By constructing an \ac{abm}, we have exemplified how close the devised model is to the individual biological processes that occur within the bacteria.
+This mechanistic single-cell level of understanding of our \ac{abm} is a perspective which can not be easily replicated by other spatial models such as \acp{pde}.
 We have investigated the effect of spatial effects in the form of varying diffusion parameters and different initial spatial distributions of bacteria.
@@
 \section{Conclusion}
-\begin{itemize}
-    \item New theoretical framework
-    \item consistent with \ac{ode} models (Baranyi-Roberts)
-    \item Better explainability
-    \item Include effects such as inhibition, toxins in understandable manner
-    \item showed spatial limitations of \ac{ode} model
-    \item \ac{abm} model allowed modeling of cellular processes; ensure that pool model is derivable
-        from the individual behvaiour of cellular processes
-    \item outlook: generalized mathematical formulation; extend inhibition model (because it is a
-        dimensional reduction of existing models)
-    \item Provide code in a pypi package as well (TODO)
-\end{itemize}
+% \begin{itemize}
+%     \item New theoretical framework
+%     \item consistent with \ac{ode} models (Baranyi-Roberts)
+%     \item Better explainability
+%     \item Include effects such as inhibition, toxins in understandable manner
+%     \item showed spatial limitations of \ac{ode} model
+%     \item \ac{abm} model allowed modeling of cellular processes; ensure that pool model is derivable
+%         from the individual behvaiour of cellular processes
+%     \item outlook: generalized mathematical formulation; extend inhibition model (because it is a
+%         dimensional reduction of existing models)
+%     \item Provide code in a pypi package as well (TODO)
+% \end{itemize}
 
@@
 It is able to capture the effects of many established models and provides a lens which is useful to generalize existing systems and results in better overall explainability.
-
-\textbf{CHRISTIAN HELP!}
+By modeling effects such as mutual inhibition on a single-cell level within an \ac{abm}, we have exemplified that the pool model directly builds on these more fundamental concepts.
+It can thus aid in the mechanistic understanding of the underlying processes, allowing researchers to deepen their konwledge about microbial growth within food science.
+Going forward, this framework can provide a starting point for a more abstract mathematical formulation which should be able to cover a much wider class of systems.
 

### 2025-09-05 paper/pool_model.tex
replace old abm figures

@@
     \begin{tikzonimage}[width=0.49\columnwidth]
-        {Figures/abm-homogenous/snapshot_00006000.png}
+        {Figures/abm-homogenous/images/snapshot_00006000.png}
         \node at (0.01, 0.99)[anchor=north west]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.49\columnwidth]
-        {Figures/abm-homogenous/snapshot_00012000.png}
+        {Figures/abm-homogenous/images/snapshot_00012000.png}
         \node at (0.01, 0.99)[anchor=north west]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.49\columnwidth]
-        {Figures/abm-homogenous/snapshot_00018000.png}
+        {Figures/abm-homogenous/images/snapshot_00018000.png}
         \node at (0.01, 0.99)[anchor=north west, white]{\textbf{C}};
@@
     \begin{tikzonimage}[width=0.49\columnwidth]
-        {Figures/abm-homogenous/snapshot_00024000.png}
+        {Figures/abm-homogenous/images/snapshot_00024000.png}
         \node at (0.01, 0.99)[anchor=north west, white]{\textbf{D}};
@@
     \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures/abm-homogenous/snapshot_00012000.png}
+        \includegraphics[width=\textwidth]{Figures/abm-homogenous/images/snapshot_00012000.png}
         \subcaption{Homogenous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
@@
     \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures/abm-homogenous-low-diffusion/snapshot_00012000.png}
+        \includegraphics[width=\textwidth]{Figures/abm-homogenous-low-diffusion/images/snapshot_00012000.png}
         \subcaption{Homogenous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
@@
     \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures/abm-inhomogenous/snapshot_00012000.png}%
+        \includegraphics[width=\textwidth]{Figures/abm-inhomogenous/images/snapshot_00012000.png}%
         \subcaption{Heterogeneous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
@@
     \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures/abm-random/snapshot_00012000.png}%
+        \includegraphics[width=\textwidth]{Figures/abm-random/images/snapshot_00012000.png}%
         \subcaption{Random initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}

### 2025-09-05 scripts/paper_spatial_plots.py
specify correct variables for saving

@@ -4,17 +4,10 @@ import argparse
 
 # Diffusion constants, randomness, homogenous
 CONFIGS = [
-    # (30, 0, True),
-    # (5, 0, True),
-    # (30, 0, False),
-    (30, 0.25, True),
-    (30, 0.20, True),
-    (30, 0.15, True),
-    (30, 0.10, True),
-    (5, 0.25, True),
-    (5, 0.20, True),
-    (5, 0.15, True),
-    (5, 0.10, True),
+    (30, 0, True),
+    (5, 0, True),
+    (30, 0, False),
+    (30, 0.30, True),
 ]
 
 

### 2025-09-05 scripts/paper_spatial_plots.py
save hand-selected snapshots

@@ -56,10 +56,11 @@ if __name__ == "__main__":
 
     if args.save_snapshots:
         for output_path in paths:
-            picked_iter = crp.get_all_iterations(output_path)[20]
-            crp.save_snapshot(output_path, picked_iter)
-        # for output_path in paths:
-        #     crp.save_all_snapshots(Path(output_path))
+            iters = crp.get_all_iterations(output_path)
+            crp.save_snapshot(output_path, iters[6])
+            crp.save_snapshot(output_path, iters[12])
+            crp.save_snapshot(output_path, iters[18])
+            crp.save_snapshot(output_path, iters[24])
 
     crp.load_style()
     for output_path in paths:

### 2025-09-24 pool_paper_plotting_code/examples/main_paper.py
Update Figures: beta=0 in Fig2/3. Separate plots for all subfigures

@@ -138,8 +138,8 @@ def pool_model_2sp_cooper(t, x, param, x0, const):
 def observable_3pool(output):
     # Output (n_variables x n_times)
     # Variables: (L, G, D,) = x
-    # => Output [n = L+G+D]
-    return [output[0]+output[1]+output[2]]
+    # => Output [n = L+G]
+    return [output[0]+output[1]]#+output[2]]
 
 def observable_2pool(output):
     # Output (n_variables x n_times)
@@ -190,10 +190,10 @@ if __name__ == "__main__":
 
 ###############33############### 3 pool models with resource pool ##############################3########333
     # Generate model data
-    const1 = [1e-2, 0., 2.5, .5, 5., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
-    const2 = [1e-3, 0., 2.5, .5, 5., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
-    const3 = [5e-5, 0., 2.5, .5, 5., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
-    const4 = [1e-3, .5, 2.5, .5, 5., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
+    const1 = [1e-2, 0., 2.5, .5, 0., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
+    const2 = [1e-3, 0., 2.5, .5, 0., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
+    const3 = [5e-5, 0., 2.5, .5, 0., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
+    const4 = [1e-3, .5, 2.5, .5, 0., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
     const_3pool_res = [const1, const2, const3, const4]
     data_3pool_res = [generate_insilico_data(pool_model_3pools_resource, [t], [], [[]], x01,
                 const=const, obs_func=observable_3pool, n_traj=1) for const in const_3pool_res]
@@ -208,8 +208,8 @@ if __name__ == "__main__":
             ax.plot(d['times'], d['obs_mean'][0], linewidth=2.5, color=color_palette_1sp[i], label=labels[i])
     fig, ax = set_labels(fig, ax, r'Time, $t$', r'Total Bacterial Count, $N$')
[…]

### 2025-09-26 paper/pool_model.tex
correct diffusion constant
migrate to only xcolor
add two more sentences for description of figure
Corrections from Christian

@@
 \usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off, font=small]{caption}
-\usepackage{color}
 \usepackage{diagbox}
@@
 
-\title{Pool Model - A unified Mathematical\\ Modelling Framework for Food Science}
+\title{Pool Model - A Mathematical Framework\\ for Modelling Bacterial Consortia in Food Science}
 \author{
@@
     \footnotesize
-    \textsuperscript{\textbf{1}}Freiburg Center for Data Analysis and Modeling
+    \textsuperscript{\textbf{1}}Freiburg Center for Data Analysis and Modeling and AI
 }
@@
     food microbiology.
-    However, the utilized models live on the popluation level, omitting any spatial effects such as
+    However, the utilized models live on the population level, omitting any spatial effects such as
     heterogeneity or diffusion.
@@
     G &\stackrel{\alpha}{\longrightarrow} & 2G\\
-    G &\stackrel{\mu'}{\longrightarrow} & D\\
-    D &\stackrel{\beta}{\longrightarrow} & \emptyset
+    G &\stackrel{\mu'}{\longrightarrow} & D
  \end{eqnarray}
@@
     \dot{G} &= \lambda L + \alpha G - \mu' G\\
-    \dot{D} &= \mu  L + \mu' G- \beta D  
+    \dot{D} &= \mu  L + \mu' G 
 \end{split}\end{align}
@@
 \begin{equation}
-    G + R_0  \stackrel{\alpha_0}{\longrightarrow} 2G.
-\end{equation}
-Note that the pools $L$, $G$ and $D$ represent the number of bacteria (or number of bacteria per unit area/volume) while $R_0$ represents an abstract resource pool.
+    G + R  \stackrel{\alpha_0}{\longrightarrow} 2G.
+\end{equation}
+Note that the pools $L$, $G$ and $D$ represent the number of bacteria (or number of bacteria per unit area/volume) while $R$ represents an abstract resource pool.
 Let say $G=N_B/V$ and $[G]=1/V$, where $N_B$ is the number of bacteria and $V$ is the volume.
-Then $R_0=N_R/V$ and $[R_0]=1/V$, where $N_R$ is the number of resource molecules (or any appropriate unit, e.g., mol).
+Then $R=N_R/V$ and $[R]=1/V$, where $N_R$ is the number of resource molecules (or any appropriate unit, e.g., mol).
 Then $[\alpha_0]=V/s$, $[N_t]=1/V$, and $[\alpha]=1/s$.
@@
 In this case $[L]=[G]=[N_t]=1$.
-For simplicity, we make a transition to the dimensionless constant $\alpha:=\alpha_0 N_t$ and dimensionless resource pool $R(t) = R_0 / N_t$ ($0 \leqslant R \leqslant 1$).
+For simplicity, we make a transition to the dimensionless constant $\alpha:=\alpha_0 N_t$ and dimensionless resource pool $\tilde{R}(t) = R/ N_t$ ($0 \leqslant \tilde{R} \leqslant 1$).
 Here the full capacity of the resource pool $N_t$ was used as a scaling constant.
-In this case the growth term can be simply written as $\alpha G R$ and the model is fully formulated in the following way:
+In this case the growth term can be simply written as $\alpha G \tilde{R}$ and the model is fully formulated in the following way:
 \begin{align}
@@
         \dot{L} &= -(\mu + \lambda) L\\
-        \dot{G} &= \lambda L + \alpha R G-\mu' G\\
-        \dot{D} &= \mu  L + \mu' G- \beta D\\
-        \dot{R} &= - \frac{\alpha}{N_t} R G
+        \dot{G} &= \lambda L + \alpha \tilde{R} G-\mu' G\\
+        \dot{D} &= \mu  L + \mu' G\\
+        \dot{\tilde{R}} &= - \frac{\alpha}{N_t} \tilde{R} G
     \end{split}.
@@
 %
-Such a model can be simplified by exploiting that any limitation of the resource pool also leads to a limit of the total bacterial count
-\begin{equation}
-    R = 1 - \frac{L+G+D}{N_t}
+This can be simplified by exploiting the conservation of the resource pool:
+\begin{equation}
+    \tilde{R} = 1 - \frac{L+G+D}{N_t}
 \end{equation}
@@
         \dot{G} &= \lambda L + \alpha G\left(1-\frac{L+G+D}{N_t}\right)-\mu' G\\
-        \dot{D} &= \mu  L + \mu' G- \beta D 
+        \dot{D} &= \mu  L + \mu' G
     \end{split}.
@@
     \caption{
-        The time dependence of the total bacterial count $N(t) = L(t) + G(t) + D(t)$ is obtained as a sum of bacteria in all states.
-        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\alpha=2.5$, $\mu'=0.5$ and $\beta=5$.
+        The time dependence of the bacterial count $N(t) = L(t) + G(t) $ is shown.
+        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\alpha=2.5$, $\mu'=0.5$.
         The blue curve shows a significant dip in the behavior of the system where the death of the bacteria is present $\mu = 0.5$.
         Other curves show the case of no bacteria dying $\mu=0$ and represent different lengths of lag phases: short $\lambda=10^{-2}$ (dark green), intermediate $\lambda=10^{-3}$ (light green) and long one $\lambda=10^{-4}$ (yellow).
+Note that the steady state for $N(t)$ and $D(t)$ is for this model $\lim_{t\to\infty} N(t)=0$ and $\lim_{t\to\infty} D(t)=N_t$, because of the depletion of the resource pool.
     }
@@
 % The dynamics of the three pool model can be captured by the coupled system of ordinary differential
-The dynamics of the three pool model is captured by equations~\ref{eq:ode_3pool_resource2} which can be used to derive the two pool model by setting $D\equiv 0$ and replacing $\mu$ with $\beta$).
-For $\mu'=0$ the equations for $L$ and $D$ can be readily integrated resulting in:
+The dynamics of the three pool model is captured by equations~\ref{eq:ode_3pool_resource2} which can be used to derive the two pool model simply by omitting $D$.
+Moreover, for $\mu'=0$ the equations for $L$ and $D$ can be readily integrated resulting in:
 \begin{align}
@@
         L(t) &= N_0 e^{-(\mu+\lambda)t}\\
-        D(t) &= N_0 \mu e^{-\beta t}\frac{1-e^{-(\mu+\lambda-\beta)t}}{\mu+\lambda-\beta}
+        D(t) &= N_0 \mu \frac{1-e^{-(\mu+\lambda)t}}{\mu+\lambda
+       }
     \end{split}.
 \end{align}
-For constant parameters $\mu$, $\lambda$, and $\beta$ the integrals for $L$ and $D$ can be inserted into the differential equation for $G$.
+For constant parameters $\mu$ and $\lambda$ the integrals for $L$ and $D$ can be inserted into the differential equation for $G$.
 However, in the case of time dependent parameters, e.g., for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$.
@@
         \dot{G} &= \lambda L + \alpha G\left(1-\frac{N}{N_t}\right)\\
-        \dot{N} &= - \beta D + \alpha G\left(1-\frac{N}{N_t}\right)
+        \dot{N} &= \alpha G\left(1-\frac{N}{N_t}\right)
     \end{split}
@@
 A notable difference is that the growth rate $\alpha$ is time independent.
-For the functions $L$ and $D$ it holds: $\lim_{t\to\infty} L(t) = \lim_{t\to\infty} D(t) = 0$.
-It follows immediately that $\lim_{t\to\infty} G(t) = N_t$.
-The equation exhibits a lag phase and an initial drop due to cell death.
-To find a substantial initial drop one needs $\beta \mu \gg \alpha \lambda$.
+For the functions $L$ and $D$ it holds: $\lim_{t\to\infty} L(t) = 0$ and  $\lim_{t\to\infty} D(t) =  N_0\frac{\mu}{\mu+\lambda} $.
+It follows immediately that $\lim_{t\to\infty} G(t) = N_t-N_0\frac{\mu}{\mu+\lambda} $.
 The lag time $t_L$ can be approximated by the time at which half of the initial population is in the growth phase, i.e., $N_0=2G(t_L)$.
@@
 \begin{align}
-    1&= 2\int_0^{t_L} \lambda L(t)e^{\alpha(t_L-t)}dt\\
+    N_0&= 2\int_0^{t_L} \lambda L(t)e^{\alpha(t_L-t)}dt\\
     1  &= 2\lambda e^{\alpha t_L}\frac{1-e^{-(\mu+\lambda+\alpha)t_L}}{\mu+\lambda+\alpha}.
@@
 The lag time $t_L$ diverges very slowly, logarithmically with $\lambda\to 0$.
-%
 %
@@
     G &\stackrel{\alpha}{\longrightarrow} & 2G\\
-    G &\stackrel{\gamma(t)}{\longrightarrow} & L\\
-    D &\stackrel{\beta}{\longrightarrow} & \emptyset
+    G &\stackrel{\gamma(t)}{\longrightarrow} & L
 \end{eqnarray}
@@
         \dot{G} &= \lambda L -\gamma(t) G + \alpha G\left(1-\frac{L+G+D}{N_t}\right)\\
-        \dot{D} &= \mu  L - \beta D
+        \dot{D} &= \mu  L 
 \end{split}
@@
 Subfigure (d) shows how a randomized homogenous initial distribution can lead to even more notable effects of inhibition.
+In contrast to the other results, here the final values of species $A$ and $B$ have larger/smaller counts respectively.
+This illustrates that spatial effects can shift the final results of growth curves towards both directions, leading to an overall increase/decrease of one species.
 We furthermore note that the total bacteria volume of the \ac{abm} model is always smaller or equal compared to the \ac{ode} model.
@@
         \includegraphics[width=\columnwidth]{Figures/abm-random/abm_ode_comparison.png}%
-        \subcaption{Rand. $D=5.0$\unit{\micro\metre\squared\per\minute}}
+        \subcaption{Rand. $D=30.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}
@@
         \includegraphics[width=\textwidth]{Figures/abm-random/images/snapshot_00012000.png}%
-        \subcaption{Random initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
+        \subcaption{Random initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}

### 2025-10-30 paper/pool_model.tex
adjust to journal latex template

@@
-\documentclass[10pt,A4paper]{article}
+% \documentclass[10pt,A4paper]{article}
+\documentclass[10pt,twocolumn,5p]{elsarticle}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{amssymb}
-\usepackage[
-    backend=biber,
-    sorting=none,
-    style=numeric-comp,
-    hyperref=true,
-    maxnames=2,
-    url=false,
-    isbn=false,
-    eprint=false,
-    date=year,
-    doi=false,
-    giveninits=true
-]{biblatex}
 \usepackage{array}
@@
 \usepackage[T1]{fontenc}
-\usepackage[a4paper,inner=2cm,outer=2cm,top=2.5cm,bottom=2.5cm]{geometry}
+% \usepackage[a4paper,inner=2cm,outer=2cm,top=2.5cm,bottom=2.5cm]{geometry}
 \usepackage{graphicx}% Include figure files
@@
 
-% Add bibliography
-\addbibresource{references.bib}
+\bibliographystyle{abbrvnat}
 
@@
 
-\title{Pool Model - A Mathematical Framework\\ for Modelling Bacterial Consortia in Food Science}
-\author{
-    Polina Gaindrik\textsuperscript{1},
-    Jonas Pleyer\textsuperscript{1}
-    and Christian Fleck\textsuperscript{1}
+\tnotetext[t1]{TODO}
+\tnotetext[t2]{TODO}
+
+\title{Pool Model - A unified Mathematical\\ Modelling Framework for Food Science}
+\author[1]{Polina Gaindrik}
+\author[1]{Jonas Pleyer\fnref{fn1}}
+\author[1]{Christian Fleck\corref{cor1}}
+\ead{christian.fleck@fdm.uni-freiburg.de}
+
+\cortext[cor1]{Corresponding Author}
+\fntext[fn1]{jonas.pleyer.org}
+\affiliation[1]{
+    organization={Freiburg Center for Data Analysis and Modeling},
+    addressline={Ernst-Zermelo Str. 1},
+    city={Freiburg im Breisgau},
+    citysep={},
+    postcode={79104},
+    country={Germany},
 }
@@
     \footnotesize
-    \textsuperscript{\textbf{1}}Freiburg Center for Data Analysis and Modeling and AI
+    \textsuperscript{\textbf{1}}
 }
@@
 \begin{document}
-\maketitle
 
@@
     growth.
-    % \textbf{Notes from Christian}
-    % Different empirical and inactivation models applied in the estimation
-    % of parameters and modelling of phenomena in food microbiology, gen-
-    % eralized Lotka Voltera approach used to describe the interaction of mi-
-    % croorganisms, mechanistic models utilized to provide more insight into
-    % the underlying dynamic can all be unified as variations of a new model
-    % known as three pool model. Compared against the aforementioned
-    % methods which are applicable in specific applications with some limi-
-    % tations, the present model turns out to be less complex with general
-    % applicability, low number of parameters and preferably to explain the
-    % mechanism of the phenomenon of interest. To illustrate the generality
-    % of the model, three different examples are reported.
 \end{abstract}
 
-\textbf{Keywords:}
-Bacterial Consortia, Competition and Cooperation, Mechanistic Modeling, Spatial Effects
-
-\newpage
-\tableofcontents
-
-\twocolumn
-
-\newpage
-%
-%
-%
+\maketitle
+
+\begin{graphicalabstract}
+% TODO
+\end{graphicalabstract}
+
+\begin{keyword}
+Bacterial Consortia \sep Competition and Cooperation \sep Mechanistic Modeling \sep Spatial Effects
+\end{keyword}
 %
@@
 Moreover, by including the flow of living bacteria to the death pool, we can model a greater variety of effects by allowing a decrease in the bacterial count.
+% TODO change the name "death pool"?
 %
@@
 
-\printbibliography
+% \printbibliography
+\bibliography{references.bib}
 

### 2025-10-30 paper/pool_model.tex
remove unused footnotes
flatten file structure of images

@@
 
-\tnotetext[t1]{TODO}
-\tnotetext[t2]{TODO}
-
 \title{Pool Model - A unified Mathematical\\ Modelling Framework for Food Science}
@@
     \begin{center}
-    \includegraphics[width=0.9\columnwidth]{Figures/TPM_fig.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures-TPM_fig.pdf}
     \caption{Schematic representation of exchange between three states or pools.}
@@
     \begin{center}
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_3pools_resource.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures-pool_model_3pools_resource.pdf}
     \caption{
@@
     \begin{center}
-    %\includegraphics[width=0.9\columnwidth]{Figures/pool_model_3pools_resource_tempshift.pdf}
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_3pools_resource_tempshift1.pdf}\\
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_3pools_resource_tempshift2.pdf}
+    %\includegraphics[width=0.9\columnwidth]{Figures-pool_model_3pools_resource_tempshift.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures-pool_model_3pools_resource_tempshift1.pdf}\\
+    \includegraphics[width=0.9\columnwidth]{Figures-pool_model_3pools_resource_tempshift2.pdf}
     \caption{
@@
     \begin{center}
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_2pools_resource_competition.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures-pool_model_2pools_resource_competition.pdf}
     \caption{
@@
     \begin{center}
-    %\includegraphics[width=\columnwidth]{Figures/pool_model_2pools_toxin.pdf}
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_2pools_toxin1.pdf}\\
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_2pools_toxin2.pdf}
+    %\includegraphics[width=\columnwidth]{Figures-pool_model_2pools_toxin.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures-pool_model_2pools_toxin1.pdf}\\
+    \includegraphics[width=0.9\columnwidth]{Figures-pool_model_2pools_toxin2.pdf}
     \caption{
@@
     \begin{center}
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_1pool_inhib1.pdf}\\
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_1pool_inhib2.pdf}\\
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_1pool_inhib3.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures-pool_model_1pool_inhib1.pdf}\\
+    \includegraphics[width=0.9\columnwidth]{Figures-pool_model_1pool_inhib2.pdf}\\
+    \includegraphics[width=0.9\columnwidth]{Figures-pool_model_1pool_inhib3.pdf}
     \caption{
@@
     \begin{center}
-    %\includegraphics[width=0.9\columnwidth]{Figures/pool_model_2sp_1pool_coop.pdf}
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_2sp_1pool_coop1.pdf}
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_2sp_1pool_coop2.pdf}
-    \includegraphics[width=0.9\columnwidth]{Figures/pool_model_2sp_1pool_coop3.pdf}
+    %\includegraphics[width=0.9\columnwidth]{Figures-pool_model_2sp_1pool_coop.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures-pool_model_2sp_1pool_coop1.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures-pool_model_2sp_1pool_coop2.pdf}
+    \includegraphics[width=0.9\columnwidth]{Figures-pool_model_2sp_1pool_coop3.pdf}
     \caption{
@@
 %     \centering
-%     \includegraphics[width=0.6\textwidth]{Figures/interaciton_potential.png}
+%     \includegraphics[width=0.6\textwidth]{Figures-interaciton_potential.png}
 %     \caption{
@@
     \begin{tikzonimage}[width=0.49\columnwidth]
-        {Figures/snapshot-2-cells-before-division.png}%
+        {Figures-snapshot-2-cells-before-division.png}%
         \node at (0.01, 0.99)[anchor=north west]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.49\columnwidth]
-        {Figures/snapshot-2-cells-after-division.png}
+        {Figures-snapshot-2-cells-after-division.png}
         \node at (0.01, 0.99)[anchor=north west]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.49\columnwidth]
-        {Figures/abm-homogenous/images/snapshot_00006000.png}
+        {Figures-abm-homogenous-images-snapshot_00006000.png}
         \node at (0.01, 0.99)[anchor=north west]{\textbf{A}};
@@
     \begin{tikzonimage}[width=0.49\columnwidth]
-        {Figures/abm-homogenous/images/snapshot_00012000.png}
+        {Figures-abm-homogenous-images-snapshot_00012000.png}
         \node at (0.01, 0.99)[anchor=north west]{\textbf{B}};
@@
     \begin{tikzonimage}[width=0.49\columnwidth]
-        {Figures/abm-homogenous/images/snapshot_00018000.png}
+        {Figures-abm-homogenous-images-snapshot_00018000.png}
         \node at (0.01, 0.99)[anchor=north west, white]{\textbf{C}};
@@
     \begin{tikzonimage}[width=0.49\columnwidth]
-        {Figures/abm-homogenous/images/snapshot_00024000.png}
+        {Figures-abm-homogenous-images-snapshot_00024000.png}
         \node at (0.01, 0.99)[anchor=north west, white]{\textbf{D}};
@@
     \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=\columnwidth]{Figures/abm-homogenous/abm_ode_comparison.png}
+        \includegraphics[width=\columnwidth]{Figures-abm-homogenous-abm_ode_comparison.png}
         \subcaption{Hom. $D=30.0$\unit{\micro\metre\squared\per\minute}}
@@
     \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=\columnwidth]{Figures/abm-homogenous-low-diffusion/abm_ode_comparison.png}
+        \includegraphics[width=\columnwidth]{Figures-abm-homogenous-low-diffusion-abm_ode_comparison.png}
         \subcaption{Hom. $D=5.0$\unit{\micro\metre\squared\per\minute}}
@@
     \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=\columnwidth]{Figures/abm-inhomogenous/abm_ode_comparison.png}%
+        \includegraphics[width=\columnwidth]{Figures-abm-inhomogenous-abm_ode_comparison.png}%
         \subcaption{Het. $D=30.0$\unit{\micro\metre\squared\per\minute}}
@@
     \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=\columnwidth]{Figures/abm-random/abm_ode_comparison.png}%
+        \includegraphics[width=\columnwidth]{Figures-abm-random-abm_ode_comparison.png}%
         \subcaption{Rand. $D=30.0$\unit{\micro\metre\squared\per\minute}}
@@
     \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures/abm-homogenous/images/snapshot_00012000.png}
+        \includegraphics[width=\textwidth]{Figures-abm-homogenous-images-snapshot_00012000.png}
         \subcaption{Homogenous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
@@
     \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures/abm-homogenous-low-diffusion/images/snapshot_00012000.png}
+        \includegraphics[width=\textwidth]{Figures-abm-homogenous-low-diffusion-images-snapshot_00012000.png}
         \subcaption{Homogenous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
@@
     \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures/abm-inhomogenous/images/snapshot_00012000.png}%
+        \includegraphics[width=\textwidth]{Figures-abm-inhomogenous-images-snapshot_00012000.png}%
         \subcaption{Heterogeneous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
@@
     \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures/abm-random/images/snapshot_00012000.png}%
+        \includegraphics[width=\textwidth]{Figures-abm-random-images-snapshot_00012000.png}%
         \subcaption{Random initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}

### 2025-10-31 paper/pool_model.tex
tex: correct mistakes, delete unnecessary commands

@@
 \usepackage{tikz}
-\usepackage[dvipsnames]{xcolor}
+%\usepackage[dvipsnames]{xcolor}
 \usepackage{xpatch}
@@
 }
-
-\newcommand{\mbP}{\textbf{P: }}
-\newcommand{\mbC}{\textbf{C: }}
 \newcommand{\mbx}{\mathbf{x}}
@@
 \newcommand{\mbH}{\mathbf{H}}
-\newcommand{\cmtb}[1]{\textcolor{blue} {#1}}
 \newcommand{\cmt}[1]{\textcolor{red} {[#1]}}
@@
     city={Freiburg im Breisgau},
-    citysep={},
     postcode={79104},
@@
     Their dynamics on a population-based level often result from interactions that lead to
-    inter-species cooperation or competition.
+    interspecies cooperation or competition.
     Various empirical and inactivation models have been conceived and applied to problems related to
@@
     arises naturally from the individual behaviour of many cellular agents.
-    We further apply this \ac{abm} to study the role of spatial effects on the dynamcis of bacterial
+    We further apply this \ac{abm} to study the role of spatial effects on the dynamics of bacterial
     growth.
@@
 \end{figure}
-To determine the system of equations which describe the systems dynamics, we consider the flows between the different states.
+To determine the system of equations which describe the system's dynamics, we consider the flows between the different states.
 The pools $G$ and $D$ are unidirectionally connected with the pool $L$.
@@
 Here species $A$ has a longer lag-phase than $B$ but the exponential growth of $A$ is stronger compared to which is experienced by bacteria $B$.
-As a result, species $A$ was able to suppress species $B$, consuminl the larger amount of nutrients and reaching a larger bacterial count at the end.
+As a result, species $A$ was able to suppress species $B$, consuming the larger amount of nutrients and reaching a larger bacterial count at the end.
 \begin{figure}
@@
 Mutual inhibition between different species can not only result from competition for a shared resource for which bacteria have developed various other active competition mechanisms.
-One of the examples of such a mechanisms is the production of antimicrobial compounds, inhibitors or toxins~\cite{wloch-salamon_effect_2008, chao_structured_1981}.
+One of the examples of such mechanisms is the production of antimicrobial compounds, inhibitors or toxins~\cite{wloch-salamon_effect_2008, chao_structured_1981}.
 On the other hand, some bacteria can produce public goods, the compounds that benefit growth of other species.
@@
 Furthermore, neighbors can also benefit from a behavior which does not require any action on an expression level, so-called 'social cheaters'~\cite{rainey_evolution_2003}.
-This mutual activation benefits all involved species and can by easily included in our mathematical framework.
+This mutual activation benefits all involved species and can be easily included in our mathematical framework.
 For the sake of clarity we omit again the lag pool and write
@@
 species to develop competitive mechanisms such as cell adhesion to surfaces in the areas rich with
-nutritients~\cite{htuson_bacteriasurface_2013}.
+nutrients~\cite{htuson_bacteriasurface_2013}.
 Another mechanism is the motility of bacteria, e.g., the flagellates \textit{E. coli} or
-\textit{B. subtilis} which are capable of chemotactic motion of driven by oxygen
+\textit{B. subtilis} which are capable of chemotactic motion driven by oxygen
 concentrations~\cite{decoene_microscopic_2011}.
-The bacterial social interactions such as releas of toxins or public goods can also
+The bacterial social interactions such as release of toxins or public goods can also
 significantly affect the spatial structure of the community~\cite{blanchard_bacterial_2015}.
@@
 To construct our \ac{abm}, we use \texttt{cellular\_raza}~\cite{Pleyer2025}.
-It allows us to define our bacterial agents from a clean slate, without having to rely and inherit
+It allows us to define our bacterial agents from a clean slate, without having to rely on and inherit
 existing work.
@@
 Therefore, we obtain the relation $\frac{u}{1+\nu_I [I]}[R_e] = a[R_i]$ which we can use to eliminate the variable monitoring internal resource concentrations $[R_i]$.
-Additionally, we require our external components to be freely diffusable with constants $D_R,D_I$.
+Additionally, we require our external components to be freely diffusible with constants $D_R,D_I$.
 Plugging this into equation~\eqref{eq:spatial-full-system-non-simplified} yields.
@@
 The combination of the two effects shows the largest difference between model predictions.
-In subfigure (c), due to the heterogenous cell distribution (see also~\ref{fig:supplement-spatial-initial-config-cells}), overall inhibition of species $B$ is significantly reduced.
+In subfigure (c), due to the heterogeneous cell distribution (see also~\ref{fig:supplement-spatial-initial-config-cells}), overall inhibition of species $B$ is significantly reduced.
 Subfigure (d) shows how a randomized homogenous initial distribution can lead to even more notable effects of inhibition.
@@
 We furthermore note that the total bacteria volume of the \ac{abm} model is always smaller or equal compared to the \ac{ode} model.
-This is expected, since processes such as diffusion and cells growing outward introduce another time component which is simply assumed to be instantanous in the \ac{ode} model.
+This is expected, since processes such as diffusion and cells growing outward introduce another time component which is simply assumed to be instantaneous in the \ac{ode} model.
 The chosen diffusion constants $D_I=D_R=30.0$\unit{\micro\metre\squared\per\minute} are in the order of small molecules such as $NaCl$ while the lower values $D_I=D_R=5.0$\unit{\micro\metre\squared\per\minute} are around values of Glucose in water.
@@
         Comparison of the \ac{ode} and \ac{abm} model for different configurations, showing
-        homogenous (Hom.), heterogenous (Het.) or randomized (Rand.) initial configurations.
+        homogenous (Hom.), heterogeneous (Het.) or randomized (Rand.) initial configurations.
         The diffusion constants are chosen such that $D_R=D_I=D$.
@@
 The inclusion of additional metabolic processes or larger number of cell-species can be achieved by extending the list of possible states and defining transitions between them.
-In addition we have shown how to treat interspecies interactions in the form of cooperation and competition.
+In addition, we have shown how to treat interspecies interactions in the form of cooperation and competition.
 By constructing an \ac{abm}, we have exemplified how close the devised model is to the individual biological processes that occur within the bacteria.
@@
 By modeling effects such as mutual inhibition on a single-cell level within an \ac{abm}, we have exemplified that the pool model directly builds on these more fundamental concepts.
-It can thus aid in the mechanistic understanding of the underlying processes, allowing researchers to deepen their konwledge about microbial growth within food science.
+It can thus aid in the mechanistic understanding of the underlying processes, allowing researchers to deepen their knowledge about microbial growth within food science.
 Going forward, this framework can provide a starting point for a more abstract mathematical formulation which should be able to cover a much wider class of systems.
@@
 Now we assume homogeneity in the sense, that we can reduce $[R]$ to only a single variable instead of considering its spatial distribution.
-This means that every contribution does not depend on the position at which it was evaluated, ie. $[R]_{x=x_c}=[R]$.
+This means that every contribution does not depend on the position at which it was evaluated, i.e. $[R]_{x=x_c}=[R]$.
 Finally, we can perform the sum to gain the total volume of all cells $V=\sum\limits_c V_c$.
@@
 The preceding equations all assume that cells are in growth phase.
-It is however straiforward, to setup these equations.
+It is however straightforward, to set up these equations.
 %
@@
 Since backend of the simulation is written in Rust, so is the implementation of cellular aspects.
-Additionally bindings are generated such that the full power of the simulation tool can be accessed.
+Additionally, bindings are generated such that the full power of the simulation tool can be accessed.
 This significantly improves handling and development time of new models.
@@
 % ................................................................................................
-\subsubsection{Homogenous and Inhomogenous Initial Cell Distributions}
-To clarify what we mean by homogenous and inhomogenous initial configurations, we include snapshots of the simulation.
+\subsubsection{Homogeneous and Inhomogeneous Initial Cell Distributions}
+To clarify what we mean by homogeneous and inhomogeneous initial configurations, we include snapshots of the simulation.
 At this point, cells have already grown such that the sizes of cell-clumps are better visible by the naked eye.
@@
     \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures-abm-homogenous-images-snapshot_00012000.png}
-        \subcaption{Homogenous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
+        \includegraphics[width=\textwidth]{Figures-abm-homogeneous-images-snapshot_00012000.png}
+        \subcaption{Homogeneous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}%
     \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures-abm-homogenous-low-diffusion-images-snapshot_00012000.png}
-        \subcaption{Homogenous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
+        \includegraphics[width=\textwidth]{Figures-abm-homogeneous-low-diffusion-images-snapshot_00012000.png}
+        \subcaption{Homogeneous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}\\

### 2025-10-31 paper/pool_model.tex
tex: update abstract

@@
     Bacterial consortia are complex systems which are often comprised of multiple distinct species.
-    Their dynamics on a population-based level often result from interactions that lead to
-    interspecies cooperation or competition.
-    Various empirical and inactivation models have been conceived and applied to problems related to
-    food microbiology.
-    However, the utilized models live on the population level, omitting any spatial effects such as
-    heterogeneity or diffusion.
-    In this work we present a novel generalized framework, called "pool model" which uses multiple
-    distinct pools that can hold bacterial states of distinct species or extracellular nutrient
-    abundance.
-    It allows us to describe these systems mechanistically as variations of an underlying, more
-    fundamental model.
-    To illustrate the usefulness of the model, multiple different examples are reported.
-    Finally, we construct an \acl{abm} for a given pool model, showcasing that the resulting model
+    Their dynamics on a population-based level often result from various intra- and inter-species interactions,
+    including cooperation or competition.
+    Various empirical and mechanistic models have been conceived and applied to problems related to food microbiology.
+    In this work we present a generalized framework, which uses multiple distinct "pools" that can distinguish
+    between microbial types in different metabolic states or describe the extracellular environment properties.
+    It allows us to incorporate the underlying biological processes and construct the \ac{ode} models of
+    different complexity levels.
+    To illustrate the application scope of the model, multiple different examples of the microbial systems
+    are reported.
+    Finally, to explore the limitations of the \ac{ode} formulation, we analyze the influence of the spatial effects.
+    For this, we construct an \acl{abm} for a two-species system, showcasing that the resulting pool model
     arises naturally from the individual behaviour of many cellular agents.
-    We further apply this \ac{abm} to study the role of spatial effects on the dynamics of bacterial
-    growth.
+    We further apply this \ac{abm} to study the role of properties such as diffusion and spatial heterogeneity
+    on the dynamics of bacterial growth.
 \end{abstract}

### 2025-10-31 paper/pool_model.tex
tex: update discussion

@@
 Figure~\ref{fig:spatial-snapshots} shows snapshots of a numerical simulation of the described system.
-The intial bacteria have been ordered on a grid such that only every diagonal neighbor is of the same species, thus resulting in a more homogenous initial distribution.
+The initial bacteria have been ordered on a grid such that only every diagonal neighbor is of the same species, thus resulting in a more homogenous initial distribution.
 %
@@
 
-The pool model presented here was used to describe the dynamics of 6 systems containing established models of bacterial growth.
-Comparing our approach with the Baranyi-Roberts model which is widely used in food science, it is possible to obtain results which provide growth curves of very similar shape.
-However, their analytical form is not completely identical.
+The presented pool model was used to describe the dynamics of one- and two-species systems containing the main growth stages and interactions between bacteria in the form of cooperation and competition.
+\cmt{Comparing our approach with the Baranyi-Roberts model which is widely used in food science, it is possible
+to obtain results that provide growth curves of very similar shape. However, their analytical form is not completely identical.}
 The inclusion of additional metabolic processes or larger number of cell-species can be achieved by extending the list of possible states and defining transitions between them.
-In addition, we have shown how to treat interspecies interactions in the form of cooperation and competition.
-By constructing an \ac{abm}, we have exemplified how close the devised model is to the individual biological processes that occur within the bacteria.
+As an example, the description of the quorum sensing phenomena can be modelled by adding the equation for the released autoinducer molecules as a part of the microenvironment.
+These molecules then affect the same-species bacteria growth controlling the size of the population.
+The described above approach provides a primary predictive model with transition rates between different pools as parameters.
+To include the influence of external factors, i.e.
+temperature, pH, pressure, etc., one can simply use the secondary models for the transition rates, e.g., the Ratkowsky model for temperature dependence~\cite{ratkowsky_relationship_1982}.\\
+By constructing an \ac{abm}, we have exemplified how close the devised model is to the individual biological 
+processes that occur within the bacteria.
 This mechanistic single-cell level of understanding of our \ac{abm} is a perspective which can not be easily replicated by other spatial models such as \acp{pde}.
-We have investigated the effect of spatial effects in the form of varying diffusion parameters and different initial spatial distributions of bacteria.
+We have investigated the impact of spatial effects in the form of varying diffusion parameters and different initial spatial distributions of bacteria.
 Our findings reinforce that the \ac{ode} model is a good simplification in the highly-diffusive regime within a mostly homogenous spatial bacterial distribution.
@@
 
-We introduced the pool model as a novel framework to unify existing approaches and allows researchers to follow a mechanistic approach during model development.
+We introduced the pool model as a framework to unify existing approaches and allows researchers to follow a mechanistic approach during model development.
 It is able to capture the effects of many established models and provides a lens which is useful to generalize existing systems and results in better overall explainability.

### 2025-10-31 paper/pool_model.tex
use documentclass from folder
fix typo in image name

@@
 % \documentclass[10pt,A4paper]{article}
-\documentclass[10pt,twocolumn,5p]{elsarticle}
+\documentclass[10pt,twocolumn,5p]{elsarticle/elsarticle}
 \usepackage[utf8]{inputenc}
@@
     \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures-abm-homogeneous-images-snapshot_00012000.png}
+        \includegraphics[width=\textwidth]{Figures-abm-homogenous-images-snapshot_00012000.png}
         \subcaption{Homogeneous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
@@
     \begin{subfigure}[c]{0.5\textwidth}
-        \includegraphics[width=\textwidth]{Figures-abm-homogeneous-low-diffusion-images-snapshot_00012000.png}
+        \includegraphics[width=\textwidth]{Figures-abm-homogenous-low-diffusion-images-snapshot_00012000.png}
         \subcaption{Homogeneous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}

### 2025-10-31 paper/pool_model.tex
format citations
add linenumbers
add frontmatter
update (inline) citations

@@
 % \documentclass[10pt,A4paper]{article}
-\documentclass[10pt,twocolumn,5p]{elsarticle/elsarticle}
+\documentclass[10pt,twocolumn,5p,authoryear]{elsarticle/elsarticle}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{listings-rust}
+\usepackage{lineno}
 \usepackage{mathtools}
@@
 
+\linenumbers
+
 % Must be loaded after tikz
@@
 
-\bibliographystyle{abbrvnat}
+\bibliographystyle{elsarticle/elsarticle-harv}
 
@@
 
+%%Graphical abstract
+\begin{graphicalabstract}
+%\includegraphics{grabs}
+\end{graphicalabstract}
+
+\begin{highlights}
+    \item @Christian fill this
+\end{highlights}
+
+\begin{keyword}
+    Bacterial Consortia
+    \sep Competition and Cooperation
+    \sep Mechanistic Modeling
+    \sep Spatial Effects
+    \sep Predictive Modeling
+\end{keyword}
+
 \maketitle
-
-\begin{graphicalabstract}
-% TODO
-\end{graphicalabstract}
-
-\begin{keyword}
-Bacterial Consortia \sep Competition and Cooperation \sep Mechanistic Modeling \sep Spatial Effects
-\end{keyword}
 %
@@
 %     \item Lag Phase, Competition, Cooperation, Limiting Resource
-%     \item Comprehensive reviews about competition/cooperation bacterial mechanisms: \cite{ghoul_ecology_2016, stubbendieck_bacterial_2016, hibbing_bacterial_2010, west_social_2007}
+%     \item Comprehensive reviews about competition/cooperation bacterial mechanisms: \citep{ghoul_ecology_2016, stubbendieck_bacterial_2016, hibbing_bacterial_2010, west_social_2007}
 %     \item Spatial Effects: adhesion, migration, chemotaxis, diffusion
 %     \item Quorum-sensing
-%     \item Biofilm formation \cite{Jin2020}
+%     \item Biofilm formation \citep{Jin2020}
 % \end{itemize}
@@
 all relevant biological phenomena arising in these bacterial systems.
-During their reproductive process, bacteria undergo multiple phases~\cite{buchanan_when_1997}, starting in a lag phase which
-is necessary to adapt to new environments and initially impedes the overall growth~\cite{rolfe_lag_2012, monod_growth_1949}.
+During their reproductive process, bacteria undergo multiple phases~\citep{buchanan_when_1997}, starting in a lag phase which
+is necessary to adapt to new environments and initially impedes the overall growth~\citep{rolfe_lag_2012, monod_growth_1949}.
 Additionally, bacteria are able to cooperate e.g. via complementary metabolic processes or inhibit
-each other e.g. by means of waste production as a metabolic byproduct~\cite{west_social_2007, rainey_evolution_2003, hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
-A limited supply of nutrients will also affect the overall reproductive capacity of the colony~\cite{monod_growth_1949}.
+each other e.g. by means of waste production as a metabolic byproduct~\citep{west_social_2007, rainey_evolution_2003, hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
+A limited supply of nutrients will also affect the overall reproductive capacity of the colony~\citep{monod_growth_1949}.
 Bacteria have thus devised various survival strategies to ensure sufficient access to nutrients.
-Effects on a single-cell level such as adhesion~\cite{htuson_bacteriasurface_2013}, migration~\cite{decoene_microscopic_2011} and chemotaxis can lead to complex
+Effects on a single-cell level such as adhesion~\citep{htuson_bacteriasurface_2013}, migration~\cite{decoene_microscopic_2011} and chemotaxis can lead to complex
 phenomena on a population-based level which can aid in the goal of survival.
 Cell signaling e.g. in the form of quorum-sensing can lead to collective emergent phenomena such as
-biofilm formation which play an important role for many mature microbial communities~\cite{Jin2020,ng_bacterial_2009}.
+biofilm formation which play an important role for many mature microbial communities~\citep{Jin2020,ng_bacterial_2009}.
 
 % \paragraph{ODE models (predictive microbiology, existing, overview, most popular)}
-% Existing models in Predictive Microbiology (Primary Models)~\cite{perez-rodriguez_predictive_2012}:
+% Existing models in Predictive Microbiology (Primary Models)~\citep{perez-rodriguez_predictive_2012}:
 % \begin{itemize}
-%     \item An overview of primary models used in food science can be found in~\cite{van_boekel_kinetic_2008}.
-%     \item Baranyi-Roberts Model~\cite{baranyi_modeling_1993, baranyi_dynamic_1994}
-%     \item modified Gompertz~\cite{Gompertz1825-wi,Gibson1987}
-%     \item Buchanan et al.: three-phase linear model for lag phase, exponential growth phase, and stationary phase~\cite{buchanan_when_1997}
-%     \item McKellar two-compartment model with bacteria in growing and non-growing state~\cite{mckellar_heterogeneous_1997}
+%     \item An overview of primary models used in food science can be found in~\citep{van_boekel_kinetic_2008}.
+%     \item Baranyi-Roberts Model~\citep{baranyi_modeling_1993, baranyi_dynamic_1994}
+%     \item modified Gompertz~\citep{Gompertz1825-wi,Gibson1987}
+%     \item Buchanan et al.: three-phase linear model for lag phase, exponential growth phase, and stationary phase~\citep{buchanan_when_1997}
+%     \item McKellar two-compartment model with bacteria in growing and non-growing state~\citep{mckellar_heterogeneous_1997}
 %     \item Inactivation models
@@
 %     \item Polynomial models 
-%     \item Square Root-Type Models, e.g., Ratkowski for temperature~\cite{ratkowsky_relationship_1982}
-%     \item The Cardinal Parameter Model~\cite{zwietering_decision_1992}
+%     \item Square Root-Type Models, e.g., Ratkowski for temperature~\citep{ratkowsky_relationship_1982}
+%     \item The Cardinal Parameter Model~\citep{zwietering_decision_1992}
 % \end{itemize}
@@
 The field of predictive microbiology has developed a variety of ODE-based models to capture
-bacterial growth dynamics~\cite{perez-rodriguez_predictive_2012}.
+bacterial growth dynamics~\citep{perez-rodriguez_predictive_2012}.
 Various \acp{ode} \textit{primary models} describe the bacterial growth curve under constant
 environmental conditions.
-Widely used examples include the modified Gompertz model~\cite{Gompertz1825-wi,Gibson1987}, Baranyi-Roberts
-model~\cite{baranyi_modeling_1993, baranyi_dynamic_1994} and McKellar two-compartment model~\cite{buchanan_when_1997} which distinguishes between growing and non-growing bacteria.
+Widely used examples include the modified Gompertz model~\citep{Gompertz1825-wi,Gibson1987}, Baranyi-Roberts
+model~\citep{baranyi_modeling_1993, baranyi_dynamic_1994} and McKellar two-compartment model~\cite{buchanan_when_1997} which distinguishes between growing and non-growing bacteria.
 To take into account changing environmental conditions, so-called \textit{secondary models}
@@
 humidity.
-These can be simple polynomial models, square root-type models~\cite{ratkowsky_relationship_1982} or
-more complex relations such as the cardinal parameter model~\cite{zwietering_decision_1992}.
+These can be simple polynomial models, square root-type models~\citep{ratkowsky_relationship_1982} or
+more complex relations such as the cardinal parameter model~\citep{zwietering_decision_1992}.
 The advantage of the \ac{ode} formulation lies in the predictive power and simplicity of usage.
@@
 % \begin{itemize}
-%     \item BSim \cite{Gorochowski2012,Dang2020}
-%     \item Bacterial Branching patterns \cite{Kawasaki1997,Matsushita1998}
+%     \item BSim \citep{Gorochowski2012,Dang2020}
+%     \item Bacterial Branching patterns \citep{Kawasaki1997,Matsushita1998}
 % \end{itemize}
@@
 lead to more complex collective phenomena.
-Tools such as BSim~\cite{Gorochowski2012,Dang2020}, AgentCell~\cite{emonet_agentcell_2005}, BacSim~\cite{kreft_bacsim_1998} or \texttt{cellular\_raza}~\cite{Pleyer2025} are
+Tools such as BSim~\citep{Gorochowski2012,Dang2020}, AgentCell~\cite{emonet_agentcell_2005}, BacSim~\cite{kreft_bacsim_1998} or \texttt{cellular\_raza}~\cite{Pleyer2025} are
 capable to describe large amounts of bacteria, performing various intracellular and extracellular
@@
 The latter has been used to adapt \ac{pde}-based models into an \ac{abm} formulation which describe
-bacterial branching patterns~\cite{Kawasaki1997,Matsushita1998}.
+bacterial branching patterns~\citep{Kawasaki1997,Matsushita1998}.
 
@@
 understood on an individual-based level in terms of single-cell processes.
-Furthermore, we assess how spatial heterogeneity~\cite{mckellar_heterogeneous_1997} can impact the shape of the growth curves.
+Furthermore, we assess how spatial heterogeneity~\citep{mckellar_heterogeneous_1997} can impact the shape of the growth curves.
 
@@
 We begin our discussion by considering a simplistic system where only one bacteria species is present $N=1$.
-Bacterial growth undergoes different consecutive phases~\cite{buchanan_when_1997}, starting with the so called lag-phase, during which no increase in the bacterial count occurs.
+Bacterial growth undergoes different consecutive phases~\citep{buchanan_when_1997}, starting with the so called lag-phase, during which no increase in the bacterial count occurs.
 This phenomenon appears due to adaptations of the microbial community, for example, to the new environmental conditions.
-This so-called preparation before the exponential growth includes synthesis of the needed compounds, initiation of the transcription~\cite{rolfe_lag_2012} and poorly studied physiological and regulatory mechanisms~\cite{monod_growth_1949}.
+This so-called preparation before the exponential growth includes synthesis of the needed compounds, initiation of the transcription~\citep{rolfe_lag_2012} and poorly studied physiological and regulatory mechanisms~\citep{monod_growth_1949}.
 After this, the bacteria which have succeeded in this adaptation stage enter the exponential growth phase, where the process of cell division is predominant.
@@
 At the same time dying cells release the nutrients for survivors to continue growing.
-As a result the balance between the growing and dying processes leads to a constant bacterial count~\cite{navarro_llorens_stationary_2010}.
+As a result the balance between the growing and dying processes leads to a constant bacterial count~\citep{navarro_llorens_stationary_2010}.
 In the described simplified version of the real-life system three possible bacterial states are mentioned, corresponding to the underlying process: adaptation, growth or death.
@@
 Based on experimental observations, it was shown that the bacterial growth is strongly influenced by the limiting resource.
-According to the works of Jacques Monod, in the system where all needed nutrients are in excess except for one, the total growth is linear to the concentration of this limiting nutrient resource~\cite{monod_growth_1949}.
+According to the works of \citet{monod_growth_1949} in the system where all needed nutrients are in excess except for one, the total growth is linear to the concentration of this limiting nutrient resource.
 The control of the population density with respect to the available nutrition is implemented using cell communication processes called quorum sensing.
-By releasing extracellular signalling molecules (autoinducer molecules) it is possible for the bacteria to synchronize the gene expression of the whole group~\cite{ng_bacterial_2009}.
+By releasing extracellular signalling molecules (autoinducer molecules) it is possible for the bacteria to synchronize the gene expression of the whole group~\citep{ng_bacterial_2009}.
 We will not dive into the processes of quorum sensing and try to keep it as simple as possible.
@@
 Bacteria which have lived under rather unfavorable conditions have developed different surviving mechanisms, one of which is adapting their metabolism and transitioning into dormant state.
-Dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\cite{kaprelyants_dormancy_1993}.
+Dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\citep{kaprelyants_dormancy_1993}.
 This survival strategy can appear in response to different external stresses, e.g., nutrient deficiency, temperature stress or extreme fluctuations of environment parameters.
-And after "difficult times" bacteria, also called viable but non-culturable (VBNC), can continue cell division~\cite{kell_viability_1998}.
-Multiple models which include transitions of bacteria to the dormant state were proposed by Bär \etal~\cite{bar_modelling_2002} and Jones with Lennon~\cite{jones_dormancy_2010}.
+And after "difficult times" bacteria, also called viable but non-culturable (VBNC), can continue cell division~\citep{kell_viability_1998}.
+Multiple models which include transitions of bacteria to the dormant state were proposed by \citet{bar_modelling_2002} and \citet{jones_dormancy_2010}.
 %
@@
 \paragraph{Maxwell type of stress - strain relation}
-One of the reasons that bacteria enter the dormant state can be due to a shift in temperature~\cite{oliver_viable_1995}.
-Moreover, a rapid shift in temperature is associated with the appearance of an additional lag phase~\cite{zwietering_modeling_1994}.
+One of the reasons that bacteria enter the dormant state can be due to a shift in temperature~\citep{oliver_viable_1995}.
+Moreover, a rapid shift in temperature is associated with the appearance of an additional lag phase~\citep{zwietering_modeling_1994}.
 To model it we propose a kind of viscoelastic Maxwell type of stress-strain relation and write down the ordinary differential equation for $\gamma$ to be:
@@
 As mentioned above, bacteria usually coexist in large interacting communities of multiple species whose growth is vastly determined by the developed survival mechanisms.
-These mechanisms mostly aim to secure as much available nutrition which lead to competitive strategies or, if it is advantageous, cooperative ones~\cite{hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
+These mechanisms mostly aim to secure as much available nutrition which lead to competitive strategies or, if it is advantageous, cooperative ones~\citep{hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
 Further we show that the model can be readily extended to the systems with more than one bacterium and describe more complex relationships between species.
@@
 \subsection{Resource Competition}
-One of the ways to describe competition between bacterial populations is as a function of the limiting resource ratio~\cite{tilman_resource_1977, smith_effects_2002,ghoul_ecology_2016}.
+One of the ways to describe competition between bacterial populations is as a function of the limiting resource ratio~\citep{tilman_resource_1977, smith_effects_2002,ghoul_ecology_2016}.
 To this end we consider for simplicity only one common limiting resource pool.
@@
 Equations (\ref{eq:LV_simple}) can be understood as competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competitive excursion.
-That principle claims that two species which compete for the same resource cannot coexist~\cite{wangersky_lotka-volterra_1978}.
-In previous studies it was shown that \ac{glv} equations with one bacterial state can be successfully applied to predict the microbial community dynamics, determine different types of the pairwise interspecies interactions~\cite{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
-In food studies the \ac{glv} was applied by Mounier \etal~\cite{mounier_microbial_2008} to study interactions of six bacteria species influenced by yeast omissions in cheese.
-Cauchie \etal~\cite{cauchie_modeling_2020} applied Lotka-Volterra model to study the three-species system of the bacteria majorly present in minced pork meat.\\
+That principle claims that two species which compete for the same resource cannot coexist~\citep{wangersky_lotka-volterra_1978}.
+In previous studies it was shown that \ac{glv} equations with one bacterial state can be successfully applied to predict the microbial community dynamics, determine different types of the pairwise interspecies interactions~\citep{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
+In food studies the \ac{glv} was applied by \citet{mounier_microbial_2008} to study interactions of six bacteria species influenced by yeast omissions in cheese.
+\citet{cauchie_modeling_2020} applied Lotka-Volterra model to study the three-species system of the bacteria majorly present in minced pork meat.\\
 %
 However, in contrast to the classical Lotka-Volterra approach, the introduced pool model allows for far more flexibility in tuning properties of the bacterial interactions within and between species.
-For further comparison of the derived model with the matrix formulation of the \ac{glv} we refer the reader to the supplement materials~\ref{ssec:supplement1}.
+For further comparison of the derived model with the matrix formulation of the \ac{glv} we refer the reader to the Appendix~\ref{ssec:supplement1}.
 %
@@
 Mutual inhibition between different species can not only result from competition for a shared resource for which bacteria have developed various other active competition mechanisms.
-One of the examples of such mechanisms is the production of antimicrobial compounds, inhibitors or toxins~\cite{wloch-salamon_effect_2008, chao_structured_1981}.
+One of the examples of such mechanisms is the production of antimicrobial compounds, inhibitors or toxins~\citep{wloch-salamon_effect_2008, chao_structured_1981}.
 On the other hand, some bacteria can produce public goods, the compounds that benefit growth of other species.
-In this way, they can suppress the non-preferred species by helping their competitors~\cite{hibbing_bacterial_2010}.
+In this way, they can suppress the non-preferred species by helping their competitors~\citep{hibbing_bacterial_2010}.
 %
@@
 Another special case of the species competition is when the produced toxin does not cause the death of the bacteria but rather inhibit its growth.
-Such situation with 2 bacteria species, a shared resource and produced inhibitor was modeled, e.g., by Lenski and Hattingh~\cite{lenski_coexistence_1986}.
+Such situation with 2 bacteria species, a shared resource and produced inhibitor was modeled, e.g., by \citet{lenski_coexistence_1986}.
 In our case, we would like to look at a rather simplified case reducing the number of variables of the system.
@@
 It is most often observed between species which are close relatives or clonal populations.
-This process is called kin selection~\cite{west_social_2007}.
-Furthermore, neighbors can also benefit from a behavior which does not require any action on an expression level, so-called 'social cheaters'~\cite{rainey_evolution_2003}.
+This process is called kin selection~\citep{west_social_2007}.
+Furthermore, neighbors can also benefit from a behavior which does not require any action on an expression level, so-called 'social cheaters'~\citep{rainey_evolution_2003}.
 This mutual activation benefits all involved species and can be easily included in our mathematical framework.
@@
 species to develop competitive mechanisms such as cell adhesion to surfaces in the areas rich with
-nutrients~\cite{htuson_bacteriasurface_2013}.
+nutrients~\citep{htuson_bacteriasurface_2013}.
 Another mechanism is the motility of bacteria, e.g., the flagellates \textit{E. coli} or
 \textit{B. subtilis} which are capable of chemotactic motion driven by oxygen
-concentrations~\cite{decoene_microscopic_2011}.
+concentrations~\citep{decoene_microscopic_2011}.
 The bacterial social interactions such as release of toxins or public goods can also
-significantly affect the spatial structure of the community~\cite{blanchard_bacterial_2015}.
+significantly affect the spatial structure of the community~\citep{blanchard_bacterial_2015}.
 These effects can only be treated with a spatial model that is able to describe the observed
 effects.\\
-Compared to other modeling techniques such as \acp{pde}, \acp{abm}~\cite{nagarajan_agent-based_2022-1} force researchers to think in
+Compared to other modeling techniques such as \acp{pde}, \acp{abm}~\citep{nagarajan_agent-based_2022-1} force researchers to think in
 terms of cellular processes, thus bridging the gap from a
@@
 %
-% - \cite{emonet_agentcell_2005, gorochowski_bsim_2012, li_nufeb_2019, kreft_bacsim_1998}: examples of another toolboxes for bacterial communities modeling (ABM).\\
-% More about the modelling of the motile bacteria can be seen in works:~\cite{decoene_microscopic_2011, rosser_modelling_2014, sokolov_physical_2012, li_amplified_2008}.
-%
-%
-% In general, bacterial communities in 2D can experience collective motion called swarming~\cite{wu_collective_2015}.
+% - \citep{emonet_agentcell_2005, gorochowski_bsim_2012, li_nufeb_2019, kreft_bacsim_1998}: examples of another toolboxes for bacterial communities modeling (ABM).\\
+% More about the modelling of the motile bacteria can be seen in works:~\citep{decoene_microscopic_2011, rosser_modelling_2014, sokolov_physical_2012, li_amplified_2008}.
+%
+%
+% In general, bacterial communities in 2D can experience collective motion called swarming~\citep{wu_collective_2015}.
 % Some works modelled 2D motile bacteria:
 % \begin{itemize}
-%     \item \cite{decoene_microscopic_2011}: 2D chemotactic self-propelled bacterial motion of the flagellated bacteria (e.g. E. coli, B. subtilis).
+%     \item \citep{decoene_microscopic_2011}: 2D chemotactic self-propelled bacterial motion of the flagellated bacteria (e.g. E. coli, B. subtilis).
 %     Chemotaxis here was driven by the oxygen concentrations.
-%     \item \cite{rosser_modelling_2014}: modeling of the bacterial reorintation, rotation.
-%     \item \cite{sokolov_physical_2012, li_amplified_2008}: modeling of the bacteria using hydrodynamic interactions and Brownian motion.
+%     \item \citep{rosser_modelling_2014}: modeling of the bacterial reorintation, rotation.
+%     \item \citep{sokolov_physical_2012, li_amplified_2008}: modeling of the bacteria using hydrodynamic interactions and Brownian motion.
 %     Fluctuations in velocity and rotations.
-%     (\cite{li_amplified_2008}: Brownian motion gives right motion observed using microscopy).
-%     \item \cite{farrell_mechanically_2013}: 2D modeling using Newtonian dynamics and only mechanical interactions + nutrient diffusion.
+%     (\citep{li_amplified_2008}: Brownian motion gives right motion observed using microscopy).
+%     \item \citep{farrell_mechanically_2013}: 2D modeling using Newtonian dynamics and only mechanical interactions + nutrient diffusion.
 %     Branching parameter defines diff limits where diffusion dominant and limits growth, and where it's irrelevant.
@@
 % \begin{itemize}
-%     \item \cite{farrell_mechanically_2013}: Showed the influence of mechanic interactions between bacteria in 2D system with bacteria consuming the nutrient.
+%     \item \citep{farrell_mechanically_2013}: Showed the influence of mechanic interactions between bacteria in 2D system with bacteria consuming the nutrient.
 %     The growth of bacterial colonies where non-motile microorganisms replicate and push each other away as they grow.
@@
 %     (Structurally in 2D it's regimes: smooth propagating of bacterial front vs. branching.)
-%     \item \cite{nagarajan_agent-based_2022-1}: review on the ABM for bacterial population.
+%     \item \citep{nagarajan_agent-based_2022-1}: review on the ABM for bacterial population.
 %     The ABM represent a microbial colony as a group of discrete agents that are governed by a set of rules, where each agent can be an individual cell or a collection of cells.
@@
 \ac{ode}-based systems.\\
-To construct our \ac{abm}, we use \texttt{cellular\_raza}~\cite{Pleyer2025}.
+To construct our \ac{abm}, we use \texttt{cellular\_raza}~\citep{Pleyer2025}.
 It allows us to define our bacterial agents from a clean slate, without having to rely on and inherit
@@
 \end{alignat}
-We assume that the cells are not sticky and consider only a repulsive interaction potential, leaving out any possible attractive forces~\cite{Garrett2008}.
+We assume that the cells are not sticky and consider only a repulsive interaction potential, leaving out any possible attractive forces~\citep{Garrett2008}.
 These mechanisms will result in the overall expansion of the colony as individual cells push each further other apart.
@@
 \subsubsection{Cell Cycle}
-Typically the cycle of a cell consist of multiple phases~\cite{Cooper2000_fn}.
+Typically the cycle of a cell consist of multiple phases~\citep{Cooper2000_fn}.
 Since our goal is to include only the necessary details of the underlying biological problem, we drastically simplify the description of the cell cycle.
@@
 The \ac{abm} model should be considered to be the more fundamental model compared to the \ac{ode} model in the sense that it incorporates more biological aspects and maps them closer to the actual behaviour of the bacteria.
-We are thus able to derive the \ac{ode} pool model from its \ac{abm} version via a process of homogenization (see supplementary material~\ref{subsec:supplement-deriving-pool-odes}).
+We are thus able to derive the \ac{ode} pool model from its \ac{abm} version via a process of homogenization (see Appendix~\ref{subsec:supplement-abm-details}).
 By doing so, we find a mapping between parameters from the \ac{abm} to parameters used in the \ac{ode} description.
@@
 Heterogeneity and slower diffusion cause larger differences between the two models as can be seen in subfigures (b) and (c).
-The combination of the two effects shows the largest difference between model predictions.
-In subfigure (c), due to the heterogeneous cell distribution (see also~\ref{fig:supplement-spatial-initial-config-cells}), overall inhibition of species $B$ is significantly reduced.
+The combination of the two effects shows the largest difference between model outputs.
+In subfigure (c), due to the heterogeneous cell distribution (see Appendix~\ref{fig:supplement-spatial-initial-config-cells}), overall inhibition of species $B$ is significantly reduced.
 Subfigure (d) shows how a randomized homogenous initial distribution can lead to even more notable effects of inhibition.
@@
         The diffusion constants are chosen such that $D_R=D_I=D$.
-        Snapshots of the spatial configurations of cells are included in the appendix~\ref{fig:supplement-spatial-initial-config-cells}.
+        Snapshots of the spatial configurations of cells are included in the Appendix~\ref{fig:supplement-spatial-initial-config-cells}.
     }
@@
 To include the influence of external factors, i.e.
-temperature, pH, pressure, etc., one can simply use the secondary models for the transition rates, e.g., the Ratkowsky model for temperature dependence~\cite{ratkowsky_relationship_1982}.\\
+temperature, pH, pressure, etc., one can simply use the secondary models for the transition rates, e.g., the Ratkowsky model for temperature dependence~\citep{ratkowsky_relationship_1982}.\\
 By constructing an \ac{abm}, we have exemplified how close the devised model is to the individual biological 
@@
 \beginsupplement
-\section{Supplementary Material}
+\section{Appendix}
 % ================================================================================================
-\subsection{General Mathematical Formalism}  \label{ssec:supplement1}
+\subsection{General Mathematical Formalism}
+\label{ssec:supplement1}
 Consider the system consisting of $N$ species in $M$ different states.
@@
 
-\subsection{ABM Details} \label{ssec:supplement2}
+\subsection{ABM Details}
+\label{subsec:supplement-abm-details}
+
 \subsubsection*{Deriving the Pool Model}
-\label{subsec:supplement-deriving-pool-odes}
 Since the \ac{abm} can be considered more fundamental than the \ac{ode} model, we should be able to derive the latter one from the former.

### 2025-10-31 paper/pool_model.tex
soften sentence wrt. Baranyi-Roberts

@@
 The presented pool model was used to describe the dynamics of one- and two-species systems containing the main growth stages and interactions between bacteria in the form of cooperation and competition.
-\cmt{Comparing our approach with the Baranyi-Roberts model which is widely used in food science, it is possible
-to obtain results that provide growth curves of very similar shape. However, their analytical form is not completely identical.}
+Our approach produces results which exhibit similarities to other well-known predictive microbiology models such as the Baranyi-Roberts model while providing better explainability.
 The inclusion of additional metabolic processes or larger number of cell-species can be achieved by extending the list of possible states and defining transitions between them.

### 2025-10-31 paper/pool_model.tex
fix formatting of i.e. and e.g.

@@
     between microbial types in different metabolic states or describe the extracellular environment properties.
-    It allows us to incorporate the underlying biological processes and construct the \ac{ode} models of
+    It allows us to incorporate the underlying biological processes and construct \ac{ode} based models of
     different complexity levels.
-    To illustrate the application scope of the model, multiple different examples of the microbial systems
+    To illustrate the application scope of the model, multiple different examples of microbial systems
     are reported.
-    Finally, to explore the limitations of the \ac{ode} formulation, we analyze the influence of the spatial effects.
-    For this, we construct an \acl{abm} for a two-species system, showcasing that the resulting pool model
+    Finally, to explore the limitations of the \ac{ode} formulation, we analyze the influence of spatial effects.
+    For this, we construct an \ac{abm} for a two-species system, showcasing that the resulting pool model
     arises naturally from the individual behaviour of many cellular agents.
@@
 is necessary to adapt to new environments and initially impedes the overall growth~\citep{rolfe_lag_2012, monod_growth_1949}.
-Additionally, bacteria are able to cooperate e.g. via complementary metabolic processes or inhibit
-each other e.g. by means of waste production as a metabolic byproduct~\citep{west_social_2007, rainey_evolution_2003, hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
+Additionally, bacteria are able to cooperate, e.g., via complementary metabolic processes or inhibit
+each other, e.g., by means of waste production as a metabolic byproduct~\citep{west_social_2007, rainey_evolution_2003, hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
 A limited supply of nutrients will also affect the overall reproductive capacity of the colony~\citep{monod_growth_1949}.
@@
 phenomena on a population-based level which can aid in the goal of survival.
-Cell signaling e.g. in the form of quorum-sensing can lead to collective emergent phenomena such as
+Cell signaling, e.g., in the form of quorum-sensing can lead to collective emergent phenomena such as
 biofilm formation which play an important role for many mature microbial communities~\citep{Jin2020,ng_bacterial_2009}.
@@
 %     \item Polynomial models 
-%     \item Square Root-Type Models, e.g., Ratkowski for temperature~\citep{ratkowsky_relationship_1982}
+%     \item Square Root-Type Models, e.g.,, Ratkowski for temperature~\citep{ratkowsky_relationship_1982}
 %     \item The Cardinal Parameter Model~\citep{zwietering_decision_1992}
@@
 The dynamics of bacterial growth can be understood in terms of transitions between the different
-bacterial states i.e. pools.
+bacterial states, i.e., pools.
 Our model also allows us to capture interactions between these bacterial states and the external
@@
 Let say $G=N_B/V$ and $[G]=1/V$, where $N_B$ is the number of bacteria and $V$ is the volume.
-Then $R=N_R/V$ and $[R]=1/V$, where $N_R$ is the number of resource molecules (or any appropriate unit, e.g., mol).
+Then $R=N_R/V$ and $[R]=1/V$, where $N_R$ is the number of resource molecules (or any appropriate unit, e.g.,, mol).
 Then $[\alpha_0]=V/s$, $[N_t]=1/V$, and $[\alpha]=1/s$.
@@
 Note that the previously introduced model could be further simplified to a two pool or state model by omitting the pool $D$ and removing cells with rate $\beta$ directly from pool $L$.
-The relevant question here is whether cells in the pool $D$ still contribute to the limitation of the resources by, e.g., occupying space or consuming important nutrients.
+The relevant question here is whether cells in the pool $D$ still contribute to the limitation of the resources by, e.g.,, occupying space or consuming important nutrients.
 If yes, the three pool model would be advantageous, if not the two pool model is sufficient with the advantages of having one parameter less.\\
@@
 For constant parameters $\mu$ and $\lambda$ the integrals for $L$ and $D$ can be inserted into the differential equation for $G$.
-However, in the case of time dependent parameters, e.g., for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$.
+However, in the case of time dependent parameters, e.g.,, for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$.
 It is interesting to see what type of equations result for the total bacterial population $N=L+G+D$:
@@
 Dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\citep{kaprelyants_dormancy_1993}.
-This survival strategy can appear in response to different external stresses, e.g., nutrient deficiency, temperature stress or extreme fluctuations of environment parameters.
+This survival strategy can appear in response to different external stresses, e.g.,, nutrient deficiency, temperature stress or extreme fluctuations of environment parameters.
 And after "difficult times" bacteria, also called viable but non-culturable (VBNC), can continue cell division~\citep{kell_viability_1998}.
@@
 Another special case of the species competition is when the produced toxin does not cause the death of the bacteria but rather inhibit its growth.
-Such situation with 2 bacteria species, a shared resource and produced inhibitor was modeled, e.g., by \citet{lenski_coexistence_1986}.
+Such situation with 2 bacteria species, a shared resource and produced inhibitor was modeled, e.g.,, by \citet{lenski_coexistence_1986}.
 In our case, we would like to look at a rather simplified case reducing the number of variables of the system.
@@
 The function $\mathcal{B}_{A/B}$ captures the inhibiting effects effectively.
-It is straight forward to make the mechanism of inhibition more explicit by, e.g., introducing a toxin pool.
+It is straight forward to make the mechanism of inhibition more explicit by, e.g.,, introducing a toxin pool.
 One possible choice for $\mathcal{B}$ is:
@@
 nutrients~\citep{htuson_bacteriasurface_2013}.
-Another mechanism is the motility of bacteria, e.g., the flagellates \textit{E. coli} or
+Another mechanism is the motility of bacteria, e.g.,, the flagellates \textit{E. coli} or
 \textit{B. subtilis} which are capable of chemotactic motion driven by oxygen
@@
 % \begin{itemize}
-%     \item \citep{decoene_microscopic_2011}: 2D chemotactic self-propelled bacterial motion of the flagellated bacteria (e.g. E. coli, B. subtilis).
+%     \item \citep{decoene_microscopic_2011}: 2D chemotactic self-propelled bacterial motion of the flagellated bacteria (e.g., E. coli, B. subtilis).
 %     Chemotaxis here was driven by the oxygen concentrations.
@@
 %     The ABM represent a microbial colony as a group of discrete agents that are governed by a set of rules, where each agent can be an individual cell or a collection of cells.
-%     Can use citation when talk about general processes we are accounting as growth, uptake of nutrient, production of inhibitor, e.g.,:
+%     Can use citation when talk about general processes we are accounting as growth, uptake of nutrient, production of inhibitor, e.g.,,:
 %     Physical interactions between agents are calculated using excluded volume interactions.
@@
 The described above approach provides a primary predictive model with transition rates between different pools as parameters.
-To include the influence of external factors, i.e.
-temperature, pH, pressure, etc., one can simply use the secondary models for the transition rates, e.g., the Ratkowsky model for temperature dependence~\citep{ratkowsky_relationship_1982}.\\
+To include the influence of external factors, i.e., temperature, pH, pressure, etc., one can simply use the secondary models for the transition rates, e.g.,, the Ratkowsky model for temperature dependence~\citep{ratkowsky_relationship_1982}.\\
 By constructing an \ac{abm}, we have exemplified how close the devised model is to the individual biological 
@@
 with $x_{j}^{i}$ describing the bacterial counts for species $i$ in state $j$.
-$\mbu$ is the vector of different inputs of the model, e.g., temperature, pressure etc.
+$\mbu$ is the vector of different inputs of the model, e.g.,, temperature, pressure etc.
 The vector $\mbm \in \mathbb{R}^{K}$ includes all variables corresponding to the microenvironment of the system which can be concentration of the inhibitors or activators:
@@
 Now we assume homogeneity in the sense, that we can reduce $[R]$ to only a single variable instead of considering its spatial distribution.
-This means that every contribution does not depend on the position at which it was evaluated, i.e. $[R]_{x=x_c}=[R]$.
+This means that every contribution does not depend on the position at which it was evaluated, i.e., $[R]_{x=x_c}=[R]$.
 Finally, we can perform the sum to gain the total volume of all cells $V=\sum\limits_c V_c$.

### 2025-10-31 paper/pool_model.tex
make section title lowercase

@@
 % ================================================================================================
-\section{One Species Modeling}
+\section{One species modeling}
 % ------------------------------------------------------------------------------------------------
@@
 % ------------------------------------------------------------------------------------------------
-\subsection{Lag-time Estimation}
+\subsection{Lag-time estimation}
 Note that the previously introduced model could be further simplified to a two pool or state model by omitting the pool $D$ and removing cells with rate $\beta$ directly from pool $L$.
@@
 % ------------------------------------------------------------------------------------------------
-\subsection{Back flow from Growth to Lag pool}
+\subsection{Back flow from growth to lag pool}
 Bacteria which have lived under rather unfavorable conditions have developed different surviving mechanisms, one of which is adapting their metabolism and transitioning into dormant state.
@@
 % ------------------------------------------------------------------------------------------------
-\subsection{Resource Competition}
+\subsection{Resource competition}
 One of the ways to describe competition between bacterial populations is as a function of the limiting resource ratio~\citep{tilman_resource_1977, smith_effects_2002,ghoul_ecology_2016}.
@@
 % ------------------------------------------------------------------------------------------------
-\subsection{Interspecies Competition}
+\subsection{Interspecies competition}
 Mutual inhibition between different species can not only result from competition for a shared resource for which bacteria have developed various other active competition mechanisms.
@@
 % ................................................................................................
-\subsubsection{Toxins Production}
+\subsubsection{Toxin production}
 One direct form of inhibition is the production of toxins.
@@
 % ................................................................................................
-\subsubsection{Mutual Inhibition}
+\subsubsection{Mutual inhibition}
 Another special case of the species competition is when the produced toxin does not cause the death of the bacteria but rather inhibit its growth.
@@
 % ------------------------------------------------------------------------------------------------
-\subsection{Interspecies Cooperation}
+\subsection{Interspecies cooperation}
 On the other hand, one can also observe some cooperative behaviors between species in cases when it facilitates the competition with others.
@@
 % ================================================================================================
-\section{Spatial Effects}
+\section{Spatial effects}
 As described above, the proposed model operates with the total number of the bacterial counts for
@@
 % ------------------------------------------------------------------------------------------------
-\subsection{Constructing an Agent-Based Pool Model}
+\subsection{Constructing an agent-based pool model}
 The following sections will demonstrate how a given \ac{ode} system can be translated into a
@@
 
-\subsubsection{Cellular Reactions}
+\subsubsection{Cellular reactions}
 To construct an \ac{abm} which resembles the given \ac{ode} system of equations~\eqref{eq:spatial_limit_F} and \eqref{eq:spatial_limit_H}, we consider multiple cellular reactions.
@@
 % ................................................................................................
-\subsubsection{Physical Mechanics \& Interactions}
+\subsubsection{Physical mechanics \& interactions}
 We represent cells as spherical objects in 2D space.
@@
 % ................................................................................................
-\subsubsection{Cell Cycle}
+\subsubsection{Cell cycle}
 Typically the cycle of a cell consist of multiple phases~\citep{Cooper2000_fn}.
@@
 % ------------------------------------------------------------------------------------------------
-\subsection{Comparing \acs{abm} and \acs{ode} Models}
+\subsection{Comparing \acs{abm} and \acs{ode} models}
 The \ac{abm} model should be considered to be the more fundamental model compared to the \ac{ode} model in the sense that it incorporates more biological aspects and maps them closer to the actual behaviour of the bacteria.
@@
 \newpage
-\section*{Data Availability \& Reproducibility}
+\section*{Data availability \& reproducibility}
 
@@
 % ================================================================================================
-\subsection{General Mathematical Formalism}
+\subsection{General mathematical formalism}
 \label{ssec:supplement1}
@@
 
-\subsection{ABM Details}
+\subsection{\acs{abm} details}
 \label{subsec:supplement-abm-details}
 
-\subsubsection*{Deriving the Pool Model}
+\subsubsection*{Deriving the pool model}
 Since the \ac{abm} can be considered more fundamental than the \ac{ode} model, we should be able to derive the latter one from the former.
@@
 % ................................................................................................
-\subsubsection{Code Samples}
+\subsubsection{Code samples}
 The code samples show sample code from the simulation that calculate cellular interactions and cell cycle.
@@
 % ................................................................................................
-\subsubsection{Full List of Parameters}
+\subsubsection{Full list of parameters}
 \label{sec:supplement-full-parameters}
@@
 % ................................................................................................
-\subsubsection{Homogeneous and Inhomogeneous Initial Cell Distributions}
+\subsubsection{Homogeneous and inhomogeneous initial cell distributions}
 To clarify what we mean by homogeneous and inhomogeneous initial configurations, we include snapshots of the simulation.

### 2025-10-31 paper/pool_model.tex
formatting first section

@@
 % ------------------------------------------------------------------------------------------------
-\subsection{Three-Pool Model}
-We begin our discussion by considering a simplistic system where only one bacteria species is present $N=1$.
+\subsection{Three-pool model}
+We begin our discussion by considering a simplistic system where only one species is present $N=1$.
 Bacterial growth undergoes different consecutive phases~\citep{buchanan_when_1997}, starting with the so called lag-phase, during which no increase in the bacterial count occurs.
 This phenomenon appears due to adaptations of the microbial community, for example, to the new environmental conditions.
-This so-called preparation before the exponential growth includes synthesis of the needed compounds, initiation of the transcription~\citep{rolfe_lag_2012} and poorly studied physiological and regulatory mechanisms~\citep{monod_growth_1949}.
+This so-called preparation before the exponential growth includes synthesis of the needed compounds, initiation of the transcription~\citep{rolfe_lag_2012} and other physiological and regulatory mechanisms~\citep{monod_growth_1949}.
 After this, the bacteria which have succeeded in this adaptation stage enter the exponential growth phase, where the process of cell division is predominant.
@@
 As a result the balance between the growing and dying processes leads to a constant bacterial count~\citep{navarro_llorens_stationary_2010}.
-In the described simplified version of the real-life system three possible bacterial states are mentioned, corresponding to the underlying process: adaptation, growth or death.
-Motivated by this we divide a bacterial population into three pools (states)~(\ref{fig:SchematicRep}), defining the fraction of the population in the lag phase $L(t)$, growing and dividing $G(t)$ and undergoing cell death $D(t)$.
+In the described simplified version of the real-life system three possible bacterial states are mentioned, corresponding to the underlying processes: adaptation, growth or death.
+Motivated by this we divide a bacterial population into three pools (states) (Figure~\ref{fig:SchematicRep}), defining the fraction of the population in the lag phase $L(t)$, growing and dividing $G(t)$ and undergoing cell death $D(t)$.
 \begin{figure}[t]
@@
 \end{figure}
-To determine the system of equations which describe the system's dynamics, we consider the flows between the different states.
+To determine the equations which describe the system's dynamics, we consider the flows between the different states.
 The pools $G$ and $D$ are unidirectionally connected with the pool $L$.
-For now, we assume for that there is no back flow between growth state $G$ and lag-state $L$.
+For now, we assume no back flow from the growth state $G$ to the lag-state $L$.
 The suggested model can be cast in the following reactions:
-\begin{eqnarray}
-    L &\stackrel{\lambda}{\longrightarrow} & G\\
-    L &\stackrel{\mu}{\longrightarrow} & D\\
-    G &\stackrel{\alpha}{\longrightarrow} & 2G\\
-    G &\stackrel{\mu'}{\longrightarrow} & D
- \end{eqnarray}
+\begin{align}
+    L &\stackrel{\lambda}{\longrightarrow} G\\
+    L &\stackrel{\mu}{\longrightarrow} D\\
+    G &\stackrel{\alpha}{\longrightarrow} 2G\\
+    G &\stackrel{\mu'}{\longrightarrow} D
+\end{align}
 Then the \acp{ode} describing such a model can be written as
@@
 According to the works of \citet{monod_growth_1949} in the system where all needed nutrients are in excess except for one, the total growth is linear to the concentration of this limiting nutrient resource.
-The control of the population density with respect to the available nutrition is implemented using cell communication processes called quorum sensing.
+The control of the population density with respect to the available nutrition by means of cell communication is called quorum sensing.
 By releasing extracellular signalling molecules (autoinducer molecules) it is possible for the bacteria to synchronize the gene expression of the whole group~\citep{ng_bacterial_2009}.
-We will not dive into the processes of quorum sensing and try to keep it as simple as possible.
-We aim to include a resource pool which is consumed by the growing bacteria using the second order kinetic reaction:
-\begin{equation}
-    G + R  \stackrel{\alpha_0}{\longrightarrow} 2G.
-\end{equation}
+Instead of modeling complex signalling networks, we choose a simpler approach to describe the dependence of growth on the nutrient availability.
+We include a resource pool which is consumed by the growing bacteria using the second order kinetic reaction:
+\begin{equation}
+    G + R  \stackrel{\alpha_0}{\longrightarrow} 2G
+\end{equation}
+allowing us to keep a minimal number of parameters.
 Note that the pools $L$, $G$ and $D$ represent the number of bacteria (or number of bacteria per unit area/volume) while $R$ represents an abstract resource pool.
-Let say $G=N_B/V$ and $[G]=1/V$, where $N_B$ is the number of bacteria and $V$ is the volume.
-Then $R=N_R/V$ and $[R]=1/V$, where $N_R$ is the number of resource molecules (or any appropriate unit, e.g.,, mol).
+Let's say $G=N_B/V$ and $[G]=1/V$, where $N_B$ is the number of bacteria and $V$ is the volume.
+Then $R=N_R/V$ and $[R]=1/V$, where $N_R$ is the number of resource molecules (or any appropriate unit, e.g., mol).
 Then $[\alpha_0]=V/s$, $[N_t]=1/V$, and $[\alpha]=1/s$.
@@
         \dot{\tilde{R}} &= - \frac{\alpha}{N_t} \tilde{R} G
-    \end{split}.
+    \end{split}
 \label{eq:ode_3pool_resource}
@@
 %
-This can be simplified by exploiting the conservation of the resource pool:
+This can be simplified by exploiting the conservation of the resource pool
 \begin{equation}
@@
         \dot{G} &= \lambda L + \alpha G\left(1-\frac{L+G+D}{N_t}\right)-\mu' G\\
-        \dot{D} &= \mu  L + \mu' G
-    \end{split}.
+        \dot{D} &= \mu  L + \mu' G.
+    \end{split}
 \label{eq:ode_3pool_resource2}
@@
 Note that the previously introduced model could be further simplified to a two pool or state model by omitting the pool $D$ and removing cells with rate $\beta$ directly from pool $L$.
-The relevant question here is whether cells in the pool $D$ still contribute to the limitation of the resources by, e.g.,, occupying space or consuming important nutrients.
-If yes, the three pool model would be advantageous, if not the two pool model is sufficient with the advantages of having one parameter less.\\
+The relevant question here is whether cells in the pool $D$ still contribute to the limitation of the resources by, e.g., occupying space or consuming important nutrients.
+If yes, the three pool model would be advantageous, if not the two pool model is sufficient with the advantage of having one parameter less.\\
 %
@@
 For constant parameters $\mu$ and $\lambda$ the integrals for $L$ and $D$ can be inserted into the differential equation for $G$.
-However, in the case of time dependent parameters, e.g.,, for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$.
+However, in the case of time dependent parameters, e.g., for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$.
 It is interesting to see what type of equations result for the total bacterial population $N=L+G+D$:

### 2025-10-31 paper/pool_model.tex
more formatting

@@
 This survival strategy can appear in response to different external stresses, e.g.,, nutrient deficiency, temperature stress or extreme fluctuations of environment parameters.
-And after "difficult times" bacteria, also called viable but non-culturable (VBNC), can continue cell division~\citep{kell_viability_1998}.
+And after "difficult times" bacteria can continue cell division~\citep{kell_viability_1998}.
 Multiple models which include transitions of bacteria to the dormant state were proposed by \citet{bar_modelling_2002} and \citet{jones_dormancy_2010}.
@@
 Within the framework of our pool model this means, that there is a back-flow from $G$ to $L$.
-\begin{eqnarray}
-    L &\stackrel{\lambda}{\longrightarrow} & G\\
-    L &\stackrel{\mu}{\longrightarrow} & D\\
-    G &\stackrel{\alpha}{\longrightarrow} & 2G\\
-    G &\stackrel{\gamma(t)}{\longrightarrow} & L
-\end{eqnarray}
+\begin{align}
+    L &\stackrel{\lambda}{\longrightarrow} G\\
+    L &\stackrel{\mu}{\longrightarrow} D\\
+    G &\stackrel{\alpha}{\longrightarrow} 2G\\
+    G &\stackrel{\gamma(t)}{\longrightarrow} L
+\end{align}
 The rate $\gamma(t)$ depends on the rate of change of the environment.
@@
 %
-The ordinary differential equations for the pool model with environmental-shock based back-flow read:
+The ordinary differential equations for the pool model with back-flow then read:
 \begin{align}
@@
 The effect of the temperature shift on the total bacterial count can be shown in Figure~\ref{fig:TempJump}.
-After the rapid temperature increase, the growth slows down till the bacteria that transitioned back to lag-phase go again to growth pool.
+After the rapid temperature increase, the growth slows down till the bacteria that transitioned back to lag-phase migrate back to the growth pool.
 If the shift happened rather early in time, when most of the bacteria were in lag-phase, the effect is not noticeable.

### 2025-10-31 paper/pool_model.tex
move files to base
tex: update figures, change subfigures numeration

@@
 % \documentclass[10pt,A4paper]{article}
-\documentclass[10pt,twocolumn,5p,authoryear]{elsarticle/elsarticle}
+\documentclass[10pt,twocolumn,5p,authoryear]{elsarticle}
 \usepackage[utf8]{inputenc}
@@
 
-\bibliographystyle{elsarticle/elsarticle-harv}
+\bibliographystyle{elsarticle-harv}
 
@@
     \caption{
-        (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stress~(\ref{eq:ode_tempshift_backlag}).
-        (b) The value of the corresponding back-flow rate $\gamma$ calculated according to (\ref{eq:gamma_tempshift}).
+        (\textbf{A}) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stress~(\ref{eq:ode_tempshift_backlag}).
+        (\textbf{B}) The value of the corresponding back-flow rate $\gamma$ calculated according to (\ref{eq:gamma_tempshift}).
         The curves represent no temperature shift (green), one temperature jump $\Delta T = 10$ at time $t=4$ (orange)
@@
         The two species system competing for the common nutrient $R$ with production of the toxin $T_B$ (brown line) calculated using Eq.~(\ref{eq:model_2sp_toxin}).
-        The growing bacteria $B$ produces toxin $T_B$ (a) constantly $\dot{T}_B^1 = \mathcal{F}^1 = \kappa G_B$ or (b) only during growth process $\dot{T}_B^2 = \mathcal{F}^2 = \kappa \alpha_B R G_B$ with
+        The growing bacteria $B$ produces toxin $T_B$ (\textbf{A}) constantly $\dot{T}_B^1 = \mathcal{F}^1 = \kappa G_B$ or (\textbf{B}) only during growth process $\dot{T}_B^2 = \mathcal{F}^2 = \kappa \alpha_B R G_B$ with
         production rate $\kappa=0.2$.
@@
         Here the red lines describes the evolution of the species $A$ and blue lines correspond to species $B$.
-        The relation between inhibition strength of species $B$ and $A$ are defined by control parameter value (a) $\psi$ = 1, (b) $\psi$ = 0.5 or (c) $\psi$ = 2.
+        The relation between inhibition strength of species $B$ and $A$ are defined by control parameter value (\textbf{A}) $\psi$ = 1, (\textbf{B}) $\psi$ = 0.5 or (\textbf{C}) $\psi$ = 2.
     }
@@
         The two-species system where each of the species release public goods $P$ is defined with equations~(\ref{eq:1pool_2sp_cooper}) and calculated for different parameters:
-        (a) $\psi=1, \phi=1$, (b) $\psi=1, \phi=2$ and (b) $\psi=2, \phi=1$.
+        (\textbf{A}) $\psi=1, \phi=1$, (\textbf{B}) $\psi=1, \phi=2$ and (\textbf{C}) $\psi=2, \phi=1$.
         Here the red lines describes the evolution of the species $G_A$ and blue lines correspond to species $G_B$.
@@
     \caption{
-        (A) A cell right before and (B) after a division event.
+        (\textbf{A}) A cell right before and (\textbf{B}) after a division event.
         The volume is equally distributed between the two cells and the new positions are picked
@@
 Figure~\ref{fig:spatial-growth-curve} shows multiple results for various configurations of the \ac{abm}, differing only in their initial spatial configuration and diffusivity.
-We see that in the highly-diffusive homogenous case of subfigure (a), the pool model is a good fit to the \ac{abm} model.
-Heterogeneity and slower diffusion cause larger differences between the two models as can be seen in subfigures (b) and (c).
+We see that in the highly-diffusive homogenous case of subfigure (\textbf{A}), the pool model is a good fit to the \ac{abm} model.
+Heterogeneity and slower diffusion cause larger differences between the two models as can be seen in subfigures (\textbf{B}) and \textbf{C}.
 The combination of the two effects shows the largest difference between model outputs.
-In subfigure (c), due to the heterogeneous cell distribution (see Appendix~\ref{fig:supplement-spatial-initial-config-cells}), overall inhibition of species $B$ is significantly reduced.
-Subfigure (d) shows how a randomized homogenous initial distribution can lead to even more notable effects of inhibition.
+In subfigure \textbf{C}, due to the heterogeneous cell distribution (see Appendix~\ref{fig:supplement-spatial-initial-config-cells}), overall inhibition of species $B$ is significantly reduced.
+Subfigure (\textbf{D}) shows how a randomized homogenous initial distribution can lead to even more notable effects of inhibition.
 In contrast to the other results, here the final values of species $A$ and $B$ have larger/smaller counts respectively.

### 2025-10-31 pool_paper_plotting_code/examples/main_paper.py
Update plotting script

@@ -211,10 +211,11 @@ if __name__ == "__main__":
     ax.set_xlim(0., 12.0)
     ax.set_ylim(3.0e0, 0.6e4)
     ax.legend()
-    plt.savefig('paper/Figures/pool_model_3pools_resource.pdf', bbox_inches='tight')
+    plt.savefig('paper/Figures-pool_model_3pools_resource.pdf', bbox_inches='tight')
     plt.close(fig)
 
 #################################### Temperature shift ##########################################
+    coord_text = (0.04, 0.88)
     x01 = [[1e1], [0.], [0.]]
     Gamma, delta= 1., 5.
     shift_cnd = [(4, 10)] #((1, 10), (3, 5), (5, 5), (7, 15))
@@ -242,9 +243,9 @@ if __name__ == "__main__":
     ax[0].set_xlim(0.7, 8.5)
     ax[0].set_ylim(6e0, 1.1e4)
     ax[0].legend()
-    ax[0].text(8., 9.9, r'(a)')
-    ax[1].text(8., 1.5, r'(b)')
-    plt.savefig('paper/Figures/pool_model_3pools_resource_tempshift.pdf', bbox_inches='tight')
+    ax[0].text(8., 9.9, r'\textbf{A}')
+    ax[1].text(8., 1.5,  r'\textbf{B}')
+    plt.savefig('paper/Figures-pool_model_3pools_resource_tempshift.pdf', bbox_inches='tight')
     plt.close(fig)
 
     fig, ax = plt.subplots(1, 1, figsize=figsize_default)
@@ -256,8 +257,8 @@ if __name__ == "__main__":
     ax.set_xlim(0.7, 8.5)
     ax.set_ylim(6e0, 1.1e4)
     ax.legend()
-    ax.text(8., 9.9, r'(a)')
-    plt.savefig('paper/Figures/pool_model_3pools_resource_tempshift1.pdf', bbox_inches='tight')
+    ax.text(0.04, 0.9, r'\textbf{A}', transform = ax.transAxes)
+    plt.savefig('paper/Figures-pool_model_3pools_resource_tempshift1.pdf', bbox_inches='tight')
     plt.close(fig)
 
     fig, ax = plt.subplots(1, 1, figsize=figsize_default_small)
@@ -266,8 +267,8 @@ if __name__ == "__main__":
     ax.plot(t, y2, color=color_palette_1sp_tempshift[i])
[…]

### 2025-10-31 cr_pool/plotting.py
update plotting methods for ABM results

@@ -7,7 +7,7 @@ import gc
 import os
 from pathlib import Path
 
-from .analysis import UNIT_DAY, UNIT_HOUR
+from .analysis import UNIT_DAY
 from .ode_models import ODEModel, observable_2pool_2species
 from .plotting_templates import COLORS_ALL, FIGSIZE_DEFAULT
 from .cr_pool import Domain, BacteriaTemplate, MetaParams
@@ -32,16 +32,26 @@ def plot_growth_curve(data_cells, output_path):
     data_cells.plot(
         x="time", y="bacteria_count_total", ax=ax1, label="Combined", color=CN, alpha=0.5
     )
-    ax1.set_ylabel("Bacteria Count")
+    ticks = mpl.ticker.FuncFormatter(lambda x, pos: "{0:g}".format(x / 1e3))
+    ax1.yaxis.set_major_formatter(ticks)
+    ax1.set_ylabel("Bacteria Count [10³]")
     ax1.set_xlabel("Time [days]")
 
     ax2 = ax1.twinx()
     data_cells.plot(x="time", y="bacteria_volume_1", ax=ax2, label="Species 1", color=CA)
     data_cells.plot(x="time", y="bacteria_volume_2", ax=ax2, label="Species 2", color=CB)
     data_cells.plot(x="time", y="bacteria_volume_total", ax=ax2, label="Combined", color=CN)
-    ax2.set_ylabel("Combined Bacteria Volume")
+    ax2.yaxis.set_major_formatter(ticks)
+    ax2.set_ylabel("Total Bacterial Area [10³µm²]")
 
-    ax1.legend()
+    handles, labels = ax1.get_legend_handles_labels()
+    handles = handles[:3] + [
+        mpl.lines.Line2D([0], [0], color="gray", linestyle="-", lw=2, alpha=0.5),
+        mpl.lines.Line2D([0], [0], color="gray", linestyle="-", lw=2),
+    ]
+    labels = labels[:3] + ["Count", "Volume"]
+    ax1.legend(handles, labels)
+    ax1.set_xlim(np.min(data_cells["time"]), np.max(data_cells["time"]))
     fig.tight_layout()
     fig.savefig(f"{output_path}/cell_growth.png")
     fig.savefig(f"{output_path}/cell_growth.pdf")
@@ -56,6 +66,7 @@ def plot_nutrients(data_voxels, output_path):
     )
[…]

### 2025-11-01 scripts/paper_spatial_plots.py
add units and titles

@@ -2,12 +2,14 @@ import cr_pool as crp
 from pathlib import Path
 import argparse
 
+DIFFUSION_UNITS = "\\text{\\textmu}\\text{m}^2\\text{min}^{-1}"
+
 # Diffusion constants, randomness, homogenous
 CONFIGS = [
-    (30, 0, True),
-    (5, 0, True),
-    (30, 0, False),
-    (30, 0.30, True),
+    (30, 0, True, f"Homogeneous $D=30{DIFFUSION_UNITS}$"),
+    (5, 0, True, f"Homogeneous $D=5{DIFFUSION_UNITS}$"),
+    (30, 0, False, f"Heterogeneous $D=30{DIFFUSION_UNITS}$"),
+    (30, 0.30, True, f"Random $D=30{DIFFUSION_UNITS}$"),
 ]
 
 
@@ -45,17 +47,16 @@ if __name__ == "__main__":
     args = parser.parse_args()
 
     paths = args.paths
+    paths = [(Path(p), None) for p in paths]
 
     if len(paths) == 0:
         paths = []
-        for diffusion_constant, randomness, homogenous in CONFIGS:
+        for diffusion_constant, randomness, homogenous, title in CONFIGS:
             output_path = calcualte_results(diffusion_constant, randomness, homogenous)
-            paths.append(output_path)
-
-    paths = [Path(p) for p in paths]
+            paths.append((Path(output_path), title))
 
     if args.save_snapshots:
-        for output_path in paths:
+        for output_path, _ in paths:
             iters = crp.get_all_iterations(output_path)
             crp.save_snapshot(output_path, iters[6])
             crp.save_snapshot(output_path, iters[12])
@@ -63,9 +64,9 @@ if __name__ == "__main__":
             crp.save_snapshot(output_path, iters[24])
 
     crp.load_style()
-    for output_path in paths:
+    for output_path, title in paths:
         domain, initial_cells, meta_params = crp.get_simulation_settings(output_path)
         data_cells, data_voxels = crp.analyze_all_cell_voxel_data(output_path, meta_params)
         crp.plot_growth_curve(data_cells, output_path)
         crp.plot_nutrients(data_voxels, output_path)
-        crp.plot_comparisons(data_cells, output_path)
+        crp.plot_comparisons(data_cells, output_path, title=title)

### 2025-11-01 paper/pool_model.tex
ensure that subcaptions have capital letters
update abm-ode comparisons

@@
 
+% Ensure that subcaptions have captial letter
+\renewcommand{\thesubfigure}{\Alph{subfigure}}
+
 % Must be loaded after tikz
@@
     \centering
-    \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=\columnwidth]{Figures-abm-homogenous-abm_ode_comparison.png}
-        \subcaption{Hom. $D=30.0$\unit{\micro\metre\squared\per\minute}}
-    \end{subfigure}%
-    \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=\columnwidth]{Figures-abm-homogenous-low-diffusion-abm_ode_comparison.png}
-        \subcaption{Hom. $D=5.0$\unit{\micro\metre\squared\per\minute}}
-    \end{subfigure}\\
-    \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=\columnwidth]{Figures-abm-inhomogenous-abm_ode_comparison.png}%
-        \subcaption{Het. $D=30.0$\unit{\micro\metre\squared\per\minute}}
-    \end{subfigure}%
-    \begin{subfigure}[c]{0.5\columnwidth}
-        \includegraphics[width=\columnwidth]{Figures-abm-random-abm_ode_comparison.png}%
-        \subcaption{Rand. $D=30.0$\unit{\micro\metre\squared\per\minute}}
-    \end{subfigure}
+    \begin{tikzonimage}[width=0.5\columnwidth]
+        {Figures-abm-homogenous-abm_ode_comparison.png}
+        \node at (0, 1)[anchor=north west]{\textbf{A}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.5\columnwidth]
+        {Figures-abm-homogenous-low-diffusion-abm_ode_comparison.png}
+        \node at (0, 1)[anchor=north west]{\textbf{B}};
+    \end{tikzonimage}\\
+    \begin{tikzonimage}[width=0.5\columnwidth]
+        {Figures-abm-inhomogenous-abm_ode_comparison.png}
+        \node at (0, 1)[anchor=north west]{\textbf{C}};
+    \end{tikzonimage}%
+    \begin{tikzonimage}[width=0.5\columnwidth]
+        {Figures-abm-random-abm_ode_comparison.png}
+        \node at (0, 1)[anchor=north west]{\textbf{D}};
+    \end{tikzonimage}%
     \caption{
-        Comparison of the \ac{ode} and \ac{abm} model for different configurations, showing
-        homogenous (Hom.), heterogeneous (Het.) or randomized (Rand.) initial configurations.
+        (A-D) Comparison of the \ac{ode} and \ac{abm} model for different configurations, showing homogenous, heterogeneous or randomized initial configurations.
         The diffusion constants are chosen such that $D_R=D_I=D$.

### 2025-11-01 cr_pool/plotting.py
use proper units within plots

@@ -23,6 +23,9 @@ CB = COLORS_ALL["N_B"]
 CN = COLORS_ALL["N"]
 
 
+MICRON_TEX_UNITS = "\\text{\\textmu m}"
+
+
 def plot_growth_curve(data_cells, output_path):
     # Growth Curve
     fig, ax1 = plt.subplots(figsize=FIGSIZE_DEFAULT)
@@ -42,7 +45,7 @@ def plot_growth_curve(data_cells, output_path):
     data_cells.plot(x="time", y="bacteria_volume_2", ax=ax2, label="Species 2", color=CB)
     data_cells.plot(x="time", y="bacteria_volume_total", ax=ax2, label="Combined", color=CN)
     ax2.yaxis.set_major_formatter(ticks)
-    ax2.set_ylabel("Total Bacterial Area [10³µm²]")
+    ax2.set_ylabel("Total Bacterial Area [$10^3{MICRON_TEX_UNITS}^2$]")
 
     handles, labels = ax1.get_legend_handles_labels()
     handles = handles[:3] + [
@@ -157,7 +160,7 @@ def calculate_difference(data_cells, output_path):
     return variance
 
 
-def plot_comparisons(data_cells, output_path):
+def plot_comparisons(data_cells, output_path, title=None):
     settings = get_simulation_settings(output_path)
 
     model = abm_to_ode(*settings)
@@ -174,10 +177,13 @@ def plot_comparisons(data_cells, output_path):
     data_cells.plot(x="time", y="bacteria_volume_2", ax=ax, color=CB, linestyle="--")
     data_cells.plot(x="time", y="bacteria_volume_total", ax=ax, color=CN, linestyle="--")
 
+    if title is not None:
+        ax.set_title(title)
+
     ticks = mpl.ticker.FuncFormatter(lambda x, pos: "{0:g}".format(x / 1e3))
     ax.yaxis.set_major_formatter(ticks)
     ax.set_xlabel("Time [days]")
-    ax.set_ylabel("Total Bacterial Area [10³µm²]")
+    ax.set_ylabel(f"Total Bacterial Area [$10^3{MICRON_TEX_UNITS}^2$]")
     # ax.set_yscale('log')
     handles, labels = ax.get_legend_handles_labels()
     handles = handles[:3] + [
@@ -237,7 +243,7 @@ def plot_lag_phase(data_cells, res, output_path):
 
     ax.legend()
[…]

### 2025-11-01 scripts/paper_spatial_plots.py
simplify units

@@ -2,7 +2,7 @@ import cr_pool as crp
 from pathlib import Path
 import argparse
 
-DIFFUSION_UNITS = "\\text{\\textmu}\\text{m}^2\\text{min}^{-1}"
+DIFFUSION_UNITS = "\\text{\\textmu m}^2\\text{min}^{-1}"
 
 # Diffusion constants, randomness, homogenous
 CONFIGS = [

### 2025-11-03 paper/pool_model.tex
Use consistent equation reference style, remove extra commas

@@
 \end{equation}
-Inserting Eq.~\ref{eq:Resource} into Eqs.~\ref{eq:model_2sp_resource_comp} yields:
+Inserting equation~\ref{eq:Resource} into system~\ref{eq:model_2sp_resource_comp} yields:
 \begin{align}
@@
     \caption{
-        The competition of the two bacterial species due to the common limiting resource calculated using Eq.~(\ref{eq:model_2sp_resource_comp}).
+        The competition of the two bacterial species due to the common limiting resource calculated using equation~(\ref{eq:model_2sp_resource_comp}).
         The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
@@
     \caption{
-        The two species system competing for the common nutrient $R$ with production of the toxin $T_B$ (brown line) calculated using Eq.~(\ref{eq:model_2sp_toxin}).
+        The two species system competing for the common nutrient $R$ with production of the toxin $T_B$ (brown line) calculated using equation~(\ref{eq:model_2sp_toxin}).
         The growing bacteria $B$ produces toxin $T_B$ (\textbf{A}) constantly $\dot{T}_B^1 = \mathcal{F}^1 = \kappa G_B$ or (\textbf{B}) only during growth process $\dot{T}_B^2 = \mathcal{F}^2 = \kappa \alpha_B R G_B$ with
@@
 Another special case of the species competition is when the produced toxin does not cause the death of the bacteria but rather inhibit its growth.
-Such situation with 2 bacteria species, a shared resource and produced inhibitor was modeled, e.g.,, by \citet{lenski_coexistence_1986}.
+Such situation with 2 bacteria species, a shared resource and produced inhibitor was modeled, e.g., by \citet{lenski_coexistence_1986}.
 In our case, we would like to look at a rather simplified case reducing the number of variables of the system.
@@
 The function $\mathcal{B}_{A/B}$ captures the inhibiting effects effectively.
-It is straight forward to make the mechanism of inhibition more explicit by, e.g.,, introducing a toxin pool.
+It is straight forward to make the mechanism of inhibition more explicit by, e.g., introducing a toxin pool.
 One possible choice for $\mathcal{B}$ is:
@@
 %     The ABM represent a microbial colony as a group of discrete agents that are governed by a set of rules, where each agent can be an individual cell or a collection of cells.
-%     Can use citation when talk about general processes we are accounting as growth, uptake of nutrient, production of inhibitor, e.g.,,:
+%     Can use citation when talk about general processes we are accounting as growth, uptake of nutrient, production of inhibitor, e.g.,:
 %     Physical interactions between agents are calculated using excluded volume interactions.
@@
 with $x_{j}^{i}$ describing the bacterial counts for species $i$ in state $j$.
-$\mbu$ is the vector of different inputs of the model, e.g.,, temperature, pressure etc.
+$\mbu$ is the vector of different inputs of the model, e.g., temperature, pressure etc.
 The vector $\mbm \in \mathbb{R}^{K}$ includes all variables corresponding to the microenvironment of the system which can be concentration of the inhibitors or activators:
@@
 
-The simpler version of one state per species (Eq.~\ref{eq:LV_simple}) can be simply written in a general mathematical form as generalized Lotka-Volterra:
+The simpler version of one state per species (equation~\ref{eq:LV_simple}) can be simply written in a general mathematical form as generalized Lotka-Volterra:
 \begin{equation}

### 2025-11-04 paper/pool_model.tex
fix line numbers
fix notation of total bacteria

@@
 \usepackage{listings-rust}
-\usepackage{lineno}
+\usepackage[switch]{lineno}
 \usepackage{mathtools}
@@
 \usepackage{xpatch}
-
-\linenumbers
 
@@
 \maketitle
+\linenumbers
 %
@@
 \subsection{Lag-time estimation}
-Note that the previously introduced model could be further simplified to a two pool or state model by omitting the pool $D$ and removing cells with rate $\beta$ directly from pool $L$.
+Note that the previously introduced model could be further simplified to a two pool or state model by omitting the pool $D$ and removing cells directly from pools $L$ and $G$.
 The relevant question here is whether cells in the pool $D$ still contribute to the limitation of the resources by, e.g., occupying space or consuming important nutrients.
-If yes, the three pool model would be advantageous, if not the two pool model is sufficient with the advantage of having one parameter less.\\
+If yes, the three pool model would be advantageous, if not the two pool model is sufficient with the benefit of having one parameter less.\\
 %
@@
 However, in the case of time dependent parameters, e.g., for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$.
-It is interesting to see what type of equations result for the total bacterial population $N=L+G+D$:
+It is interesting to see what type of equations result for the total bacterial population $\mathcal{N}=L+G+D$:
 \begin{align}
     \begin{split}
-        \dot{G} &= \lambda L + \alpha G\left(1-\frac{N}{N_t}\right)\\
-        \dot{N} &= \alpha G\left(1-\frac{N}{N_t}\right)
+        \dot{G} &= \lambda L + \alpha G\left(1-\frac{\mathcal{N}}{N_t}\right)\\
+        \dot{\mathcal{N}} &= \alpha G\left(1-\frac{\mathcal{N}}{N_t}\right)
     \end{split}
 \end{align}
-with initial conditions $G(0)=0$ and $N(0)=N_0$.
+with initial conditions $G(0)=0$ and $\mathcal{N}(0)=N_0$.
 These equations have a clear interpretation, are derived from first principles, and are not identical to the popular Barayni-Roberts and modified-Gompertz models.

### 2025-11-04 paper/pool_model.tex
include results from shared session
- formulations
- equation nunmbering
- consitency

Co-authored-by: Polina Gaindrik <polinagaindrik@users.noreply.github.com>

@@
 As mentioned above, bacteria usually coexist in large interacting communities of multiple species whose growth is vastly determined by the developed survival mechanisms.
-These mechanisms mostly aim to secure as much available nutrition which lead to competitive strategies or, if it is advantageous, cooperative ones~\citep{hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
+These mechanisms mostly aim to secure a constant availability of nutrition which leads to competitive strategies or, if it is advantageous, to cooperative ones~\citep{hibbing_bacterial_2010, stubbendieck_bacterial_2016}.
 Further we show that the model can be readily extended to the systems with more than one bacterium and describe more complex relationships between species.
@@
 To this end we consider for simplicity only one common limiting resource pool.
-The simplest pool model for bacterial species $A$ and $B$ encompassing lag-phases, growth-phases and a common resource pool is given by (no back-flow to the lag-pool):
+The simplest pool model for bacterial species $A$ and $B$ encompassing lag-phases, growth-phases and a common resource pool which is given by (no back-flow to the lag-pool):
 \begin{align}
@@
 \end{align}
-We again rescaled the resource pool with $N_t$ and find
+We again rescaled the resource pool with its total capacity $N_t$ and find
 \begin{equation}
@@
 \begin{align}
-    \dot{L}_A &= -\lambda_A  L_A + \frac{\lambda_A}{N_t}L_A\left<LG\right>\\
-    \dot{L}_B &= -\lambda_B L_B + \frac{\lambda_B}{N_t}L_B \left<LG\right>\\
-    \dot{G}_A &=  \lambda_A  L_A + \alpha_A G_A - \frac{\lambda_A L_A + \alpha_A G_A}{N_t} \left<LG\right>\\
-    \dot{G}_B &=  \lambda_B L_B - \alpha_B G_B \frac{\lambda_B L_B + \alpha_B G_B}{N_t}\left<LG\right>
-\end{align}
-where we used the short notation $\left<LG\right>=(L_A+G_A+L_B+G_B)$.
+    \dot{L}_A &= -\lambda_A  L_A + \frac{\lambda_A}{N_t}L_AN\\
+    \dot{L}_B &= -\lambda_B L_B + \frac{\lambda_B}{N_t}L_B N\\
+    \dot{G}_A &=  \lambda_A  L_A + \alpha_A G_A - \frac{\lambda_A L_A + \alpha_A G_A}{N_t} N\\
+    \dot{G}_B &=  \lambda_B L_B - \alpha_B G_B \frac{\lambda_B L_B + \alpha_B G_B}{N_t}N
+\end{align}
+where we used the total bacterial count $N=(L_A+G_A+L_B+G_B)$.
 These equations have the structure of a \ac{glv} model expanded with different internal states for one species.
-As one can see in Figure~\ref{fig:2pool_resource_2sp} even without direct interaction between species the bacterial growth was inhibited due to the limitation of the resource.
-Here species $A$ has a longer lag-phase than $B$ but the exponential growth of $A$ is stronger compared to which is experienced by bacteria $B$.
-As a result, species $A$ was able to suppress species $B$, consuming the larger amount of nutrients and reaching a larger bacterial count at the end.
+As one can see in Figure~\ref{fig:2pool_resource_2sp} even without any direct interaction between species the bacterial growth was inhibited due to the limitation of the resource.
+Here species $A$ has a longer lag-phase than $B$ but the exponential growth of $A$ is faster compared to $B$.
+As a result, species $A$ was able to suppress $B$, consuming the larger amount of nutrients and reaching a larger bacterial count at the end.
 \begin{figure}
@@
 %
-In order to get the commonly used form of the \ac{glv}, we need to simply eliminate of the bacterial states by setting $L_A=L_B=0$:
+In order to get the commonly used form of the \ac{glv}, we need to simply omit the bacterial states $L_A=L_B=0$:
 \begin{align}
@@
 %
-Equations (\ref{eq:LV_simple}) can be understood as competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competitive excursion.
+Equations (\ref{eq:LV_simple}) can be understood as competitive Lotka-Volterra equations that are widely used in ecological studies and show the principle of competitive exclusion.
 That principle claims that two species which compete for the same resource cannot coexist~\citep{wangersky_lotka-volterra_1978}.
-In previous studies it was shown that \ac{glv} equations with one bacterial state can be successfully applied to predict the microbial community dynamics, determine different types of the pairwise interspecies interactions~\citep{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
-In food studies the \ac{glv} was applied by \citet{mounier_microbial_2008} to study interactions of six bacteria species influenced by yeast omissions in cheese.
-\citet{cauchie_modeling_2020} applied Lotka-Volterra model to study the three-species system of the bacteria majorly present in minced pork meat.\\
+In previous studies it was shown that \ac{glv} equations with one bacterial state can be successfully applied to predict the microbial community dynamics and determine different types of the pairwise interspecies interactions~\citep{dedrick_when_2023, stein_ecological_2013, venturelli_deciphering_2018, hoffmann_power_2007}.
+In food studies the \ac{glv} was applied by \citet{mounier_microbial_2008} to study interactions of six bacterial species influenced by yeast omissions in cheese.
+\citet{cauchie_modeling_2020} applied the Lotka-Volterra model in their study of the three-species bacterial system of minced pork meat.\\
 %
@@
 \subsection{Interspecies competition}
-Mutual inhibition between different species can not only result from competition for a shared resource for which bacteria have developed various other active competition mechanisms.
-One of the examples of such mechanisms is the production of antimicrobial compounds, inhibitors or toxins~\citep{wloch-salamon_effect_2008, chao_structured_1981}.
+Mutual inhibition is often a result of active competition mechanisms which have been developed by bacteria to contend for a shared resource.
+Some examples of such mechanisms are the production of antimicrobial compounds, inhibitors or toxins~\citep{wloch-salamon_effect_2008, chao_structured_1981}.
 On the other hand, some bacteria can produce public goods, the compounds that benefit growth of other species.
@@
 The function $\mathcal{F}$ denotes the model for the toxin production rate.
-It can be, for example, the accumulation of the bacteria waste which is responsible for the spoilage of food products and affecting another species in negative way.
-Consider the system for two species competing for resource $R$ with the production of the toxin $T$ by species $B$.
-Assume that the toxin causes the death of the species $A$ with rate $\mu$.
+One example is the accumulation of bacterial waste which is responsible for the spoilage of food products and suppressing growth of other species.
+Consider the system of two species competing for resource $R$ with the production of the toxin $T$ by species $B$.
+We assume that the toxin causes the death of the species $A$ with rate $\mu$.
 Then the full \ac{ode} system is
@@
         \dot{T}_{B}^{1,2} &= \mathcal{F}^{1,2} (R, G_B) \\
-    \end{split}.
+    \end{split}
     \label{eq:model_2sp_toxin}
@@
 \end{align}
-Here $\kappa$ is the toxin production rate.
+with $\kappa$ as the toxin production rate.
 The difference in the solutions of the given \acp{ode} between the two toxin production models is presented in Figure~\ref{fig:2pool_2sp_toxin}.
@@
 Model $\mathcal{F}^1$ results in toxin concentration $T_B^1(t)=\kappa\int_{t_0}^tG(t')dt'$.
-That means that as long as there are bacteria in the growth phase $G_B>0$, toxin is produced, no matter whether the total bacterial count $N_B$ is still growing.
-On the other hand, model $\mathcal{F}^2$ results in toxin accumulation in steady state $T_B^{2s}=\lim_{t\to\infty}T_B^2(t)=\kappa(N_t-N_0)$.
+The production of toxin only depends on the presence of bacteria $G_B$ and is independent of their growth dynamics.
+% That means that as long as there are bacteria in the growth phase $G_B>0$, toxin is produced, no matter whether the bacterial count $N_B$ is still growing.
+On the other hand, model $\mathcal{F}^2$ assumes that toxin is only produced during bacterial growth and division, hence depends on the growth dynamics of $B$.
+This leads to a toxin accumulation in the steady state $T_B^{2s}=\lim_{t\to\infty}T_B^2(t)=\kappa(N_t-N_0)$.
 To see this we note $\dot{T}_B^2=\kappa(\dot{L_B}+\dot{G_B})$, which yields: $T_B^2(t)=\kappa(L_B(t)+G_B(t)-N_{B0})$ where we used $T_B^2(0)=G_B(0)=0$, $L_B(0)=N_{B0}$.
-This model assumes that toxin is only produced when the total bacterial grow and divide, hence, the total bacterial count $N_B$ increases.
-%
 %
@@
 \subsubsection{Mutual inhibition}
-Another special case of the species competition is when the produced toxin does not cause the death of the bacteria but rather inhibit its growth.
-Such situation with 2 bacteria species, a shared resource and produced inhibitor was modeled, e.g., by \citet{lenski_coexistence_1986}.
-In our case, we would like to look at a rather simplified case reducing the number of variables of the system.
-The easiest way to introduce inhibition of species $A$ by species $B$ is via a growth rate of species $A$ depending on the abundance of species $B$.
-For sake of simplicity we ignore the lag pool and write for a mutual inhibition:
+Another special case of inter-species competition is when the produced toxin does not cause the death of the bacteria but rather inhibit their growth.
+Such a situation with 2 bacteria species, a shared resource and produced inhibitor was modeled by \citet{lenski_coexistence_1986}.
+The same effect can be modeled by utilizing our pool-based approach.
+In our case, we would like to look at a rather simplified case reducing the number of variables in the system.
+The easiest way to introduce inhibition of species $A$ by species $B$ is by assuming that the growth rate of species $A$ depends on the abundance of species $B$.
+For the sake of simplicity we ignore the lag pool and describe the growth components via:
 \begin{align}
     \dot{G}_A &= \mathcal{B}_A(G_B)G_A\left(1 - \frac{G_A+G_B}{N_t}\right)\\
-    \dot{G}_B &= \mathcal{B}_B(G_A) G_B\left(1-\frac{G_A+G_B}{N_t}\right).
-\end{align}
-The function $\mathcal{B}_{A/B}$ captures the inhibiting effects effectively.
-It is straight forward to make the mechanism of inhibition more explicit by, e.g., introducing a toxin pool.
+    \dot{G}_B &= \mathcal{B}_B(G_A) G_B\left(1-\frac{G_A+G_B}{N_t}\right)
+\end{align}
+The functions $\mathcal{B}_{A/B}$ are a phenomenological choice to capture the inhibiting effects.
+It is straightforward to make the mechanism of inhibition more explicit by, e.g., introducing a toxin pool.
 One possible choice for $\mathcal{B}$ is:
 \begin{equation}
-    \mathcal{B}(G) = \frac{\alpha}{1+KI(G)},
-\end{equation}
-where $I$ is the concentration of the inhibiting metabolic byproduct producing during bacterial growth.
+    \mathcal{B}(G) = \frac{\alpha}{1+KI(G)}
+\end{equation}
+where $I$ is the concentration of the inhibiting metabolic byproduct of the bacterial growth.
+This leads to
 \begin{align}\begin{split}
@@
 \end{split}\end{align}
-Note that if the toxin accumulates, i.e., $\dot I \sim G$,  one needs to replace $G$ by $\int G dt$.
+% Note that if the toxin accumulates, i.e., $\dot I \sim G$,  one needs to replace $G$ by $\int G dt$.
 One can distinguish two limiting cases: {\it i}) $K_{A/B}\ll N_t$ and {\it ii}) $K_{A/B}\gg N_t$.
@@
 which are displayed in Figure~\ref{fig:1pool_2sp_inhibit}.
+The advantage of this system is its simplicity, allowing us to describe the effect of mutual inhibition between two bacterial species with only a single parameter.
 %
@@
 On the other hand, one can also observe some cooperative behaviors between species in cases when it facilitates the competition with others.
-This phenomenon occurs when bacterial metabolism produces compounds that are beneficial for the growth of each other, so-called public goods.
+This phenomenon occurs due to complementary metabolic pathways, i.e., production of so-called public goods which are beneficial for the growth of other species.
 It is most often observed between species which are close relatives or clonal populations.
@@
 \begin{align}
+    \begin{split}
         \dot{G}_A &= P_B G_A\left(1 - G_A-G_B\right)\\
@@
         \dot{P}_B &=\phi G_B\left(1 - G_A-G_B\right).
+    \end{split}
     \label{eq:1pool_2sp_cooper}
@@
 The solution of the system~\ref{eq:1pool_2sp_cooper} is presented in Figure~\ref{fig:1pool_2sp_cooper}.
-Here we can see that if $\psi=1$ and $\phi=1$ (see Figure~\ref{fig:1pool_2sp_cooper}a) the production of the public goods of the two species is happening with the same rate as the growth of bacteria.
-Increasing the production rate of the public good (see Figure~\ref{fig:1pool_2sp_cooper}b), leads to an increase in the growth of species $A$ that profit from the presence of $B$.
-On the other hand, choosing $\psi=2$ (see Figure~\ref{fig:1pool_2sp_cooper}c), we increase the profit that species $B$ has from public goods $P_A$.
+Here we can see that if $\psi=1$ and $\phi=1$ (see Figure~\ref{fig:1pool_2sp_cooper}~A) the production rate of the public goods of both species is identical to the growth rate of bacteria.
+Increasing the production rate of the public good (see Figure~\ref{fig:1pool_2sp_cooper}~B), leads to an increase in the growth of species $A$ that profit from the presence of $B$.
+On the other hand, choosing $\psi=2$ (see Figure~\ref{fig:1pool_2sp_cooper}~C), we increase the profit that species $B$ has from public goods $P_A$.
 That indirectly leads to larger amount of $P_B$ produced by species $B$.

### 2025-11-04 paper/pool_model.tex
make subfigure letters not bold

@@
     \caption{
-        (\textbf{A}) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stress~(\ref{eq:ode_tempshift_backlag}).
-        (\textbf{B}) The value of the corresponding back-flow rate $\gamma$ calculated according to (\ref{eq:gamma_tempshift}).
+        (A) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stress~(\ref{eq:ode_tempshift_backlag}).
+        (B) The value of the corresponding back-flow rate $\gamma$ calculated according to (\ref{eq:gamma_tempshift}).
         The curves represent no temperature shift (green), one temperature jump $\Delta T = 10$ at time $t=4$ (orange)
@@
         The two species system competing for the common nutrient $R$ with production of the toxin $T_B$ (brown line) calculated using equation~(\ref{eq:model_2sp_toxin}).
-        The growing bacteria $B$ produces toxin $T_B$ (\textbf{A}) constantly $\dot{T}_B^1 = \mathcal{F}^1 = \kappa G_B$ or (\textbf{B}) only during growth process $\dot{T}_B^2 = \mathcal{F}^2 = \kappa \alpha_B R G_B$ with
+        The growing bacteria $B$ produces toxin $T_B$ (A) constantly $\dot{T}_B^1 = \mathcal{F}^1 = \kappa G_B$ or (B) only during growth process $\dot{T}_B^2 = \mathcal{F}^2 = \kappa \alpha_B R G_B$ with
         production rate $\kappa=0.2$.
@@
         Here the red lines describes the evolution of the species $A$ and blue lines correspond to species $B$.
-        The relation between inhibition strength of species $B$ and $A$ are defined by control parameter value (\textbf{A}) $\psi$ = 1, (\textbf{B}) $\psi$ = 0.5 or (\textbf{C}) $\psi$ = 2.
+        The relation between inhibition strength of species $B$ and $A$ are defined by control parameter value (A) $\psi$ = 1, (B) $\psi$ = 0.5 or (C) $\psi$ = 2.
     }
@@
         The two-species system where each of the species release public goods $P$ is defined with equations~(\ref{eq:1pool_2sp_cooper}) and calculated for different parameters:
-        (\textbf{A}) $\psi=1, \phi=1$, (\textbf{B}) $\psi=1, \phi=2$ and (\textbf{C}) $\psi=2, \phi=1$.
+        (A) $\psi=1, \phi=1$, (B) $\psi=1, \phi=2$ and (C) $\psi=2, \phi=1$.
         Here the red lines describes the evolution of the species $G_A$ and blue lines correspond to species $G_B$.
@@
     \caption{
-        (\textbf{A}) A cell right before and (\textbf{B}) after a division event.
+        (A) A cell right before and (B) after a division event.
         The volume is equally distributed between the two cells and the new positions are picked

### 2025-11-04 paper/pool_model.tex
2nd part of shared session
Co-authored-by: Polina Gaindrik <polinagaindrik@users.noreply.github.com>

@@
 nutrients~\citep{htuson_bacteriasurface_2013}.
-Another mechanism is the motility of bacteria, e.g.,, the flagellates \textit{E. coli} or
+Another mechanism is the motility of bacteria, e.g., the flagellates of \textit{E. coli} or
 \textit{B. subtilis} which are capable of chemotactic motion driven by oxygen
@@
 significantly affect the spatial structure of the community~\citep{blanchard_bacterial_2015}.
-These effects can only be treated with a spatial model that is able to describe the observed
+These effects can only be treated within a spatial framework that is able to describe the observed
 effects.\\
@@
 single-cell behaviour of the bacteria.
-To this end, we consider the system consisting from two bacterial species that can be described by two internal states: lag and growth states.
+To this end, we consider the system consisting of two bacterial species that can be described by the two internal states: lag phase and growth.
 Both of these species grow using the same resource pool $R$ but with different growth constants
-$\alpha$ and different lengths of lag phase $\lambda$.
-Moreover, the species $A$ is able to release a certain antimicrobial component $I$ that shows an
-inhibitory effect on species 2 (amensalism).
+$\alpha_{A/B}$ and different lag phase transition rates $\lambda_{A/B}$.
+Moreover, species $A$ is able to release a certain antimicrobial component $I$ that shows an
+inhibitory effect on species $B$ (amensalism).
 We assume that $L,G$ are measured in terms of the overall volume and $R,I$ are measured as total
 amount.
-The \acp{ode} the bacterial population can be written as:
+The \acp{ode} system of the bacterial population can be written as:
 %
@@
 \begin{align}
+    \begin{split}
         \dot{L_A} &= -\lambda_A R L_A\\
@@
         \dot{L_B} &= -\lambda_B R L_B\\
-    \dot{G_B} &= \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B
+        \dot{G_B} &= \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B.
+    \end{split}
     \label{eq:spatial_limit_F}
@@
 %
-The microenvironment of our system contains the resource and the inhibitor molecules:
-\begin{align}
+Moreover, the microenvironment of our system contains the resource and the inhibitor molecules
+\begin{align}
+    \begin{split}
         \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \mu_I I) N_t} R G_B \\
-    \dot{I} &= \kappa G_A
+        \dot{I} &= \kappa G_A.
+    \end{split}
     \label{eq:spatial_limit_H}
@@
 \ac{ode}-based systems.\\
-To construct our \ac{abm}, we use \texttt{cellular\_raza}~\citep{Pleyer2025}.
-It allows us to define our bacterial agents from a clean slate, without having to rely on and inherit
-existing work.
+To construct our \ac{abm}, we use the modeling framework \texttt{cellular\_raza}~\citep{Pleyer2025}.
+ allows us to define our bacterial agents from a clean slate.
 With this flexibility, researchers are able to construct minimal models using only as many parameters which are necessary.
@@
 To construct an \ac{abm} which resembles the given \ac{ode} system of equations~\eqref{eq:spatial_limit_F} and \eqref{eq:spatial_limit_H}, we consider multiple cellular reactions.
-The first is uptake of an extracellular resource $[R_e]$ into the cell $[R_i]$ with rate $u$ in \unit{\per\minute}.
-Nutrients are only taken up at the position of the cell $x_c$
-\begin{align}
-    \dot{[R_i]} &= u [R_e](x_c)\\
-    \dot{[R_e]}(x) &= -u \frac{V_c}{V_D} [R_e] \delta(x-x_c)
-\end{align}
-where $\delta(x-x_\text{cell})$ is the Dirac delta-distribution.
+We denote cells of species $A,B$ with $c_{A/B}$ respectively and indicate shared behavior by simply using $c$.
+Cells $c_A$ of species $A$ take up extracellular resources $[R_e]$ into the cell $[R_{i,c_A}]$ with rate $u$ in \unit{\per\minute}.
+The uptake only occurs at the position of the cell $x_{c_A}$.
+Furthermore, we let cells of species $A$ produce a freely diffusable inhibitor $I$ with production rate $b$ and diffusion constant $D_I$ which inhibits cells of species $B$ in their uptake of nutrients.
+\begin{equation}
+    \dot{[I]} = b\frac{V_{c_A}}{V_D}\delta(x-x_{c_A}) + D_I \Delta [I]
+\end{equation}
+We model this inhibiting effect by modifying the uptake term of species $B$ with a factor of $\frac{1}{1+\nu_I [I]}$.
+% Note that we do not need to distinguish between the two species because we can set parameters individually on agent-level.
+% To distinguish between the two species, it suffices to set parameters for every individual agent accordingly.
+% This means, we can implement the most general version of the equations and by choosing appropriate parameters, this defines the species of the respective cell-agent.
+If we combine these effects and consider multiple cells, we obtain the system
+\begin{alignat}{3}
+    \dot{[R_{i,c_A}]} &=& u &[R_e] - a[R_{i,c_A}]\\
+    \label{eq:spatial-intracellular-a}
+    \dot{[R_{i,c_B}]} &=& \frac{u}{1+\nu_I [I]} &[R_e] - a[R_{i,c_B}]\\
+    \label{eq:spatial-intracellular-b}
+    % TODO fix this
+    % \begin{split}
+        \dot{[R_e]} &=& u &\sum\limits_{c_A}\frac{V_{c_A}}{V_D} [R_e] \delta(x-x_{c_A})\\
+            &&-\frac{u}{1+\nu_I [I]} &\sum\limits_{c_B}\frac{V_{c_B}}{V_D} [R_e] \delta(x-x_{c_B})
+    % \end{split}
+    \label{eq:spatial-extracellular-resources-pde}
+\end{alignat}
+where equations~\eqref{eq:spatial-intracellular-a} and \eqref{eq:spatial-intracellular-b} describe the change of intracellular resources for every individual cell of of species $A,B$ and the equation~\eqref{eq:spatial-extracellular-resources-pde} contains extracellular resources coupled to each individual agent.
+They contain the uptake of resources and term used grow the cell, converting resources into volume which will be explained below.
+We use the Dirac delta-distribution $\delta(x-x_c)$ to enforce exchange of nutrients at the positions of the cells $x_{c}$.
 Note that $[R_e]=[R_e](x)$ explicitly depends on the spatial coordinate but we choose to omit its notation.
-In the second equation, we need to account for the ratio of cellular $V_c$ and domain volume $V_D$ to calculate the correct change.
-Further, we want to modify the cellular volume by growing it increasingly with rate $a$.
+In equation~\eqref{eq:spatial-extracellular-resources-pde} we need to account for the ratio of cellular $V_c$ and domain volume $V_D$ to calculate the correct exchange.
+In addition, we want to modify the cellular volume by growing it increasingly with rate $a$.
 Therefore, we need to specify how many resource molecules are turned into volumetric units $\sigma$ in \unit{\micro\metre\cubed\per\mole}.
-The total amount of resource molecules present inside the cell is given by the current concentration multiplied by the cell-volume $V_c [R_i]$.
+The total amount of resource molecules present inside the cell is given by the current concentration multiplied by the cell-volume $V_c [R_{i,c}]$.
 With a rate $a$ and conversion $\sigma$, the cell turns these resource molecules into volume.
-\begin{align}
-    \dot{V_c} &= a\sigma [R_i] V_c\\
-    \dot{[R_i]} &= - a[R_i]
-\end{align}
-Furthermore, we let cells produce an inhibitor with production rate $b$ which inhibits the cell's ability to uptake and thus use the resource by a factor of $1+\nu_I [I]$
-\begin{align}
-    \dot{[I]} &= b\frac{V_c}{V_D}\delta(x-x_c)\\
-    u &\longrightarrow \frac{u}{1+\nu_I [I]}.
-\end{align}
-% Note that we do not need to distinguish between the two species because we can set parameters individually on agent-level.
-To distinguish between the two species, it suffices to set parameters for every individual agent accordingly.
-This means, we can implement the most general version of the equations and by choosing appropriate parameters, this defines the species of the respective cell-agent.
-If we combine all equations and consider multiple cells, we obtain the system
-\begin{align}
-    \dot{V_c} &= a\sigma [R_i] V_c\\
-    \dot{[R_i]} &= \frac{u}{1+\nu_I [I]} [R_e] - a[R_i]\\
-    \dot{[R_e]} &= -\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R_e] \delta(x-x_c)\\
-    \dot{[I]} &= b\frac{V_c}{V_D}\delta(x-x_c).
-    \label{eq:spatial-full-system-non-simplified}
-\end{align}
+\begin{equation}
+    \dot{V_c} = a\sigma [R_{i,c}] V_c
+\end{equation}
 In order to simplify this system, we can assume that the internal resources are constantly being depleted by the growth and the limiting process is the uptake of the resource.
-This means, we can effectively set $\dot{[R_i]}=0$.
-Therefore, we obtain the relation $\frac{u}{1+\nu_I [I]}[R_e] = a[R_i]$ which we can use to eliminate the variable monitoring internal resource concentrations $[R_i]$.
-Additionally, we require our external components to be freely diffusible with constants $D_R,D_I$.
-Plugging this into equation~\eqref{eq:spatial-full-system-non-simplified} yields.
-\begin{align}
-    \dot{V_c} &= \frac{u}{1+\nu_I [I]} \sigma [R]V_c\\
-    \dot{[R]} &= \sum\limits_c -\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c) + D_R\Delta [R]\\
-    \dot{[I]} &= \sum\limits_c b\frac{V_c}{V_D}\delta(x-x_c) + D_I\Delta [I].
+This means, we can effectively set $\dot{[R_{i,c}]}=0$.
+Therefore, we obtain the relations $\frac{u}{1+\nu_I [I]}[R_e] = a[R_{i,cB}]$ and $u[R_e]=a[R_{i,c_A}]$ which we can use to eliminate the variable monitoring internal resource concentrations $[R_{i,c}]$.
+Additionally, we also require the nutrients to be freely diffusible with constants $D_R$.
+With these considerations, we can update equations~\eqref{eq:spatial-intracellular-a} to \eqref{eq:spatial-extracellular-resources-pde} and obtain:
+\begin{align}
+    \begin{split}
+        % TODO format these equations
+        \dot{V_{c_B}} &= u \sigma [R]V_c\\
+        \dot{V_{c_A}} &= \frac{u}{1+\nu_I [I]} \sigma [R]V_c\\
+        % \dot{[R_e]} &= \sum\limits_c -\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c) + D_R\Delta [R]\\
+        \dot{[R]} &= u \sum\limits_{c_A}\frac{V_{c_A}}{V_D} [R_e] \delta(x-x_{c_A})\\
+            &-\frac{u}{1+\nu_I [I]} \sum\limits_{c_B}\frac{V_{c_B}}{V_D} [R_e] \delta(x-x_{c_B})\\
+            &+D_R \Delta [R_e]\\
+        \dot{[I]} &= \sum\limits_c b\frac{V_c}{V_D}\delta(x-x_c) + D_I\Delta [I]
+    \end{split}
     \label{eq:spatial-full-system-simplified}
@@
 For the diffusion, we require Neumann (zero-flux) boundary conditions.
-Since, we will not explicitly model intracellular concentrations $[R_i]$, we will drop the index entirely, only considering the extracellular resource.
+Since, we will not explicitly model intracellular concentrations $[R_{i,c}]$, we will drop the index entirely, only considering the extracellular resource.
 %
@@
 \subsubsection{Physical mechanics \& interactions}
-We represent cells as spherical objects in 2D space.
-They can be represented by their position $x_c$ and radius $r_c$ which is of course related to their (2D) volume $V_c = \pi r_c^2$.
+We model cells as spherical objects in 2D space.
+They can be represented by their position $x_c$ and radius $r_c$ which is related to their (2D) volume $V_c = \pi r_c^2$.
 As the coupled reactions from equations~\eqref{eq:spatial-full-system-simplified} change $V_c$, we also automatically adapt the cell radius accordingly.
-Notice however, that how the radius is updated does not correspond to the cellular reactions defined previously but is a purely mechanical consideration.
+% Notice however, that how the radius is updated does not correspond to the cellular reactions defined previously but is a purely mechanical consideration.
 The dimensionality of the problem does not find its way into the cellular reactions but only manifests itself in these mechanical considerations.
-The underlying equation of motion is that of a standard damped particle
-\begin{equation}
-    m \ddot{\vec{x}} = -\eta\dot{\vec{x}} + \vec{F}
+The dynamics are described by newtondian motions for particle damping.
+\begin{equation}
+    m \ddot{\vec{x}} = -\eta\dot{\vec{x}} + \vec{F}.
 \end{equation}
@@
 We assume that the cells are not sticky and consider only a repulsive interaction potential, leaving out any possible attractive forces~\citep{Garrett2008}.
-These mechanisms will result in the overall expansion of the colony as individual cells push each further other apart.
+These mechanisms will result in the overall expansion of the colony as individual cells push each other further apart.
 %
@@
 Since our goal is to include only the necessary details of the underlying biological problem, we drastically simplify the description of the cell cycle.
-Cell growth is already being described the system of coupled \acp{ode} given in equations~\eqref{eq:spatial-full-system-simplified}.
+Cell growth is already being described by the system of coupled \acp{ode} given in equations~\eqref{eq:spatial-full-system-simplified}.
 Thus, we introduce an additional rule that performs the division process when a certain volume threshold is reached.
@@
 \subsection{Comparing \acs{abm} and \acs{ode} models}
-The \ac{abm} model should be considered to be the more fundamental model compared to the \ac{ode} model in the sense that it incorporates more biological aspects and maps them closer to the actual behaviour of the bacteria.
+The \ac{abm} model should be considered to be more fundamental compared to the \ac{ode} description in the sense that it incorporates more biological aspects and maps them closer to the actual behaviour of the bacteria.
 We are thus able to derive the \ac{ode} pool model from its \ac{abm} version via a process of homogenization (see Appendix~\ref{subsec:supplement-abm-details}).
@@
     \caption{Mappings between \ac{ode} and \ac{abm} parameters.}
+    \label{tab:mappings-ode-abm}
 \end{table*}

### 2025-11-04 paper/pool_model.tex
fix some formulations/typos

@@
     \caption{
-        Simulation snapshots at $1.04$\unit{\day},$2.08$\unit{\day},$3.13$\unit{\day} and $4.17$\unit{\day} corresponding to the parameters given in table~table~\ref{tab:spatial_limit_init}.
+        Simulation snapshots at $1.04$\unit{\day},$2.08$\unit{\day},$3.13$\unit{\day} and $4.17$\unit{\day} corresponding to the parameters given in Table~\ref{tab:spatial_limit_init}.
         Species $B$ is colored blue while species $A$ is colored in red.
@@
 Figure~\ref{fig:spatial-growth-curve} shows multiple results for various configurations of the \ac{abm}, differing only in their initial spatial configuration and diffusivity.
-We see that in the highly-diffusive homogenous case of subfigure (\textbf{A}), the pool model is a good fit to the \ac{abm} model.
-Heterogeneity and slower diffusion cause larger differences between the two models as can be seen in subfigures (\textbf{B}) and \textbf{C}.
+We see that in the highly-diffusive homogenous case of subfigure (A), the pool model is a good fit to the \ac{abm} model.
+Heterogeneity and slower diffusion cause larger differences between the two models as can be seen in subfigures (B) and (C).
 The combination of the two effects shows the largest difference between model outputs.
-In subfigure \textbf{C}, due to the heterogeneous cell distribution (see Appendix~\ref{fig:supplement-spatial-initial-config-cells}), overall inhibition of species $B$ is significantly reduced.
-Subfigure (\textbf{D}) shows how a randomized homogenous initial distribution can lead to even more notable effects of inhibition.
+In subfigure (C), due to the heterogeneous cell distribution (see Appendix~\ref{fig:supplement-spatial-initial-config-cells}), overall inhibition of species $B$ is significantly reduced.
+Subfigure (D) shows how a randomized homogenous initial distribution can lead to even more notable effects of inhibition.
 In contrast to the other results, here the final values of species $A$ and $B$ have larger/smaller counts respectively.
@@
 These molecules then affect the same-species bacteria growth controlling the size of the population.
-The described above approach provides a primary predictive model with transition rates between different pools as parameters.
-To include the influence of external factors, i.e., temperature, pH, pressure, etc., one can simply use the secondary models for the transition rates, e.g.,, the Ratkowsky model for temperature dependence~\citep{ratkowsky_relationship_1982}.\\
+The approach described above provides a primary predictive model with transition rates between different pools as parameters.
+To include the influence of external factors, i.e., temperature, pH, pressure, etc., one can simply use the secondary models for the transition rates, e.g., the Ratkowsky model for temperature dependence~\citep{ratkowsky_relationship_1982}.\\
 By constructing an \ac{abm}, we have exemplified how close the devised model is to the individual biological 
@@
 
-We introduced the pool model as a framework to unify existing approaches and allows researchers to follow a mechanistic approach during model development.
+We introduced the pool model as a framework to unify existing approaches and allow researchers to follow a mechanistic approach during model development.
 It is able to capture the effects of many established models and provides a lens which is useful to generalize existing systems and results in better overall explainability.

### 2025-11-04 paper/pool_model.tex
relabel supplementary sections
remove code samples

@@
 \usepackage{lipsum}
-\usepackage{listings}
-\usepackage{listings-rust}
 \usepackage[switch]{lineno}
@@
 
-% Define Rust lisintgs
-\lstset{
-    language=Rust,
-    showstringspaces=false
-}
-\DeclareFloatingEnvironment[
-  name=Listing,
-  listname=List of Listings,
-  within=section,
-]{listing}
-
 \restylefloat{table}
@@
   \renewcommand{\thesection}{}
-  \setcounter{subsection}{0}
-  \renewcommand{\thesubsection}{S\arabic{subsection}}
-  \setcounter{table}{0}
+  \setcounter{section}{0}
+  \renewcommand{\thesection}{S\arabic{section}}
+  % \setcounter{table}{0}
   \renewcommand{\thetable}{S\arabic{table}}%
-  \setcounter{figure}{0}
+  % \setcounter{figure}{0}
   \renewcommand{\thefigure}{S\arabic{figure}}%
-  \setcounter{listing}{0}
-  \renewcommand{\thelisting}{S\arabic{listing}}
 }
@@
 \colorlet{numb}{magenta!60!black}
-
-\lstdefinelanguage{json}{
-    basicstyle=\normalfont\ttfamily,
-    numbers=left,
-    numberstyle=\scriptsize,
-    stepnumber=1,
-    numbersep=8pt,
-    showstringspaces=false,
-    breaklines=true,
-    frame=lines,
-    backgroundcolor=\color{background},
-    literate=
-     *{0}{{{\color{numb}0}}}{1}
-      {1}{{{\color{numb}1}}}{1}
-      {2}{{{\color{numb}2}}}{1}
-      {3}{{{\color{numb}3}}}{1}
-      {4}{{{\color{numb}4}}}{1}
-      {5}{{{\color{numb}5}}}{1}
-      {6}{{{\color{numb}6}}}{1}
-      {7}{{{\color{numb}7}}}{1}
-      {8}{{{\color{numb}8}}}{1}
-      {9}{{{\color{numb}9}}}{1}
-      {:}{{{\color{punct}{:}}}}{1}
-      {,}{{{\color{punct}{,}}}}{1}
-      {\{}{{{\color{delim}{\{}}}}{1}
-      {\}}{{{\color{delim}{\}}}}}{1}
-      {[}{{{\color{delim}{[}}}}{1}
-      {]}{{{\color{delim}{]}}}}{1},
-}
 
@@
 It is however straightforward, to set up these equations.
-%
-%
-% ................................................................................................
-\subsubsection{Code samples}
-The code samples show sample code from the simulation that calculate cellular interactions and cell cycle.
-Since backend of the simulation is written in Rust, so is the implementation of cellular aspects.
-Additionally, bindings are generated such that the full power of the simulation tool can be accessed.
-This significantly improves handling and development time of new models.
-\begin{listing}
-\begin{lstlisting}[language=rust]
-let z = ext_pos - own_pos;
-let r = z.norm();
-let sigma = r / (self.cell_radius + ext_radius);
-if sigma < 1.0 {
-    let q = 0.2;
-    let dir = z.normalize();
-    let modifier = (1.0 + q) / (q + sigma);
-    return self.potential_strength * dir * modifier;
-} else {
-    return Vector2::zero();
-}
-\end{lstlisting}
-\caption{
-    Sample code for calculation of interactions between two bacteria cells.
-}
-\label{listing:spatial-interaction-strength}
-\end{listing}
-%
-%
-\begin{listing}
-\begin{lstlisting}[language=rust]
-// Clone existing cell
-let mut c2 = c1.clone();
-
-let r = c1.interaction.cell_radius;
-
-// Make both cells smaller
-c1.interaction.cell_radius /= std::f64::consts::SQRT_2;
-c2.interaction.cell_radius /= std::f64::consts::SQRT_2;
-
-// Generate cellular splitting direction randomly
-let angle_1 = rng.gen_range(0.0..2.0 * std::f64::consts::PI);
-let dir_vec = nalgebra::Rotation2::new(angle_1) * Vector2::from([1.0, 0.0]);
-
-// Define new positions for cells
-// It is randomly chosen if the old cell is left or right
-let offset = dir_vec * r / std::f64::consts::SQRT_2;
-let old_pos = c1.pos();
-
-// Increase the remaining food in the bacteria by the fraction of volumes
-// from before and after the division. This exactly coincides with factor 2
-// since we half the volume of the cell
-c1.cellular_reactions.intracellular_concentrations[0] *= 2.0;
-c2.cellular_reactions.intracellular_concentrations[0] *= 2.0;
-
-c1.set_pos(&(old_pos + offset));
-c2.set_pos(&(old_pos - offset));
-\end{lstlisting}
-\caption{
-    Sample code for dividing one bacteria cell into two.
-}
-\end{listing}
 %

### 2025-11-05 paper/pool_model.tex
reformulate derivation of new equations for ABM
- remove table in favor of list of equations for mappings
- also adapt supplement
- add derivation for transition from lag-phase to growth phase for ABM

@@
         % \dot{[R_e]} &= \sum\limits_c -\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c) + D_R\Delta [R]\\
-        \dot{[R]} &= u \sum\limits_{c_A}\frac{V_{c_A}}{V_D} [R_e] \delta(x-x_{c_A})\\
-            &-\frac{u}{1+\nu_I [I]} \sum\limits_{c_B}\frac{V_{c_B}}{V_D} [R_e] \delta(x-x_{c_B})\\
-            &+D_R \Delta [R_e]\\
+        \dot{[R]} &= u \sum\limits_{c_A}\frac{V_{c_A}}{V_D} [R] \delta(x-x_{c_A})\\
+            &-\frac{u}{1+\nu_I [I]} \sum\limits_{c_B}\frac{V_{c_B}}{V_D} [R] \delta(x-x_{c_B})\\
+            &+D_R \Delta [R]\\
         \dot{[I]} &= \sum\limits_c b\frac{V_c}{V_D}\delta(x-x_c) + D_I\Delta [I]
@@
 The \ac{abm} model should be considered to be more fundamental compared to the \ac{ode} description in the sense that it incorporates more biological aspects and maps them closer to the actual behaviour of the bacteria.
-We are thus able to derive the \ac{ode} pool model from its \ac{abm} version via a process of homogenization (see Appendix~\ref{subsec:supplement-abm-details}).
+We are thus able to derive the \ac{ode} pool model from its \ac{abm} version via a process of homogenization (see Appendix~\ref{sec:supplement-abm-derivation}).
 By doing so, we find a mapping between parameters from the \ac{abm} to parameters used in the \ac{ode} description.
@@
 \begin{align}
-    \dot{R} &= -\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V\\
-    \dot{I} &= b V\\
-    \dot{V} &= \frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V.
+    \dot{L_A} &= -\lambda L_A\frac{R}{R_0}\\
+    \dot{L_B} &= -\lambda L_B\frac{R}{R_0}\\
+    \dot{V_A} &= u\sigma R \frac{V_A}{V_D} + \lambda L_A\frac{R}{R_0}\\
+    \dot{V_B} &= \frac{u}{1+\nu_I V_D^{-1} I} \sigma R \frac{V_B}{V_D} + \lambda L_B\frac{R}{R_0}\\
+    \dot{R} &= - R \frac{V_A}{V_D} - \frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V_B\\
+    \dot{I} &= b V_A
 \end{align}
@@
 \end{align}
-We compare the most general version and thus omit species $A$ for the moment.
-\begin{table*}
-    \centering
-    \renewcommand{\arraystretch}{1.5}
-    \begin{tabularx}{0.8\textwidth}{@{}l *3{>{\centering\arraybackslash}X}@{}}
-        & \textbf{\acs{ode}}
-        & \textbf{\acs{abm}}
-        & \textbf{Mappings}\\
-
-        \toprule
-
-        \multirow{2}{*}{Resource}
-        & \multirow{2}{*}{$\displaystyle \dot{R} = - \frac{\alpha}{(1 + \mu_I I) N_t} R G$}
-        & \multirow{2}{*}{$\displaystyle \dot{R} = -\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V$}
-        & $\displaystyle N_t = [R_0] V_D \sigma$\\
-
-        &&& $\displaystyle \mu_I = \frac{\nu_I}{V_D}$\\
-
-        Volume
-        & $\displaystyle \dot{G} = \frac{\alpha}{1 + \mu_I I} R G$
-        & $\displaystyle \dot{V} = \frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V$
-        & $\displaystyle \alpha = u[R_0]\sigma$\\
-
-        Inhibitor
-        & $\displaystyle \dot{I} = \kappa G_A$
-        & $\displaystyle \dot{I} = b V$
-        & $\displaystyle \kappa = b$\\
-        \bottomrule
-    \end{tabularx}
-    \caption{Mappings between \ac{ode} and \ac{abm} parameters.}
-    \label{tab:mappings-ode-abm}
-\end{table*}
-We can use these mappings to numerically compare both approaches.
-Table~\ref{tab:spatial_limit_init} shows a comparison of parameters of two simulations for the \ac{ode} and \ac{abm} model which have been mapped onto each other.
+When comparing these equations, we can map them onto each other via the relations
+\begin{align}
+    \begin{split}
+        N_t &= [R_0]V_D\sigma\\
+        \mu_I &= \frac{\nu_I}{V_D}\\
+        \alpha &= u[R_0]\sigma\\
+        \kappa &= \sigma.
+    \end{split}
+    \label{eq:ode-abm-mappings}
+\end{align}
+Furthermore, we can use these mappings to numerically compare both approaches.
+Table~\ref{tab:spatial_limit_init} shows an example comparison of parameters of the \ac{ode} and \ac{abm} which have been mapped onto each other.
 Figure~\ref{fig:spatial-snapshots} shows snapshots of a numerical simulation of the described system.
@@
 The chosen diffusion constants $D_I=D_R=30.0$\unit{\micro\metre\squared\per\minute} are in the order of small molecules such as $NaCl$ while the lower values $D_I=D_R=5.0$\unit{\micro\metre\squared\per\minute} are around values of Glucose in water.
-\begin{figure}
+\begin{figure*}
     \centering
-    \begin{tikzonimage}[width=0.5\columnwidth]
+    \begin{tikzonimage}[width=\columnwidth]
         {Figures-abm-homogenous-abm_ode_comparison.png}
@@
     \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.5\columnwidth]
+    \begin{tikzonimage}[width=\columnwidth]
         {Figures-abm-homogenous-low-diffusion-abm_ode_comparison.png}
@@
     \end{tikzonimage}\\
-    \begin{tikzonimage}[width=0.5\columnwidth]
+    \begin{tikzonimage}[width=\columnwidth]
         {Figures-abm-inhomogenous-abm_ode_comparison.png}
@@
     \end{tikzonimage}%
-    \begin{tikzonimage}[width=0.5\columnwidth]
+    \begin{tikzonimage}[width=\columnwidth]
         {Figures-abm-random-abm_ode_comparison.png}
         \node at (0, 1)[anchor=north west]{\textbf{D}};
-    \end{tikzonimage}%
+    \end{tikzonimage}
     \caption{
@@
     \label{fig:spatial-growth-curve}
-\end{figure}
+\end{figure*}
 These results show how variations in properties that can not be described by simple \ac{ode} models can lead to large deviations in the growth curves of the species.
@@
 \beginsupplement
-\section{Appendix}
 % ================================================================================================
-\subsection{General mathematical formalism}
+\section{General mathematical formalism}
 \label{ssec:supplement1}
@@
 \end{equation}
-
-\subsection{\acs{abm} details}
-\label{subsec:supplement-abm-details}
-
-\subsubsection*{Deriving the pool model}
+%
+%
+\vfill
+\pagebreak
+\section{Deriving the pool model}
+\label{sec:supplement-abm-derivation}
+\paragraph{Reactions}
 Since the \ac{abm} can be considered more fundamental than the \ac{ode} model, we should be able to derive the latter one from the former.
@@
 We integrate the spatial equations over the total domain starting with the resource concentration $[R]$.
+\begin{alignat}{8}
+    \int\limits_D\dot{[R]}\text{dVol}
+    &= \int\limits_D\bigg[&-&u \sum\limits_{c_A}\frac{V_{c_A}}{V_D} [R] \delta(x-x_{c_A})
+        &-&\frac{u}{1+\nu_I [I]} \sum\limits_{c_B}\frac{V_{c_B}}{V_D} [R] \delta(x-x_{c_B})
+        +D_R \Delta [R]\bigg]\text{dVol}\\
+    &= &-&u\sum\limits_{c_A}V_{c_A}\int\limits_D [R] \frac{\delta(x-x_{c_A})}{V_D} \text{dVol}
+        &-&\frac{u}{1+\nu_I [I]} \sum\limits_{c_B}V_{c_B} \int\limits_D [R] \frac{\delta(x-x_{c_B})}{V_D}\text{dVol}
+        +D_R \int\limits_{\partial D} \nabla[R]\text{dVol}\\
+    &= &-&u\sum\limits_{c_A} V_{c_A} [R]_{x=x_{c_A}}
+        &-& \frac{u}{1+\nu_I [I]} \sum\limits_{c_B} V_{c_B} [R]_{x=x_{c_B}}\\
+    &= &-&u\sum\limits_{c_A} V_{c_A} [R]
+        &-& \frac{u}{1+\nu_I [I]} \sum\limits_{c_B} V_{c_B} [R]\\
+    &= &-&uV_A [R] &-&\frac{u}{1+\nu_I[I]} V_B [R]\\
+    &= &-&u \frac{V_A}{V_D}R &-&\frac{u}{1+\nu_I V_D^{-1} I} \frac{V_B}{V_D}R
+\end{alignat}
+We can split up the integral into three parts.
+The last part can then be transformed into an integral over the flux of the resource along the boundary.
+Since we chose Neumann boundary conditions, this contribution vanishes identically.
+Afterwards, the remaining parts can be evaluated by using the delta-distribution.
+Note that we additionally need to consume a factor of $V_D^{-1}$ which is the inverse of the domain
+volume when evaluating the delta function.
+We evaluate the resource concentration $[R]$ at the position of the individual cells and sum over all contributions.
+Now we assume homogeneity in the sense, that we can reduce $[R]$ to only a variable with time-dependence, with homogeneous spatial distribution.
+This means that every contribution does not depend on the position at which it was evaluated, i.e., $[R]_{x=x_c}=[R]$.
+Finally, we can perform the sum to gain the total volume of all cells $V_{A/B}=\sum\limits_c V_{c_{A/B}}$.
+Following the same principle as before, we can integrate the inhibitor equation.
 \begin{alignat}{7}
-    \int\limits_D\dot{[R]}\text{dVol} 
-    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c) &&+ D_R\Delta [R]\bigg]\text{dVol}\\
-    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c)\bigg]\text{dVol} &&+ \int\limits_D D_R\Delta [R]\text{dVol}\\
-    &= \int\limits_D\bigg[&&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c)\bigg]\text{dVol} &&+ \int\limits_{\partial D} D_R\vec{\nabla} [R]\vec{\text{dA}}\\
-    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c \int\limits_D \frac{1}{V_D} [R] \delta(x-x_c)\text{dVol}\\
-    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c [R]_{x=x_c}\\
-    &= &&\sum\limits_c &&-\frac{u}{1+\nu_I [I]} V_c [R]\\
-    &= &&&&-\frac{u}{1+\nu_I [I]V_D V_D^{-1}} V [R]\frac{V_D}{V_D}\\
-    \dot{R} &= &&&&-\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V\\
+    \int\limits_D\dot{[I]} &= \int\limits_D \bigg[&&\sum\limits_{c_A} b\frac{V_{c_A}}{V_D}&&\delta(x-x_{c_A}) &&+ D_I\Delta [I]\bigg]\text{dVol}\\
+    &= \int\limits_D&&\sum\limits_{c_A} b\frac{V_{c_A}}{V_D}&&\delta(x-x_{c_A}) \text{dVol} &&+ \int\limits_D D_I\Delta [I]\text{dVol}\\
+    &= \int\limits_D&&\sum\limits_{c_A} b\frac{V_{c_A}}{V_D}&&\delta(x-x_{c_A}) \text{dVol} &&+ \int\limits_{\partial D} D_I\nabla [I]\vec{\text{dA}}\\
+    &= &&\sum\limits_{c_A} b V_{c_A}\int\limits_D \frac{1}{V_D} &&\delta(x-x_{c_A}) \text{dVol}
 \end{alignat}
-We can split up the integral into two parts.
-The second part can then be transformed into an integral over the flux of the resource along the boundary.
-Since we chose Neumann boundary conditions, this contribution vanishes identically.
-Afterwards, the first part can be evaluated by using the delta-distribution.
-Note that we additionally need to consume a $V_D^{-1}$ term since our volume is finite.
-We evaluate the resource concentration $[R]$ at the position of the individual cells and sum over all contributions.
-Now we assume homogeneity in the sense, that we can reduce $[R]$ to only a single variable instead of considering its spatial distribution.
-This means that every contribution does not depend on the position at which it was evaluated, i.e., $[R]_{x=x_c}=[R]$.
-Finally, we can perform the sum to gain the total volume of all cells $V=\sum\limits_c V_c$.
-\begin{alignat}{7}
-    \int\limits_D\dot{[I]} &= \int\limits_D \bigg[&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) &&+ D_I\Delta [I]\bigg]\text{dVol}\\
-    &= \int\limits_D&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) \text{dVol} &&+ \int\limits_D D_I\Delta [I]\text{dVol}\\
-    &= \int\limits_D&&\sum\limits_c b\frac{V_c}{V_D}&&\delta(x-x_c) \text{dVol} &&+ \int\limits_{\partial D} D_I\nabla [I]\vec{\text{dA}}\\
-    &= &&\sum\limits_c b V_c\int\limits_D \frac{1}{V_D} &&\delta(x-x_c) \text{dVol}\\
-    \dot{I} &= b V
+%
+\paragraph{Lag-Phase}
+Cells are initially assumed to be in a lag-phase which they leave with probability $P(\text{transition})$.
+Dividing this probability by the time increment $\Delta t$, gives us the rate at which the transition is occuring.
+This expression can be transfered to a collective rate of change by multiplying with the volume of cells in lag-phase $L_{A/B}$.
+\begin{align}
+    P(\text{transition}) &= \Delta t \lambda \frac{[R]}{[R_0]}\\
+    \frac{1}{\Delta t}P(\text{transition}) &= \lambda \frac{[R]}{[R_0]}\\
+    \frac{L_{A/B}}{\Delta t}P(\text{transition}) &= \lambda L_{A/B}\frac{R}{R_0}\\
+    \partial_t L_{A/B} &= -\lambda L_{A/B}\frac{R}{R_0}
+\end{align}
+In the last step, we take the expectation value and equate it with the negative of its derivative $\partial_t L_{A/B}$.
+%
+%
+\paragraph{Growth}
+For the equation describing the growth of the cell volume, we simply add up cells for the respective species to obtain their total volume.
+\begin{alignat}{5}
+    \sum\limits_{c_A}\dot{V_{c_A}}
+    &= &\sum\limits_{c_A} u& \sigma [R]V_{c_A} + \lambda L_A\frac{R}{R_0} &= &&\sigma [R]V_A + \lambda L_A\frac{R}{R_0}\\
+    \sum\limits_{c_B}\dot{V_{c_B}}
+    &= &\frac{u}{1+\nu_I [I]} &\sigma [R]V_B + \lambda L_B\frac{R}{R_0} &= &\sum\limits_{c_B} \frac{u}{1+\nu_I [I]} &\sigma [R]V_{c_B} + \lambda L_B\frac{R}{R_0}
 \end{alignat}
-Following the same principle as before, we can integrate the inhibitor equation.
-For the equation describing the growth of the cell volume, we simply add up all cells to obtain the total volume of all cells.
-\begin{alignat}{3}
-    \sum\limits_c\dot{V_c}
-    &= \sum\limits_c&&\frac{u}{1+\nu_I [I]} \sigma [R]V_c\\
-    &= &&\frac{u}{1+\nu_I [I]} \sigma [R]V\\
-    \dot{V} &= &&\frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V
-\end{alignat}
+In addition, we can reformulate $\tilde{\sigma}=\sigma/V_D$ thus redefining the conversion rate $\sigma$ in terms of the total volume $V_D$.
 We can thus summarize the equations obtained
 \begin{align}
-    \dot{R} &= -\frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V\\
-    \dot{I} &= b V\\
-    \dot{V} &= \frac{u}{1+\nu_I V_D^{-1} I} \frac{1}{V_D}\sigma R V.
-\end{align}
-It is clear that the original model is recovered this way.
-To obtain the full equations, we additionally need to consider transition between states of lag-phase and growth phase not depicted here.
-The preceding equations all assume that cells are in growth phase.
-It is however straightforward, to set up these equations.
+    \dot{L_A} &= -\lambda L_A\frac{R}{R_0}\\
+    \dot{L_B} &= -\lambda L_B\frac{R}{R_0}\\
+    \dot{V_A} &= u\sigma R \frac{V_A}{V_D} + \lambda L_A\frac{R}{R_0}\\
+    \dot{V_B} &= \frac{u}{1+\nu_I V_D^{-1} I} \sigma R \frac{V_B}{V_D} + \lambda L_B\frac{R}{R_0}\\
+    \dot{R} &= - R \frac{V_A}{V_D} - \frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V_B\\
+    \dot{I} &= b V_A
+\end{align}
+In this way, we are able to recover the \ac{ode} model.
 %
@@
 % ................................................................................................
-\subsubsection{Full list of parameters}
-\label{sec:supplement-full-parameters}
-
-\begin{table}
+\vfill
+\pagebreak
+\section{Full list of parameters}
+We utilize the calculate mappings (see equations~\eqref{eq:ode-abm-mappings}) between the \ac{ode} formulation and \ac{abm} to compare values for parameters which serve a similar purpose within the two simulations.
+Table~\ref{tab:spatial_limit_init} lists these parameters together with their values.
+Most parameters have clear $1:1$ mappings to ones within the other model.
+However, some such as the ones related to spatial dynamics (damping $\eta$, mass $m$, potential
+strength $\epsilon$, diffusion $D_R,D_I$) do not.
+\begin{table}[H]
     \centering
@@
 % ................................................................................................
-\subsubsection{Homogeneous and inhomogeneous initial cell distributions}
+\vfill
+\pagebreak
+\section{Homogeneous and inhomogeneous initial cell distributions}
 To clarify what we mean by homogeneous and inhomogeneous initial configurations, we include snapshots of the simulation.
@@
 The snapshots were taken after $2.08$\unit{\day}.
-\begin{figure}
+\begin{figure}[H]
     \centering

### 2025-11-05 paper/pool_model.tex
add missing uptake rate to eq 4.19
Corrections in equations and language style

@@
 
-\title{Pool Model - A unified Mathematical\\ Modelling Framework for Food Science}
+\title{Pool Model - A unified Mathematical\\ Modeling Framework for Food Science}
 \author[1]{Polina Gaindrik}
@@
     For this, we construct an \ac{abm} for a two-species system, showcasing that the resulting pool model
-    arises naturally from the individual behaviour of many cellular agents.
+    arises naturally from the individual behavior of many cellular agents.
     We further apply this \ac{abm} to study the role of properties such as diffusion and spatial heterogeneity
@@
 %     \item Polynomial models 
-%     \item Square Root-Type Models, e.g.,, Ratkowski for temperature~\citep{ratkowsky_relationship_1982}
+%     \item Square Root-Type Models, e.g., Ratkowski for temperature~\citep{ratkowsky_relationship_1982}
 %     \item The Cardinal Parameter Model~\citep{zwietering_decision_1992}
@@
 cells that can be approximated as a continuum.
-On the other hand, \acp{abm} excel at describing the individual behaviour of cells which can then
+On the other hand, \acp{abm} excel at describing the individual behavior of cells which can then
 lead to more complex collective phenomena.
@@
 The control of the population density with respect to the available nutrition by means of cell communication is called quorum sensing.
-By releasing extracellular signalling molecules (autoinducer molecules) it is possible for the bacteria to synchronize the gene expression of the whole group~\citep{ng_bacterial_2009}.
+By releasing extracellular signalling molecules (autoinducers) it is possible for the bacteria to synchronize the gene expression of the whole group~\citep{ng_bacterial_2009}.
 Instead of modeling complex signalling networks, we choose a simpler approach to describe the dependence of growth on the nutrient availability.
@@
 Note that the pools $L$, $G$ and $D$ represent the number of bacteria (or number of bacteria per unit area/volume) while $R$ represents an abstract resource pool.
-Let's say $G=N_B/V$ and $[G]=1/V$, where $N_B$ is the number of bacteria and $V$ is the volume.
-Then $R=N_R/V$ and $[R]=1/V$, where $N_R$ is the number of resource molecules (or any appropriate unit, e.g., mol).
-Then $[\alpha_0]=V/s$, $[N_t]=1/V$, and $[\alpha]=1/s$.
+Let's say $G=N_B/V$ and $[G]=\unit{1/V}$, where $N_B$ is the number of bacteria and $V$ is the volume.
+Then $R=N_R/V$ and $[R]=\unit{1/V}$, where $N_R$ is the number of resource molecules (or any appropriate unit, e.g., \unit{mol}).
+Then $[\alpha_0]=\unit{V/s}$, $[N_t]=\unit{1/V}$, and $[\alpha]= \unit{1/s}$.
 $N_t$ denotes the maximal size of the total bacterial population due to environmental limitations (Figure \ref{fig:3pool_resource_plots}).
@@
 Dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\citep{kaprelyants_dormancy_1993}.
-This survival strategy can appear in response to different external stresses, e.g.,, nutrient deficiency, temperature stress or extreme fluctuations of environment parameters.
+This survival strategy can appear in response to different external stresses, e.g., nutrient deficiency, temperature stress or extreme fluctuations of environment parameters.
 And after "difficult times" bacteria can continue cell division~\citep{kell_viability_1998}.
@@
     \begin{center}
-    %\includegraphics[width=0.9\columnwidth]{Figures-pool_model_3pools_resource_tempshift.pdf}
     \includegraphics[width=0.9\columnwidth]{Figures-pool_model_3pools_resource_tempshift1.pdf}\\
@@
     \begin{center}
-    %\includegraphics[width=\columnwidth]{Figures-pool_model_2pools_toxin.pdf}
     \includegraphics[width=0.9\columnwidth]{Figures-pool_model_2pools_toxin1.pdf}\\
@@
 \subsubsection{Mutual inhibition}
-Another special case of inter-species competition is when the produced toxin does not cause the death of the bacteria but rather inhibit their growth.
+Another special case of interspecies competition is when the produced toxin does not cause the death of the bacteria but rather inhibit their growth.
 Such a situation with 2 bacteria species, a shared resource and produced inhibitor was modeled by \citet{lenski_coexistence_1986}.
@@
 In this way, we can show that the bulk-based \ac{ode} model is directly related to the mechanistic
-single-cell behaviour of the bacteria.
+single-cell behavior of the bacteria.
 To this end, we consider the system consisting of two bacterial species that can be described by the two internal states: lag phase and growth.
@@
 % - \citep{emonet_agentcell_2005, gorochowski_bsim_2012, li_nufeb_2019, kreft_bacsim_1998}: examples of another toolboxes for bacterial communities modeling (ABM).\\
-% More about the modelling of the motile bacteria can be seen in works:~\citep{decoene_microscopic_2011, rosser_modelling_2014, sokolov_physical_2012, li_amplified_2008}.
+% More about the modeling of the motile bacteria can be seen in works:~\citep{decoene_microscopic_2011, rosser_modelling_2014, sokolov_physical_2012, li_amplified_2008}.
 %
@@
     \begin{split}
-        \dot{L_A} &= -\lambda_A R L_A\\
-        \dot{G_A} &= \lambda_A R L_A + \alpha_A R G_A\\
-        \dot{L_B} &= -\lambda_B R L_B\\
-        \dot{G_B} &= \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B.
+        \dot{L}_A &= -\lambda_A R L_A\\
+        \dot{G}_A &= \lambda_A R L_A + \alpha_A R G_A\\
+        \dot{L}_B &= -\lambda_B R L_B\\
+        \dot{G}_B &= \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B.
     \end{split}
@@
 \ac{ode}-based systems.\\
-To construct our \ac{abm}, we use the modeling framework \texttt{cellular\_raza}~\citep{Pleyer2025}.
- allows us to define our bacterial agents from a clean slate.
+To construct our \ac{abm}, we use the modeling framework \texttt{cellular\_raza}~\citep{Pleyer2025} that allows us to define our bacterial agents from a clean slate.
 With this flexibility, researchers are able to construct minimal models using only as many parameters which are necessary.
@@
 The uptake only occurs at the position of the cell $x_{c_A}$.
-Furthermore, we let cells of species $A$ produce a freely diffusable inhibitor $I$ with production rate $b$ and diffusion constant $D_I$ which inhibits cells of species $B$ in their uptake of nutrients.
+Furthermore, we let cells of species $A$ produce a freely diffusible inhibitor $I$ with production rate $b$ and diffusion constant $D_I$ which inhibits cells of species $B$ in their uptake of nutrients.
 \begin{equation}
@@
 \begin{equation}
-    \dot{V_c} = a\sigma [R_{i,c}] V_c
+    \dot{V}_c = a\sigma [R_{i,c}] V_c
 \end{equation}
@@
         % TODO format these equations
-        \dot{V_{c_B}} &= u \sigma [R]V_c\\
-        \dot{V_{c_A}} &= \frac{u}{1+\nu_I [I]} \sigma [R]V_c\\
+        \dot{V}_{c_B} &= u \sigma [R]V_c\\
+        \dot{V}_{c_A} &= \frac{u}{1+\nu_I [I]} \sigma [R]V_c\\
         % \dot{[R_e]} &= \sum\limits_c -\frac{u}{1+\nu_I [I]} \frac{V_c}{V_D} [R] \delta(x-x_c) + D_R\Delta [R]\\
@@
 The dimensionality of the problem does not find its way into the cellular reactions but only manifests itself in these mechanical considerations.
-The dynamics are described by newtondian motions for particle damping.
-\begin{equation}
-    m \ddot{\vec{x}} = -\eta\dot{\vec{x}} + \vec{F}.
+The dynamics are described by Newton's law of motion for particle damping.
+\begin{equation}
+    m \ddot{\mathbf{x}} = -\eta\dot{\mathbf{x}} + \mathbf{F}.
 \end{equation}
@@
 \begin{alignat}{3}
-    \vec{z} &= \vec{x_1} - \vec{x_2} \hspace{2em}&& r = ||\vec{z}||\\
+    \mathbf{z} &= \mathbf{x_1} - \mathbf{x_2} \hspace{2em}&& r = ||\mathbf{z}||\\
     \sigma &= \frac{r}{r_1+r_2} && q=0.2\\
-    \vec{F} &= \epsilon \frac{1+q}{q+\sigma} \theta(\sigma - 1) \frac{1}{r}\vec{z}
+    \mathbf{F} &= \epsilon \frac{1+q}{q+\sigma} \theta(\sigma - 1) \frac{1}{r}\mathbf{z}
 \end{alignat}
@@
 If this boolean is true, we move the cell from the inactive state to the active state where it starts to consume nutrients and grow.
-The concatenation of many such stochastic processes on a population-based level converges identically to the lag-phase behaviour of the \ac{ode} model.
+The concatenation of many such stochastic processes on a population-based level converges identically to the lag-phase behavior of the \ac{ode} model.
 %
@@
 \subsection{Comparing \acs{abm} and \acs{ode} models}
-The \ac{abm} model should be considered to be more fundamental compared to the \ac{ode} description in the sense that it incorporates more biological aspects and maps them closer to the actual behaviour of the bacteria.
+The \ac{abm} model should be considered to be more fundamental compared to the \ac{ode} description in the sense that it incorporates more biological aspects and maps them closer to the actual behavior of the bacteria.
 We are thus able to derive the \ac{ode} pool model from its \ac{abm} version via a process of homogenization (see Appendix~\ref{sec:supplement-abm-derivation}).
@@
 \begin{align}
-    \dot{L_A} &= -\lambda L_A\frac{R}{R_0}\\
-    \dot{L_B} &= -\lambda L_B\frac{R}{R_0}\\
-    \dot{V_A} &= u\sigma R \frac{V_A}{V_D} + \lambda L_A\frac{R}{R_0}\\
-    \dot{V_B} &= \frac{u}{1+\nu_I V_D^{-1} I} \sigma R \frac{V_B}{V_D} + \lambda L_B\frac{R}{R_0}\\
+    \dot{L}_A &= -\lambda L_A\frac{R}{R_0}\\
+    \dot{L}_B &= -\lambda L_B\frac{R}{R_0}\\
+    \dot{V}_A &= u\sigma R \frac{V_A}{V_D} + \lambda L_A\frac{R}{R_0}\\
+    \dot{V}_B &= \frac{u}{1+\nu_I V_D^{-1} I} \sigma R \frac{V_B}{V_D} + \lambda L_B\frac{R}{R_0}\\
     \dot{R} &= - R \frac{V_A}{V_D} - \frac{u}{1+\nu_I V_D^{-1}I} \frac{1}{V_D} R V_B\\
@@
 \begin{align}
-    \dot{L_A} &= -\lambda_A R L_A\\
-    \dot{G_A} &= \lambda_A R L_A + \alpha_A R G_A\\
-    \dot{L_B} &= -\lambda_B R L_B\\
-    \dot{G_B} &= \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B\\
+    \dot{L}_A &= -\lambda_A R L_A\\
+    \dot{L}_B &= -\lambda_B R L_B\\
+    \dot{G}_A &= \lambda_A R L_A + \alpha_A R G_A\\
+    \dot{G}_B &= \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B\\
     \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \mu_I I) N_t} R G_B \\
@@
     \caption{
-        Simulation snapshots at $1.04$\unit{\day},$2.08$\unit{\day},$3.13$\unit{\day} and $4.17$\unit{\day} corresponding to the parameters given in Table~\ref{tab:spatial_limit_init}.
+        Simulation snapshots at $1.04$~\unit{\day}, $2.08$~\unit{\day}, $3.13$~\unit{\day} and $4.17$~\unit{\day} corresponding to the parameters given in Table~\ref{tab:spatial_limit_init}.
         Species $B$ is colored blue while species $A$ is colored in red.
@@
 This is expected, since processes such as diffusion and cells growing outward introduce another time component which is simply assumed to be instantaneous in the \ac{ode} model.
-The chosen diffusion constants $D_I=D_R=30.0$\unit{\micro\metre\squared\per\minute} are in the order of small molecules such as $NaCl$ while the lower values $D_I=D_R=5.0$\unit{\micro\metre\squared\per\minute} are around values of Glucose in water.
+The chosen diffusion constants $D_I=D_R=30.0$~\unit{\micro\metre\squared\per\minute} are in the order of small molecules such as $NaCl$ while the lower values $D_I=D_R=5.0$~\unit{\micro\metre\squared\per\minute} are around values of Glucose in water.
 \begin{figure*}
@@
 The approach described above provides a primary predictive model with transition rates between different pools as parameters.
-To include the influence of external factors, i.e., temperature, pH, pressure, etc., one can simply use the secondary models for the transition rates, e.g., the Ratkowsky model for temperature dependence~\citep{ratkowsky_relationship_1982}.\\
+To include the influence of external factors, i.e., temperature, pH, pressure, etc., one can simply use the secondary models for the transition rates, e.g., the square-root temperature dependence~\citep{ratkowsky_relationship_1982}.\\
 By constructing an \ac{abm}, we have exemplified how close the devised model is to the individual biological 
@@
     &= \int\limits_D&&\sum\limits_{c_A} b\frac{V_{c_A}}{V_D}&&\delta(x-x_{c_A}) \text{dVol} &&+ \int\limits_D D_I\Delta [I]\text{dVol}\\
-    &= \int\limits_D&&\sum\limits_{c_A} b\frac{V_{c_A}}{V_D}&&\delta(x-x_{c_A}) \text{dVol} &&+ \int\limits_{\partial D} D_I\nabla [I]\vec{\text{dA}}\\
+    &= \int\limits_D&&\sum\limits_{c_A} b\frac{V_{c_A}}{V_D}&&\delta(x-x_{c_A}) \text{dVol} &&+ \int\limits_{\partial D} D_I\nabla [I]\mathbf{\text{dA}}\\
     &= &&\sum\limits_{c_A} b V_{c_A}\int\limits_D \frac{1}{V_D} &&\delta(x-x_{c_A}) \text{dVol}
@@
 \begin{alignat}{5}
-    \sum\limits_{c_A}\dot{V_{c_A}}
+    \sum\limits_{c_A}\dot{V}_{c_A}
     &= &\sum\limits_{c_A} u& \sigma [R]V_{c_A} + \lambda L_A\frac{R}{R_0} &= &&\sigma [R]V_A + \lambda L_A\frac{R}{R_0}\\
-    \sum\limits_{c_B}\dot{V_{c_B}}
+    \sum\limits_{c_B}\dot{V}_{c_B}
     &= &\frac{u}{1+\nu_I [I]} &\sigma [R]V_B + \lambda L_B\frac{R}{R_0} &= &\sum\limits_{c_B} \frac{u}{1+\nu_I [I]} &\sigma [R]V_{c_B} + \lambda L_B\frac{R}{R_0}
@@
         \toprule
-        Production $G_A$            & $\alpha_A$    & $0.002$\unit{\per\minute}
-        & Uptake                    & $u_A$         & $0.01$\unit{\per\minute}\\
-        Production $G_B$            & $\alpha_B$    & $0.002$\unit{\per\minute}
-        & Uptake                    & $u_B$         & $0.01$\unit{\per\minute}\\
-        Production $I$              & $\kappa$      & $0.1$\unit{\per\minute}
-        & Production $I$            & $b$           & $0.1$\unit{\per\minute}\\
-        Inhibition                  & $\mu_I$       & $1.0\cdot 10^{-7}$\unit{\per\micro\metre\cubed}
-        &                           & $\nu_I$       & $0.1$\unit{\micro\metre\cubed\per\mol}\\
+        Production $G_A$            & $\alpha_A$    & $0.002$~\unit{\per\minute}
+        & Uptake                    & $u_A$         & $0.01$~\unit{\per\minute}\\
+        Production $G_B$            & $\alpha_B$    & $0.002$~\unit{\per\minute}
+        & Uptake                    & $u_B$         & $0.01$~\unit{\per\minute}\\
+        Production $I$              & $\kappa$      & $0.1$~\unit{\per\minute}
+        & Production $I$            & $b$           & $0.1$~\unit{\per\minute}\\
+        Inhibition                  & $\mu_I$       & $1.0\cdot 10^{-7}$~\unit{\per\micro\metre\cubed}
+        &                           & $\nu_I$       & $0.1$~\unit{\micro\metre\cubed\per\mol}\\
         &&
-        & Conversion                & $\sigma$      & $0.1$\unit{\micro\metre\cubed\per\mol}\\
-        Total volume                & $N_t$& $2.0\cdot 10^5$\unit{\micro\metre\cubed}
+        & Conversion                & $\sigma$      & $0.1$~\unit{\micro\metre\cubed\per\mol}\\
+        Total volume                & $N_t$& $2.0\cdot 10^5$~\unit{\micro\metre\cubed}
         &&&\\
         &&
-        & Diffusion $R$             & $D_R$         & $30.0$\unit{\micro\metre\squared\per\minute}\\
+        & Diffusion $R$             & $D_R$         & $30.0$~\unit{\micro\metre\squared\per\minute}\\
         &&
-        & Diffusion $I$             & $D_I$         & $30.0$\unit{\micro\metre\squared\per\minute}\\
+        & Diffusion $I$             & $D_I$         & $30.0$~\unit{\micro\metre\squared\per\minute}\\
         % Note: this diffusion roughly corresponds to the one of ethanol in water
@@
         \toprule
-        &&& Damping                 & $\eta$        & $0.5$\unit{\per\minute}\\
-        &&& Mass                    & $m$           & $7.70$\unit{\pico\gram}\\
-        &&& Potential               & $\epsilon$    & $0.5$\unit{\pico\gram\micro\metre\per\minute\squared}\\
+        &&& Damping                 & $\eta$        & $0.5$~\unit{\per\minute}\\
+        &&& Mass                    & $m$           & $7.70$~\unit{\pico\gram}\\
+        &&& Potential               & $\epsilon$    & $0.5$~\unit{\pico\gram\micro\metre\per\minute\squared}\\
         \toprule
@@
         \toprule
-        Rate $L_A\rightarrow G_A$   & $\lambda_A$   & $0.0025$\unit{\per\minute}
-        &                           & $\lambda_A$   & $0.0025$\unit{\per\minute}\\
-        Rate $L_B\rightarrow G_B$   & $\lambda_B$   & $0.0005$\unit{\per\minute}
-        &                           & $\lambda_B$   & $0.0005$\unit{\per\minute}\\
-        &&& Division Vol.           & $V_t$         & $14.14$\unit{\micro\metre\cubed}\\
+        Rate $L_A\rightarrow G_A$   & $\lambda_A$   & $0.0025$~\unit{\per\minute}
+        &                           & $\lambda_A$   & $0.0025$~\unit{\per\minute}\\
+        Rate $L_B\rightarrow G_B$   & $\lambda_B$   & $0.0005$~\unit{\per\minute}
+        &                           & $\lambda_B$   & $0.0005$~\unit{\per\minute}\\
+        &&& Division Vol.           & $V_t$         & $14.14$~\unit{\micro\metre\cubed}\\
         \toprule
@@
         \toprule
-        Volume of $L_A$             & $L_A$         & $127.2$\unit{\micro\metre\cubed}
+        Volume of $L_A$             & $L_A$         & $127.2$~\unit{\micro\metre\cubed}
         & Cell Number               & $C_A$         & 18\\
-        Volume of $G_A$             & $G_A$         & $0.0$\unit{\micro\metre\cubed}
-        & Cell Volume               & $V_{c,A}$     & $7.07$\unit{\micro\metre\cubed}\\
-        Volume of $L_B$             & $L_B$         & $127.23$\unit{\micro\metre\cubed}
+        Volume of $G_A$             & $G_A$         & $0.0$~\unit{\micro\metre\cubed}
+        & Cell Volume               & $V_{c,A}$     & $7.07$~\unit{\micro\metre\cubed}\\
+        Volume of $L_B$             & $L_B$         & $127.23$~\unit{\micro\metre\cubed}
         & Cell Number               & $C_B$         & 18\\
-        Volume of $G_B$             & $G_B$         & $0.0$\unit{\micro\metre\cubed}
-        & Cell Volume               & $V_{c,B}$     & $7.07$\unit{\micro\metre\cubed}\\
+        Volume of $G_B$             & $G_B$         & $0.0$~\unit{\micro\metre\cubed}
+        & Cell Volume               & $V_{c,B}$     & $7.07$~\unit{\micro\metre\cubed}\\
         Relative Nutrients          & $R$         & $1.0$
-        & Nutrients                 & $[R_0]$       & $1.0$\unit{\mol\per\micro\metre\cubed}\\
-        Inhibitor                   & $I_0$         & $0.0$\unit{\mol}
-        & Inhibitor                 & $[I_0]$       & $0.0$\unit{\mol\per\micro\metre\cubed}\\
+        & Nutrients                 & $[R_0]$       & $1.0$~\unit{\mol\per\micro\metre\cubed}\\
+        Inhibitor                   & $I_0$         & $0.0$~\unit{\mol}
+        & Inhibitor                 & $[I_0]$       & $0.0$~\unit{\mol\per\micro\metre\cubed}\\
         \bottomrule
@@
 At this point, cells have already grown such that the sizes of cell-clumps are better visible by the naked eye.
-The snapshots were taken after $2.08$\unit{\day}.
+The snapshots were taken after $2.08$~\unit{\day}.
 \begin{figure}[H]
@@
         \includegraphics[width=\textwidth]{Figures-abm-homogenous-images-snapshot_00012000.png}
-        \subcaption{Homogeneous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
+        \subcaption{Homogeneous initial cell distribution, $D_R=D_I=30.0$~\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}%
@@
         \includegraphics[width=\textwidth]{Figures-abm-homogenous-low-diffusion-images-snapshot_00012000.png}
-        \subcaption{Homogeneous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
+        \subcaption{Homogeneous initial cell distribution, $D_R=D_I=5.0$~\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}\\
@@
         \includegraphics[width=\textwidth]{Figures-abm-inhomogenous-images-snapshot_00012000.png}%
-        \subcaption{Heterogeneous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
+        \subcaption{Heterogeneous initial cell distribution, $D_R=D_I=30.0$~\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}%
@@
         \includegraphics[width=\textwidth]{Figures-abm-random-images-snapshot_00012000.png}%
-        \subcaption{Random initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
+        \subcaption{Random initial cell distribution, $D_R=D_I=30.0$~\unit{\micro\metre\squared\per\minute}}
     \end{subfigure}

### 2025-11-06 paper/pool_model.tex
Change order of terms for better comparison of ABM and ODE
Correct mistake in definition of the toxin production (2 type)

@@
 \usepackage{xpatch}
+\usepackage{orcidlink}
 
@@
 \title{Pool Model - A unified Mathematical\\ Modeling Framework for Food Science}
-\author[1]{Polina Gaindrik}
+\author[1]{Polina Gaindrik\orcidlink{0000-0003-0928-3951}}
 \author[1]{Jonas Pleyer\fnref{fn1}}
@@
     \mathcal{F}^1(R,G)&=\kappa G_B\\
-    \mathcal{F}^2(R,G)&=\kappa\alpha_B R G_B.
+    \mathcal{F}^2(R,G)&=\kappa R G_B.
 \end{align}
@@
     \caption{
-        The two species system competing for the common nutrient $R$ with production of the toxin $T_B$ (brown line) calculated using equation~(\ref{eq:model_2sp_toxin}).
-        The growing bacteria $B$ produces toxin $T_B$ (A) constantly $\dot{T}_B^1 = \mathcal{F}^1 = \kappa G_B$ or (B) only during growth process $\dot{T}_B^2 = \mathcal{F}^2 = \kappa \alpha_B R G_B$ with
-        production rate $\kappa=0.2$.
-        The toxin $T_B$ causes the cell death of bacteria $A$ with rate constant $\mu = 0.2$.
-        The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
-        The blue curve describes the growth of the bacteria $B$ ($N_B = L_B+G_B$) with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
+        The two species system competing for the common nutrient $R$ with production of the toxin $T_B$ (brown line) calculated using equations~(\ref{eq:model_2sp_toxin}).
+        The growing bacteria $B$ produces toxin $T_B$ (A) constantly $\dot{T}_B^1 = \mathcal{F}^1 = \kappa G_B$ or (B) only during growth process $\dot{T}_B^2 = \mathcal{F}^2 = \kappa R G_B$ 
+        with production rate $\kappa=0.2~\mathrm{s}^{-1}$.
+        The toxin $T_B$ causes the cell death of bacteria $A$ with rate constant $\mu = 0.2~\mathrm{s}^{-1}$.
+        The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}~\mathrm{s}^{-1}$, $\alpha_A=3.0~\mathrm{s}^{-1}$.
+        The blue curve describes the growth of the bacteria $B$ ($N_B = L_B+G_B$) with model parameters $\lambda_B=5\cdot 10^{-2}~\mathrm{s}^{-1}$, $\alpha_B=2.5~\mathrm{s}^{-1}$.
         The gray line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
@@
     \dot{L}_B &= -\lambda_B R L_B\\
-    \dot{G}_A &= \lambda_A R L_A + \alpha_A R G_A\\
-    \dot{G}_B &= \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B\\
-    \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \mu_I I) N_t} R G_B \\
+    \dot{G}_A &= \alpha_A R G_A + \lambda_A R L_A\\
+    \dot{G}_B &= \frac{\alpha_B}{1 + \mu_I I} R G_B + \lambda_B R L_B\\
+    \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \mu_I I) N_t} R G_B\\
     \dot{I} &= \kappa G_A

### 2025-11-06 pool_paper_plotting_code/examples/main_paper.py

@@ -113,7 +113,7 @@ def pool_model_2sp_comp_toxin2(t, x, param, x0, const):
         - lambd_B * R * L_B,
           lambd_B * R * L_B + (alpha_B * R) * G_B,
         - (alpha_A / N_t) * R * G_A - (alpha_B / N_t) * R * G_B,
-          k * G_B * alpha_B * R
+          k * G_B  * R # * alpha_B
         ]
 
 def pool_model_2sp_comp_inhib(t, x, param, x0, const):

### 2025-11-06 paper/pool_model.tex
Add units to captions for pool model

@@
 Then $[\alpha_0]=\unit{V/s}$, $[N_t]=\unit{1/V}$, and $[\alpha]= \unit{1/s}$.
-$N_t$ denotes the maximal size of the total bacterial population due to environmental limitations (Figure \ref{fig:3pool_resource_plots}).
+$N_t$ denotes the maximal size of the total bacterial population due to environmental limitations.
 For homogenous densities it is equivalent to consider the total amount, i.e., multiplying $L$ and $G$ by the volume.
@@
         The time dependence of the bacterial count $N(t) = L(t) + G(t) $ is shown.
-        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\alpha=2.5$, $\mu'=0.5$.
-        The blue curve shows a significant dip in the behavior of the system where the death of the bacteria is present $\mu = 0.5$.
-        Other curves show the case of no bacteria dying $\mu=0$ and represent different lengths of lag phases: short $\lambda=10^{-2}$ (dark green), intermediate $\lambda=10^{-3}$ (light green) and long one $\lambda=10^{-4}$ (yellow).
+        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\alpha=2.5~\mathrm{s}^{-1}$, $\mu'=0.5~\mathrm{s}^{-1}$, $N_t = 10^4$.
+        The blue curve shows a significant dip in the behavior of the system where the death of the bacteria is present $\mu = 0.5~\mathrm{s}^{-1}$.
+        Other curves show the case of no bacteria dying $\mu=0~\mathrm{s}^{-1}$ and represent different lengths of lag phases: short $\lambda=10^{-2}~\mathrm{s}^{-1}$ (dark green), intermediate $\lambda=10^{-3}~\mathrm{s}^{-1}$ (light green) and long one $\lambda=10^{-4}~\mathrm{s}^{-1}$ (yellow).
 Note that the steady state for $N(t)$ and $D(t)$ is for this model $\lim_{t\to\infty} N(t)=0$ and $\lim_{t\to\infty} D(t)=N_t$, because of the depletion of the resource pool.
@@
         (B) The value of the corresponding back-flow rate $\gamma$ calculated according to (\ref{eq:gamma_tempshift}).
-        The curves represent no temperature shift (green), one temperature jump $\Delta T = 10$ at time $t=4$ (orange)
-        and the series of different temperature shifts of value 10, 5 and 15 at the time points 1, 3 and 7, respectively (rose).
-        The parameters of the back-flow rate are $\Gamma=1$ and $\delta=5$.
-        The other model parameters are $\alpha=2.5$, $\mu'=0.5$, $\beta=5$, $\lambda=10^{-2}$, $\mu = 0$ for each curve.
+        The curves represent no temperature shift (green), one temperature jump $\Delta T = 10^\circ C$ at time $t=4~\mathrm{s}$ (orange)
+        and the series of different temperature shifts of value $10^\circ C$, $5^\circ C$ and $15^\circ C$ at the time points $1, 3$ and $7$, respectively (rose).
+        The parameters of the back-flow rate are $\Gamma=1^\circ C / s$ and $\delta=5~\mathrm{s}^{-1}$.
+        The other model parameters are $\alpha=2.5~\mathrm{s}^{-1}$, $\mu'=0.5~\mathrm{s}^{-1}$, $\beta=5~\mathrm{s}^{-1}$, $\lambda=10^{-2}~\mathrm{s}^{-1}$, $\mu = 0~\mathrm{s}^{-1}$ for each curve.
     }
@@
     \caption{
-        The competition of the two bacterial species due to the common limiting resource calculated using equation~(\ref{eq:model_2sp_resource_comp}).
-        The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
-        The blue curve describes the growth of the bacteria $B$ ($N_B = L_B+G_B$) with model parameters $\lambda_B=5\cdot 10^{-2}$, $\alpha_B=2.5$.
+        The competition of the two bacterial species due to the common limiting resource calculated using equations~(\ref{eq:model_2sp_resource_comp}).
+        The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}~\mathrm{s}^{-1}$, $\alpha_A=3.0~\mathrm{s}^{-1}$.
+        The blue curve describes the growth of the bacteria $B$ ($N_B = L_B+G_B$) with model parameters $\lambda_B=5\cdot 10^{-2}~\mathrm{s}^{-1}$, $\alpha_B=2.5~\mathrm{s}^{-1}$.
         The purple line shows the total bacterial count of the system  $N(t)=N_A+N_B$.
@@
     \caption{
-        Spatial configuration of cells corresponding to the pictures shown in figure~\ref{fig:spatial-growth-curve}.
+        Spatial configuration of cells corresponding to the pictures shown in Figure~\ref{fig:spatial-growth-curve}.
     }

### 2025-11-06 paper/pool_model.tex
delete commented old text

@@
 
-% \paragraph{General Introduction to Bacterial Consortia}
-% \begin{itemize}
-%     \item Complex system; many players (cell-types, interactions)
-%     \item processes on different scales (population, cellular, molecular, gene-level)
-%     \item Food Safety, Production
-%     \item Healthcare: Drug development
-%     \item Environmental Sustainabiliy
-% \end{itemize}
-% Motivation: general, flexible approach for bacterial systems that allows easily modify the system depending on processes on different scales.
-% Allows us to describe the processes within populations using different bacterial states as well as interactions between different populations using ODEs.
-
 Bacterial consortia are highly complex systems which consist of multiple distinct cell types.
@@
 tasks such as nutrient cycling, decomposition of organic matter and oxygen production.
-It is of great interest to understand the dynamics of these various processes.
-
-% \paragraph{Experimental/Biological evidence/phenomena}
-% \begin{itemize}
-%     \item Lag Phase, Competition, Cooperation, Limiting Resource
-%     \item Comprehensive reviews about competition/cooperation bacterial mechanisms: \citep{ghoul_ecology_2016, stubbendieck_bacterial_2016, hibbing_bacterial_2010, west_social_2007}
-%     \item Spatial Effects: adhesion, migration, chemotaxis, diffusion
-%     \item Quorum-sensing
-%     \item Biofilm formation \citep{Jin2020}
-% \end{itemize}
-
+It is of great interest to understand the dynamics of these various processes.\\
+%
 In order to achieve a comprehensive description of the target system, it is important to incorporate
@@
 Cell signaling, e.g., in the form of quorum-sensing can lead to collective emergent phenomena such as
-biofilm formation which play an important role for many mature microbial communities~\citep{Jin2020,ng_bacterial_2009}.
-
-% \paragraph{ODE models (predictive microbiology, existing, overview, most popular)}
-% Existing models in Predictive Microbiology (Primary Models)~\citep{perez-rodriguez_predictive_2012}:
-% \begin{itemize}
-%     \item An overview of primary models used in food science can be found in~\citep{van_boekel_kinetic_2008}.
-%     \item Baranyi-Roberts Model~\citep{baranyi_modeling_1993, baranyi_dynamic_1994}
-%     \item modified Gompertz~\citep{Gompertz1825-wi,Gibson1987}
-%     \item Buchanan et al.: three-phase linear model for lag phase, exponential growth phase, and stationary phase~\citep{buchanan_when_1997}
-%     \item McKellar two-compartment model with bacteria in growing and non-growing state~\citep{mckellar_heterogeneous_1997}
-%     \item Inactivation models
-% \end{itemize}
-% %
-% Secondary models:
-% \begin{itemize}
-%     \item Polynomial models 
-%     \item Square Root-Type Models, e.g., Ratkowski for temperature~\citep{ratkowsky_relationship_1982}
-%     \item The Cardinal Parameter Model~\citep{zwietering_decision_1992}
-% \end{itemize}
-
+biofilm formation which play an important role for many mature microbial communities~\citep{Jin2020,ng_bacterial_2009}.\\
+%
 The first step in the construction of any mathematical model is to determine the required level of
@@
 However, due to their coarse-grained nature, they lack the ability to describe heterogeneity and
-other spatial effects.
-
-% \paragraph{Models for spatial effects; PDEs, \acp{abm}, others}
-% \begin{itemize}
-%     \item BSim \citep{Gorochowski2012,Dang2020}
-%     \item Bacterial Branching patterns \citep{Kawasaki1997,Matsushita1998}
-% \end{itemize}
-
-% To incorporate spatial heterogeneity or other effects other modeling approaches have been devised.
-% \acp{pde} follow a similar pattern as \acp{ode} but also distribute the described variables through
-% space.
+other spatial effects.\\
+%
 When dealing with spatial effects such as heterogeneity, we are required to use more extensive
@@
 The latter has been used to adapt \ac{pde}-based models into an \ac{abm} formulation which describe
-bacterial branching patterns~\citep{Kawasaki1997,Matsushita1998}.
-
-% \paragraph{What is the problem/what is the question? How do we address this?}
-
+bacterial branching patterns~\citep{Kawasaki1997,Matsushita1998}.\\
+%
 In this work, we introduce a new \ac{ode}-based modeling framework which describes different
@@
 Furthermore, we assess how spatial heterogeneity~\citep{mckellar_heterogeneous_1997} can impact the shape of the growth curves.
-
+%
+%
 % ================================================================================================
@@
 Moreover, by including the flow of living bacteria to the death pool, we can model a greater variety of effects by allowing a decrease in the bacterial count.
-% TODO change the name "death pool"?
-%
 %
@@
 %
-% The dynamics of the three pool model can be captured by the coupled system of ordinary differential
 The dynamics of the three pool model is captured by equations~\ref{eq:ode_3pool_resource2} which can be used to derive the two pool model simply by omitting $D$.
@@
 The production of toxin only depends on the presence of bacteria $G_B$ and is independent of their growth dynamics.
-% That means that as long as there are bacteria in the growth phase $G_B>0$, toxin is produced, no matter whether the bacterial count $N_B$ is still growing.
 On the other hand, model $\mathcal{F}^2$ assumes that toxin is only produced during bacterial growth and division, hence depends on the growth dynamics of $B$.
@@
 %
-% - \citep{emonet_agentcell_2005, gorochowski_bsim_2012, li_nufeb_2019, kreft_bacsim_1998}: examples of another toolboxes for bacterial communities modeling (ABM).\\
-% More about the modeling of the motile bacteria can be seen in works:~\citep{decoene_microscopic_2011, rosser_modelling_2014, sokolov_physical_2012, li_amplified_2008}.
-%
-%
-% In general, bacterial communities in 2D can experience collective motion called swarming~\citep{wu_collective_2015}.
-% Some works modelled 2D motile bacteria:
-% \begin{itemize}
-%     \item \citep{decoene_microscopic_2011}: 2D chemotactic self-propelled bacterial motion of the flagellated bacteria (e.g., E. coli, B. subtilis).
-%     Chemotaxis here was driven by the oxygen concentrations.
-%     \item \citep{rosser_modelling_2014}: modeling of the bacterial reorintation, rotation.
-%     \item \citep{sokolov_physical_2012, li_amplified_2008}: modeling of the bacteria using hydrodynamic interactions and Brownian motion.
-%     Fluctuations in velocity and rotations.
-%     (\citep{li_amplified_2008}: Brownian motion gives right motion observed using microscopy).
-%     \item \citep{farrell_mechanically_2013}: 2D modeling using Newtonian dynamics and only mechanical interactions + nutrient diffusion.
-%     Branching parameter defines diff limits where diffusion dominant and limits growth, and where it's irrelevant.
-%     Mechanical pressure (due to bacteria-bacteria interactions) affects growth, can limit it.(!!!)
-% \end{itemize}
-%
-% For simulation results:
-% \begin{itemize}
-%     \item \citep{farrell_mechanically_2013}: Showed the influence of mechanic interactions between bacteria in 2D system with bacteria consuming the nutrient.
-%     The growth of bacterial colonies where non-motile microorganisms replicate and push each other away as they grow.
-%     Found a transition between two different growth regimes, controlled by the balance between growth and uptake of nutrient:\\
-%     Exponential growth where diffusion of the nutrient does not play any role. Sublinear growth where growth is limited by nutrient diffusion.
-%     (Structurally in 2D it's regimes: smooth propagating of bacterial front vs. branching.)
-%     \item \citep{nagarajan_agent-based_2022-1}: review on the ABM for bacterial population.
-%     The ABM represent a microbial colony as a group of discrete agents that are governed by a set of rules, where each agent can be an individual cell or a collection of cells.
-%     Can use citation when talk about general processes we are accounting as growth, uptake of nutrient, production of inhibitor, e.g.,:
-%     Physical interactions between agents are calculated using excluded volume interactions.
-%     Chemical interactions as releasing inhibitors are described using concentrations and solved using PDEs.
-% \end{itemize}
-%
 \begin{align}
@@
 This ability, to precisely define and control the exact behavior of our cell-agents, enables the following parts.
-
+%
 \subsubsection{Cellular reactions}
@@
 \section{Discussion}
-
-% How well does the pool model do to encapsulate all the different growth models
-% \begin{itemize}
-%     \item All 6 models can be represented by pool model
-%     \item Baranyi-Roberts results can be reproduced, but not identical; requires some work
-%     \item what are some general limitations within this framework? how can they be circumvented?
-%     \item why is the pool model "better" than the other \ac{ode} models? (why did we do this?);
-%         generalization of many models; better explainability;
-%     \item link to supplement; we can generalize this and prove theorems for the abstract,
-%         generalized model; this allows us to discuss many cases at once
-% \end{itemize}
-%
-% How can we cover more effects of bacteria?
-% \begin{itemize}
-%     \item More states for metabolic processes
-%     \item More states for more species
-%     \item Other cooperiation/inhibition effects
-%     \item external influences
-% \end{itemize}
-%
-% How justified was the construction of the \ac{abm}?
-% \begin{itemize}
-%     \item Discuss simplified physical interaction model
-%     \item Heterogeneity plays a role; as we have shown
-%     \item why did we not simply use a PDE?; because we have to think in terms of cellular processes
-%         and this ensures that our model is actually realistic
-% \end{itemize}
-%
-% What are limitations of our approach?
-% \begin{itemize}
-%     \item spatial stuff; heterogeneity (because \ac{ode})
-%     \item larger models require more parameters, harder to estimate
-%     \item only simple models have analytical solutions
-% \end{itemize}
-
+%
 The presented pool model was used to describe the dynamics of one- and two-species systems containing the main growth stages and interactions between bacteria in the form of cooperation and competition.
@@
 However, deviations to these assumptions can lead to non-negligible effects, influencing the growth of the colony.
-
+%
 % ------------------------------------------------------------------------------------------------
 \section{Conclusion}
-% \begin{itemize}
-%     \item New theoretical framework
-%     \item consistent with \ac{ode} models (Baranyi-Roberts)
-%     \item Better explainability
-%     \item Include effects such as inhibition, toxins in understandable manner
-%     \item showed spatial limitations of \ac{ode} model
-%     \item \ac{abm} model allowed modeling of cellular processes; ensure that pool model is derivable
-%         from the individual behvaiour of cellular processes
-%     \item outlook: generalized mathematical formulation; extend inhibition model (because it is a
-%         dimensional reduction of existing models)
-%     \item Provide code in a pypi package as well (TODO)
-% \end{itemize}
-
+%
 We introduced the pool model as a framework to unify existing approaches and allow researchers to follow a mechanistic approach during model development.
@@
 %
-%The three pool system described with equations~\ref{eq:ode_3pools} will be written as
-%\begin{equation}
-%    \dot \mbx  = A \mbx, \hspace{3mm} A = \begin{pmatrix}
-%        -(\mu + \lambda) & 0             & 0      \\
-%        \lambda          & \alpha - \mu' & 0      \\
-%        \mu              & \mu'          & -\beta 
-%    \end{pmatrix}.
-%\label{eq:x_repres_3pools}
-%\end{equation}
 %

### 2025-11-07 paper/pool_model.tex
add citation to book our chapter
add Christian's e-mail
Use pdf images for spatial curves

@@
 
-\cortext[cor1]{Corresponding Author}
 \fntext[fn1]{jonas.pleyer.org}
+\cortext[cor1]{Corresponding Author\\ \textit{E-mail address:} christian.fleck@fdm.uni-freiburg.de}
 \affiliation[1]{
@@
 The field of predictive microbiology has developed a variety of ODE-based models to capture
-bacterial growth dynamics~\citep{perez-rodriguez_predictive_2012}.
+bacterial growth dynamics~\citep{perez-rodriguez_predictive_2012, gaindrik_experimental_2025}.
 Various \acp{ode} \textit{primary models} describe the bacterial growth curve under constant
@@
     \begin{tikzonimage}[width=\columnwidth]
-        {Figures-abm-homogenous-abm_ode_comparison.png}
+        {Figures-abm-homogenous-abm_ode_comparison.pdf}
         \node at (0, 1)[anchor=north west]{\textbf{A}};
@@
     \begin{tikzonimage}[width=\columnwidth]
-        {Figures-abm-homogenous-low-diffusion-abm_ode_comparison.png}
+        {Figures-abm-homogenous-low-diffusion-abm_ode_comparison.pdf}
         \node at (0, 1)[anchor=north west]{\textbf{B}};
@@
     \begin{tikzonimage}[width=\columnwidth]
-        {Figures-abm-inhomogenous-abm_ode_comparison.png}
+        {Figures-abm-inhomogenous-abm_ode_comparison.pdf}
         \node at (0, 1)[anchor=north west]{\textbf{C}};
@@
     \begin{tikzonimage}[width=\columnwidth]
-        {Figures-abm-random-abm_ode_comparison.png}
+        {Figures-abm-random-abm_ode_comparison.pdf}
         \node at (0, 1)[anchor=north west]{\textbf{D}};

### 2025-11-28 paper/pool_model.tex
adapt figure description to new parameters

@@
         The time dependence of the bacterial count $N(t) = L(t) + G(t) $ is shown.
-        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\alpha=2.5~\mathrm{s}^{-1}$, $\mu'=0.5~\mathrm{s}^{-1}$, $N_t = 10^4$.
-        The blue curve shows a significant dip in the behavior of the system where the death of the bacteria is present $\mu = 0.5~\mathrm{s}^{-1}$.
+        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\alpha=2.5~\mathrm{s}^{-1}$, $\mu'=0.1~\mathrm{s}^{-1}$, $N_t = 10^4$.
+        The blue curve includes cell-death of the bacteria in the lag phase $\mu = 0.15~\mathrm{s}^{-1}$ and thus exhibits a small dip before its exponential growth.
         Other curves show the case of no bacteria dying $\mu=0~\mathrm{s}^{-1}$ and represent different lengths of lag phases: short $\lambda=10^{-2}~\mathrm{s}^{-1}$ (dark green), intermediate $\lambda=10^{-3}~\mathrm{s}^{-1}$ (light green) and long one $\lambda=10^{-4}~\mathrm{s}^{-1}$ (yellow).

### 2025-11-28 paper/pool_model.tex
replace \mu with \omega in all equations

--- 2025-11-07 paper/pool_model.tex
@@
     L &\stackrel{\lambda}{\longrightarrow} G\\
-    L &\stackrel{\mu}{\longrightarrow} D\\
+    L &\stackrel{\omega}{\longrightarrow} D\\
     G &\stackrel{\alpha}{\longrightarrow} 2G\\
-    G &\stackrel{\mu'}{\longrightarrow} D
+    G &\stackrel{\omega'}{\longrightarrow} D
 \end{align}
@@
 \begin{align}\begin{split}
-    \dot{L} &= -(\mu + \lambda) L\\
-    \dot{G} &= \lambda L + \alpha G - \mu' G\\
-    \dot{D} &= \mu  L + \mu' G 
+    \dot{L} &= -(\omega + \lambda) L\\
+    \dot{G} &= \lambda L + \alpha G - \omega' G\\
+    \dot{D} &= \omega  L + \omega' G
 \end{split}\end{align}
@@
     \begin{split}
-        \dot{L} &= -(\mu + \lambda) L\\
-        \dot{G} &= \lambda L + \alpha \tilde{R} G-\mu' G\\
-        \dot{D} &= \mu  L + \mu' G\\
+        \dot{L} &= -(\omega + \lambda) L\\
+        \dot{G} &= \lambda L + \alpha \tilde{R} G-\omega' G\\
+        \dot{D} &= \omega  L + \omega' G\\
         \dot{\tilde{R}} &= - \frac{\alpha}{N_t} \tilde{R} G
@@
     \begin{split}
-        \dot{L} &= -(\mu + \lambda) L\\
-        \dot{G} &= \lambda L + \alpha G\left(1-\frac{L+G+D}{N_t}\right)-\mu' G\\
-        \dot{D} &= \mu  L + \mu' G.
+        \dot{L} &= -(\omega + \lambda) L\\
+        \dot{G} &= \lambda L + \alpha G\left(1-\frac{L+G+D}{N_t}\right)-\omega' G\\
+        \dot{D} &= \omega  L + \omega' G.
     \end{split}
@@
         The time dependence of the bacterial count $N(t) = L(t) + G(t) $ is shown.
-        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\alpha=2.5~\mathrm{s}^{-1}$, $\mu'=0.5~\mathrm{s}^{-1}$, $N_t = 10^4$.
-        The blue curve shows a significant dip in the behavior of the system where the death of the bacteria is present $\mu = 0.5~\mathrm{s}^{-1}$.
-        Other curves show the case of no bacteria dying $\mu=0~\mathrm{s}^{-1}$ and represent different lengths of lag phases: short $\lambda=10^{-2}~\mathrm{s}^{-1}$ (dark green), intermediate $\lambda=10^{-3}~\mathrm{s}^{-1}$ (light green) and long one $\lambda=10^{-4}~\mathrm{s}^{-1}$ (yellow).
+        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\alpha=2.5~\mathrm{s}^{-1}$, $\omega'=0.1~\mathrm{s}^{-1}$, $N_t = 10^4$.
+        The blue curve includes cell-death of the bacteria in the lag phase $\omega = 0.15~\mathrm{s}^{-1}$ and thus exhibits a small dip before its exponential growth.
+        Other curves show the case of no bacteria dying $\omega=0~\mathrm{s}^{-1}$ and represent different lengths of lag phases: short $\lambda=10^{-2}~\mathrm{s}^{-1}$ (dark green), intermediate $\lambda=10^{-3}~\mathrm{s}^{-1}$ (light green) and long one $\lambda=10^{-4}~\mathrm{s}^{-1}$ (yellow).
 Note that the steady state for $N(t)$ and $D(t)$ is for this model $\lim_{t\to\infty} N(t)=0$ and $\lim_{t\to\infty} D(t)=N_t$, because of the depletion of the resource pool.
@@
 The dynamics of the three pool model is captured by equations~\ref{eq:ode_3pool_resource2} which can be used to derive the two pool model simply by omitting $D$.
-Moreover, for $\mu'=0$ the equations for $L$ and $D$ can be readily integrated resulting in:
+Moreover, for $\omega'=0$ the equations for $L$ and $D$ can be readily integrated resulting in:
 \begin{align}
     \begin{split}
-        L(t) &= N_0 e^{-(\mu+\lambda)t}\\
-        D(t) &= N_0 \mu \frac{1-e^{-(\mu+\lambda)t}}{\mu+\lambda
+        L(t) &= N_0 e^{-(\omega+\lambda)t}\\
+        D(t) &= N_0 \omega \frac{1-e^{-(\omega+\lambda)t}}{\omega+\lambda
        }
@@
 \end{align}
-For constant parameters $\mu$ and $\lambda$ the integrals for $L$ and $D$ can be inserted into the differential equation for $G$.
+For constant parameters $\omega$ and $\lambda$ the integrals for $L$ and $D$ can be inserted into the differential equation for $G$.
 However, in the case of time dependent parameters, e.g., for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$.
@@
 A notable difference is that the growth rate $\alpha$ is time independent.
-For the functions $L$ and $D$ it holds: $\lim_{t\to\infty} L(t) = 0$ and  $\lim_{t\to\infty} D(t) =  N_0\frac{\mu}{\mu+\lambda} $.
-It follows immediately that $\lim_{t\to\infty} G(t) = N_t-N_0\frac{\mu}{\mu+\lambda} $.
+For the functions $L$ and $D$ it holds: $\lim_{t\to\infty} L(t) = 0$ and  $\lim_{t\to\infty} D(t) =  N_0\frac{\omega}{\omega+\lambda} $.
+It follows immediately that $\lim_{t\to\infty} G(t) = N_t-N_0\frac{\omega}{\omega+\lambda} $.
 The lag time $t_L$ can be approximated by the time at which half of the initial population is in the growth phase, i.e., $N_0=2G(t_L)$.
@@
     N_0&= 2\int_0^{t_L} \lambda L(t)e^{\alpha(t_L-t)}dt\\
-    1  &= 2\lambda e^{\alpha t_L}\frac{1-e^{-(\mu+\lambda+\alpha)t_L}}{\mu+\lambda+\alpha}.
-\end{align}
-For $\alpha > \mu+\lambda$  we can approximate $t_L$ by:
-\begin{align}
-    t_L &\approx -\frac{1}{\alpha}\ln\left(\frac{2\lambda}{\mu+\lambda+\alpha}\right).
+    1  &= 2\lambda e^{\alpha t_L}\frac{1-e^{-(\omega+\lambda+\alpha)t_L}}{\omega+\lambda+\alpha}.
+\end{align}
+For $\alpha > \omega+\lambda$  we can approximate $t_L$ by:
+\begin{align}
+    t_L &\approx -\frac{1}{\alpha}\ln\left(\frac{2\lambda}{\omega+\lambda+\alpha}\right).
 \end{align}
@@
     L &\stackrel{\lambda}{\longrightarrow} G\\
-    L &\stackrel{\mu}{\longrightarrow} D\\
+    L &\stackrel{\omega}{\longrightarrow} D\\
     G &\stackrel{\alpha}{\longrightarrow} 2G\\
@@
     \begin{split}
-        \dot{L} &=\gamma(t) G - (\mu + \lambda) L\\
+        \dot{L} &=\gamma(t) G - (\omega + \lambda) L\\
         \dot{G} &= \lambda L -\gamma(t) G + \alpha G\left(1-\frac{L+G+D}{N_t}\right)\\
-        \dot{D} &= \mu  L
+        \dot{D} &= \omega  L
     \end{split}
@@
         The parameters of the back-flow rate are $\Gamma=1^\circ C / s$ and $\delta=5~\mathrm{s}^{-1}$.
-        The other model parameters are $\alpha=2.5~\mathrm{s}^{-1}$, $\mu'=0.5~\mathrm{s}^{-1}$, $\beta=5~\mathrm{s}^{-1}$, $\lambda=10^{-2}~\mathrm{s}^{-1}$, $\mu = 0~\mathrm{s}^{-1}$ for each curve.
+        The other model parameters are $\alpha=2.5~\mathrm{s}^{-1}$, $\omega'=0.5~\mathrm{s}^{-1}$, $\beta=5~\mathrm{s}^{-1}$, $\lambda=10^{-2}~\mathrm{s}^{-1}$, $\omega = 0~\mathrm{s}^{-1}$ for each curve.
     }
@@
 Consider the system of two species competing for resource $R$ with the production of the toxin $T$ by species $B$.
-We assume that the toxin causes the death of the species $A$ with rate $\mu$.
+We assume that the toxin causes the death of the species $A$ with rate $\omega$.
 Then the full \ac{ode} system is
@@
         \dot{L}_B &= - \lambda_B R L_B \\
-        \dot{G}_A &= \lambda_A R L_A + \alpha_A R G_A - \frac{\mu}{\kappa N_t} T G_A\\
+        \dot{G}_A &= \lambda_A R L_A + \alpha_A R G_A - \frac{\omega}{\kappa N_t} T G_A\\
         \dot{G}_B &= \lambda_B R L_B + \alpha_B R G_B\\ 
@@
         with production rate $\kappa=0.2~\mathrm{s}^{-1}$.
-        The toxin $T_B$ causes the cell death of bacteria $A$ with rate constant $\mu = 0.2~\mathrm{s}^{-1}$.
+        The toxin $T_B$ causes the cell death of bacteria $A$ with rate constant $\omega = 0.2~\mathrm{s}^{-1}$.
         The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}~\mathrm{s}^{-1}$, $\alpha_A=3.0~\mathrm{s}^{-1}$.
@@
         \dot{L}_B &= -\lambda_B R L_B\\
-        \dot{G}_B &= \lambda_B R L_B + \frac{\alpha_B}{1 + \mu_I I} R G_B.
+        \dot{G}_B &= \lambda_B R L_B + \frac{\alpha_B}{1 + \omega_I I} R G_B.
     \end{split}
@@
     \begin{split}
-        \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \mu_I I) N_t} R G_B \\
+        \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \omega_I I) N_t} R G_B \\
         \dot{I} &= \kappa G_A.
@@
     \dot{G}_A &= \alpha_A R G_A + \lambda_A R L_A\\
-    \dot{G}_B &= \frac{\alpha_B}{1 + \mu_I I} R G_B + \lambda_B R L_B\\
-    \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \mu_I I) N_t} R G_B\\
+    \dot{G}_B &= \frac{\alpha_B}{1 + \omega_I I} R G_B + \lambda_B R L_B\\
+    \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \omega_I I) N_t} R G_B\\
     \dot{I} &= \kappa G_A
@@
         N_t &= [R_0]V_D\sigma\\
-        \mu_I &= \frac{\nu_I}{V_D}\\
+        \omega_I &= \frac{\nu_I}{V_D}\\
         \alpha &= u[R_0]\sigma\\
@@
     \dot \mbx  &= \mbF(\mbx, \mbm, \mbp, \mbu) = \begin{pmatrix}
-        -(\mu + \lambda) & 0               & 0      \\
-        \lambda          & \alpha R - \mu' & 0      \\
-        \mu              & \mu'            & -\beta 
+        -(\omega + \lambda) & 0               & 0      \\
+        \lambda          & \alpha R - \omega' & 0      \\
+        \omega              & \omega'            & -\beta
     \end{pmatrix} \mbx\\
@@
     \dot \mbx = \begin{pmatrix}
-        -(\mu + \lambda) & 0       & 0 \\
+        -(\omega + \lambda) & 0       & 0 \\
          \lambda         & \alpha  & 0 \\
-         \mu &  \mu'& - \beta 
+         \omega &  \omega'& - \beta
     \end{pmatrix} 
@@
         & Production $I$            & $b$           & $0.1$~\unit{\per\minute}\\
-        Inhibition                  & $\mu_I$       & $1.0\cdot 10^{-7}$~\unit{\per\micro\metre\cubed}
+        Inhibition                  & $\omega_I$    & $1.0\cdot 10^{-7}$~\unit{\per\micro\metre\cubed}
         &                           & $\nu_I$       & $0.1$~\unit{\micro\metre\cubed\per\mol}\\

### 2025-11-28 paper/pool_model.tex
rename variable for ODE in ABM section
change alpha to mu as growth rate

@@
     L &\stackrel{\omega}{\longrightarrow} D\\
-    G &\stackrel{\alpha}{\longrightarrow} 2G\\
+    G &\stackrel{\mu}{\longrightarrow} 2G\\
     G &\stackrel{\omega'}{\longrightarrow} D
@@
     \dot{L} &= -(\omega + \lambda) L\\
-    \dot{G} &= \lambda L + \alpha G - \omega' G\\
+    \dot{G} &= \lambda L + \mu G - \omega' G\\
     \dot{D} &= \omega  L + \omega' G
@@
 \begin{equation}
-    G + R  \stackrel{\alpha_0}{\longrightarrow} 2G
+    G + R  \stackrel{\mu_0}{\longrightarrow} 2G
 \end{equation}
@@
 Then $R=N_R/V$ and $[R]=\unit{1/V}$, where $N_R$ is the number of resource molecules (or any appropriate unit, e.g., \unit{mol}).
-Then $[\alpha_0]=\unit{V/s}$, $[N_t]=\unit{1/V}$, and $[\alpha]= \unit{1/s}$.
+Then $[\mu_0]=\unit{V/s}$, $[N_t]=\unit{1/V}$, and $[\mu]= \unit{1/s}$.
 $N_t$ denotes the maximal size of the total bacterial population due to environmental limitations.
@@
 In this case $[L]=[G]=[N_t]=1$.
-For simplicity, we make a transition to the dimensionless constant $\alpha:=\alpha_0 N_t$ and dimensionless resource pool $\tilde{R}(t) = R/ N_t$ ($0 \leqslant \tilde{R} \leqslant 1$).
+For simplicity, we make a transition to the dimensionless constant $\mu:=\mu_0 N_t$ and dimensionless resource pool $\tilde{R}(t) = R/ N_t$ ($0 \leqslant \tilde{R} \leqslant 1$).
 Here the full capacity of the resource pool $N_t$ was used as a scaling constant.
-In this case the growth term can be simply written as $\alpha G \tilde{R}$ and the model is fully formulated in the following way:
+In this case the growth term can be simply written as $\mu G \tilde{R}$ and the model is fully formulated in the following way:
 \begin{align}
@@
         \dot{L} &= -(\omega + \lambda) L\\
-        \dot{G} &= \lambda L + \alpha \tilde{R} G-\omega' G\\
+        \dot{G} &= \lambda L + \mu \tilde{R} G-\omega' G\\
         \dot{D} &= \omega  L + \omega' G\\
-        \dot{\tilde{R}} &= - \frac{\alpha}{N_t} \tilde{R} G
+        \dot{\tilde{R}} &= - \frac{\mu}{N_t} \tilde{R} G
     \end{split}
@@
         \dot{L} &= -(\omega + \lambda) L\\
-        \dot{G} &= \lambda L + \alpha G\left(1-\frac{L+G+D}{N_t}\right)-\omega' G\\
+        \dot{G} &= \lambda L + \mu G\left(1-\frac{L+G+D}{N_t}\right)-\omega' G\\
         \dot{D} &= \omega  L + \omega' G.
@@
         The time dependence of the bacterial count $N(t) = L(t) + G(t) $ is shown.
-        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\alpha=2.5~\mathrm{s}^{-1}$, $\omega'=0.1~\mathrm{s}^{-1}$, $N_t = 10^4$.
+        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\mu=2.5~\mathrm{s}^{-1}$, $\omega'=0.1~\mathrm{s}^{-1}$, $N_t = 10^4$.
         The blue curve includes cell-death of the bacteria in the lag phase $\omega = 0.15~\mathrm{s}^{-1}$ and thus exhibits a small dip before its exponential growth.
@@
     \begin{split}
-        \dot{G} &= \lambda L + \alpha G\left(1-\frac{\mathcal{N}}{N_t}\right)\\
-        \dot{\mathcal{N}} &= \alpha G\left(1-\frac{\mathcal{N}}{N_t}\right)
+        \dot{G} &= \lambda L + \mu G\left(1-\frac{\mathcal{N}}{N_t}\right)\\
+        \dot{\mathcal{N}} &= \mu G\left(1-\frac{\mathcal{N}}{N_t}\right)
     \end{split}
@@
 These equations have a clear interpretation, are derived from first principles, and are not identical to the popular Barayni-Roberts and modified-Gompertz models.
-A notable difference is that the growth rate $\alpha$ is time independent.
+A notable difference is that the growth rate $\mu$ is time independent.
 For the functions $L$ and $D$ it holds: $\lim_{t\to\infty} L(t) = 0$ and  $\lim_{t\to\infty} D(t) =  N_0\frac{\omega}{\omega+\lambda} $.
@@
 \begin{align}
-    N_0&= 2\int_0^{t_L} \lambda L(t)e^{\alpha(t_L-t)}dt\\
-    1  &= 2\lambda e^{\alpha t_L}\frac{1-e^{-(\omega+\lambda+\alpha)t_L}}{\omega+\lambda+\alpha}.
-\end{align}
-For $\alpha > \omega+\lambda$  we can approximate $t_L$ by:
-\begin{align}
-    t_L &\approx -\frac{1}{\alpha}\ln\left(\frac{2\lambda}{\omega+\lambda+\alpha}\right).
+    N_0&= 2\int_0^{t_L} \lambda L(t)e^{\mu(t_L-t)}dt\\
+    1  &= 2\lambda e^{\mu t_L}\frac{1-e^{-(\omega+\lambda+\mu)t_L}}{\omega+\lambda+\mu}.
+\end{align}
+For $\mu > \omega+\lambda$  we can approximate $t_L$ by:
+\begin{align}
+    t_L &\approx -\frac{1}{\mu}\ln\left(\frac{2\lambda}{\omega+\lambda+\mu}\right).
 \end{align}
@@
     L &\stackrel{\omega}{\longrightarrow} D\\
-    G &\stackrel{\alpha}{\longrightarrow} 2G\\
+    G &\stackrel{\mu}{\longrightarrow} 2G\\
     G &\stackrel{\gamma(t)}{\longrightarrow} L
@@
         \dot{L} &=\gamma(t) G - (\omega + \lambda) L\\
-        \dot{G} &= \lambda L -\gamma(t) G + \alpha G\left(1-\frac{L+G+D}{N_t}\right)\\
+        \dot{G} &= \lambda L -\gamma(t) G + \mu G\left(1-\frac{L+G+D}{N_t}\right)\\
         \dot{D} &= \omega  L
@@
         The parameters of the back-flow rate are $\Gamma=1^\circ C / s$ and $\delta=5~\mathrm{s}^{-1}$.
-        The other model parameters are $\alpha=2.5~\mathrm{s}^{-1}$, $\omega'=0.5~\mathrm{s}^{-1}$, $\beta=5~\mathrm{s}^{-1}$, $\lambda=10^{-2}~\mathrm{s}^{-1}$, $\omega = 0~\mathrm{s}^{-1}$ for each curve.
+        The other model parameters are $\mu=2.5~\mathrm{s}^{-1}$, $\omega'=0.5~\mathrm{s}^{-1}$, $\beta=5~\mathrm{s}^{-1}$, $\lambda=10^{-2}~\mathrm{s}^{-1}$, $\omega = 0~\mathrm{s}^{-1}$ for each curve.
     }
@@
         \dot{L}_B &= - \lambda_B R L_B \\
-        \dot{G}_A &= \lambda_A R L_A +\alpha_A R G_A\\
-        \dot{G}_B &= \lambda_B R L_B +\alpha_B R G_B\\
-        \dot{R} &=-\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B
+        \dot{G}_A &= \lambda_A R L_A +\mu_A R G_A\\
+        \dot{G}_B &= \lambda_B R L_B +\mu_B R G_B\\
+        \dot{R} &=-\frac{\mu_A}{N_t} R G_A-\frac{\mu_B}{N_t} R G_B
     \end{split}
@@
     \dot{L}_B &= -\lambda_B L_B + \frac{\lambda_B}{N_t}L_B N\\
-    \dot{G}_A &=  \lambda_A  L_A + \alpha_A G_A - \frac{\lambda_A L_A + \alpha_A G_A}{N_t} N\\
-    \dot{G}_B &=  \lambda_B L_B - \alpha_B G_B \frac{\lambda_B L_B + \alpha_B G_B}{N_t}N
+    \dot{G}_A &=  \lambda_A  L_A + \mu_A G_A - \frac{\lambda_A L_A + \mu_A G_A}{N_t} N\\
+    \dot{G}_B &=  \lambda_B L_B - \mu_B G_B \frac{\lambda_B L_B + \mu_B G_B}{N_t}N
 \end{align}
@@
         The competition of the two bacterial species due to the common limiting resource calculated using equations~(\ref{eq:model_2sp_resource_comp}).
-        The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}~\mathrm{s}^{-1}$, $\alpha_A=3.0~\mathrm{s}^{-1}$.
-        The blue curve describes the growth of the bacteria $B$ ($N_B = L_B+G_B$) with model parameters $\lambda_B=5\cdot 10^{-2}~\mathrm{s}^{-1}$, $\alpha_B=2.5~\mathrm{s}^{-1}$.
+        The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}~\mathrm{s}^{-1}$, $\mu_A=3.0~\mathrm{s}^{-1}$.
+        The blue curve describes the growth of the bacteria $B$ ($N_B = L_B+G_B$) with model parameters $\lambda_B=5\cdot 10^{-2}~\mathrm{s}^{-1}$, $\mu_B=2.5~\mathrm{s}^{-1}$.
         The purple line shows the total bacterial count of the system  $N(t)=N_A+N_B$.
@@
     \begin{split}
-        \dot{G}_A &= \alpha_A G_A\left(1 - \frac{G_A}{N_t}\right) - \frac{\alpha_A}{N_t}G_AG_B\\
-        \dot{G}_B &= \alpha_B G_B\left(1-\frac{G_B}{N_t}\right) -\frac{\alpha_B}{N_t}G_AG_B. 
+        \dot{G}_A &= \mu_A G_A\left(1 - \frac{G_A}{N_t}\right) - \frac{\mu_A}{N_t}G_AG_B\\
+        \dot{G}_B &= \mu_B G_B\left(1-\frac{G_B}{N_t}\right) -\frac{\mu_B}{N_t}G_AG_B. 
     \label{eq:LV_simple}
@@
         \dot{L}_B &= - \lambda_B R L_B \\
-        \dot{G}_A &= \lambda_A R L_A + \alpha_A R G_A - \frac{\omega}{\kappa N_t} T G_A\\
-        \dot{G}_B &= \lambda_B R L_B + \alpha_B R G_B\\ 
-        \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B\\
+        \dot{G}_A &= \lambda_A R L_A + \mu_A R G_A - \frac{\omega}{\kappa N_t} T G_A\\
+        \dot{G}_B &= \lambda_B R L_B + \mu_B R G_B\\
+        \dot{R} &= -\frac{\mu_A}{N_t} R G_A-\frac{\mu_B}{N_t} R G_B\\
         \dot{T}_{B}^{1,2} &= \mathcal{F}^{1,2} (R, G_B) \\
@@
         The toxin $T_B$ causes the cell death of bacteria $A$ with rate constant $\omega = 0.2~\mathrm{s}^{-1}$.
-        The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}~\mathrm{s}^{-1}$, $\alpha_A=3.0~\mathrm{s}^{-1}$.
-        The blue curve describes the growth of the bacteria $B$ ($N_B = L_B+G_B$) with model parameters $\lambda_B=5\cdot 10^{-2}~\mathrm{s}^{-1}$, $\alpha_B=2.5~\mathrm{s}^{-1}$.
+        The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}~\mathrm{s}^{-1}$, $\mu_A=3.0~\mathrm{s}^{-1}$.
+        The blue curve describes the growth of the bacteria $B$ ($N_B = L_B+G_B$) with model parameters $\lambda_B=5\cdot 10^{-2}~\mathrm{s}^{-1}$, $\mu_B=2.5~\mathrm{s}^{-1}$.
         The gray line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
@@
 \begin{equation}
-    \mathcal{B}(G) = \frac{\alpha}{1+KI(G)}
+    \mathcal{B}(G) = \frac{\mu}{1+KI(G)}
 \end{equation}
@@
 \begin{align}\begin{split}
-    \dot{G}_A &= \frac{\alpha_A G_A}{1+K_BG_B}\left(1 - \frac{G_A+G_B}{N_t}\right)\\
-    \dot{G}_B &= \frac{\alpha_B G_B}{1+K_AG_A}\left(1-\frac{G_A+G_B}{N_t}\right).
+    \dot{G}_A &= \frac{\mu_A G_A}{1+K_BG_B}\left(1 - \frac{G_A+G_B}{N_t}\right)\\
+    \dot{G}_B &= \frac{\mu_B G_B}{1+K_AG_A}\left(1-\frac{G_A+G_B}{N_t}\right).
 \end{split}\end{align}
@@
 \begin{align}\begin{split}
-    \dot{G}_A &=\tilde{\alpha}_A\frac{G_A}{G_B}\left(1 - G_A-G_B\right)\\
-    \dot{G}_B &= \tilde{\alpha}_B\frac{G_B}{G_A}\left(1 - G_A-G_B\right).
+    \dot{G}_A &=\tilde{\mu}_A\frac{G_A}{G_B}\left(1 - G_A-G_B\right)\\
+    \dot{G}_B &= \tilde{\mu}_B\frac{G_B}{G_A}\left(1 - G_A-G_B\right).
 \end{split}\end{align}
-We introduced the rescaled growth rates $\tilde{\alpha}_A=\alpha_A/(N_tK_B)$ and $\tilde{\alpha}_B=\alpha_B/(N_tK_A)$.
-Note that $[\tilde{\alpha}]=\mathrm{s}^{-1}$.
-We rescale the time with $\tilde{\alpha}_A$ and define the dimensionless parameter
-\begin{equation}
-    \psi = \tilde{\alpha}_B/\tilde{\alpha}_A = \frac{K_B\alpha_B}{K_A\alpha_A}
+We introduced the rescaled growth rates $\tilde{\mu}_A=\mu_A/(N_tK_B)$ and $\tilde{\mu}_B=\mu_B/(N_tK_A)$.
+Note that $[\tilde{\mu}]=\mathrm{s}^{-1}$.
+We rescale the time with $\tilde{\mu}_A$ and define the dimensionless parameter
+\begin{equation}
+    \psi = \tilde{\mu}_B/\tilde{\mu}_A = \frac{K_B\mu_B}{K_A\mu_A}
 \end{equation}
@@
 The functions $\mathcal{B}_A(P_B,R)$ and $\mathcal{B}_B(P_A,R)$ capture the resource dependent growth.
-Setting $\mathcal{B}_A(P_B,R)=\alpha_A P_BR$, $\mathcal{B}_B(P_A,R)=\alpha_B P_A R$, $\mathcal{F}_A(R,G_A)=\kappa_A RG_A$, $\mathcal{F}_B(R,G_B)=\kappa_B RG_B$, rescaling $G_A$, $G_B$, and $P_B$ with $N_t$, $P_A$ with $\kappa_A/\alpha_A$, time with $\alpha_AN_t^2$, defining $\psi=\kappa_A\alpha_B/(\alpha_A^2N_t)$, and $\phi=\kappa_B/(\alpha_A N_t)$ results in
+Setting $\mathcal{B}_A(P_B,R)=\mu_A P_BR$, $\mathcal{B}_B(P_A,R)=\mu_B P_A R$, $\mathcal{F}_A(R,G_A)=\kappa_A RG_A$, $\mathcal{F}_B(R,G_B)=\kappa_B RG_B$, rescaling $G_A$, $G_B$, and $P_B$ with $N_t$, $P_A$ with $\kappa_A/\mu_A$, time with $\mu_AN_t^2$, defining $\psi=\kappa_A\mu_B/(\mu_A^2N_t)$, and $\phi=\kappa_B/(\mu_A N_t)$ results in
 \begin{align}
@@
 Both of these species grow using the same resource pool $R$ but with different growth constants
-$\alpha_{A/B}$ and different lag phase transition rates $\lambda_{A/B}$.
+$\mu_{A/B}$ and different lag phase transition rates $\lambda_{A/B}$.
 Moreover, species $A$ is able to release a certain antimicrobial component $I$ that shows an
@@
         \dot{L}_A &= -\lambda_A R L_A\\
-        \dot{G}_A &= \lambda_A R L_A + \alpha_A R G_A\\
+        \dot{G}_A &= \lambda_A R L_A + \mu_A R G_A\\
         \dot{L}_B &= -\lambda_B R L_B\\
-        \dot{G}_B &= \lambda_B R L_B + \frac{\alpha_B}{1 + \omega_I I} R G_B.
+        \dot{G}_B &= \lambda_B R L_B + \frac{\mu_B}{1 + K_I I} R G_B.
     \end{split}
@@
     \begin{split}
-        \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \omega_I I) N_t} R G_B \\
+        \dot{R} &= -\frac{\mu_A}{N_t} R G_A-\frac{\mu_B}{(1 + K_I I) N_t} R G_B \\
         \dot{I} &= \kappa G_A.
@@
     \dot{L}_B &= -\lambda_B R L_B\\
-    \dot{G}_A &= \alpha_A R G_A + \lambda_A R L_A\\
-    \dot{G}_B &= \frac{\alpha_B}{1 + \omega_I I} R G_B + \lambda_B R L_B\\
-    \dot{R} &= -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{(1 + \omega_I I) N_t} R G_B\\
+    \dot{G}_A &= \mu_A R G_A + \lambda_A R L_A\\
+    \dot{G}_B &= \frac{\mu_B}{1 + K_I I} R G_B + \lambda_B R L_B\\
+    \dot{R} &= -\frac{\mu_A}{N_t} R G_A-\frac{\mu_B}{(1 + K_I I) N_t} R G_B\\
     \dot{I} &= \kappa G_A
@@
         N_t &= [R_0]V_D\sigma\\
-        \omega_I &= \frac{\nu_I}{V_D}\\
-        \alpha &= u[R_0]\sigma\\
+        K_I &= \frac{\nu_I}{V_D}\\
+        \mu &= u[R_0]\sigma\\
         \kappa &= \sigma.
@@
         -(\omega + \lambda) & 0               & 0      \\
-        \lambda          & \alpha R - \omega' & 0      \\
+        \lambda          & \mu R - \omega' & 0      \\
         \omega              & \omega'            & -\beta
     \end{pmatrix} \mbx\\
-    \dot \mbm &= \mbH (\mbx, \mbm, \mbp, \mbu) = -\frac{\alpha}{N_t} R G.
+    \dot \mbm &= \mbH (\mbx, \mbm, \mbp, \mbu) = -\frac{\mu}{N_t} R G.
 \end{align}
@@
         -(\omega + \lambda) & 0       & 0 \\
-         \lambda         & \alpha  & 0 \\
+         \lambda         & \mu  & 0 \\
          \omega &  \omega'& - \beta
     \end{pmatrix} 
-    \mbx + \frac{\alpha}{N_t} x^2 \sum_i^3 x^i \begin{pmatrix} 0 \\ 1 \\ 0  \end{pmatrix}.
+    \mbx + \frac{\mu}{N_t} x^2 \sum_i^3 x^i \begin{pmatrix} 0 \\ 1 \\ 0  \end{pmatrix}.
 \label{eq:x_repres_3pools_resource2}
@@
         \toprule
-        Production $G_A$            & $\alpha_A$    & $0.002$~\unit{\per\minute}
+        Production $G_A$            & $\mu_A$    & $0.002$~\unit{\per\minute}
         & Uptake                    & $u_A$         & $0.01$~\unit{\per\minute}\\
-        Production $G_B$            & $\alpha_B$    & $0.002$~\unit{\per\minute}
+        Production $G_B$            & $\mu_B$    & $0.002$~\unit{\per\minute}
         & Uptake                    & $u_B$         & $0.01$~\unit{\per\minute}\\
@@
         & Production $I$            & $b$           & $0.1$~\unit{\per\minute}\\
-        Inhibition                  & $\omega_I$    & $1.0\cdot 10^{-7}$~\unit{\per\micro\metre\cubed}
-        &                           & $\nu_I$       & $0.1$~\unit{\micro\metre\cubed\per\mol}\\
+        Inhibition                  & $K_I$         & $1.0\cdot 10^{-7}$~\unit{\per\micro\metre\cubed}
+        & Inhibition                & $\nu_I$       & $0.1$~\unit{\micro\metre\cubed\per\mol}\\
         &&

### 2025-11-28 paper/pool_model.tex
fix supplement references
add reformulations, citations and other comments by fernando

@@
 
-\newcommand{\beginsupplement}{%
-  \renewcommand{\thesection}{}
-  \setcounter{section}{0}
-  \renewcommand{\thesection}{S\arabic{section}}
-  % \setcounter{table}{0}
-  \renewcommand{\thetable}{S\arabic{table}}%
-  % \setcounter{figure}{0}
-  \renewcommand{\thefigure}{S\arabic{figure}}%
-}
 \newcommand{\mbx}{\mathbf{x}}
@@
 virtually omnipresent around any living system.
-They play an especially important role within the context of food safety and production
-being a key contributor to the spoilage of food products as well as fermentation processes.
-Furthermore, effects such as biofilm formation can hinder drug treatment by shielding the bacterial
-colony from antibiotic substances or immune response of the organism.
-The natural bacterial processes are a crucial part of our planet's ecosystem, being responsible for
-tasks such as nutrient cycling, decomposition of organic matter and oxygen production.
+In particular, food matrices represent highly relevant microbial ecosystems, where these communities
+play a key role in both food safety and spoilage as well as fermentation processes.
+Biofilms are biological structures formed by complex and dynamic microbial communities adhered to
+surfaces that develop different communication mechanisms (e.g. quorum sensing), and coordinate to
+improve nutrient accessibility and provide protection against antimicrobial chemicals and physical
+stress~\citep{Zhao2023}.
+These structures are an important problem in food environments, and healthcare settings, being a
+source of contamination and increased risk of infection.
+% Furthermore, effects such as biofilm formation can hinder drug treatment by shielding the bacterial
+% colony from antibiotic substances or immune response of the organism.
+In addition, the natural bacterial processes are a crucial part of our planet's ecosystem, being
+responsible for tasks such as nutrient cycling, decomposition of organic matter and oxygen production.
 It is of great interest to understand the dynamics of these various processes.\\
@@
 all relevant biological phenomena arising in these bacterial systems.
-During their reproductive process, bacteria undergo multiple phases~\citep{buchanan_when_1997}, starting in a lag phase which
-is necessary to adapt to new environments and initially impedes the overall growth~\citep{rolfe_lag_2012, monod_growth_1949}.
+% During their reproductive process, bacteria undergo multiple phases~\citep{buchanan_when_1997}, starting in a lag phase which
+% is necessary to adapt to new environments and initially impedes the overall growth~\citep{rolfe_lag_2012, monod_growth_1949}.
+During the growth process, bacteria undergo different physiological and metabolic
+states~\citep{buchanan_when_1997}, that determine growth dynamics and response to the environment.
+Typically, when bacteria colonize a new environment, they enter an adaptation phase, commonly
+referred to as lag time~\citep{rolfe_lag_2012, monod_growth_1949}.
+Under this phase, bacteria prepare the enzymatic and metabolic machinery to initiate cell division
+and multiplication.
+According to Baranyi and Roberts~\citep{baranyi_dynamic_1994}, lag time can
+be defined as a function of the concentration of a limiting component that cells should synthesise
+in order to start growing.
 Additionally, bacteria are able to cooperate, e.g., via complementary metabolic processes or inhibit
@@
 environmental conditions.
-Widely used examples include the modified Gompertz model~\citep{Gompertz1825-wi,Gibson1987}, Baranyi-Roberts
-model~\citep{baranyi_modeling_1993, baranyi_dynamic_1994} and McKellar two-compartment model~\cite{buchanan_when_1997} which distinguishes between growing and non-growing bacteria.
+Widely used examples include the modified Gompertz model~\citep{Zwietering1990,Gompertz1825-wi,Gibson1987}, Baranyi-Roberts
+model~\citep{baranyi_modeling_1993, baranyi_dynamic_1994} and McKellar~\cite{McKellar1997} two-compartment model~\cite{buchanan_when_1997} which distinguishes between growing and non-growing bacteria.
 To take into account changing environmental conditions, so-called \textit{secondary models}
@@
 The former represent an intermediate step between the population-based \ac{ode} models and the
-individually-based \acp{abm} and can be effective, especially when dealing with large crowds of
+individually-based \acp{abm} and can be effective, especially when dealing with large population of
 cells that can be approximated as a continuum.
@@
 We begin our discussion by considering a simplistic system where only one species is present $N=1$.
-Bacterial growth undergoes different consecutive phases~\citep{buchanan_when_1997}, starting with the so called lag-phase, during which no increase in the bacterial count occurs.
+Bacterial growth undergoes different consecutive phases~\citep{buchanan_when_1997}, starting with the so called lag-phase, which is a specific case of quiescence, during which no increase in the bacterial count occurs.
 This phenomenon appears due to adaptations of the microbial community, for example, to the new environmental conditions.
@@
 For now, we assume no back flow from the growth state $G$ to the lag-state $L$.
+It is a logical assumption of the speicific case of cell adaption if environmental conditions do not
+change.
 The suggested model can be cast in the following reactions:
@@
 According to the works of \citet{monod_growth_1949} in the system where all needed nutrients are in excess except for one, the total growth is linear to the concentration of this limiting nutrient resource.
-The control of the population density with respect to the available nutrition by means of cell communication is called quorum sensing.
-By releasing extracellular signalling molecules (autoinducers) it is possible for the bacteria to synchronize the gene expression of the whole group~\citep{ng_bacterial_2009}.
-Instead of modeling complex signalling networks, we choose a simpler approach to describe the dependence of growth on the nutrient availability.
+% The control of the population density with respect to the available nutrition by means of cell communication is called quorum sensing.
+We choose a minimalistic approach to describe the dependence of growth on the nutrient availability.
 We include a resource pool which is consumed by the growing bacteria using the second order kinetic reaction:
@@
 Note that the previously introduced model could be further simplified to a two pool or state model by omitting the pool $D$ and removing cells directly from pools $L$ and $G$.
-The relevant question here is whether cells in the pool $D$ still contribute to the limitation of the resources by, e.g., occupying space or consuming important nutrients.
+The relevant question here is whether cells in the pool $D$ provide any significant impact on resource availability, e.g., by freeing previously occupied spaces or releasing previously consumed nutrients.
 If yes, the three pool model would be advantageous, if not the two pool model is sufficient with the benefit of having one parameter less.\\
@@
 For the functions $L$ and $D$ it holds: $\lim_{t\to\infty} L(t) = 0$ and  $\lim_{t\to\infty} D(t) =  N_0\frac{\omega}{\omega+\lambda} $.
-It follows immediately that $\lim_{t\to\infty} G(t) = N_t-N_0\frac{\omega}{\omega+\lambda} $.
+It follows immediately that $\lim_{t\to\infty} G(t) = N_t-N_0\frac{\omega}{\omega+\lambda}$.
 The lag time $t_L$ can be approximated by the time at which half of the initial population is in the growth phase, i.e., $N_0=2G(t_L)$.
+This approach differs to other concepts of lag time such as the one by Gompertz (visual lag time) or Baranyi (work to be done).
 Assuming $L\ll N_t$, we can ignore the non-linear term in the equation for $G$, which gives rise to:
@@
 Bacteria which have lived under rather unfavorable conditions have developed different surviving mechanisms, one of which is adapting their metabolism and transitioning into dormant state.
-Dormancy in bacteria is a reversible state where cell continue living with low metabolic activity while the cell division stops~\citep{kaprelyants_dormancy_1993}.
+Dormancy in bacteria is a reversible state where cells continue living with low metabolic activity while the cell division stops~\citep{kaprelyants_dormancy_1993}.
 This survival strategy can appear in response to different external stresses, e.g., nutrient deficiency, temperature stress or extreme fluctuations of environment parameters.
@@
 Mutual inhibition is often a result of active competition mechanisms which have been developed by bacteria to contend for a shared resource.
-Some examples of such mechanisms are the production of antimicrobial compounds, inhibitors or toxins~\citep{wloch-salamon_effect_2008, chao_structured_1981}.
+Some examples of such mechanisms are the production of antimicrobial compounds, inhibitors (section~\ref{section:toxin-production}) or toxins~\citep{wloch-salamon_effect_2008, chao_structured_1981}.
 On the other hand, some bacteria can produce public goods, the compounds that benefit growth of other species.
@@
 \subsubsection{Toxin production}
+\label{section:toxin-production}
 One direct form of inhibition is the production of toxins.
@@
 We see that in the highly-diffusive homogenous case of subfigure (A), the pool model is a good fit to the \ac{abm} model.
+It means that ODE could still be suitable for aqueous environments, where homogeneity is more probable.
 Heterogeneity and slower diffusion cause larger differences between the two models as can be seen in subfigures (B) and (C).
@@
 As an example, the description of the quorum sensing phenomena can be modelled by adding the equation for the released autoinducer molecules as a part of the microenvironment.
+By releasing these extracellular signalling molecules (autoinducers) it is possible for the bacteria to synchronize the gene expression of the whole group~\citep{ng_bacterial_2009}.
 These molecules then affect the same-species bacteria growth controlling the size of the population.
@@
 \newpage
-\beginsupplement
+
+\newcounter{suppsection}
+\newcounter{supptable}
+\newcounter{suppfigure}
+\newcounter{suppequation}
+
+\renewcommand{\thesection}{S\arabic{suppsection}}
+\renewcommand{\thetable}{S\arabic{supptable}}%
+\renewcommand{\thefigure}{S\arabic{suppfigure}}%
+\renewcommand{\theequation}{S\arabic{suppequation}}%
+
+\let\oldsection\section
+\let\oldequation\equation
+\let\oldtable\table
+\let\oldfigure\figure
+% \renewcommand{\section}{\@ifstar{\@oldsection}{\@stepcounter{suppsection}\@oldsection}}
+\makeatletter
+\renewcommand{\section}{\@ifstar{\oldsection*}{\stepcounter{suppsection}\oldsection}}
+\renewcommand{\equation}{\@ifstar{\oldequation*}{\stepcounter{suppequation}\oldequation}}
+\renewcommand{\table}{\@ifstar{\oldtable*}{\stepcounter{supptable}\oldtable}}
+\renewcommand{\figure}{\@ifstar{\oldfigure*}{\stepcounter{suppfigure}\oldfigure}}
+\makeatother
+
+\addcontentsline{toc}{Section}{Supplementary Material}% Add section to ToC
+
 % ================================================================================================

### 2025-11-28 pool_paper_plotting_code/examples/main_paper.py
Update model parameteters for 3 pool model plotting

@@ -190,26 +190,26 @@ if __name__ == "__main__":
 
 ###############33############### 3 pool models with resource pool ##############################3########333
     # Generate model data
-    const1 = [1e-2, 0., 2.5, .5, 0., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
-    const2 = [1e-3, 0., 2.5, .5, 0., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
-    const3 = [5e-5, 0., 2.5, .5, 0., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
-    const4 = [1e-3, .5, 2.5, .5, 0., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
+    const1 = [1e-2, 0., 2.5, .1, 0., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
+    const2 = [1e-3, 0., 2.5, .1, 0., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
+    const3 = [5e-5, 0., 2.5, .1, 0., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
+    const4 = [1e-3, .15, 2.5, .1, 0., Nt] # (lambd, mu1, alpha, mu2, beta, N_t, )
     const_3pool_res = [const1, const2, const3, const4]
     data_3pool_res = [generate_insilico_data(pool_model_3pools_resource, [t], [], [[]], x01,
                 const=const, obs_func=observable_3pool, n_traj=1) for const in const_3pool_res]
 
-    labels = [r'$\lambda=10^{-2}, \mu=0$', r'$\lambda=10^{-3}, \mu=0$', r'$\lambda=10^{-4}, \mu=0$', r'$\lambda=10^{-3}, \mu=0.5$']
+    labels = [r'$\lambda=10^{-2}, \omega=0$', r'$\lambda=10^{-3}, \omega=0$', r'$\lambda=10^{-4}, \omega=0$', r'$\lambda=10^{-3}, \omega=0.15$']
     color_palette_1sp = [colors_all['N_lambd_1e-3_mu_0'], colors_all['N_wo_tempshift'], colors_all['N_lambd_1e-2_mu_0'], colors_all['N_lambd_1e-3_mu_0_5']]
-
+    lnst = ['solid', 'solid', 'solid', 'dotted']
     # Plot the model
     fig, ax = plt.subplots(1, 1, figsize=figsize_default)
[…]

### 2025-12-08 paper/pool_model.tex
add fernando to authors
add orcids; remove website of jonas
Add corrections from Christian

@@
 
-\title{Pool Model - A unified Mathematical\\ Modeling Framework for Food Science}
+\title{Pool Model - A Mathematical Framework for Modelling Bacterial Consortia in Food Science}
 \author[1]{Polina Gaindrik\orcidlink{0000-0003-0928-3951}}
-\author[1]{Jonas Pleyer\fnref{fn1}}
-\author[1]{Christian Fleck\corref{cor1}}
+\author[1]{Jonas Pleyer\orcidlink{0000-0003-0928-3951}}
+\author[2]{Fernando Pérez Rodríguez}
+\author[1]{Christian Fleck\orcidlink{0000-0002-6371-4495}\corref{cor1}}
 \ead{christian.fleck@fdm.uni-freiburg.de}
 
-\fntext[fn1]{jonas.pleyer.org}
 \cortext[cor1]{Corresponding Author\\ \textit{E-mail address:} christian.fleck@fdm.uni-freiburg.de}
@@
     country={Germany},
+}
+\affiliation[2]{
+    organization={Department of Food Science and Technology, UIC Zoonosis y Enfermedades Emergentes ENZOEM, ceiA3, Universidad de Córdoba},
+    addressline={Campus Rabanales s/n. Darwin building- Annex},
+    city={Córdoba},
+    postcode={14014},
+    country={Spain},
 }
@@
 The rate $\gamma(t)$ depends on the rate of change of the environment.
-Stronger and quicker changes will lead to a higher rate $\gamma$.
+Stronger and quicker changes will lead to a higher rate $\gamma$, but the effect will relax to zero after some time.
 %
@@
 \begin{equation}
-    \dot{\gamma} = \Gamma |\dot{T}|-\delta \gamma.
-\end{equation}
-We assumed that the direction of the temperature change does not matter, i.e., a change from $T=2$ to $T=12$ has the same effect as a change from $T=14$ to $T=4$.
+    \dot{\gamma} = \Gamma f\left(\dot{T}\right)-\delta \gamma.
+\end{equation}
+The dimensionless function $f$, with $f(0)=0$ and $f: \mathbb{R}\to\mathbb{R}_{\ge 0}$, evaluates the effect of a temperature change, i.e., a change from $T=2$ to $T=12$ could have a different effect as a change from $T=14$ to $T=4$.
 $\Gamma$ is a scaling factor and $\delta$ is the relaxation time for the temperature disturbance.
@@
 \begin{equation}
-    \gamma(t) = \Gamma \int_0^t |\dot{T}(t^{\prime})|e^{-\delta (t-t^{\prime})}dt^{\prime}.
-\end{equation}
-After some manipulation of this equation we find:
-\begin{equation}
-    \gamma(t) = \Gamma\delta\left |T[t]-T[0]e^{-\delta t}-\delta
-        \int\limits_0^t T(t^{\prime})e^{-\delta (t-t^{\prime})}dt^{\prime}\right |.
-\end{equation}
-This representation of the rate $\gamma(t)$ has the advantage that it only involves $T(t)$ and does not require the calculation of the derivative of $T$.
+    \gamma(t) = \Gamma \int_0^t f\left(\dot{T}(t^{\prime})\right)e^{-\delta (t-t^{\prime})}dt^{\prime}.
+\end{equation}
+One possible choice for $f$ reads:
+\begin{align}
+f(x)&=\begin{cases}\alpha x& x\ge 0\\ -(1-\alpha) x& x < 0,\end{cases}
+\end{align}
+$\alpha\in [0,1]$. For $\alpha=0.5$ it follows $f(x)=\frac{1}{2}|x|$. Using this choice for $f$ we find after some manipulation:
+\begin{equation}
+    \gamma(t) = \Gamma\alpha\left [T(t)-T(0)e^{-\delta t}-\delta
+        \int\limits_0^t T(t^{\prime})e^{-\delta (t-t^{\prime})}dt^{\prime}\right ],
+\end{equation}
+for $\dot{T}\ge 0, \,\forall t$ and
+\begin{equation}
+    \gamma(t) = \Gamma(1-\alpha)\left [T(0)e^{-\delta t}-T(t)+\delta
+        \int\limits_0^t T(t^{\prime})e^{-\delta (t-t^{\prime})}dt^{\prime}\right ].
+\end{equation}
+for $\dot{T}< 0, \,\forall t$. 
+This representation of the rate $\gamma(t)$ has the advantage that it only involves $T(t)$ and does not require the calculation of a derivative.
 If $T$ exhibits $n$ temperature jumps, one finds:
 \begin{eqnarray}
-    \gamma(t) &=& \Gamma\sum_{i=1}^n \left |\Delta T_i \right |\theta(t-t_i)e^{-\delta(t-t_i)}.
+    \gamma(t) &=& \Gamma\sum_{i=1}^n f\left(\Delta T_i \right)\theta(t-t_i)e^{-\delta(t-t_i)}.
 \label{eq:gamma_tempshift}

### 2025-12-11 paper/pool_model.tex
Christian's changes to maxwell stress section

@@
 \begin{equation}
-    \dot{\gamma} = \Gamma f\left(\dot{T}\right)-\delta \gamma.
+    \dot{\gamma} = \sigma f\left(\zeta\dot{T}\right)-\delta \gamma.
 \end{equation}
 The dimensionless function $f$, with $f(0)=0$ and $f: \mathbb{R}\to\mathbb{R}_{\ge 0}$, evaluates the effect of a temperature change, i.e., a change from $T=2$ to $T=12$ could have a different effect as a change from $T=14$ to $T=4$.
-$\Gamma$ is a scaling factor and $\delta$ is the relaxation time for the temperature disturbance.
+$\zeta$ is a scaling factor, $\sigma$ is a kind of stiffness telling how strongly the bacteria react to temperature stress, and $\delta^{-1}$ is the relaxation time for the temperature disturbance.
 Integration of this equation yields (with $\gamma(0)=0$):
 \begin{equation}
-    \gamma(t) = \Gamma \int_0^t f\left(\dot{T}(t^{\prime})\right)e^{-\delta (t-t^{\prime})}dt^{\prime}.
+    \gamma(t) = \sigma\int_0^t f\left(\zeta\dot{T}(t^{\prime})\right)e^{-\delta (t-t^{\prime})}dt^{\prime}.
 \end{equation}
@@
 \begin{equation}
-    \gamma(t) = \Gamma\alpha\left [T(t)-T(0)e^{-\delta t}-\delta
+    \gamma(t) = \sigma\alpha\zeta\left [T(t)-T(0)e^{-\delta t}-\delta
         \int\limits_0^t T(t^{\prime})e^{-\delta (t-t^{\prime})}dt^{\prime}\right ],
@@
 \begin{equation}
-    \gamma(t) = \Gamma(1-\alpha)\left [T(0)e^{-\delta t}-T(t)+\delta
+    \gamma(t) = \sigma(1-\alpha)\zeta\left [T(0)e^{-\delta t}-T(t)+\delta
         \int\limits_0^t T(t^{\prime})e^{-\delta (t-t^{\prime})}dt^{\prime}\right ].
@@
 \begin{eqnarray}
-    \gamma(t) &=& \Gamma\sum_{i=1}^n f\left(\Delta T_i \right)\theta(t-t_i)e^{-\delta(t-t_i)}.
+    \gamma(t) &=& \sigma\sum_{i=1}^n f\left(\zeta\Delta T_i \right)\theta(t-t_i)e^{-\delta(t-t_i)}.
 \label{eq:gamma_tempshift}
@@
         and the series of different temperature shifts of value $10^\circ C$, $5^\circ C$ and $15^\circ C$ at the time points $1, 3$ and $7$, respectively (rose).
-        The parameters of the back-flow rate are $\Gamma=1^\circ C / s$ and $\delta=5~\mathrm{s}^{-1}$.
+        The parameters of the back-flow rate are $\zeta=1 s/K $ and $\delta=5~\mathrm{s}^{-1}$.
         The other model parameters are $\mu=2.5~\mathrm{s}^{-1}$, $\omega'=0.5~\mathrm{s}^{-1}$, $\beta=5~\mathrm{s}^{-1}$, $\lambda=10^{-2}~\mathrm{s}^{-1}$, $\omega = 0~\mathrm{s}^{-1}$ for each curve.

### 2025-12-11 paper/pool_model.tex
Replace seconds units to hours

@@
 Then $R=N_R/V$ and $[R]=\unit{1/V}$, where $N_R$ is the number of resource molecules (or any appropriate unit, e.g., \unit{mol}).
-Then $[\mu_0]=\unit{V/s}$, $[N_t]=\unit{1/V}$, and $[\mu]= \unit{1/s}$.
+Then $[\mu_0]=\unit{V/h}$, $[N_t]=\unit{1/V}$, and $[\mu]= \unit{1/h}$.
 $N_t$ denotes the maximal size of the total bacterial population due to environmental limitations.
@@
         The time dependence of the bacterial count $N(t) = L(t) + G(t) $ is shown.
-        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\mu=2.5~\mathrm{s}^{-1}$, $\omega'=0.1~\mathrm{s}^{-1}$, $N_t = 10^4$.
-        The blue curve includes cell-death of the bacteria in the lag phase $\omega = 0.15~\mathrm{s}^{-1}$ and thus exhibits a small dip before its exponential growth.
-        Other curves show the case of no bacteria dying $\omega=0~\mathrm{s}^{-1}$ and represent different lengths of lag phases: short $\lambda=10^{-2}~\mathrm{s}^{-1}$ (dark green), intermediate $\lambda=10^{-3}~\mathrm{s}^{-1}$ (light green) and long one $\lambda=10^{-4}~\mathrm{s}^{-1}$ (yellow).
+        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\mu=2.5~\mathrm{h}^{-1}$, $\omega'=0.1~\mathrm{h}^{-1}$, $N_t = 10^4$.
+        The blue curve includes cell-death of the bacteria in the lag phase $\omega = 0.15~\mathrm{h}^{-1}$ and thus exhibits a small dip before its exponential growth.
+        Other curves show the case of no bacteria dying $\omega=0~\mathrm{h}^{-1}$ and represent different lengths of lag phases: short $\lambda=10^{-2}~\mathrm{h}^{-1}$ (dark green), intermediate $\lambda=10^{-3}~\mathrm{h}^{-1}$ (light green) and long one $\lambda=10^{-4}~\mathrm{h}^{-1}$ (yellow).
 Note that the steady state for $N(t)$ and $D(t)$ is for this model $\lim_{t\to\infty} N(t)=0$ and $\lim_{t\to\infty} D(t)=N_t$, because of the depletion of the resource pool.
@@
         (B) The value of the corresponding back-flow rate $\gamma$ calculated according to (\ref{eq:gamma_tempshift}).
-        The curves represent no temperature shift (green), one temperature jump $\Delta T = 10^\circ C$ at time $t=4~\mathrm{s}$ (orange)
+        The curves represent no temperature shift (green), one temperature jump $\Delta T = 10^\circ C$ at time $t=4~\mathrm{h}$ (orange)
         and the series of different temperature shifts of value $10^\circ C$, $5^\circ C$ and $15^\circ C$ at the time points $1, 3$ and $7$, respectively (rose).
-        The parameters of the back-flow rate are $\zeta=1 s/K $ and $\delta=5~\mathrm{s}^{-1}$.
-        The other model parameters are $\mu=2.5~\mathrm{s}^{-1}$, $\omega'=0.5~\mathrm{s}^{-1}$, $\beta=5~\mathrm{s}^{-1}$, $\lambda=10^{-2}~\mathrm{s}^{-1}$, $\omega = 0~\mathrm{s}^{-1}$ for each curve.
+        The parameters of the back-flow rate are $\zeta=1 s/K $ and $\delta=5~\mathrm{h}^{-1}$.
+        The other model parameters are $\mu=2.5~\mathrm{h}^{-1}$, $\omega'=0.5~\mathrm{h}^{-1}$, $\beta=5~\mathrm{h}^{-1}$, $\lambda=10^{-2}~\mathrm{h}^{-1}$, $\omega = 0~\mathrm{h}^{-1}$ for each curve.
     }
@@
         The competition of the two bacterial species due to the common limiting resource calculated using equations~(\ref{eq:model_2sp_resource_comp}).
-        The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}~\mathrm{s}^{-1}$, $\mu_A=3.0~\mathrm{s}^{-1}$.
-        The blue curve describes the growth of the bacteria $B$ ($N_B = L_B+G_B$) with model parameters $\lambda_B=5\cdot 10^{-2}~\mathrm{s}^{-1}$, $\mu_B=2.5~\mathrm{s}^{-1}$.
+        The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}~\mathrm{h}^{-1}$, $\mu_A=3.0~\mathrm{h}^{-1}$.
+        The blue curve describes the growth of the bacteria $B$ ($N_B = L_B+G_B$) with model parameters $\lambda_B=5\cdot 10^{-2}~\mathrm{h}^{-1}$, $\mu_B=2.5~\mathrm{h}^{-1}$.
         The purple line shows the total bacterial count of the system  $N(t)=N_A+N_B$.
@@
         The growing bacteria $B$ produces toxin $T_B$ (A) constantly $\dot{T}_B^1 = \mathcal{F}^1 = \kappa G_B$ or (B) only during growth process $\dot{T}_B^2 = \mathcal{F}^2 = \kappa R G_B$ 
-        with production rate $\kappa=0.2~\mathrm{s}^{-1}$.
-        The toxin $T_B$ causes the cell death of bacteria $A$ with rate constant $\omega = 0.2~\mathrm{s}^{-1}$.
-        The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}~\mathrm{s}^{-1}$, $\mu_A=3.0~\mathrm{s}^{-1}$.
-        The blue curve describes the growth of the bacteria $B$ ($N_B = L_B+G_B$) with model parameters $\lambda_B=5\cdot 10^{-2}~\mathrm{s}^{-1}$, $\mu_B=2.5~\mathrm{s}^{-1}$.
+        with production rate $\kappa=0.2~\mathrm{h}^{-1}$.
+        The toxin $T_B$ causes the cell death of bacteria $A$ with rate constant $\omega = 0.2~\mathrm{h}^{-1}$.
+        The red curve describes the growth of the bacteria $A$ ($N_A = L_A+G_A$) with model parameters $\lambda_A=10^{-2}~\mathrm{h}^{-1}$, $\mu_A=3.0~\mathrm{h}^{-1}$.
+        The blue curve describes the growth of the bacteria $B$ ($N_B = L_B+G_B$) with model parameters $\lambda_B=5\cdot 10^{-2}~\mathrm{h}^{-1}$, $\mu_B=2.5~\mathrm{h}^{-1}$.
         The gray line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
@@
 We introduced the rescaled growth rates $\tilde{\mu}_A=\mu_A/(N_tK_B)$ and $\tilde{\mu}_B=\mu_B/(N_tK_A)$.
-Note that $[\tilde{\mu}]=\mathrm{s}^{-1}$.
+Note that $[\tilde{\mu}]=\mathrm{h}^{-1}$.
 We rescale the time with $\tilde{\mu}_A$ and define the dimensionless parameter

### 2025-12-15 pool_paper_plotting_code/digital_twin/model.py
plots code: rename parameters accordung to paper notation

@@ -10,21 +10,6 @@ def pool_model(t, x, param, x0, const):
     (x1, ) = x
     return [alpha * (x1 - x01 * np.exp(- lambd * t)) * (1 - x1 / n_max)]
 
-# Define the pool model ODEs for 2 species model with resource competition between species
-# With the waste production ofspecies 1, with waste-induced death of the species 2 
-def pool_model_2species_resource_waste(t, x, param, x0, const):
-    (L1, L2, G1, G2, R, W1,) = x
-    (alph1, alph2, lambd1, lambd2, chi1, muW2, Nt, ) = const
-    
-    return [
-        - lambd1 * R * L1,
-        - lambd2 * R * L2,
-          lambd1 * R * L1 + alph1 * R * G1,
-          lambd2 * R * L2 + alph2 * R * G2 - muW2/(chi1*Nt) * W1 * G2,
-        - alph1/Nt * R * G1 - alph2/Nt * R * G2,
-          chi1 * G1
-    ]
-
 
 # The model solution for given ODEs
 def model_ODE_solution(model, t, param, x0, const):

### 2025-12-15 pool_paper_plotting_code/digital_twin/plotting_templates.py

@@ -10,9 +10,9 @@ colors_all = {
         'N':'#7E57A5',
         'T':'#99582A',
         'T_A':'#c79758',
-        'N_lambd_1e-2_mu_0':'#E2B100',
-        'N_lambd_1e-3_mu_0':'#386641',
-        'N_lambd_1e-3_mu_0_5':'#0982A4',
+        'N_lambd_1e-2_omega_0':'#E2B100',
+        'N_lambd_1e-3_omega_0':'#386641',
+        'N_lambd_1e-3_omega_0_5':'#0982A4',
         'N_wo_tempshift':'#679E48',
         'N_tempshift_10':'#ED733E',
         'N_tempshift_10_5_15':'#C3568A',

### 2025-12-15 pool_paper_plotting_code/examples/main_paper.py

@@ -10,61 +10,61 @@ from digital_twin import *
 
 
 def pool_model_3pools(t, x, param, x0, const):
-    (lambd, mu1, alpha, mu2, beta, ) = const
+    (lambd, omega1, mu, omega2, ) = const
     (L0, G0, D0, ) = x0
     (L, G, D, ) = x
 
     return [
-        - (mu1 + lambd) * L,
-                  lambd * L + (alpha - mu2) * G,
-                    mu1 * L +           mu2 * G - beta * D
+        - (omega1 + lambd) * L,
+        lambd * L + (mu - omega2) * G,
+        omega1 * L + omega2 * G
         ]
 
 def pool_model_3pools_resource(t, x, param, x0, const):
-    (lambd, mu1, alpha, mu2, beta, N_t, ) = const
+    (lambd, omega1, mu, omega2, N_t, ) = const
     (L0, G0, D0, ) = x0
     (L, G, D, ) = x
 
     return [
-        - (mu1 + lambd) * L,
-                  lambd * L + (alpha - alpha * ((L + G + D)/N_t) - mu2) * G,
-                    mu1 * L +                                       mu2 * G - beta * D
+        - (omega1 + lambd) * L,
+        lambd * L + (mu - mu * ((L + G + D)/N_t) - omega2) * G,
+        omega1 * L + omega2 * G
         ]
 
 def pool_model_2pools_resource_lag(t, x, param, x0, const):
-    (lambd, alpha, N_t, ) = const
+    (lambd, mu, N_t, ) = const
     (L0, G0, R0, ) = x0
     (L, G, R ) = x
 
     return [
         - lambd * R * L,
-          lambd * R * L + alpha * R * G,
-                - (alpha / N_t) * R * G
+        lambd * R * L + mu * R * G,
+        - (mu / N_t) * R * G
         ]
 
 def pool_model_2pools_resource_lag2(t, x, param, x0, const):
-    (lambd, alpha, N_t, ) = const
+    (lambd, mu, N_t, ) = const
     (L0, G0, ) = x0
     (L, G, ) = x
     r_term = 1 - (L+G)/N_t
     return [
         - lambd * r_term * L,
-          lambd * r_term * L + alpha * r_term * G,
[…]

### 2025-12-15 pool_paper_plotting_code/examples/main_paper.py
plot code: add dormant state ODE model

@@ -53,7 +53,16 @@ def pool_model_2pools_resource_lag2(t, x, param, x0, const):
         ]
 
 def pool_model_dormant(t, x, param, x0, const):
-    pass
+    (lambd, omega1, mu, omega2, xi, N_t, Gamma, delta, shift_cnd,) = const
+    (L0, G0, D0, S0, ) = x0
+    (L, G, D, S, ) = x
+    gamma = temp_stress(t, Gamma, delta, shift_cnd)
+    return [
+        - (omega1 + lambd) * L,
+        lambd * L + (mu - mu * ((L + G + D)/N_t) - omega2 - gamma) * G + xi * S,
+        omega1 * L + omega2 * G,
+        gamma * G - xi * S
+        ]
 
 def pool_model_tempstress(t, x, param, x0, const):
     (lambd, omega1, mu, omega2, N_t, Gamma, delta, shift_cnd,) = const

### 2025-12-15 paper/pool_model.tex
Update Fig2 : add line with no death pool

@@
         The time dependence of the bacterial count $N(t) = L(t) + G(t) $ is shown.
-        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\mu=2.5~\mathrm{h}^{-1}$, $\omega'=0.1~\mathrm{h}^{-1}$, $N_t = 10^4$.
-        The blue curve includes cell-death of the bacteria in the lag phase $\omega = 0.15~\mathrm{h}^{-1}$ and thus exhibits a small dip before its exponential growth.
-        Other curves show the case of no bacteria dying $\omega=0~\mathrm{h}^{-1}$ and represent different lengths of lag phases: short $\lambda=10^{-2}~\mathrm{h}^{-1}$ (dark green), intermediate $\lambda=10^{-3}~\mathrm{h}^{-1}$ (light green) and long one $\lambda=10^{-4}~\mathrm{h}^{-1}$ (yellow).
+        All curves were calculated using the model (\ref{eq:ode_3pool_resource2}) with parameters $\mu=2.5~\mathrm{h}^{-1}$, $\omega=0~\mathrm{h}^{-1}$, $N_t = 10^4$.
+        The blue dashed curve shows the case of no cell death $\omega' = 0~\mathrm{h}^{-1}$ and thus exhibits a steady state without the decrease of the bacterial population.
+        Other curves represent the growing bacteria dying with the rate $\omega'=0.1~\mathrm{h}^{-1}$ and different lengths of lag phases: short $\lambda=10^{-2}~\mathrm{h}^{-1}$ (dark green), intermediate $\lambda=10^{-3}~\mathrm{h}^{-1}$ (light green) and long one $\lambda=10^{-4}~\mathrm{h}^{-1}$ (yellow).
         Note that the steady state for $N(t)$ and $D(t)$ is for this model $\lim_{t\to\infty} N(t)=0$ and $\lim_{t\to\infty} D(t)=N_t$, because of the depletion of the resource pool.

### 2025-12-15 pool_paper_plotting_code/examples/main_paper.py

@@ -187,24 +187,24 @@ if __name__ == "__main__":
     const1 = [1e-2, 0., 2.5, .1, Nt] # (lambd, omega1, mu, omega2, N_t, )
     const2 = [1e-3, 0., 2.5, .1, Nt] # (lambd, omega1, mu, omega2, N_t, )
     const3 = [5e-5, 0., 2.5, .1, Nt] # (lambd, omega1, mu, omega2, N_t, )
-    const4 = [1e-3, .15, 2.5, .1, Nt] # (lambd, omega1, mu, omega2, N_t, )
+    const4 = [1e-3, .0, 2.5, .0, Nt] # (lambd, omega1, mu, omega2, N_t, )
     const_3pool_res = [const1, const2, const3, const4]
     data_3pool_res = [generate_insilico_data(pool_model_3pools_resource, [t], [], [[]], x01,
                 const=const, obs_func=observable_3pool, n_traj=1) for const in const_3pool_res]
 
-    labels = [r'$\lambda=10^{-2}, \omega=0$', r'$\lambda=10^{-3}, \omega=0$', r'$\lambda=10^{-4}, \omega=0$', r'$\lambda=10^{-3}, \omega=0.15$']
+    labels = [r"$\lambda=10^{-2}, \omega'= 0.1$", r"$\lambda=10^{-3}, \omega'= 0.1$", r"$\lambda=10^{-4}, \omega'= 0.1$", r"$\lambda=10^{-3}, \omega'= 0$"]
     color_palette_1sp = [colors_all['N_lambd_1e-3_omega_0'], colors_all['N_wo_tempshift'], colors_all['N_lambd_1e-2_omega_0'], colors_all['N_lambd_1e-3_omega_0_5']]
-    lnst = ['solid', 'solid', 'solid', 'dotted']
+    lnst = ['solid', 'solid', 'solid', 'dashed']
     # Plot the model
     fig, ax = plt.subplots(1, 1, figsize=figsize_default)
     for i, data in enumerate(data_3pool_res):
         for d in data:
             ax.plot(d['times'], d['obs_mean'][0], linewidth=2.5, color=color_palette_1sp[i], label=labels[i], linestyle=lnst[i])
-    fig, ax = set_labels(fig, ax, r'Time, $t$ [h]', r'Total Bacterial Count, $N$')
[…]

### 2025-12-15 pool_paper_plotting_code/examples/main_paper.py
Add units to figure labels

@@ -231,7 +231,7 @@ if __name__ == "__main__":
     ax[1].plot(t, y, color=color_palette_1sp_tempshift[i-1])
     y2 = [temp_stress(tt, *const6[-3:]) for tt in t]
     ax[1].plot(t, y2, color=color_palette_1sp_tempshift[i])
-    fig, ax[0] = set_labels(fig, ax[0], r'', r'Bacterial Count, $N$')
+    fig, ax[0] = set_labels(fig, ax[0], r'', r'Bacterial Count, $N$ [CFU/mL]')
     fig, ax[1] = set_labels(fig, ax[1], r'Time, $t$ [h]', r'Rate, $\gamma$')
     ax[0].set_yscale('log')
     ax[0].set_xlim(0.7, 8.5)
@@ -246,7 +246,7 @@ if __name__ == "__main__":
     for i, data in enumerate(data_Tempshift):
         for d in data:
             ax.plot(d['times'], d['obs_mean'][0], color=color_palette_1sp_tempshift[i], label=labels[i])
-    fig, ax = set_labels(fig, ax, r'', r'Bacterial Count, $N$')
+    fig, ax = set_labels(fig, ax, r'', r'Bacterial Count, $N$ [CFU/mL]')
     ax.set_yscale('log')
     ax.set_xlim(0.7, 8.5)
     ax.set_ylim(6e0, 1.1e4)
@@ -275,7 +275,7 @@ if __name__ == "__main__":
     lnst = ['solid', 'solid', 'dashed', 'solid']
 
     fig, ax = plt.subplots(1, 1, figsize=figsize_default)
-    fig, ax = set_labels(fig, ax, r'Time, $t$ [h]', r'Bacterial Count, $N$')
+    fig, ax = set_labels(fig, ax, r'Time, $t$ [h]', r'Bacterial Count, $N$ [CFU/mL]')
     for d in data_gLV:
         for i, obs in enumerate(d['obs_mean']):
             if i == 3:
@@ -316,7 +316,7 @@ if __name__ == "__main__":
         ax[i].set_xlim(2., 9.0)
         ax[i].set_ylim(-100, 1.01e4)
         ax[i].legend()
-        fig, ax[i] = set_labels(fig, ax[i], r'Time, $t$ [h]', r'Bacterial Count, $N$')
+        fig, ax[i] = set_labels(fig, ax[i], r'Time, $t$ [h]', r'Bacterial Count, $N$ [CFU/mL]')
[…]

### 2025-12-22 pool_paper_plotting_code/examples/main_paper.py
update dormancy model code/plots

@@ -53,35 +53,40 @@ def pool_model_2pools_resource_lag2(t, x, param, x0, const):
         ]
 
 def pool_model_dormant(t, x, param, x0, const):
-    (lambd, omega1, mu, omega2, xi, N_t, Gamma, delta, shift_cnd,) = const
+    T_const = 1.
+    (lambd, omega1, mu, omega2, xi, N_t, sigma, delta, zeta, alpha, shift_cnd,) = const
     (L0, G0, D0, S0, ) = x0
     (L, G, D, S, ) = x
-    gamma = temp_stress(t, Gamma, delta, shift_cnd)
+    gamma = temp_stress(t, sigma, delta, zeta, alpha, shift_cnd)
     return [
-        - (omega1 + lambd) * L,
-        lambd * L + (mu - mu * ((L + G + D)/N_t) - omega2 - gamma) * G + xi * S,
+        - (omega1 + lambd*T_const) * L,
+        lambd*T_const * L + (mu*T_const - mu*T_const * ((L + G + D)/N_t) - omega2 - gamma) * G + xi * S,
         omega1 * L + omega2 * G,
         gamma * G - xi * S
         ]
 
 def pool_model_tempstress(t, x, param, x0, const):
-    (lambd, omega1, mu, omega2, N_t, Gamma, delta, shift_cnd,) = const
+    (lambd, omega1, mu, omega2, N_t, sigma, delta, zeta, alpha, shift_cnd,) = const
     (L0, G0, D0, ) = x0
     (L, G, D, ) = x
 
-    gamma = temp_stress(t, Gamma, delta, shift_cnd)
+    gamma = temp_stress(t, sigma, delta, zeta, alpha, shift_cnd)
     return [
         - (omega1 + lambd) * L + gamma * G,
         lambd * L + (mu - mu * ((L + G + D)/N_t) - omega2 - gamma) * G,
         omega1 * L + omega2 * G
         ]
 
-def temp_stress(t, Gamma, delta, shift_cnd):
+def temp_stress(t, sigma, delta, zeta, alpha, shift_cnd):
     #shift_cnd = ((t, T), (t, T))
     gamma = 0
     for tsh, Tsh in shift_cnd:
         if t >= tsh:
-            gamma += Gamma * np.abs(Tsh) * np.exp(-delta * (t - tsh))
+            if Tsh > 0:
+                f = alpha*zeta*Tsh
+            else:
[…]
